### 2015-12-29 draft.tex
add copy of draft010515.tex from exante folder in dropbox as draft.tex

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.3in}
\addtolength{\oddsidemargin}{-.65in} %left margin
\addtolength{\evensidemargin}{-.65in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
\linespread{1.5}

%dell laptop
\usepackage[backend=biber,block=space,style=authoryear-comp]{biblatex}

%campus computer
%\usepackage[backend=bibtex,block=space,style=authoryear]{biblatex}

%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}

%campus computer
%\addbibresource{C:/Apps-SU/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\usepackage{csquotes}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true,hyperfootnotes=false]{hyperref}
\usepackage[pdftex]{graphicx}

\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
\usepackage[hang,flushmargin]{footmisc} 
\usepackage[usenames,dvipsnames]{color}
%\usepackage{cite}
\usepackage{times, verbatim,bm,pifont,pdfsync}
%\usepackage[hang,flushmargin]{footmisc}%unindents footnotes

% disables chapter, section and subsection numbering
%\setcounter{secnumdepth}{-1}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{assumption}{Assumption}
\newtheorem{fact}{Fact}
\newtheorem{result}{Result}
\newtheorem{proposition}{Proposition}

\newcommand{\ve}{\theta}
\newcommand{\ta}{\theta}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\bte}{\bm{\tau^E}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}
\newcommand{\btw}{\bm{\tau^{tw}}}
\newcommand{\de}{\delta}

%This version was begun on May 10, 2014 for first public release (to ISNIE posting and first posting to my website); I anticipate those main changes will be some significant editing since first draft for Midwest was hurried, plus some expansion to escape clause section.

\begin{document}
\title{\vskip-0.6in Endogenous Politics and the Design of Trade Institutions}
\author{Kristy Buzard\thanks{Syracuse University, Economics Department. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard.}}
\date{\vskip-.1in \today}
\maketitle

\vskip.3in
\begin{center} {\bf Abstract} \end{center}

\begin{quote}
{\small Political pressure is undoubtedly an important influence in the setting of trade policy and the formulation of trade agreements. Most of the literature in this area models the political pressure that governments face as resulting from an exogenous, stochastic process. This paper shows that when political pressure arises endogenously, important results can be overturned and new insights into the motivation for features of the trade agreements we observe and rules of organizations such as the WTO come to light. Developing a model that integrates both exogenous and endogenous political pressure and taking a general approach to the government objective function, I show that governments may want to use tariff caps both to force special interest groups to continue lobbying after a trade agreement is signed and to reduce the magnitude of that lobbying effort. The presence of endogenous politics can also destroy an escape clause's ability to provide flexibility in times of large negative political shocks when lobbies use the flexibility to seek rents. This can explain why use of WTO Safeguards are conditioned on measurable economic indicators as well as why Safeguard levels of protection are not regulated.}
\end{quote}

%JEL Codes: F13,F53,C73,D72

\bigskip
\section{Introduction}
\input{intro}

\section{The Model}
\label{sec:stage}
I employ a two good partial equilibrium model with two countries: home (no asterisk) and foreign (asterisk).  The countries trade two goods, $X$ and $Y$, where $P_i$ denotes the home price of good $i \in \left\{X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.

The fundamentals here are chosen to match those of \Textcite{bs2001,bs2005}. Home country demand, supply and profits are given by $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$. %Note that production function consistent with these supply functions are $F_X = \sqrt{l}$ and $F_Y = \sqrt{2l}$ where we assume $w=1$. 
Foreign is taken to be symmetric.

This implies Home-country imports of $X$ and exports of $Y$ of $M_X(P_X)= 1 - \frac{3}{2}P_X$ and $E_Y(P_Y)= 2P_Y -1$, with foreign imports of $Y$ and exports of $X$ given by $M_Y^*(P_Y^*)= 1 - \frac{3}{2}P_Y^*$ and $E_X(P_X^*)= 2P_X^* -1$. With the only trade policy instruments being tariffs on import-competing goods, world prices are $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y^* = P_Y^W + \tau^*$, and $P_Y = P_Y^W$. Market clearing implies that world and home prices of $X$ are $P_X^W = \frac{4-3\tau}{7}$ and $P_X = \frac{4+4\tau}{7}$, symmetric for $Y$.

As is standard in the literature, it is assumed that the production of each good requires the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factor is tied to the price of the good in whose production their factor is used.

Note that $P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective importing country's tariff. As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff. This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.

I next describe the politically-relevant actors, which are a government and import-competing lobby in each country. In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization.\footnote{Adding a pro-trade lobby for the exporting industry would serve to strengthen some results.} Each country's government can set trade policy either unilaterally or through a negotiated trade agreement and then choose an applied tariff that is potentially different from that agreed upon in a trade negotiation.

The stage-game timing is as follows. First, the governments cooperatively form a trade agreement.\footnote{See Section~\ref{sec:rigid} for a discussion of why lobbies are assumed not to be involved at the trade agreement formation stage.} After the trade agreement is concluded, the special interest group representing the import-competing industry in each country lobbies the government for protection in the form of tariffs and any exogenous shock is realized. Finally, given the trade agreement, enforcement conditions, and political pressure it experiences, each government chooses the applied tariff level for its import-competing good.

As this game is solved by backward induction, it is intuitive to start by describing the incentives of the government in setting the applied tariff in the final stage. As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector. The details are analogous for $Y$ and foreign.

The per-period welfare function of the home government is given by
\begin{equation}
  W(\ga(s,e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(s,e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:wel}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\pi$ represents profits, $\ga(s,e)$ is the weight placed on profits (producer surplus) in the import-competing industry, $s$ is an exogenous state variable, $e$ is lobbying effort, and $\mathit{TR}$ is tariff revenue. Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, is affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{The standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them. An isomorphism can be made between the two forms as discussed in \Textcite{buzard2013b}.}

%\begin{assumption}
 % $\ga(s,e)$ is continuously differentiable, strictly increasing and concave in $e$.
  %\label{as:ga_c3}
%\end{assumption}

%Assumption~\ref{as:ga_c3} formalizes the intuition that the government favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.

Given the government's preferences, the home lobby chooses its lobbying effort $e$ to maximize profits net of lobbying effort:
\begin{equation}
  U_\text{L} = \pi(\tau(\ga(e))) - e
  \label{eq:lv3}
\end{equation}
where $\pi(\cdot)$ is the current-period profit and $\tau$ is the home country's tariff on the import good. %I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.

I assume the lobby's contribution is not observable to the foreign government. The implication is that the lobby can directly influence only the home government, and so the influence of one country's lobby on the other country's government occurs only through the tariffs selected.\footnote{cfr. \Textcite{gh95}, page 685.} $\ga(s)$ is assumed to be private information of the government, with specifics about its distributions given below. This makes the environment one of asymmetric information, with asymmetries potentially coming from two sources: the exogenous shocks and the endogenous behavior of the special interest group.

In the first stage, the governments choose the trade agreement tariffs via a negotiating process that I assume to be efficient. This process therefore maximizes the joint payoffs produced by the trade agreement:
\begin{equation}
  W(\ga(s,e),\ga^*(s^*,e^*),\tau,\tau^*) = W(\ga(s,e),\tau,\tau^*) + W^*(\ga^*(s^*,e^*),\tau,\tau^*)
  \label{eq:jv}
\end{equation}
Note that in this symmetric environment, this is the Nash bargaining solution where the disagreement point is the welfare resulting from the Nash equilibrium in the non-cooperative game.

\subsection{Nash Tariffs and Internationally Efficient Tariffs}
\label{sec:nasheff}
The home government's welfare is $W(\ga(s,e),\tau,\tau^*) = W_X(\ga(s,e),\tau) +W_Y(\tau^*)$, where
\[
  W_X(\ga(s,e),\tau) = \frac{9}{98} - \frac{5}{49}\tau - \frac{34}{49}\tau^2 +\frac{1}{98}\ga(s,e)\left[ 8 + 16\tau + 8\tau^2 \right]
\]
\[
  W_Y(\tau^*) = \frac{25}{98} - \frac{3}{49}\tau^* + \frac{9}{49}(\tau^*)^2 
\]
$W_X(\ga(s,e),\tau)$ is the utility derived from consumer surplus, producer surplus and tariff revenues in the import-competing industry and $W_Y(\tau^*)$ is the utility derived from consumer and producer surplus in the exporting industry.

When setting the tariff unilaterally, the government simply maximizes $W(\ga(s,e),\tau, \tau^*)$ by choice of $\tau$ given $\ga(s,e)$ and $\tau^*$. As there are no interactions between $\tau$ and $\tau^*$, the government simply maximizes $W_X(\ga,(s,e),\tau)$ and sets the non-cooperative tariff
\begin{equation}
  \tau^N(\ga(s,e)) = \frac{8\ga(s,e)-5}{68-8\ga(s,e)}.
  \label{eq:nash}
\end{equation}
I refer to this as the Nash tariff because it is the result of Nash equilibrium in the non-cooperative game between the governments. $\tau^N$ is increasing in $\ga$ and the second order condition is satisfied for all values of $\ga < 17/2$. Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$ it seems reasonable to assume that this condition will be satisfied in equilibrium.

In order to derive the jointly efficient tariff choice---that is, the tariff that maximizes the joint welfare of the home and foreign government, we require the welfare of the foreign government, $W^*(\ga^*(s^*,e^*),\tau,\tau^*)$ = $W_Y^*(\ga^*(s^*,e^*),\tau^*) + W_X^*(\tau)$, where
\[
  W_Y^*(\ga^*(s^*,e^*),\tau^*) = \frac{9}{98} - \frac{5}{49}\tau^* - \frac{34}{49}(\tau^*)^2 +\frac{1}{98}\ga^*(s^*,e^*)\left[ 8 + 16\tau^* + 8(\tau^*)^2 \right]
\]
\[
  W_X^*(\tau) = \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2 
\]

The internationally efficient home tariff is the $\tau$ that maximizes $W(\ga(s,e),\tau,\tau^*) + W^*(\ga^*(s^*,e^*),\tau,\tau^*)$. Since government welfare in both countries is additively separable in $\tau$, this is equivalent to maximizing $W_X(\ga(s,e),\tau) + W_X^*(\tau)$. The solution to this problem is the internationally efficient tariff
\begin{equation}
  \tau^E(\ga(s,e)) = \frac{4\left[\ga(s,e)-1\right]}{25-4\ga(s,e)}
  \label{eq:eff}
\end{equation}
The second order condition is satisfied for $\ga < \frac{25}{4}$, so again for all values of $\ga$ that lead to a non-prohibitive tariff.

%In the event of a trade war and facing this tariff-setting behavior by the legislature, the lobby maximizes $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$. In order to predict the Nash tariff, the political weighting function and form of political uncertainty must be specified. Take for instance $\ga(e,\ve) = 1.25 + C^{.2}$ with $\ve$ distributed uniformly on $[-0.25,0.25]$. Facing this specification of the political process, the lobby maximizes its objective function at $e_n = 0.00166$, which produces a Nash tariff of $0.129$.

%Therefore, in line with both the theoretical and empirical literature, I will assume that $\ga(e) \geq 1$ for all $e$. That is, even for the least favorable outcome of the lobbying process, the legislature will be at least weakly more protectionist than the executive.

%\begin{assumption}
%  $\ga(e) \geq 1 \ \forall e$.
%  \label{as:ga_l_e3}
%\end{assumption}

%Assumption~\ref{as:ga_l_e3} ensures that $\tau^a < \tau^{tw}$.

\subsection{Exogenous Political Pressure and Incentive Compatibility}
In order to compare the results for the case of endogenous political pressure to \Textcite{bs2005} results for exogenous political shocks, I provide the essentials of their setup.

They assume that $\ga$ and $\ga^*$ are each drawn independently from probability distributions with cumulative distribution functions $H(\ga)$ with $h(\ga)=H'(\ga)$. The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga} < \frac{7}{4}$. They also assume $\ga$ and $\ga^*$ are private information.

This leads to a concern about incentive compatibility. It must be in a government's best interest to truthfully reveal its $\ga$ or else it will misrepresent its private information about $\ga$ in order to raise tariffs and create a terms-of-trade gain, thereby improving its unilateral welfare. They provide conditions that guarantee incentive compatibility on page 481.

On a separate note, it is important to make clear that in this strand of the literature, it is assumed that trade agreements are negotiated to maximize the expected welfare of politically motivated governments given the exogenously-determined political pressure they expect to face at the time of the agreement's implementation. I will follow this convention throughout.

\section{Rigid Tariffs with Endogenous Political Pressure}
\label{sec:rigid}
In this section and the next, I demonstrate the implications of taking account of endogenous political pressure on some central design features of trade agreements. I do this by first summarizing the essential results from \Textcite{bs2005} for the design of trade agreements when political considerations are exogenous and then compare them to the case under consideration here of endogenous political activity aimed at gaining protectionist tariffs.

In Section~\ref{sec:escape2} I will examine the more realistic case where endogenous and exogenous political forces coexist. Until then, I will consider the cases separately to establish clearly the different dynamics that arise from the two distinct types of influence. I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to political pressure that results from the rent-seeking behavior of lobbies.\footnote{See Section~\ref{sec:escape2} for a detailed discussion of the relationship and interactions between these two extreme cases.} I begin with the case of rigid tariffs---that is, trade agreements that have no provision for flexibility. Rigid tariffs cannot be made to depend on the realized level of $\ga$ and so incentive compatibility considerations will not come into play.

%Importantly, both cases suffer from asymmetric information, but we will see that the nature of the informational lacuna differs. In the  exogenous case, there is a difference between the ex-ante and ex-post information of each government that has very important implications; this difference does not exist in the endogenous case. The implications for the features of trade agreements that are desirable---such as escape clauses---result directly from this difference.

Because the trade agreement is set ex-ante---that is, before political pressure is realized, regardless of its source---the process of choosing the optimal tariffs differs across the two cases. When $\ga$ is exogenous, the government must plan for level of $\ga$ it will face \textit{in expectation}. When political pressure in purely endogenous, the government has perfect foresight to plan for the $\ga$ it will face but must confront the fact that its decisions affect lobbying incentives.

With the introduction of endogenous political pressure, one must take a stand on whether lobbies influence the formation of the trade agreement or are only active once the trade agreement is in place. Since a central goal here is to determine the impact of relaxing the assumption that political pressure is exogenously determined, I assume the latter to make the comparison most direct to this literature in which by definition there can be no endogenous political pressure at the trade agreement formation stage. The former, which is undoubtedly an important possibility in many institutional settings, is left as an important extension.\footnote{Note that the endogenous trade policy literature that involves trade agreements (e.g. \Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure while answering a different set of questions.}

\subsection{Perfect External Enforcement}
I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement. Although it is widely agreed that this is unrealistic, this is an important baseline case. Section~\ref{sec:self} will relax this assumption and require that trade agreements be self-enforcing.

I compare strong bindings, in which the terms of the agreement are that tariffs must be set at the precise value stipulated and weak bindings---commonly referred to as tariff caps---in which the applied tariff can take any value as long as it is no greater than that which is agreed upon. To reiterate, with either form of rigid tariff, there is no incentive compatibility problem because the contractually-allowed tariff may not vary with $\ga$.

\vskip.2in
\subsubsection{Strong Bindings}
\label{sec:strong}
When there is perfect external enforcement of an agreement that specifies the exact tariff each government must adhere to, \Textcite{bs2005} derive the result that the optimal rigid, strong binding $\tau_S^R$ will be set at the level that is politically efficient for the expected realization of political pressure, $\ga(s)$. In their environment, $\ga(s)$ is exogenously determined and so the optimal trade agreement tariff is chosen as
\[
  \tau_S^R = \frac{4\left[\expect \ga(s) - 1\right]}{25 - 4\expect \ga(s)}
\]
which is simply Equation~\ref{eq:eff} evaluated at the expected $\ga$.

The optimal strong binding is very different if we instead allow political pressure to be endogenously determined. In this simple setting with only endogenous political pressure and no exogenous shocks, there is no uncertainty; the level of political pressure can be determined through backward induction.

In the final stage, the external enforcement agency compels the government to apply the precise tariff listed in the trade agreement contract regardless of circumstances. Call this $\tau_{S,e}^R$ to distinguish the strong (rigid) binding in the case of endogenous pressure from $\tau_S^R$ when political pressure is exogenous above.

Moving back to the lobbying stage and given that the government will be compelled to apply $\tau_{S,e}^R$ regardless of the $\ga(e)$ it faces, the lobby maximizes its payoffs by choosing $e=0$. Making the binding strong removes any incentives for the lobby to exert effort since that effort will not change the level of tariff protection that it will receive. To be clear: $\tau_{S,e}^R$ does not depend on lobbying effort, so no effort is exerted.

Then when setting the trade agreement tariffs to maximize its political welfare, the government can predict that $e$ will be zero. If it's the case that $\ga(0) = 1$, the trade agreement will specify free trade.

\begin{result}
  When political pressure is entirely endogenous, under the equilibrium trade agreement with strong bindings the lobbies exert no effort and are afforded no protection if governments have no inherent bias toward protection. 
	\label{res:strong}
\end{result}

This very strong form of commitment for governments removes the lobby's commitment to pay for the protection it desires. Unless one assumes that the government is inherently biased toward the import-competing industry---that is, would provide protection in the absence of any political pressure being exerted---this case results in an outcome that is equivalent to the outcome under a social welfare maximizing government. This contrasts starkly with the prediction when political pressure is assumed to be completely exogenously determined.

\subsubsection{Weak Bindings}
\label{sec:weak}
Recall that a weak binding, call it $\tau_W^R$, permits the government to apply any tariff less than or equal to $\tau_W^R$. Weak bindings are often referred to as \textit{tariff caps}. \Textcite{bs2005} show several interesting results regarding weak bindings when there is perfect external enforcement and political pressure is in the form of exogenous shocks.

To be optimal, the weak binding must be set at the level that is efficient for the realization of political pressure that is expected conditional on tariff setting being constrained by the cap. Their Proposition 1 follows:

\begin{proposition}[Bagwell and Stagier 2005]
Governments prefer negotiating commitments that take the form of weak bindings over strong bindings. When governments negotiate commitments that take the form of weak bindings, they choose to adopt tariff bindings that (a) are higher than those they would choose if their negotiated commitments instead took the form of strong bindings and (b) imply that governments with low realizations of political pressure will set their applied tariffs strictly below the bound level.
\end{proposition}

I now turn to the model under examination here with endogenous political effort.\footnote{As above, in this section I consider endogenous political pressure alone in place of exogenously-determined $\ga$.} As in the analysis of strong bindings, in place of expectations over a range of probabilistically-determined levels of political pressure, we can backward induct to determine the lobby's effort decision and the government's optimal choice of weak binding in anticipation of the lobby's behavior.

In the final stage, the government will seek to maximize its welfare by choice of $\tau$ given the political pressure that it faces and the enforcement of the tariff cap, here denoted $\tau_{W,e}^R$ for the case of endogenous political pressure, as distinct from $\tau_W^R$ in the case of exogenous political pressure. Thus, the home government unilaterally maximizes Expression~\ref{eq:wel}, so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as  $\tau^N(\ga(e)) \leq \tau_{W,e}^R$. Otherwise, it must set $\tau = \tau_{W,e}^R$ since the weak binding $\tau_{W,e}^R$ is externally enforced.

Knowing how the government will set the applied tariff, in the second stage the lobby makes its effort decision to maximize net profits according to Expression~\ref{eq:lv3}. If the lobby's optimal effort level in the unconstrained problem (label it $e^L$) would lead to $\tau^N(\ga(e)) > \tau_{W,e}^R$, the lobby will reduce its choice so that no effort is wasted. I label the lobbying effort choice that leads to the weak binding level as $e^R_{W,e}$ so that $\tau^N(\ga(e^R_{W,e})) = \tau_{W,e}^R$.

We are now in a position to determine how the governments set trade agreement tariffs. Their goal is to maximize their joint welfare as in Expression~\ref{eq:jv} given the behavior they expect from the lobbies. Again, we can restrict attention to the home country and the $X$ sector because of the symmetry and separability of the economy.

We know that the lobby will either choose its unconstrained optimal effort $e^L$, or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding. The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$. The government is unable to use a weak binding to encourage extra lobbying---at least in the case with only one import-competing lobby under consideration here.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}

I assume for simplicity of exposition that the government does not set the tariff binding $\tau^R_{W,e}$ higher than that which would result from the lobby's optimal effort $e^L$, as setting a higher tariff cap would not change lobbying behavior or outcomes in any way. This is in contrast to the case of exogenous political pressure, where low realizations of $\ga$ lead to applied tariff levels below the cap in an economically meaning sense.

The optimal home tariff $\tau^R_{W,e}$ under the trade agreement is the one that maximizes joint government welfare. As argued in Section~\ref{sec:nasheff}, joint government welfare (with variables that do not affect the $X$ sector suppressed) is
  \begin{equation}
		W_x \left(\ga(e),\tau) \right) + W_x^*\left(\tau \right)
	  \label{exp:1}
	\end{equation}
where the lobby will exert no more effort than that which produces the capped $\tau^R_{W,e}=\tau^N(\ga(e))$. Recall that the Nash tariff, $\tau^N(\ga(e))$ is the solution to the unilateral optimization problem the home government faces when choosing the applied tariff level in the final stage of the game. It is given by Expression~\ref{eq:nash}, which contains the mapping between the political pressure the lobby exerts and the government's unilateral tariff response. We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e))$ equal to the tariff cap $\tau^R_{W,e}$. This $e$ is that which is labeled $e^R_{W,e}$ above.

Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage according to the chosen tariff cap $\tau^R_{W,e}$. The governments get to choose the $\tau^R_{W,e}$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.

Notice that the maximand in Expression~\ref{exp:1} is exactly the same as that in this third-stage maximization problem except it adds  foreign government welfare. This implies that $\tau^R_{W,e}$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments do not want as much lobbying as possible. To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare. See Section~\ref{sec:objfcn} for further discussion.}

A second motive is operative here that is absent in the case of exogenous political pressure: the trade agreement can act as a domestic political commitment device, allowing the home government to manipulate ex-post lobbying incentives. In fact, whenever the cap is strictly below the non-cooperative level, that is when $\tau^R_{W,e} < \tau^N$, the weak binding serves to restrain political pressure whether or not the governments desire it.

On the other hand, if the level of the optimal weak binding $\tau^R_{W,e}$ is strictly positive, the weak binding is at least in part serving to generate political involvement---most often thought of as campaign contributions---that the government finds beneficial from its politically-motivated point of view. The government could have chosen a cap of zero and eliminated lobbying activity altogether. In fact, any strong binding accomplishes this very feat, but is also \textit{unable} to generate lobbying activity because of its effect on lobbying incentives. Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is a related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.}

\begin{result}
    (a) When political pressure is entirely endogenous, governments prefer negotiating commitments that take the form of weak bindings over strong bindings. When governments negotiate commitments that take the form of weak bindings, they choose to adopt tariff bindings that are higher than those they would choose if their negotiated commitments instead took the form of strong bindings. (b) Governments will not set applied tariffs strictly below the bound level, but they may use the weak tariff binding to restrain and/or encourage endogenous political pressure.
		\label{res:weak}
\end{result}

Note that part (a) of Result~\ref{res:weak} matches Proposition 1 of \Textcite{bs2005}. The similarities are cosmetic only; the result holds in the case of endogenous $\ga(e)$ because of the extreme result for strong bindings that no lobbying effort is exerted and the trade agreement is at free trade. This can be replicated with a tariff ceiling of zero. Part (b) points out that the governments can use a tariff cap to manipulate lobbying incentives so that political effort is neither too low or too high.

\subsection{Self-Enforcing Trade Agreements}
\label{sec:self}
Here I relax the assumption of perfect external enforcement, since this is not widely available in the context of international trade relations. External enforcement is replaced with promises of future cooperation and punishment is modeled via a repeated game. I will first describe this repeated game and then present the analysis for the case of (rigid) weak bindings.

\subsubsection{Repeated Game}
For the setting of exogenous political pressure, Bagwell and Staiger assume that $\ga$ is independently and identically distributed across governments and time. Because the governments are faced with asymmetric information, perfect public equilibrium (PPE) is the appropriate solution concept. Here attention is restricted to symmetric, stationary PPE.

In order to establish an equilibrium, we must ensure that both the on-schedule (static) incentive constraint and the off-schedule (repeated-game) incentive constraint is satisfied. We know that the former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here. Following \Textcite{bs2005} and most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} and \Textcite{buzard2013a} for alternatives that take seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.}
	%see KRW, Buzard 2014 for alternatives

\subsubsection{Self Enforcement with Weak Bindings}
\label{sec:selfweak}
When political pressure is exogenous, \Textcite{bs2005} perform a standard repeated-game prisoner's dilemma analysis and show that if governments are patient enough ($\de$ is sufficiently high), the optimal weak binding from the perfect external enforcement setting can always be sustained with repeated-game incentives.

When $\ga$ is endogenously-determined, repeated-game enforcement is altered relative to the case of exogenously-given political pressure. This is because, in place of a stochastic process that determines $\ga$, $\ga$ is determined by a new repeated-game player who has incentives relative to whether the trade agreement is followed or reversion to the non-cooperative Nash outcome takes place.

When $\ga$ is exogenous, the optimal trade agreement tariff is the same under perfect external enforcement and self-enforcement because the change in enforcement conditions cannot affect the realization of $\ga$. We will see that moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem. However, it does not alter the within-period costs and benefits and so in the case of endogenous political pressure we will see the same result that the optimal trade agreement tariff in invariant to the enforcement conditions.

Recall that the government's most preferred trade agreement tariff in this setting is the one that would allow it to provide the level of protection that is demanded \textit{ex-post}. Enforcement considerations do not change the protection demands when $\ga$ is exogenous. Similarly, because the government's objective function does not change and it is able to use the tariff cap to control the lobby's effort level, as long as it is sufficiently patient, it will set the same tariff cap.

Let us see how this works. Again, because of symmetry and separability, we can focus on the home government's choice of $\tau$ and the $X$ industry.

The government's incentive constraint is given by
  \begin{multline}
    \frac{\de}{1-\de} \left\{W_x(\ga(e^R),\tau^R_{W,e}) + W_x^*(\tau^R_{W,e}) - \left[W_x(\ga(e^N),\tau^N(\ga(e^N)) + W_x^*(\tau^N(\ga(e^N)) \right] \right\} \\ \geq W_X(\ga(e^B),\tau^B(\ga(e^B)) - W_X(\ga(e^R),\tau^R_{W,e})
		\label{exp:govincent}
  \end{multline}
where $\de$ is the discount factor assumed common to the government and lobby, $\tau^R_{W,e}$ is the same tariff cap from Section~\ref{sec:weak}, and $\tau^B$ is the ``break'' tariff. Any $\tau > \tau^R_{W,e}$ is sufficient to breach the trade agreement and trigger Nash reversion; $\tau^B(\ga(e^B))$ is defined as the tariff the government would choose as unilaterally optimal if faced with $e^B$, where $e^B$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.

On the left side of the inequality is the discounted per-period gain from maintaining the trade agreement relative to Nash reversion. In order for the government to have the incentive to abide by the agreement, this must be at least as large as the benefit to cheating. This is the current period increase in unilateral welfare from the tariff applied when breaking the agreement and is given on the righthand side of Inequality~\ref{exp:govincent}.

For a given $\de$, $\tau^B$ and $e^B$ are derived using Equation~\ref{eq:nash} and Expression~\ref{exp:govincent} evaluated at equality. $e^B$ can be interpreted as the minimum level of lobbying effort that will persuade the government to abrogate the agreement. It must provide significantly more unilateral welfare than the trade agreement tariff in order to compensate the government for the loss of cooperation in every future period.

The lobby's incentives must also be satisfied, as the government's decision on whether to abide by the agreement or break it depends on the amount of lobbying effort it encounters. In order for the lobby to prefer the trade agreement tariff to the option of causing the agreement to be broken and tariffs to revert to the non-cooperative level, the following must hold
\begin{equation}
  \frac{1}{1-\de}\left[\pi(\tau^R_{W,e}) - e^R\right] \geq \pi(\tau^B) - e^B + \frac{\de}{1-\de}\left[\pi(\tau^N) - e^N\right]
	\label{exp:lobby}
\end{equation}
That is, the present discounted value of net profits under the trade agreement must be weakly higher than one period of net profits from the ``cheater'' tariffs and the discounted future Nash profits.

Here we have the possibility of a starkly different result from that under exogenous political pressure, where \Textcite{bs2005} show that sufficiently patient governments can always sustain the optimal trade agreement tariffs. Instead, the addition of a second incentive constraint can make it impossible to sustain the politically efficient tariffs.

To see this, start by noticing that the lefthand side of Expression~\ref{exp:govincent} in increasing in $\de$, while the righthand side is constant in $\de$. That is, as the government becomes more patient, it is easier to satisfy the government's constraint. Equivalently, it requires a larger $e^B$ to violate the constraint.

Shifting attention to the lobby's incentive constraint in Expression~\ref{exp:lobby}, this becomes harder to satisfy as $\de$ increases. That is, the lefthand side is decreasing in $\de$ and the righthand side is increasing in $\de$, and this relationship holds even taking into account the influence from the government's constraint on $e^B$ as long as $e^B$ is to the left of the lobby's optimal choice of effort. Once $e^B$ passes $e^L$, $\pi(\tau^B) - e^B$ begins to decrease and it's possible for the effect of $\de$ on the lobby's constraint to eventually be overturned.

In general, however, we cannot be guaranteed that even the most patient government can sustain cooperation at the politically optimal trade agreement level. At issue is the behavior of the lobby, which is now a formal player in its own right and must be incentivized to keep its behavior on the equilibrium path. More patient lobbies are willing to work harder to encourage the government to break the trade agreement so that they can enjoy infinite periods of Nash tariffs, so increasing the discount factor makes it more difficult rather than easier to satisfy the lobby's constraint.\footnote{The essential point underlying this result was first made by \Textcite{buzard2013a}.}

\begin{result}
  The presence of the lobby as a repeated-game player may imply that the politically optimal self-enforcing trade agreement tariff is strictly greater than $\tau^R_{W,e}$. 
\end{result}

If there is no $\de$ such that Inequalities~\ref{exp:govincent} and \ref{exp:lobby} hold simultaneously, the governments must raise the trade agreement tariff, which loosens both incentives constraints simultaneously.

\section{Endogenous Political Pressure and the Escape Clause}
\label{sec:escape}
I follow \Textcite{bs2005} in examining a trade agreement with an escape clause by adding a second negotiated binding to the agreement so that there is one (weak) binding for when political pressure is low / normal and a higher (weak) binding for exceptional circumstances of high levels of political pressure. Escape clauses are common features of trade agreements in practice, and when one models political pressure as exogenous, an escape clause seems attractive: in the presence of a particularly large negative political shock, being bound to a tariff designed for normal times would cause significant welfare losses to the politician.

However, because the tariff allowed under the trade agreement with an escape clause varies with the announced level of political pressure, the incentive compatibility constraint now comes into play: a government may gain an advantage by misreporting the level of $\ga$ it experiences. In their Proposition 4, Bagwell and Staiger show that a trade agreement with a costless escape clause cannot improve welfare because it cannot be made incentive compatible. No matter the realization of the stochastic $\ga$, it is always in the government's interest to announce that $\ga$ is high, which allows it to apply any tariff up to the higher weak binding. This deviation improves unilateral welfare while imposing a terms-of-trade externality on the trading partner.

A similar problem arises in the case of endogenous political pressure, but with no parallel potential gain. The problem of asymmetric information remains, as lobbying effort is not observable to the trading partner. The appeal of an escape clause is to provide the flexibility of a higher binding in exceptional circumstances when political pressure is randomly high. But in the stark case under examination here with purely endogenous politics, an exceptionally high $\ga$ can only derive from higher effort exerted by the lobby. If one chooses an optimal tariff cap, a second, higher cap can only encourage excess, sub-optimal lobbying.

Thus, while with a costless escape clause there is no way to make truth-telling incentive compatible, this is of little consequence. The government would actually be \textit{truthfully} reporting the higher level of $\ga(e)$ as long as it is worthwhile for the lobby to increase its effort to this level. The appeal of an escape clause is missing in this case, but the potential for governments to be forced ex-post to exploit the escape clause in a way that damages their ex-ante welfare remains.\footnote{Notice that the ex-post rent-seeking that creates this ex-ante loss is different in the two cases. With exogenous shocks, the government increases its welfare through the terms-of-trade channel. With endogenous political pressure, the lobby extracts the rents.}

\subsection{Escape Clauses with Strong Bindings}
Suppose that political pressure derives only from endogenous sources but an escape clause is to be implemented anyway. As \Textcite{bs2005} point out, in order for an escape clause to be useful to the government, it should be costly so that its use can be made incentive compatible. They suggest several avenues for introducing a cost for the use of the escape clause.

One possibility is to make the escape clause tariff a strong binding instead of a weak binding. The cost imposed here is the following: if the realization of $\ga$ is such that the optimal tariff is above the agreed-upon weak binding for normal times but below the strong binding for exceptionally high realizations of political pressure, the government must choose between applying the lower weak binding and the precise escape-clause tariff. In this case, there are welfare losses from implementing a sub-optimal tariff level that are not present if the escape clause tariff is a weak binding.

When $\ga$ is endogenous, the government is not subject to such random, unpredictable realizations of political pressure. It can avoid such costs by setting the escape clause tariff precisely at the level the lobby will find optimal to ask for; it thus avoids the mechanism that allows for incentive compatibility as well. In addition, given that the lower, normal binding could be set optimally for the government, the government only stands to lose from entering into a trade agreement with such an escape clause. This is, again, assuming that there is no exogenous source of political shocks for which flexibility is desired.

Thus, in this case the strong binding does not really create an incentive-compatibility-inducing cost, but again the central problem is that of lobbying incentives. This could explain, as Bagwell and Staiger point out, why the WTO does not incorporate strong bindings for the escape clause.

\subsection{Escape Clause with Side Payments}
Suppose instead that the use of the escape clause must be accompanied by a side payment when the higher tariff is applied. In efficiency terms, this will only redistribute surplus from the trading partner who invokes the escape clause to the partner whose goods are targeted by the higher tariff. This is not entirely realistic as cash transfers are rarely observed in the WTO and other trade agreements; instead compensation is in the form of retaliatory tariffs---which are not efficient.\footnote{Note also that the WTO Safeguards agreement has a dynamic use constraint but removes the requirement for compensatory action in the first three years.}

\Textcite{bs2005} show that, again, the transfer function must satisfy two incentive compatibility conditions: the static condition that makes it worthwhile to truthfully reveal one's true $\ga$, and the repeated-game incentive constraint required for self-enforcement. Given these requirements, their Proposition 5 establishes that an appropriate transfer scheme can make a trade agreement with an escape clause incentive compatible.

Given the limited nature of the result in Section~\ref{sec:selfweak}, for the case of endogenous $\ga$ I restrict attention to economies in which the jointly optimal weak binding is supportable. The conditions given in Lemma 1 of \Textcite{bs2005} are not materially changed by substituting endogenous $\ga$ for exogenous $\ga$, so incentive compatibility in the static sense is ensured.

It is the repeated-game, or off-schedule, constraint that is significantly altered. The new element compared to self-enforcement without the escape clause is that the government can cheat on both its tariff and its transfer payment. With no lobby, the Bagwell and Staiger result shows that this temptation can be overcome; the question is whether the lobby makes this more difficult.

The answer is that it does not. Assuming the lobby represents a negligible share of the population, it does not bear the cost of the transfer. In fact, because it receives all the benefit of the tariff increase under the escape clause with none of the burden of paying for it, the lobby is much better off under the trade agreement with escape than the trade agreement without escape.

%	\item BUT need to check equilibrium impact on lobby's utility--does something in this constraint make it cheaper / more expensive for lobby to get the tariff it wants?

Here, we have a mechanism that is incentive compatible despite the presence of lobbying. It makes the lobby exert effort in order to receive the higher level of protection under the escape clause just as in the other mechanism. And although the burden of the transfer falls on others in the economy, at least the trading partner is compensated. Here the government is truthfully revealing the level of political pressure it faces, but that pressure is created by endogenous forces that are intensified by the existence of the escape clause.

In the next section, I ask the question of whether this type of outcome is likely to be consistent with the ex-ante best interests of the governments themselves as a prelude to exploring the likely intentions behind the design of such escape clauses.

\section{Political Objective Functions and the Role of Trade Agreements}
\label{sec:objfcn}
The previous sections analyze in general the effects of various forms of trade agreements when political pressure is endogenous. I now use the specific functional forms for the economy given in Section~\ref{sec:stage} and compare government objective functions to understand the impact of strong bindings, weak bindings and escape clauses on government welfare.

To reiterate, in line with the dominant strain of the literature, I have assumed that the governments' choice of trade agreement tariff is optimized with respect to the political pressure it expects to experience after the trade agreement is signed. In the case of exogenous political pressure, the best one can do is take an expectation given the distribution of $\ga$, or possibly build some kind of flexibility into the agreement through tariff caps or an incentive compatible escape clause.

With endogenously-determined political pressure with no shocks---the simple case examined so far here---each government can predict the political pressure it will face and in some cases use the trade agreement to alter the incentives of the lobby. Since the government has some control through the trade agreement over the level of $\ga$ it will face, it becomes important to understand how the government's welfare varies in $\ga$.

Let's begin with the \Textcite{baldwin}-style government objective function that is most common in the recent literature on trade agreements and that has been in use throughout this paper so far, specialized to the case of endogenous pressure only.
\begin{equation}
  W(\ga(e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
  \tag{\ref{eq:wel}}
\end{equation}
What we would like to know is at what $\ga(e)$ the government's welfare is maximized, given that it will choose $\tau$ according to either Equation~\ref{eq:nash} when setting tariffs unilaterally or Equation~\ref{eq:eff} when setting tariffs in the context of a trade agreement.

The answer turns out to be the same for the unilateral and joint problems. Starting with the non-cooperative maximization problem with respect to $\ga$ (there is no loss in suppressing the dependence of $\ga$ on $e$ for the time being):
\[
 \max_{\ga} W_x(\ga,\tau) + W_y(\tau^*)
\]
The first order condition is
\[
	\frac{\partial W_x}{\partial \ga} + \frac{\partial W_x}{\partial \tau}\frac{\partial \tau}{\partial \ga} = 0
\]
In an envelope-theorem style result, because the Nash tariff is the result of the government's optimization with respect to $\tau$, $\frac{\partial W_x}{\partial \tau}=0$ so that the first order condition reduces to $\frac{\partial W_x}{\partial \ga} = 0$.

Turning to the maximization of joint welfare in the trade agreement, we have
\[
  \max_{\ga} W_x(\ga,\tau) + W_y(\tau^*) + W_x^*(\tau)+ W_y^*(\tau^*)
\]
The first order condition is 
\[
  \frac{\partial W_x}{\partial \ga} + \frac{\partial W_x}{\partial \tau}\frac{\partial \tau}{\partial \ga} + \frac{\partial W_x^*}{\partial \tau}\frac{\partial \tau}{\partial \ga} = \frac{\partial W_x}{\partial \ga} + \left[ \frac{\partial W_x}{\partial \tau} + \frac{\partial W_x^*}{\partial \tau}\right]\frac{\partial \tau}{\partial \ga} = 0
\]
Here, it is the efficient joint tariff that is chosen, and it is chosen precisely by setting the term in brackets equal to zero so that we have the first order condition simplifying to the same expression as in the unilateral case.

An interior solution does not exist to this problem. By inspection, the partial derivative of government welfare with respect to $\ga$ is simply producer surplus in the import sector. This is always positive. Thus, under the government welfare function in Expression~\ref{eq:wel}, government welfare is maximized when $\ga$ takes on the highest value achievable.

This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the partially weighted government welfare function from Expression~\ref{eq:wel} under the parameterization of \Textcite{bs2005} used throughout this paper. Graphed in terms of tariffs, one can see that from $\tau=0.05$---the tariff the home government optimally sets when making unilateral policy with no political pressure---government welfare steadily increases as the trade agreement tariff increases. This is because the governments can expect that ex-post political pressure will increase in the trade agreement tariff, and government welfare increases everywhere in political pressure.

\begin{figure}
\begin{center}
\includegraphics[height=3.5in, width=5in]{unweight.jpg}
\end{center}
\caption{Government Welfare with Partially Weighted Objective Function \label{fig:unweight}}
\end{figure}

This result that government welfare is strictly increasing in the political pressure parameter would seem to be related to the fact that, although $\ga$ is often referred to as a ``political economy weight,'' Expression~\ref{eq:wel} is only partially weighted. That the extra importance given to profits of importers when $\ga$ increases is pure additional welfare with no concomitant reduction elsewhere may be an assumption worth weakening, particularly in the context of endogenous political activity.

An alternative formulation for the government objective function is one in which $\ga$ is a true weight in the following straightforward sense:
\begin{equation}
  \frac{1}{4+\ga(e)}CS_X(\tau) + \frac{\ga(e)}{4+\ga(e)}PS_X(\tau) + \frac{1}{4+\ga(e)}CS_Y(\tau^*) + \frac{1}{4+\ga(e)}PS_Y(\tau^*) + \frac{1}{4+\ga(e)}TR(\tau)
  \label{eq:weighted}
\end{equation}

See the \hyperlink{sol_weighted}{Appendix} for the mathematical derivations corresponding to Figure~\ref{fig:weight}. They show that, particularly for the trade agreement tariffs, government welfare first declines and then increases, with the minimum value occurring around $\tau = 0.057$ and $\ga = 1.05$. The maximum still unquestionably occurs at the maximum value of $\ga$, which leads to the prohibitive tariff of $\frac{1}{6}$. But if a government's objective function is closer to Expression~\ref{eq:weighted} than Expression~\ref{eq:wel}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.

\begin{figure}
\begin{center}
\includegraphics[height=3.5in, width=5in]{weight.jpg}
\end{center}
\caption{Government Welfare with Weighted Objective Function \label{fig:weight}}
\end{figure}

Whether or not the optimal trade agreement would set a tariff cap aimed at reducing lobbying depends on the lobby's incentives, which depends in turn on the shape of $\ga(e)$. Let's look at two parameterizations in a simple family of functions for $\ga(e)$ that demonstrate drastically different outcomes.

I begin with $\ga(e) = 1 + e^{.1}$. With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0016$, which results in $\ga=1.525$ and a Nash tariff of $0.129$. Import-competing producer (net) profits are $0.1025$, while government welfare is $0.0862$. Because joint government welfare is convex in the tariff, the solution will be at a corner: either at $\tau^R_{W,e} =0$ or at the highest level for which the lobby will exert effort, i.e. the Nash tariff. Since joint welfare at the zero tariff is $0.0857$, the tariff cap is set at the trivial level of the non-cooperative Nash tariff.

Here, given the way in which lobbying effort is translated into political support as represented by $\ga$, the government cannot improve its welfare by manipulating lobbying incentives. Discouraging lobbying would reduce welfare; in this parameter range, the government's interest is solely in encouraging lobbying effort, and there is nothing it can do in this regard.

%This is does through the introduction of the trade agreement: by making protection more expensive, on the margin the lobby pays more in order to get even less protection than it did in the absence of the agreement. But there is nothing further the government can do to increase $\ga$.

This example stands in contrast to a less concave (in the sense of Arrow Pratt) $\ga$. Take $\ga(e) = 1 + e^{.3}$. With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0010$, which results in $\ga=1.126$ and a unilateral tariff of $0.068$. Import-competing producer (net) profits are $0.0921$, while government welfare is $0.0855$. Since joint government welfare is higher at free trade, the trade agreement will set $\tau^R_{W,e}=0$, profits decrease to $0.0816$, and government welfare increases to $0.0857$.

In this case, again the government does not have means to encourage any additional lobbying, but it can improve its situation by discouraging lobbying. By setting a tariff cap of zero---assuming the conditions for self enforcement outlined in Section~\ref{sec:selfweak} are met---the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters. The stark prediction that the best the government can do is by setting the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has a interior global maximum in $\ga$.

With this weighted welfare function, we see that the trade agreement is employed to different ends depending on $\ga(e)$. If the lobby's ability to make an impact on the political process is strong enough, the government will be chiefly concerned with the profits of the import-competing industry and will use the trade agreement solely to internalize the terms-of-trade externality. However, if the lobby's impact is below a given threshold, the distortions that protection create are weighed more heavily and the government will use the trade agreement to rein in lobbying activity.

%In either case, we see that the government is able to provide reduced protection while extracting more effort from the lobby under a trade agreement: internalizing the terms of trade externality increases the marginal cost of protection.

It is not claimed that the government welfare function in Expression~\ref{eq:weighted} perfectly represents the preferences of those who negotiate trade agreements. However, examining this alternative in combination with endogenous lobbying points out that the idea that trade agreements can function as domestic political commitment devices may be more general than previously demonstrated. It can also provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.

In the next section, I turn to the implications for the structure of the WTO procedures for administered protection, and the escape clause in particular, in a more realistic setting where both exogenous political shocks and endogenous lobbying are present.

\section{The Purpose and Design of the Escape Clause}
\label{sec:escape2}
The title of this section pays direct homage to \Textcite{bs2005} for their above-referenced, foundational work on the escape clause in a setting with exogenous political pressure. Here, I ask how the addition of endogenous political pressure alters the optimal design of an escape clause.

When $\ga$ is exogenous, the purpose of the escape clause is clear: it improves welfare by allowing the government to apply a higher tariff when the realization of $\ga$ is particularly high. In this case, the government would suffer significant welfare losses from abiding by the lower, normal tariff binding, and may find its interests are better served by abrogating the trade agreement if no escape were permitted.

In Section~\ref{sec:escape}, I show that there is no such benefit when political pressure is purely endogenous in nature. All political pressure can be anticipated by the government and accounted for in the normal binding, so any higher tariff provided for escape would either not be used or would reduce government welfare by encouraging excess political pressure in the case where the trade agreement tariff is being employed to reduce lobbying.

In reality, it is likely that both exogenous shocks and endogenous forces contribute to the pressure to which policy makers are subject. Thus, the desire for flexibility that derives from the purely exogenous case remains and we desire to see if it is possible to implement an escape clause in the face of the additional endogenous source of political pressure.

So we go back to the general formulation where $\ga$ is a function of both lobbying effort $e$ and exogenously-determined events that create political pressure as in \Textcite{ms2011}, $s$. Let us take a simple case where the political pressure from the two sources are additively separable so that $\ga(s,e) = \ga(s) + \ga(e)$. Various interpretations are possible, including that some part of $\ga(s)$ derives from lobbying associated with the exogenous shock, while $\ga(e)$ is pure rent-seeking. Assume the interesting case that $\ga(\un{s}) < \ga(\ov{s}) < \ga(e^L)$; that is, the optimal political economy parameter from the lobby's point of view is greater than that which results from even the highest value of the exogenous shock variable. This implies that the lobby has the incentive for all realizations strictly below $\ov{s}$ to add pressure to that which comes directly from exogenous sources.

Imagine first that $s$ and its mapping $\ga(s)$ are completely unverifiable so as to make the analogy to the exogenous shocks examined above complete. There is a large parameter space over which an escape clause could not function in the way it is intended. When the realization of the shock is high, the government will report the high shock and apply the escape clause tariff. But when the shock is low, the lobby will exert effort so that the sum of $\ga(s)$ and $\ga(e)$ equals the high realization. From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section~\ref{sec:objfcn}, the governments' ex-ante welfare is reduced by excess lobbying.

Given that the intent of the escape clause is to provide flexibility for high realizations of $\ga(s)$, we see that the presence of endogenous lobbying can easily destroy its efficacy for this purpose as the lower binding will never be used. The water in the escape clause will be filled in by endogenous political pressure.

This is not a problem of incentive compatibility. Similar to \Textcite{beshkar2010b}, it can be shown [formal result coming soon] that a dispute settlement body that provides an independent signal about the value of $\ga(s,e)$ on which the governments condition punishments for misreporting can make truth-telling incentive compatible. That is, a government will not over-report its political-economy pressure in order to impose a terms-of-trade externality on its trading partner. But this is of little consequence because the political-economy parameter will always take on its highest value.

Given the inability of achieving the first best in this environment, what can be done? Unless political pressure that comes from a shock can be distinguished from political pressure that is purely rent-seeking, there appears to be little hope that the two can be disentangled through the investigation of a body such as the WTO's DSB. In fact, in its rulings the DSB does not appear to take into account political conditions at all.

It seems plausible that this kind of dynamic is the reason that use of the WTO Safeguards are predicated on verifiable economic indicators--the $s$'s. This would appear to be the only way to prevent lobbies from exploiting the escape clause for uses for which it was not intended. However, once a shock has occurred, given that the DSB cannot distinguish between $\ga(s)$ and $\ga(e)$, determining the tariff that should be applied could only be done if (a) the DSB has the mapping $s \rightarrow \ga$ \textit{and} this strong form of separability holds. These conditions seem unlikely to be met in reality and we observe, again, that the DSB makes no attempt to determine the political parameter. Without this information, it thus can only certify the legitimacy of a Safeguard measure; it cannot determine the correct level at which the Safeguard measure should be applied.

\section{Ex-Ante Lobbying}
\label{sec:exante}

\input{ex_ante}

\section{Conclusion}
\label{sec:concl}
I have shown that accounting for endogenous lobbying has important implications for the design of trade agreements. Tariff caps may be used alternatively to reduce lobbying activity, or to incentivize lobbies to remain active after a trade agreement is in place, whereas a strong binding would remove lobbies' incentives to provide political support ex-post.

The addition of the lobby can make it harder to sustain cooperation and complicates the incentive compatibility problem associated with the use of escape clauses. The temptation for lobbies to exploit the opportunity presented by an escape clause provides a justification for the WTO requirements that verifiable economic conditions are met in order to legally invoke these protections.

I introduce a fully-weighted government utility function that is a slight modification of the standard Baldwin-style government objective function to demonstrate that it may be in even a politically-motivated government's interest to use a trade agreement to restrain political activity. The fact that the WTO Safeguards Agreement replaced compensation in the first three years of the invocation of a safeguard with a dynamic use constraint is consistent with the idea that the organization and the governments that constitute it are working to reduce rent seeking by lobbies. Section~\ref{sec:escape} demonstrates that lobbies do not care about the compensation since the costs do not fall on them, whereas a dynamic use constraint directly affects their ability to receive protection in the future.

There are many exciting avenues to extend this work. This simple framework demonstrates a tractable way to introduce endogenous political pressure into many of the important questions that have been and are currently being explored concerning the design of trade agreements and the institutions that facilitate them.

The discussion in Section~\ref{sec:exante} presumes both that there are no significant interactions between lobbies and that governments are free to choose which sectors are part of a trade agreement, but these decisions are part of a more complicated game that is, in particular, influenced by WTO rules. Taking into account the interactions between multiple lobbies, the influence of WTO rules, and differences in the $\ga$ functions across industries---that is, in how political pressure translates into weight in the policy-making process---seem like important and interesting avenues for further study.

In particular, here there is no scope for the government to encourage lobbying above the level that is optimal from the special interest group's point of view; adding additional lobbies may create insights to this effect. There is also strong potential to gain insights concerning very realistic dynamics between political and economics shocks and lobbying, especially insofar as they impact questions of enforcement and administered protection.

\section{Appendix}

\noindent \textbf{\hypertarget{sol_weighted}{Optimal $\ga$ for Weighted Government Welfare Function}}: \\
The first order condition for the weighted welfare function in Expression~\ref{eq:weighted}
			  \[
				  \frac{\partial W}{\partial \ga} = \frac{1}{\left(4+\ga(e)\right)^2}\left[ - CS_X + 4 \cdot PS_X - CS_Y - PS_Y - TR \right] = 0
				\]
Examining the expression in brackets:				
				\[
				  - CS_X + 4 \cdot PS_X - CS_Y - PS_Y - TR = 
				\]
				\[
				  -\frac{9-24\tau + 16\tau^2}{98} + 4\frac{4+8\tau + 4\tau^2}{49} - \frac{7\tau - 42\tau^2}{49} - \frac{9+18\tau + 9\tau^2}{98} - \frac{16-24\tau + 9\tau^2}{98}
				\]
				\[
				  = \frac{-9 + 32 -9 -16}{98} + \frac{24 + 64 -14 -18 +24}{98} \tau + \frac{-16 + 32 + 84 - 9 - 9}{98}\tau^2
				\]
				\[
				  = \frac{-2}{98} + \frac{80}{98} \tau + \frac{82}{98}\tau^2
				\]
			This expression is negative at $\tau=0$ and increases monotonically with an interior minimum; note the second derivative is strictly positive.

				Because $\tau$ is an increasing function of $\ga$, this pattern holds for $\ga$ as well. $\frac{1}{\left(4+\ga(e)\right)^2}$ is positive (and decreasing as a function of $\ga$), so the pattern holds for the derivative as a whole. Thus the FOC delivers a minimum point and the maximum must be at one of the endpoints. Calculations show that $W(\ga=1.75) > W(\ga=1)$. $\hfill\blacksquare$

\vskip.2in
\noindent \textbf{\hypertarget{lem1}{Proof of Lemma~\ref{lem:1}}}: \\
In the case of strong bindings, this is immediate from Result~\ref{res:strong}. In the case of weak bindings, first examine the first order condition for unilateral tariff setting:
\[
  \frac{\partial W_X(\ga,\tau)}{\partial \tau} = - \frac{5}{49} - \frac{68}{49}\tau +\frac{1}{98}\ga\left[ 16 + 16\tau \right] = 0
\]
Substituting in the inverse of the solution function $\tau^N(\ga) = \frac{8\ga-5}{68-8\ga}$ for $\ga$, this is
\[
  \frac{\partial W_X(\ga,\tau)}{\partial \tau} = - \frac{5}{49} - \frac{68}{49}\tau +\frac{1}{98} \frac{68\tau+5}{8+8\tau} \left[ 16 + 16\tau \right] = 0
\]
This same solution for $\ga$ will be substituted into the first order condition for the joint maximization problem in the ex-post trade agreement game; the only difference is the foreign welfare $W_X^*(\tau) = \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2$ is added to the objective function. The derivative with respect to $\tau$, $- \frac{3}{49} + \frac{18}{49}\tau$ is strictly negative for all non-prohibitive levels of the tariff ($\tau \in \left[0,1/6\right)$) and 0 at the prohibitive tariff. Thus the first order condition for the joint problem is negative at the unilateral tariff; a (weakly) lower tariff must be chosen to achieve the maximal value of this objective function. As demonstrated in Section{sec:objfcn}, the domestic commitment motive may also drive the solution to the free trade corner solution. $\hfill\blacksquare$

\vskip.2in
\noindent \textbf{\hypertarget{lem2}{Proof of Lemma~\ref{lem:2}}}: \\
Denote the tariff with no agreement as $\tau^N(e^N)$. The lobby's effort level in the absence of a trade agreement is determined by the first order condition
	\begin{equation}
		\frac{\partial \pi}{\partial \tau} \frac{\partial \tau^N}{\partial \ga} \frac{\partial \ga}{\partial e^N} = 1
			\label{eq:nta}
	\end{equation}
In order to compare this with Equation (\ref{eq:ea}) for weak bindings, note that
  \[
	  \frac{\partial e^{EP}}{\partial e^{EA}} = \frac{\partial e^{EP}}{\partial \tau^{EA}} \frac{\partial \tau^{EA}}{\partial e^{EA}} = \frac{1}{\frac{\partial \tau^N}{\partial e^{EP}}} \frac{\partial \tau^{E}}{\partial e^{EA}} =  \frac{\frac{\partial \tau^{E}}{\partial e^{EA}}}{\frac{\partial \tau^N}{\partial e^{EP}}} =  \frac{\frac{\partial \tau^{E}}{\partial \ga}\frac{\partial \ga}{\partial e}}{\frac{\partial \tau^N}{\partial \ga}\frac{\partial \ga}{\partial e}} =  \frac{\frac{\partial \tau^{E}}{\partial \ga}}{\frac{\partial \tau^N}{\partial \ga}} 
	\]
where the first equality follows by the fact that the impact of a marginal unit of ex-ante lobbying effort on ex-post effort is to raise the trade agreement tariff. This is a tariff cap, and so we need to calculate the effect of this increased cap on ex-post effort. The second equality represents that this is just the inverse of the marginal impact of ex-post lobbying effort on the non-cooperative tariff; the super-script on the $\tau$ in the denominator has been changed to $N$ to denote that ex-post tariff setting is non-cooperative; that on the $\tau$ in the numerator has been change to $E$ to denote that trade agreement tariff setting is cooperative. The fourth equality unpacks the partial derivatives to show the intermediate impact of $\ga$. This leads to the final expression in which the common terms have been canceled.

Multiplying Equation (\ref{eq:ea}) by $\frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^E}{\partial \ga}}$, we have
	\begin{equation}
	  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^N}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = \frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^E}{\partial \ga}} + 1.
			\label{eq:ea2}
		\end{equation}
Because $\frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^E}{\partial \ga}} > 0$, the solution to the ex-ante weak binding first order condition as given in Equation~(\ref{eq:ea2}) is smaller than the solution to Equation~(\ref{eq:nta}), so that the lobby exerts less effort in the ex-ante stage of trade agreement negotiations than when there is no trade agreement. Since $\tau^E(\cdot) < \tau^N(\cdot)$ for all $e$, $\tau^{EA}\left(e^{EA}\right) < \tau^N\left(e^N\right)$ when bindings are weak.

In contrast, the last term of the first order condition when bindings are strong is zero since the lobby does not exert any effort at the ex-post stage. The right hand side of Equation~\ref{eq:ea2} then has just the one term, and it can be either greater than or less than 1, so we cannot bound the ex-ante effort level. $\hfill\blacksquare$

\newpage
\printbibliography[title=References]

\end{document}

### 2015-12-29 ex_ante.tex
add ex_ante.tex and intro.tex from exante folder so that draft.tex will compile

%\documentclass[12pt]{article}

%\addtolength{\textwidth}{1.4in}
%\addtolength{\oddsidemargin}{-.7in} %left margin
%\addtolength{\evensidemargin}{-.7in}
%\setlength{\textheight}{8.5in}
%\setlength{\topmargin}{0.0in}
%\setlength{\headsep}{0.0in}
%\setlength{\headheight}{0.0in}
%\setlength{\footskip}{.5in}
%\renewcommand{\baselinestretch}{1.0}
%\setlength{\parindent}{0pt}
%\linespread{1.1}

%\usepackage{amssymb, amsmath, amsthm, bm}
%\usepackage{graphicx,csquotes,verbatim}
%\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
%\setlength{\bibitemsep}{\baselineskip}
%\usepackage[american]{babel}
%dell laptop
%\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
%\renewcommand{\newunitpunct}{,}
%\renewbibmacro{in:}{}

%\DeclareMathOperator*{\argmax}{arg\,max}
%\usepackage{xcolor}
%\hbadness=10000

%\newcommand{\ve}{\varepsilon}
%\newcommand{\ov}{\overline}
%\newcommand{\un}{\underline}
%\newcommand{\al}{\alpha}
%\newcommand{\expect}{\mathbb{E}}
%\newcommand{\ga}{\gamma}
%\newcommand{\Ga}{\Gamma}
%\newcommand{\de}{\delta}

%\begin{document}

In many situations, it is realistic to assume that, in addition to the ex-post lobbying examined throughout this text, lobbying also takes place before the trade agreement is signed to affect its terms. Sections~\ref{sec:rigid} and \ref{sec:objfcn} have shown that, although ex-post lobbying occurs after the trade agreement is signed in order to impact the level of the applied tariff, it alters the incentives of governments when they negotiate the trade agreement. Similarly, ex-ante lobbying is aimed directly at changing the level of the trade agreement tariff, but here lobbies only engage in this activity in order to alter what will be possible ex-post.

One of the main implications of the possibility of ex-ante lobbying is the addition of a new time at which the lobby may have to exert influence. As in Section~\ref{sec:strong}, when facing a trade agreement in which commitments take the form of strong bindings, the lobby's optimal ex-post effort choice is zero. However, when facing a tariff cap, protection will only be supplied at the ex-post stage if the lobby exerts effort. Thus, ex-ante lobbying adds a second time at which the lobby may have an incentive to exert effort in the case of a tariff cap; it introduces a first time at which the lobby may supply non-zero effort in the case of a strong binding.

For simplicity, here I consider ex-ante lobbying when political pressure is endogenous only, i.e. there is no exogenous, probabilistic influence on $\ga$. Since there is no possibility of ex-ante lobbying in the \Textcite{bs2005} framework this paper follows, it must be determined how ex-ante lobbying can best be integrated into the model.

The \Textcite{bs2005} framework posits that governments negotiate the trade agreement so that they will be able to set the applied tariff that best matches the political pressure they experience ex-post. In the case of endogenous political pressure, I argue that the governments will tune the trade agreement to deliver the amount of lobbying effort that is in their best interest ex-post. How does this change when lobbies can influence the terms of the trade agreement?

A lobby knows that the trade agreement will bind its future action, so exerting effort at the ex-ante stage only matters insofar as it improves ex-post profits net of both ex-ante and ex-post lobbying effort.

Looking at the policy-makers, with ex-post pressure only, the governments are completely forward looking: they set the trade agreement with a view toward maximizing expected joint ex-post welfare. In order for ex-ante lobbying to have an effect, it must be that governments also receive some utility from the setting of the trade agreement itself: perhaps they receive or lose political capital from various interested parties, and can be persuaded to forgo future benefits if ex-ante lobbying is strong enough to outweigh those future concerns.

Thus I model the influence of ex-ante lobbying on the governments' tariff-setting behavior as operating through a threshold. Recall how the trade agreement tariff is determined under ex-post lobbying only. Because the trade agreement can act as a commitment device, the governments choose the tariff level $\tau^R_{\cdot,e}$, and implicitly ex-post effort $e^{EP}=e^R_{\cdot,e}$, to maximize their joint political welfare as in Equation~\ref{exp:1}, subject only to the constraint that $\ga$ will be realized according to the ex-post lobbying process.

As discussed in Section~\ref{sec:objfcn}, depending on how government welfare changes as political pressure increases, the governments may be looking ahead and manipulating ex-post lobbying incentives when there is only ex-post lobbying. When ex-ante lobbying is possible the governments may have to raise the trade agreement tariff above $\tau^R_{\cdot,e}$ in order to respond to that pressure.

This is modeled here with a simple cut-off. The trade agreement tariff cap will be set according to $\ga\left(e^{EA}\right)$ instead of at $\tau^R_{\cdot,e}$ if $\ga\left(e^{EA}\right)$ would lead to a higher trade agreement tariff. If, on the other hand, $\tau = \frac{4(\ga(e^{EA})-1)}{25-4\ga(e^{EA})} < \tau^R_{\cdot,e}$, the governments respond to the higher ex-post pressure and set $\tau^R_{\cdot,e}$.

Thus the governments lose some measure of control with ex-ante lobbying. If $e^{EA}$ is low enough, ex-ante lobbying has no impact. But if $e^{EA}$ is above the threshold, the trade agreement tariff is strictly greater than $\tau^R_{\cdot,e}$ and so joint government welfare will not be at its globally optimal level. The trade agreement the governments choose will be constrained by ex-ante pressure.

We will see that this change to the government's preferences has the impacts one would expect, at least in the case of weak bindings (tariff caps): net profits for the lobby (weakly) increase, and net political utility for the governments (weakly) decrease. We begin with the less clear-cut case of strong bindings.

\subsection{Strong Bindings}
Let us first look at how ex-ante lobbying impacts trade agreements that involve strong bindings, or exact tariff commitments. We know that when there is only ex-post lobbying, $e^{EP} =0$ and $\tau^{EP} = \tau^R_{S,e}$. Since ex-ante effort will be sunk ex-post, the ex-post stage of the game that includes ex-ante lobbying is identical to the game without ex-ante lobbying. In the case of exact tariff commitments, because $e^{EP} = 0 \ \forall e^{EA}$, the lobby does not consider the impact of ex-ante lobbying on the ex-post effort level.

This thus looks like a problem with no ex-post lobbying, but where there is perfect enforcement of the tariff commitment. This collapses to a simple problem of maximizing joint government welfare with $e=e^{EA}$, and the lobby maximizing profits net of ex-ante effort facing the joint maximization of the governments.

Without knowing how government welfare varies in $\ga$, we can't say for sure the effect on government welfare at the ex-ante stage. As with the example in Figure~\ref{fig:weight}, joint welfare could fall but could also rise. Ex-post political pressure will not match ex-ante political pressure: because of the incentives involved with strong bindings, the home government will experience $\ga(0)$ but be forced to apply the ex-ante tariff, so this is a welfare loss relative to the no ex-ante case at this stage.\footnote{This is a government welfare loss from exact tariff commitments that is not accounted for in \Textcite{mrc2007} because of their assumption that the government has no bargaining power. It seems to be closely related that, in their model,  exact tariff commitments and tariff caps are equivalent in the case of perfectly immobile capital. As will be shown below, that is not the case here.} There is a very clear gain for the lobby relative to the case where there is no ex-ante lobbying; it can avoid any loss by setting $e^{EA}=0$.

\subsection{Weak Bindings}
In the case of weak bindings, or tariff caps, the ex-post stage will also work as in the no-ex-ante-lobbying case, that is, as in Section~\ref{sec:weak}, again because ex-ante effort is sunk at the ex-post stage. To recap, when there is no ex-ante lobbying---or ex-ante-lobbying is less important than ex-post lobbying---the governments choose the $\tau^R_{W,e}$ that leads to the level of ex-post political pressure that maximizes ex-ante joint welfare
		\[
		  W(\ga(e^R_{W,e})),\tau^R_{W,e}) + W^*(\tau^R_{W,e})
		\]
		knowing that the home government will set the applied tariff in the third stage unilaterally (i.e. $\tau^N(\ga(e^R_{W,e})) = \frac{8\ga(e^R_{W,e}) - 5}{68 - 8 \ga(e^R_{W,e})} = \tau^R_{W,e}$).

What is different here is that the tariff cap could be set higher than this level because of the presence of ex-ante lobbying. The lobby chooses its ex-ante effort level $e^{EA}$ taking into account that the governments will set either the optimal ex-post tariff cap $\tau^R_{W,e}$ or a higher cap in ex-ante stage given the $e^{EA}$ with which it is confronted.

The lobby maximizes $\pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA})$, with a first order condition of 
	\begin{equation}
	  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^E}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = 1 +\frac{\partial e^{EP}}{\partial e^{EA}}
			\label{eq:ea}
	\end{equation}
This condition illustrates that, when it is operative, the governments choose the tariff cap according to Expression~\ref{exp:1} with $e^{EA}$, but the lobby still has to exert effort ex-post to receive the applied tariff. $e^{EP}$ is the amount the lobby pays according to Equation~\ref{eq:nash} to achieve $\tau^E(e^{EA})$ (see Equation~\ref{eq:eff}).

The lobby sets $e^{EA}$ according to this condition as long as 
	\begin{equation}
		\pi(\tau^E(e^{EA})) - e^{EA} - e^{EP}(e^{EA}) > \pi(\tau^N(e^R_{W,e})) - e^R_{W,e}.
		\label{ine:ea}
	\end{equation}
Otherwise, it maximizes net profits by setting $e^{EA} =0$.

Thus if the lobby's incentives to exert pressure ex-ante are strong enough, they overpower the government's ability to use the trade agreement to directly manipulate lobbying incentives ex-post. The lobby can force the government to set a cap higher than the level that would optimize its welfare when there is no ex-ante lobbying. The lobby has no interest in a lower cap and can choose a zero effort level whenever positive effort would reduce its nets profits below the level with ex-post lobbying only. Therefore we have the following result:

\begin{result}
	When trade agreements take the form of weak bindings, the binding under ex-ante lobbying can only be (weakly) higher than when ex-ante lobbying is not possible. This can only (weakly) reduce the joint political welfare of the governments and it can only (weakly) increase net profits for the lobby.
\end{result}

\subsection{The Role of Trade Agreements with Ex-Ante Lobbying}				
To understand the function trade agreements serve in the presence of ex-ante lobbying when lobbying is endogenous, it is helpful to establish some facts about the levels at which tariff are set under various scenarios.

\begin{lemma}
  Trade agreements with ex-post lobbying lead to weakly lower tariffs (when bindings are either strong or weak) than those that are applied in the absence of an agreement.
	\label{lem:1}
\end{lemma}

See the \hyperlink{lem1}{Appendix} for the proof. Intuitively, both the terms-of-trade and domestic-commitment motives drive the governments to chose a (weakly) lower tariff than that which the home government chooses when there is no agreement.

It has been established above that the presence of ex-ante lobbying can only weakly increase tariffs from the ex-post agreement level with both strong and weak bindings. Lemma~\ref{lem:1} only establishes that the tariffs with ex-post lobbying are lower than the tariffs with no agreement; we would like to be able to derive a clear result comparing the tariff level with ex-ante lobbying to that with no agreement as well. Lemma~\ref{lem:2} does just that, but this is only possible in the case of weak bindings.

\begin{lemma}
  When bindings are weak, trade agreements with ex-ante lobbying lead to lower tariffs than those that are applied in the absence of an agreement. When bindings are strong, tariffs with ex-ante lobbying may be either higher or lower than those that are applied in the absence of an agreement.
	\label{lem:2}
\end{lemma}

See the \hyperlink{lem1}{Appendix} for the proof. The key insight here is that every unit of effort ex ante results in more political support than it would ex post because producer profits are counted the same while the terms of trade motivation is absent. With stronger incentives on the margin ex ante, we must rule out that the lobby exerts sufficiently more effort to overcome the internalization of the terms of trade externality that occurs through the trade agreement negotiation. The fact that the lobby must pay a second time, as well as from a lower base, is sufficient to mute its incentives at the ex-ante stage in the case of weak bindings; because it pays nothing at the ex-post stage with strong bindings, it is possible (although not likely) that the heightened incentives on the margin could lead to sufficient effort to produce a higher trade agreement tariff than that which the lobby would achieve with unilateral tariff setting.

\begin{result}
	When trade agreements take the form of weak bindings and political pressure is endogenous, tariffs are always lower under trade agreements than in the absence of a trade agreement. 
\end{result}

This follows directly from Lemmas~\ref{lem:1} and \ref{lem:2}, since tariffs under ex-ante lobbying are set either in line with lobbying effort according to the first order condition (\ref{eq:ea}) or via ex-post incentives. Although Lemma~\ref{lem:2} shows that ex-ante effort will be strictly lower than effort under no trade agreement, it is possible that ex-ante effort will be nil and the trade agreement level will be set by ex-post incentives that provide for a tariff at the non-cooperative level (see Section~\ref{sec:objfcn} for an example).\footnote{Note that it is possible to have positive ex-ante lobbying when $e^R_{W,e}$ is restrictive, notably in the case where joint government welfare is concave in $\ga$ with a strict interior maximum. This implies that government welfare must eventually decrease in $\ga$, as in \Textcite{ethier2012} where there are diminishing returns to lobbying as tariffs increase. In the current model, one way this can arise is if $\ga$ is taken to be a decreasing function of $\tau$.}

A similar result clearly does not hold for strong bindings, and adds to the list of potential reasons why we do not see trade agreements take this form in reality.

This contrasts with \Textcite{mrc2007}. In their model, which also features endogenous political pressure with ex-ante lobbying, tariff caps and exact tariff commitments are equivalent in the case of perfectly immobile capital. This seems to be the result of their assumption, within the context of a cooperative bargaining solution for the trade agreement phase, that the lobby has all the bargaining power and thus receives all rents. Because the lobby receives protection ex post without exerting effort under exact tariff commitments, these two types of bindings lead to very different decisions ex-ante and splits of the surplus ex-post in the model under consideration here.\footnote{One reason the lobby is not included in the cooperative bargaining formulation here is that it leads to the strange outcome that the governments overweight the lobby's profits even when the lobby exerts no effort ex ante.}

Focusing on the case of weak bindings (i.e. tariff caps), we see that, as long as governments want to restrain lobbying, relative to non-cooperative levels, trade agreements will deliver this result even when ex-ante lobbying is possible. If a government wants to encourage political pressure, it may want to consider eschewing trade agreements, although the tradeoff is losing the ability to internalize the terms of trade externality. If a government wants to restrain political pressure, a trade agreement is doubly good.

There are several avenues through which trade agreements mute lobbying incentives. First, according to Condition~\ref{ine:ea}, ex-ante lobbying may be too costly since it adds a second time at which the lobby has to exert effort. When the lobby does find it advantageous to participate in ex-ante lobbying, although on the margin the incentives are stronger, the same effort produces a lower trade agreement tariff ex-ante than ex-post because $\ga$ is translated via Equation~\ref{eq:eff} instead of Equation~\ref{eq:nash} as when there is only ex-post lobbying. Thus not only must the lobby exert effort a second time, but it must work harder for the same level of protection ex-ante.

This points out an important role of trade agreements vis-\`{a}-vis political pressure groups: relative to ex-post lobbying, ex-ante lobbying gives them more ability to exert influence, but at a significantly higher cost. If either that higher cost or some institutional constraint makes only ex-post lobbying feasible, the governments can use the trade agreement to exert more significant control over the trade policy outcome.

%\end{document}

### 2015-12-29 intro.tex
add ex_ante.tex and intro.tex from exante folder so that draft.tex will compile

Much of the work on the political economy of trade agreements focuses on questions of the optimal design of trade agreements, trade agreement negotiations, and trade dispute settlement that arise in the presence of asymmetric information about shocks to an exogenous political economy parameter.

One of the basic ideas that emerges from this literature is that, in the presence of asymmetric information about the strength of the ex-post political economy shocks, it is often advantageous to grant governments a period of relief from trade commitments. That is, one would rather allow a short period of ``escape'' from the agreement rather than have the agreement abandoned forever because domestic political opposition is temporarily too strong to be resisted.

Various trade agreement and institutional design features have been proposed in the literature that can help to make an escape clause work, since it is not incentive compatible to allow governments to take advantage of an escape clause whenever they wish. If some cost can be associated with the use of the escape clause or a dispute settlement body (DSB) such as that at the World Trade Organization (WTO) can procure a signal about the strength of $\ga$, it is often advantageous to grant this period of relief from trade commitments if the signal is strong enough.

%In this vein, \Textcite{bs2005} find that governments prefer tariff caps to strong bindings when there is no escape clause, but that impatient governments may need an escape clause in order to make cooperation sustainable.

This is an intuitively appealing story, but the logic can break down in the presence of endogenous political pressure. An escape clause allows a government to apply a higher tariff barrier when it experiences intense political pressure. But if a government gets a free pass from the WTO whenever it feels sufficient political pressure from domestic interest groups, those interest groups have a strong incentive to exert the required level of pressure, eviscerating the escape clause.

This is one example of a design question whose answers are sensitive to assumptions about the endogeneity of political pressure. In order to examine it and others, I build a model that closely follows \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights. I find that the presence of endogenous political pressure changes the optimal choice of both strong and weak tariff bindings (exact tariff commitments and tariff caps, respectively) in quite significantly ways compared with political pressure that is assumed to be exogenous.

As in \Textcite{mrc2007}, I show that one of the uses of tariff caps is to incentivize the lobby to engage in the political process after the trade agreement is in place. Once the trade agreement is in place, the lobby loses all incentive to exert effort under a strong binding and so a government who prefers some level of ex-post political engagement from the lobby will not use a strong binding. This result extends \Textcite{mrc2007} to the standard case, examined here, where specific factors are completely immobile.

This, perhaps, provides an explanation for the ubiquity of tariff caps. Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.

In order to see this, I compare welfare under the standard \Textcite{baldwin}-style government objective function and a similar, fully-weighted version of the government's objective function. At issue is the fact that the standard version is everywhere increasing in political pressure. In an environment with endogenous political pressure, this leads inexorably to the conclusion that governments never want to discourage special interest groups from exerting pressure. The slight modification to the government objective function demonstrates that governments may indeed want to use trade agreements to reduce lobbying.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.}

Thus, examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.

This example shows that it may be important to consider how the features of the government objective function interact with the question being asked. For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the shape of the objective function is not crucial, there may be no loss in employing the simple form. However, for other questions that directly involve the government maximizing with respect to political pressure, a more general treatment may be prudent.\footnote{Note that, although \Textcite{gh94} provide microfoundations for one particular \Textcite{baldwin}-style objective function, it is a very specific form with fixed weights in which $1+a$ is attached to the surplus of groups who lobby, and $a$ to the surplus of those who don't lobby. This cannot microfound a flexible model as in \Textcite{longvousden} in which weights depend on lobbying activity or in which weights are the result of shocks.}

In a more general setting, I show that, depending on the government's objectives, in some cases extra political pressure would reduce the government's political welfare \textit{and} trade agreements can help them alleviate this pressure. I demonstrate that this is true whether or not lobbying is possible ex ante---that is, whether or not special interests are able to influence the formation of the trade agreement. Even in the face of ex-ante lobbying, a trade agreement that takes the form of tariff caps works to restrict lobbying in many cases, but also to keep lobbyists ``in the game.'' This broadens the result from \Textcite{mrc2007}, in which the political commitment role of trade agreements is present only when capital has some degree of mobility; here it holds with specific factors that are perfectly immobile. I present an extensive-form game that helps us to see that trade agreements serve this role even when ex-ante lobbying is possible. Although ex-ante lobbying removes some degree of autonomy from the government, it also screens out groups for whom the extra stage of lobbying becomes unprofitable.

In another example, I show that the same combination---endogenously-determined political pressure and a general approach to the government's objective function---can explain why the conditions for invoking the WTO Safeguards measure rely on purely observable economic variables, and why the level of protection governments can choose when invoking a Safeguard is not restricted. The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role that is intended for it.

Through the lens of these examples, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective function have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.

A rich literature has been developed to address questions concerning the design and enforcement of trade agreements. Repeated non-cooperative game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.

Tariff caps have received particular attention in the literature that takes political pressure to be exogenous. The basic result on the political efficiency of tariff caps established in \Textcite{bs2005} has been extended in numerous directions. \Textcite{hms} show that the basic logic concerning tariff caps is unaltered in a model with contracting costs and multiple policy instruments, while \Textcite{ls} show that the result holds when fines are allowed as punishments, although the politically efficient tariffs may not be achievable when enforcement of the punishment is taken into consideration.

\Textcite{ab2012} consider that the weight that tariff revenue receives in the government welfare function may be private information and show that tariff caps remain optimal under certain conditions, while \Textcite{ab2013} show that the result goes through when the \Textcite{bs2005} model is generalized to include monopolistic competition as well as more general payoff and distribution function. \Textcite{bb} extend the theory to an environment with asymmetric country sizes with costly state verification and show that tariff caps and escape clauses can both emerge endogenously in optimal trade agreements. \Textcite{bbr} and \Textcite{nos} provide evidence on the relationship between tariff caps, binding overhang and market power that confirms the predictions of the terms of trade theory.

Maggi and Staiger have a series of papers that employ an exogenous political economy force to study questions about the design of trade agreements and trade dispute settlement. \Textcite{ms2012a} study the conditions under which property versus liability rules will be optimal when renegotiation of agreements is possible. They find that when property rules are optimal, agreements are never renegotiated; only liability rules are renegotiated in equilibrium, and when this renegotiation occurs, it always results in trade liberalization. \Textcite{ms2013} builds on this work to answer questions about when governments will settle disputes and how this relates to the contracting environment. \Textcite{ms2011} has a more sophisticated set-up for the exogenous shock that allows the authors to speak to issues of the role of the dispute settlement body as interpreter and completer of incomplete contracts.

\Textcite{beshkar2010a} shows that when one assumes that utility is not transferable between countries as has become common in the literature, the optimal mechanism involves less-than-proportional retaliation against parties who have defected from the agreement. \Textcite{beshkar2010b} compares the GATT escape clause to the WTO Safeguards agreement and shows that the DSB as a non-binding arbitrator can assist governments in self-enforcing their trade agreements. \Textcite{martinvergote} demonstrate that future punishment provides for higher welfare than contemporaneous punishment when governments are sufficiently patient. Indeed, they show that retaliation is a necessary feature of any efficient equilibrium in this environment. \Textcite{hungerford} and \Textcite{riezman1991} also consider the impact of different assumptions about reactions and timing of punishments for deviations from agreements.

This work is also related to the literature on the endogenous political economy of trade. The foundational work is \Textcite{gh94}; the insights are applied to trade agreements in \Textcite{gh95}. \Textcite{mrc2007} advanced the literature by demonstrating that there is a domestic commitment role for trade agreements. \Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government has a separation-of-powers structure, while \Textcite{coatesludema} demonstrate that, in the presence of opposition to a trade agreement from foreign lobbies, it may be optimal to liberalize trade unilaterally.

In the next section, I present the model and some preliminary results. Section~\ref{sec:rigid} contains the analysis of rigid tariffs while Section~\ref{sec:escape} explores trade agreements with escape clauses. The political welfare function of the government is examined in depth in Section~\ref{sec:objfcn} before returning to examine the escape clause in an environment with both endogenous and exogenous political pressure in Section~\ref{sec:escape2}. Section~\ref{sec:exante} details the extension to ex-ante lobbying. Section~\ref{sec:concl} concludes.

### 2015-12-29 feedback.tex
add copy of feedback.tex from exante and compile

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
Feedback on Jan. 10 draft
\end{center}

Chad Bown, Jan. 22, 2015
\begin{itemize}
	\item What it's missing: intersection with reality
		\begin{itemize}
			\item Anecdotes
			\item How policy is used
				\begin{itemize}
					\item I thought that escape clause implications \textit{were} this
				\end{itemize}
			\item Something like how GH predicts export taxes but it doesn't work that way in the real world
			\item Well articulated question / paradox
		\end{itemize}
	\item Look at ``Theory of Managed Trade'' model motivation (BS)
	\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
		\begin{itemize}
			\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
			\item You can have one good idea in a paper, the rest get lost
		\end{itemize}
	\item How did Devashish structure his AER paper: the argument, the model?
		\begin{itemize}
			\item This paper is structured differently than the norm---is there a good reason to do it?
			\item Don't want to challenge what people are used to unless I really need to
		\end{itemize}
	\item In general, how do AER papers \textit{look}?
	\item Intro doesn't reflect body of paper
	\item What kinds of empirical questions might we ask with this model?
		\begin{itemize}
			\item What would be different because we use this model?
		\end{itemize}
\end{itemize}

\end{document}

### 2015-12-29 feedback.tex
add maurizio and ben's comments to feedback.tex

@@
 
+\newpage
+Ben Zissimos, May 11, 2015
+
+\begin{itemize}
+	\item I think the topic of tariff caps and how they can be motivated by endogenising lobbying effort is a very interesting one and this literature seems to be wide open.
+	\item I also think it is a nice approach to extend the Bagwell-Staiger-2005 framework by endogenising their gamma parameter.
+	\item You do a lot in the paper, but for me the really interesting contribution is the one you highlight in the last couple of sentences of the abstract:
+\begin{quote}
+The presence of endogenous politics can also destroy an escape clause's ability to provide flexibility in times of large negative political shocks when lobbies use the flexibility to seek rents.
+This can explain why use of WTO Safeguards are conditioned on measurable economic indicators as well as why Safeguard levels of protection are not regulated.
+\end{quote}
+\end{itemize}
+
+Here are some thoughts on which particular aspects of your paper I got excited about and why.
+\begin{enumerate}
+	\item I started to get really excited about the paper around page 19, especially with Result 3.
+The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.
+In fact I was itching to get to that part right from when you discussed it in the introduction.
+I also found the results in the static framework reasonably predictable.
+So I'm wondering whether you could dispense with most of the static analysis and get into the repeated game more quickly.
+
+	\item What I found most interesting was the effect you discussed on the 2nd paragraph of page 28.
+
+``There is a large parameter space over which an escape clause could not function in the way it is intended.
+When the realization of the shock is high, the government will report the high shock and apply the escape clause tariff.
+But when the shock is low, the lobby will exert effort so that the sum of gamma(s) and gamma(e) equals the high realization.
+From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section 5, the governments' ex-ante welfare is reduced by excess lobbying.''
+
+I think the effect the paper identifies here is critical to the operation of trade policy making under the rules of the WTO and to the best of my knowledge it has not received attention in the literature.
+I would like to see the paper do much more with this.
+Here are some thoughts on how you could do this.
+
+First, the paper makes the extremely important point above almost in passing.
+I would be interested to see a more detailed examination of how this works.
+It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.
+I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
+
+Then the paper begins to address the equally important question of what can be done about this problem, but again stops short of providing a really compelling answer.
+I think it would be worth exploring this in detail.
+The paper does get to the heart of the issue, which is that `s' can be observed by the DSB but e cannot.
+But the paper then moves on to ex-ante lobbying before getting to the bottom of how to design a dispute settlement system that resolves the conflict created by `e vs s'.
+
+	\item Following on from the previous point, the one institutional feature I kept expecting the paper to talk about was a `withdrawal of equivalent concessions' (WEC).
+As you probably know, the only legal response to a breach of an agreement under the GATT/WTO is WEC: if your partner denies you $\$$1m worth of market access promised to you under an agreement then that's what you're allowed to withdraw from them.
+One interesting feature of WEC highlighted by Zissimos (2007) is that a government `chooses the severity of its own punishment' by the extent of its initial deviation.
+In that paper, I took WEC for granted and then played out trade liberalization according to its rules.
+But this left open the question of why WEC made sense as an approach to punishment.
+Now I'm wondering whether your framework offers an answer.
+Under WEC, it is worth deviating from the agreement in proportion to `s' because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
+But WEC might eliminate the incentive to respond to lobbying pressure `e' because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
+
+\textbf{I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated: one that applies if the trade partner keeps to the agreement and one that applies if they deviate.
+As far as I know this latter approach does not mirror at all the institutional features actually set out by the GATT/WTO.}
+
+	\item I like the way you use Bagwell and Staiger (2005) as your benchmark.
+But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.
+I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.
+To keep things manageable, I wonder whether you could restrict attention to the case of ex post lobbying.
+This is where all the action is, right?
+In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.
+But your mechanisms seem quite different.
+They have capital mobility whereas your factors are specific to each sector.
+If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.
+What drives this feature in your model?
+Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.
+What is the relationship between lobby effort level and tariff policy?
+Is effort fixed first, after which the government sets its tariff?
+If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.
+It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.
+Having done that, you could bring in ex-ante lobbying and treat it the same way.
+
+	\item I'm wondering how you view the time-frame over which your model is set.
+On one hand a specific-factors set-up is normally associated with the short-run.
+But on the other hand you have an infinite time horizon which suggests very long run.
+
+	\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
+I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  Their idea is that if you want to use the escape clause then this comes at the cost of providing information to the GATT panel to persuade them that you really have been hit by a shock.
+This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
+How does your approach relate to costly state verification?
+
+	\item I'm wondering whether footnote 18 provides a general way of addressing the issues arising from the endogeneity of gamma that you illustrate in Figures 1$\&$2.
+\end{enumerate}
+
+\vskip.3in
+Minor comments
+\begin{enumerate}
+	\item Page 1:  Should define gamma before using it, especially given how critical it is to the paper.
+	\item Page 17: `is invariant'
+	\item Page 26: `an interior'
+	\item I'm wondering why are the results referred to as `Results' and not `Propositions'?
+Odd question when I put it like that!
+But since we're used to `Propositions', saying `Results' is a bit distracting.
+Or maybe I'm just being old fashioned ;-)
+\end{enumerate}
+
+\newpage
+Maurizio Zanardi, June 22, 2015
+\begin{itemize}
+	\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
+	\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper.
+However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government.
+Is this aspect crucial for the point you want to make?
+If yes, it should be integrate with the previous object from the beginning.
+If not, it should be postponed a bit.
+	\item I also find the introduction too long.
+You could/should split the literature review that begins on page 4 to a separate section.
+And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it.
+Again, your main point seems to get lost into the discussion instead of standing out.
+	\item If above I suggest to add a section, I think that the paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
+The analysis is also split between strict and weak commitments.
+Which scenario clearly develops the point/contribution you want to make?
+Shouldn't the other one occupy a smaller place (i.e.
+space) of the paper?)
+	\item Finally, introducing ex-ante lobbying is not simply an "extension"...
+it seems to me a more fundamental issue in itself: should it be developed in a separate paper?
+Should the intuition only developed in the current paper (and the formal analysis moved to an appendix)?
+	\item In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
+Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point (and in general it seems to me that you seem to "dependent" on Bagwell and Staiger...
+(e.g.
+beginning of Sections 4 and 6).
+	\item In a sort of conclusion, I think that you need to think a bit more about the objective(s) of the paper and how best to achieve them. And I hope that some of the above is useful!
+\end{itemize}
+
 \end{document}

### 2015-12-30 draft.tex
fix two typos identified by Ben Z.

@@
 We will see that moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem.
-However, it does not alter the within-period costs and benefits and so in the case of endogenous political pressure we will see the same result that the optimal trade agreement tariff in invariant to the enforcement conditions.
+However, it does not alter the within-period costs and benefits and so in the case of endogenous political pressure we will see the same result that the optimal trade agreement tariff is invariant to the enforcement conditions.
 
@@
 By setting a tariff cap of zero---assuming the conditions for self enforcement outlined in Section~\ref{sec:selfweak} are met---the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.
-The stark prediction that the best the government can do is by setting the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has a interior global maximum in $\ga$.
+The stark prediction that the best the government can do is by setting the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has an interior global maximum in $\ga$.
 

### 2015-12-30 draft.tex
rename results to propositions

@@
 
-\begin{result}
+\begin{proposition}
   When political pressure is entirely endogenous, under the equilibrium trade agreement with strong bindings the lobbies exert no effort and are afforded no protection if governments have no inherent bias toward protection.
 	\label{res:strong}
-\end{result}
+\end{proposition}
 
@@
 
-\begin{result}
+\begin{proposition}
     (a) When political pressure is entirely endogenous, governments prefer negotiating commitments that take the form of weak bindings over strong bindings.
@@
 		\label{res:weak}
-\end{result}
+\end{proposition}
 
-Note that part (a) of Result~\ref{res:weak} matches Proposition 1 of \Textcite{bs2005}.
+Note that part (a) of Proposition~\ref{res:weak} matches Proposition 1 of \Textcite{bs2005}.
 The similarities are cosmetic only; the result holds in the case of endogenous $\ga(e)$ because of the extreme result for strong bindings that no lobbying effort is exerted and the trade agreement is at free trade.
@@
 
-\begin{result}
+\begin{proposition}
   The presence of the lobby as a repeated-game player may imply that the politically optimal self-enforcing trade agreement tariff is strictly greater than $\tau^R_{W,e}$. 
-\end{result}
+\end{proposition}
 
@@
 \noindent \textbf{\hypertarget{lem1}{Proof of Lemma~\ref{lem:1}}}: \\
-In the case of strong bindings, this is immediate from Result~\ref{res:strong}. In the case of weak bindings, first examine the first order condition for unilateral tariff setting:
+In the case of strong bindings, this is immediate from Proposition~\ref{res:strong}. In the case of weak bindings, first examine the first order condition for unilateral tariff setting:
 \[

### 2015-12-30 ex_ante.tex
rename results to propositions

@@
 
-\begin{result}
+\begin{proposition}
 	When trade agreements take the form of weak bindings, the binding under ex-ante lobbying can only be (weakly) higher than when ex-ante lobbying is not possible.
 This can only (weakly) reduce the joint political welfare of the governments and it can only (weakly) increase net profits for the lobby.
-\end{result}
+\end{proposition}
 
@@
 
-\begin{result}
+\begin{proposition}
 	When trade agreements take the form of weak bindings and political pressure is endogenous, tariffs are always lower under trade agreements than in the absence of a trade agreement. 
-\end{result}
+\end{proposition}
 

### 2015-12-30 to-do.tex
create to-do.tex with just WEC parts of feedback.tex

--- 2015-12-29 feedback.tex
@@
 \usepackage{amssymb, amsmath, amsthm, bm}
+\usepackage[normalem]{ulem}
 \usepackage{graphicx,csquotes,verbatim}
@@
 \begin{center}
-Feedback on Jan. 10 draft
+PE Shocks To-Do list
 \end{center}
 
-Chad Bown, Jan. 22, 2015
+December 30, 2015
 \begin{itemize}
-	\item What it's missing: intersection with reality
+	\item From Chad: it's missing intersection with reality
 		\begin{itemize}
-			\item Anecdotes
-			\item How policy is used
-				\begin{itemize}
-					\item I thought that escape clause implications \textit{were} this
-				\end{itemize}
-			\item Something like how GH predicts export taxes but it doesn't work that way in the real world
 			\item Well articulated question / paradox
 		\end{itemize}
-	\item Look at ``Theory of Managed Trade'' model motivation (BS)
-	\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
+	\item From Ben
 		\begin{itemize}
-			\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
-			\item You can have one good idea in a paper, the rest get lost
+			\item It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.
+I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
+			\item `Withdrawal of equivalent concessions' (WEC).
+				\begin{itemize}
+					\item Zissimos (2007): a government `chooses the severity of its own punishment' by the extent of its initial deviation.
+In that paper, I took WEC for granted and then played out trade liberalization according to its rules.
+But this left open the question of why WEC made sense as an approach to punishment.
+Now I'm wondering whether your framework offers an answer.
+					\item Under WEC, it is worth deviating from the agreement in proportion to $s$ because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
+					\item But WEC might eliminate the incentive to respond to lobbying pressure $e$ because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
+					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated: one that applies if the trade partner keeps to the agreement and one that applies if they deviate.
+As far as I know this latter approach does not mirror at all the institutional features actually set out by the GATT/WTO.
 				\end{itemize}
-	\item How did Devashish structure his AER paper: the argument, the model?
-		\begin{itemize}
-			\item This paper is structured differently than the norm---is there a good reason to do it?
-			\item Don't want to challenge what people are used to unless I really need to
-		\end{itemize}
-	\item In general, how do AER papers \textit{look}?
-	\item Intro doesn't reflect body of paper
-	\item What kinds of empirical questions might we ask with this model?
-		\begin{itemize}
-			\item What would be different because we use this model?
 		\end{itemize}

### 2015-12-30 to-do.tex
add notes on WEC to to-do.tex

@@
 
+\vskip.5in
+Possible puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round
+\begin{itemize}
+	\item Read Chad's ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), at \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
+		\begin{itemize}
+			\item Need to understand better what changed, both with safeguard and with other policy instruments
+		\end{itemize}
+	\item There is this drop off in usage of escape clause (EC) after Uruguay Round (UR) reforms, even though retaliation prohibited for first three years. Why, when retaliation goes \textit{away}?
+	\item I'm not sure there's a way to make sense of Ben's point: in my set-up, government doesn't ``get'' something different from lobby than it does directly from the shock. Maybe it should.
+		\begin{itemize}
+			\item Government feels $\ga$ the same whether it's elevated because of $s$ or because of $e$
+			\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
+		\end{itemize}
+	\item Need to go to continuous value of EC tariff
+		\begin{itemize}
+			\item There will be optimal tariff for whatever value of $s$ is realized
+				\begin{itemize}
+					\item Not sure how fine I want to go with $s$
+				\end{itemize}
+			\item Lobby will choose its optimal level of $e$. Will this be constrained in any way? Probably not, if there isn't a fixed EC-binding.
+		\end{itemize}
+\end{itemize}
+
 \end{document}

### 2016-01-01 draft.tex
remove ex ante section (and proofs); just reference in intro remains

@@
 
-\section{Ex-Ante Lobbying}
-\label{sec:exante}
+%\section{Ex-Ante Lobbying}
+%\label{sec:exante}
 
-\input{ex_ante}
+%\input{ex_ante}
 
@@
 
-The discussion in Section~\ref{sec:exante} presumes both that there are no significant interactions between lobbies and that governments are free to choose which sectors are part of a trade agreement, but these decisions are part of a more complicated game that is, in particular, influenced by WTO rules.
-Taking into account the interactions between multiple lobbies, the influence of WTO rules, and differences in the $\ga$ functions across industries---that is, in how political pressure translates into weight in the policy-making process---seem like important and interesting avenues for further study.
+%The discussion in Section~\ref{sec:exante} presumes both that there are no significant interactions between lobbies and that governments are free to choose which sectors are part of a trade agreement, but these decisions are part of a more complicated game that is, in particular, influenced by WTO rules.
+
+Taking into account the interactions between multiple lobbies, the influence of WTO rules, and differences in the $\ga$ functions across industries---that is, in how political pressure translates into weight in the policy-making process---seems like an important and interesting avenues for further study.
 
@@
 
+\begin{comment}
 \vskip.2in
@@
 $\hfill\blacksquare$
+\end{comment}
 

### 2016-01-01 draft.tex
revise intro to subsection{Perfect External Enforcement}

@@
 I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement.
-Although it is widely agreed that this is unrealistic, this is an important baseline case.
+Although it is widely agreed that this is unrealistic, this is an important baseline case, as there is no incentive compatibility problem because the contractually-allowed tariff may not vary with $\ga$.
 Section~\ref{sec:self} will relax this assumption and require that trade agreements be self-enforcing.
 
-I compare strong bindings, in which the terms of the agreement are that tariffs must be set at the precise value stipulated and weak bindings---commonly referred to as tariff caps---in which the applied tariff can take any value as long as it is no greater than that which is agreed upon.
-To reiterate, with either form of rigid tariff, there is no incentive compatibility problem because the contractually-allowed tariff may not vary with $\ga$.
+I restrict attention to the case of weak bindings throughout as this is the more interesting and realistic case.\footnote{There \textit{is} an extreme difference in outcome for strong bindings between exogenous and endogenous cases.
+When the terms of the agreement are that tariffs must be set at the precise value stipulated, the lobby has no incentive to exert effort and the trade agreement tariff is set at zero.
+Results are available upon request from the author.} Weak bindings are commonly referred to as tariff caps.
+When bindings are weak, the applied tariff can take any value as long as it is no greater than that which is agreed upon.
 

### 2016-01-01 draft.tex
remove sec:strong subsection

@@
 \vskip.2in
-\subsubsection{Strong Bindings}
-\label{sec:strong}
-When there is perfect external enforcement of an agreement that specifies the exact tariff each government must adhere to, \Textcite{bs2005} derive the result that the optimal rigid, strong binding $\tau_S^R$ will be set at the level that is politically efficient for the expected realization of political pressure, $\ga(s)$.
-In their environment, $\ga(s)$ is exogenously determined and so the optimal trade agreement tariff is chosen as
-\[
-  \tau_S^R = \frac{4\left[\expect \ga(s) - 1\right]}{25 - 4\expect \ga(s)}
-\]
-which is simply Equation~\ref{eq:eff} evaluated at the expected $\ga$.
-
-The optimal strong binding is very different if we instead allow political pressure to be endogenously determined.
-In this simple setting with only endogenous political pressure and no exogenous shocks, there is no uncertainty; the level of political pressure can be determined through backward induction.
-
-In the final stage, the external enforcement agency compels the government to apply the precise tariff listed in the trade agreement contract regardless of circumstances.
-Call this $\tau_{S,e}^R$ to distinguish the strong (rigid) binding in the case of endogenous pressure from $\tau_S^R$ when political pressure is exogenous above.
-
-Moving back to the lobbying stage and given that the government will be compelled to apply $\tau_{S,e}^R$ regardless of the $\ga(e)$ it faces, the lobby maximizes its payoffs by choosing $e=0$.
-Making the binding strong removes any incentives for the lobby to exert effort since that effort will not change the level of tariff protection that it will receive.
-To be clear: $\tau_{S,e}^R$ does not depend on lobbying effort, so no effort is exerted.
-
-Then when setting the trade agreement tariffs to maximize its political welfare, the government can predict that $e$ will be zero.
-If it's the case that $\ga(0) = 1$, the trade agreement will specify free trade.
-
-\begin{proposition}
-  When political pressure is entirely endogenous, under the equilibrium trade agreement with strong bindings the lobbies exert no effort and are afforded no protection if governments have no inherent bias toward protection.
-	\label{res:strong}
-\end{proposition}
-
-This very strong form of commitment for governments removes the lobby's commitment to pay for the protection it desires.
-Unless one assumes that the government is inherently biased toward the import-competing industry---that is, would provide protection in the absence of any political pressure being exerted---this case results in an outcome that is equivalent to the outcome under a social welfare maximizing government.
-This contrasts starkly with the prediction when political pressure is assumed to be completely exogenously determined.
-
 \subsubsection{Weak Bindings}

### 2016-01-01 draft.tex
remove sec:weak; name perfect external enforcement section sec:perfect and change reference to sec:weak

@@
 \subsection{Perfect External Enforcement}
+\label{sec:perfect}
 I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement.
@@
 
-\vskip.2in
-\subsubsection{Weak Bindings}
-\label{sec:weak}
 Recall that a weak binding, call it $\tau_W^R$, permits the government to apply any tariff less than or equal to $\tau_W^R$.
@@
   \end{multline}
-where $\de$ is the discount factor assumed common to the government and lobby, $\tau^R_{W,e}$ is the same tariff cap from Section~\ref{sec:weak}, and $\tau^B$ is the ``break'' tariff.
+where $\de$ is the discount factor assumed common to the government and lobby, $\tau^R_{W,e}$ is the same tariff cap from Section~\ref{sec:perfect}, and $\tau^B$ is the ``break'' tariff.
 Any $\tau > \tau^R_{W,e}$ is sufficient to breach the trade agreement and trigger Nash reversion; $\tau^B(\ga(e^B))$ is defined as the tariff the government would choose as unilaterally optimal if faced with $e^B$, where $e^B$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.

### 2016-01-01 draft.tex
revise sec:perfect to take out comparison to strong bindings

@@
 I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement.
-Although it is widely agreed that this is unrealistic, this is an important baseline case, as there is no incentive compatibility problem because the contractually-allowed tariff may not vary with $\ga$.
+Although it is widely agreed that this is unrealistic, this is an important baseline case in which there is no incentive compatibility problem because the contractually-allowed tariff may not vary with $\ga$.
 Section~\ref{sec:self} will relax this assumption and require that trade agreements be self-enforcing.
@@
 
-Recall that a weak binding, call it $\tau_W^R$, permits the government to apply any tariff less than or equal to $\tau_W^R$.
-Weak bindings are often referred to as \textit{tariff caps}.
-\Textcite{bs2005} show several interesting results regarding weak bindings when there is perfect external enforcement and political pressure is in the form of exogenous shocks.
-
-To be optimal, the weak binding must be set at the level that is efficient for the realization of political pressure that is expected conditional on tariff setting being constrained by the cap.
-Their Proposition 1 follows:
-
-\begin{proposition}[Bagwell and Stagier 2005]
-Governments prefer negotiating commitments that take the form of weak bindings over strong bindings.
-When governments negotiate commitments that take the form of weak bindings, they choose to adopt tariff bindings that (a) are higher than those they would choose if their negotiated commitments instead took the form of strong bindings and (b) imply that governments with low realizations of political pressure will set their applied tariffs strictly below the bound level.
-\end{proposition}
-
-I now turn to the model under examination here with endogenous political effort.\footnote{As above, in this section I consider endogenous political pressure alone in place of exogenously-determined $\ga$.} As in the analysis of strong bindings, in place of expectations over a range of probabilistically-determined levels of political pressure, we can backward induct to determine the lobby's effort decision and the government's optimal choice of weak binding in anticipation of the lobby's behavior.
+In the model under examination here with endogenous political effort,\footnote{As above, in this section I consider endogenous political pressure alone in place of exogenously-determined $\ga$.} in place of taking expectations over a range of probabilistically-determined levels of exogenous political pressure, we can backward induct to determine the lobby's effort decision and the government's optimal choice of tariff caps in anticipation of the lobby's behavior.
 
@@
 I assume for simplicity of exposition that the government does not set the tariff binding $\tau^R_{W,e}$ higher than that which would result from the lobby's optimal effort $e^L$, as setting a higher tariff cap would not change lobbying behavior or outcomes in any way.
-This is in contrast to the case of exogenous political pressure, where low realizations of $\ga$ lead to applied tariff levels below the cap in an economically meaning sense.
+This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.
+In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.
 
-The optimal home tariff $\tau^R_{W,e}$ under the trade agreement is the one that maximizes joint government welfare.
+To understand how the optimal home tariff $\tau^R_{W,e}$ under the trade agreement is set, notice that it must maximize joint government welfare.
 As argued in Section~\ref{sec:nasheff}, joint government welfare (with variables that do not affect the $X$ sector suppressed) is
@@
 \begin{proposition}
-    (a) When political pressure is entirely endogenous, governments prefer negotiating commitments that take the form of weak bindings over strong bindings.
-When governments negotiate commitments that take the form of weak bindings, they choose to adopt tariff bindings that are higher than those they would choose if their negotiated commitments instead took the form of strong bindings.
-(b) Governments will not set applied tariffs strictly below the bound level, but they may use the weak tariff binding to restrain and/or encourage endogenous political pressure.
+    When political pressure is entirely endogenous and governments negotiate commitments that take the form of weak bindings, they will not set applied tariffs strictly below the bound level.
+Governments may use the weak tariff binding to restrain and/or encourage endogenous political pressure.
 		\label{res:weak}
@@
 
-Note that part (a) of Proposition~\ref{res:weak} matches Proposition 1 of \Textcite{bs2005}.
-The similarities are cosmetic only; the result holds in the case of endogenous $\ga(e)$ because of the extreme result for strong bindings that no lobbying effort is exerted and the trade agreement is at free trade.
-This can be replicated with a tariff ceiling of zero.
-Part (b) points out that the governments can use a tariff cap to manipulate lobbying incentives so that political effort is neither too low or too high.
+Thus we see that the governments can use a tariff cap to manipulate lobbying incentives so that political effort is neither too low or too high.
 

### 2016-01-02 draft.tex
fix reference to my repeated game paper (sec:self)

@@
 We know that the former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here.
-Following \Textcite{bs2005} and most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} and \Textcite{buzard2013a} for alternatives that take seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.}
+Following \Textcite{bs2005} and most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternatives that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.}
 	%see KRW, Buzard 2014 for alternatives

### 2016-01-02 draft.tex
remove references to strong bindings from into paragraphs to sec:rigid and sec:objfcn

@@
 In this section and the next, I demonstrate the implications of taking account of endogenous political pressure on some central design features of trade agreements.
-I do this by first summarizing the essential results from \Textcite{bs2005} for the design of trade agreements when political considerations are exogenous and then compare them to the case under consideration here of endogenous political activity aimed at gaining protectionist tariffs.
+I begin by developing some basic results about the design of trade agreements with tariff caps when political pressure is endogenous.
+I then compare these results to those from from \Textcite{bs2005} for the design of trade agreements when political considerations are exogenous.
 
@@
 The previous sections analyze in general the effects of various forms of trade agreements when political pressure is endogenous.
-I now use the specific functional forms for the economy given in Section~\ref{sec:stage} and compare government objective functions to understand the impact of strong bindings, weak bindings and escape clauses on government welfare.
+I now use the specific functional forms for the economy given in Section~\ref{sec:stage} and compare government objective functions to understand the impact of trade agreements on government welfare.
 

### 2016-01-02 draft.tex
remove reference to strong bindings from conclusion

@@
 I have shown that accounting for endogenous lobbying has important implications for the design of trade agreements.
-Tariff caps may be used alternatively to reduce lobbying activity, or to incentivize lobbies to remain active after a trade agreement is in place, whereas a strong binding would remove lobbies' incentives to provide political support ex-post.
+For instance, tariff caps may be used alternatively to reduce lobbying activity, or to incentivize lobbies to remain active after a trade agreement is in place.
 

### 2016-01-02 intro.tex
make one expositional improvement in intro
remove ex ante section (and proofs); just reference in intro remains
revise intro for ex-ante and strong binding excisions

@@
 In order to examine it and others, I build a model that closely follows \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights.
-I find that the presence of endogenous political pressure changes the optimal choice of both strong and weak tariff bindings (exact tariff commitments and tariff caps, respectively) in quite significantly ways compared with political pressure that is assumed to be exogenous.
+I find that the presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significantly ways compared with political pressure that is assumed to be exogenous.
 
 As in \Textcite{mrc2007}, I show that one of the uses of tariff caps is to incentivize the lobby to engage in the political process after the trade agreement is in place.
-Once the trade agreement is in place, the lobby loses all incentive to exert effort under a strong binding and so a government who prefers some level of ex-post political engagement from the lobby will not use a strong binding.
-This result extends \Textcite{mrc2007} to the standard case, examined here, where specific factors are completely immobile.
-
 This, perhaps, provides an explanation for the ubiquity of tariff caps.
-Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.
+Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.\footnote{Both points extend results from \Textcite{mrc2007} to the case of perfectly immobile specific factors considered in this paper.
+For this, it must be shown that the results hold when lobbying is possible ex ante---that is, when special interests are able to influence the formation of the trade agreement.
+The results with ex-ante lobbying are available from the author upon request.}
 
 In order to see this, I compare welfare under the standard \Textcite{baldwin}-style government objective function and a similar, fully-weighted version of the government's objective function.
-At issue is the fact that the standard version is everywhere increasing in political pressure.
+At issue is the fact that the standard government objective function is everywhere increasing in political pressure.
 In an environment with endogenous political pressure, this leads inexorably to the conclusion that governments never want to discourage special interest groups from exerting pressure.
@@
 This cannot microfound a flexible model as in \Textcite{longvousden} in which weights depend on lobbying activity or in which weights are the result of shocks.}
-
-In a more general setting, I show that, depending on the government's objectives, in some cases extra political pressure would reduce the government's political welfare \textit{and} trade agreements can help them alleviate this pressure.
-I demonstrate that this is true whether or not lobbying is possible ex ante---that is, whether or not special interests are able to influence the formation of the trade agreement.
-Even in the face of ex-ante lobbying, a trade agreement that takes the form of tariff caps works to restrict lobbying in many cases, but also to keep lobbyists ``in the game.'' This broadens the result from \Textcite{mrc2007}, in which the political commitment role of trade agreements is present only when capital has some degree of mobility; here it holds with specific factors that are perfectly immobile.
-I present an extensive-form game that helps us to see that trade agreements serve this role even when ex-ante lobbying is possible.
-Although ex-ante lobbying removes some degree of autonomy from the government, it also screens out groups for whom the extra stage of lobbying becomes unprofitable.
 
@@
 The political welfare function of the government is examined in depth in Section~\ref{sec:objfcn} before returning to examine the escape clause in an environment with both endogenous and exogenous political pressure in Section~\ref{sec:escape2}.
-Section~\ref{sec:exante} details the extension to ex-ante lobbying.
 Section~\ref{sec:concl} concludes.

### 2016-03-30 to-do.tex
add major categories to to-do

@@
 
-December 30, 2015
+March 30, 2016
+\begin{enumerate}
+	\item Motivation
 		\begin{itemize}
@@
 				\end{itemize}
-	\item From Ben
+		\end{itemize}
+	\item More on escape clause / WEC
 		\begin{itemize}
@@
 				\end{itemize}
-		\end{itemize}
-\end{itemize}
-
-\vskip.5in
-Possible puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round
-\begin{itemize}
+			\item Possible puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round
 			\item Read Chad's ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), at \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
@@
 			\end{itemize}
+	\item One good idea / do too much
+	\item BS2005 vs. MRC2007
+	\item Structured differently from most papers
+	\item Costly state verification
+	\item Define $\ga$ before using it
+	\item Take out repeats, especially of references to BS2005, rely to heavily on BS2005
+	\item Scrub unnecessary references to weak bindings (since strong are now mostly gone)
+\end{enumerate}
 

### 2016-03-30 to-do.tex
organize comments from feedback.tex into to-do.tex

@@
 		\begin{itemize}
-			\item From Chad: it's missing intersection with reality
+			\item What it's missing: intersection with reality
 				\begin{itemize}
+					\item Anecdotes
+					\item How policy is used
+						\begin{itemize}
+							\item I thought that escape clause implications \textit{were} this
+						\end{itemize}
+					\item Something like how GH predicts export taxes but it doesn't work that way in the real world
 					\item Well articulated question / paradox
 				\end{itemize}
+			\item Look at ``Theory of Managed Trade'' model motivation (BS)
+			\item Intro doesn't reflect body of paper
+			\item What kinds of empirical questions might we ask with this model?
+			\item What would be different because we use this model?
+			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper.
+However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government.
+Is this aspect crucial for the point you want to make?
+If yes, it should be integrate with the previous object from the beginning.
+If not, it should be postponed a bit.
+			\item I also find the introduction too long.
+You could/should split the literature review that begins on page 4 to a separate section.
+And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it.
+Again, your main point seems to get lost into the discussion instead of standing out.
+			\item In a sort of conclusion, I think that you need to think a bit more about the objective(s) of the paper and how best to achieve them. And I hope that some of the above is useful!
 		\end{itemize}
@@
 		\begin{itemize}
-			\item It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.
+			\item I started to get really excited about the paper around page 19, especially with Result 3.
+The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.
+In fact I was itching to get to that part right from when you discussed it in the introduction.
+I also found the results in the static framework reasonably predictable.
+So I'm wondering whether you could dispense with most of the static analysis and get into the repeated game more quickly.
+				\begin{itemize}
+					\item First, the paper makes the extremely important point above almost in passing.
+I would be interested to see a more detailed examination of how this works.
+It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.
 I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
+					\item Then the paper begins to address the equally important question of what can be done about this problem, but again stops short of providing a really compelling answer.
+I think it would be worth exploring this in detail.
+The paper does get to the heart of the issue, which is that `s' can be observed by the DSB but e cannot.
+But the paper then moves on to ex-ante lobbying before getting to the bottom of how to design a dispute settlement system that resolves the conflict created by `e vs s'.
+				\end{itemize}
 			\item `Withdrawal of equivalent concessions' (WEC).
@@
 					\item But WEC might eliminate the incentive to respond to lobbying pressure $e$ because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
+						\begin{itemize}
+							\item \textbf{Think about how WEC may mitigate LOBBY'S incentives through reaction of gov't}
+							\item Ben doesn't have any need for escape, so have to shoehorn WEC into this framework
+						\end{itemize}
 					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated: one that applies if the trade partner keeps to the agreement and one that applies if they deviate.
@@
 			\item There is this drop off in usage of escape clause (EC) after Uruguay Round (UR) reforms, even though retaliation prohibited for first three years. Why, when retaliation goes \textit{away}?
+				\begin{itemize}
+					\item \textbf{perhaps better abel to respond to $\ga(e)$ before dispute settlement?}
+				\end{itemize}
 			\item I'm not sure there's a way to make sense of Ben's point: in my set-up, government doesn't ``get'' something different from lobby than it does directly from the shock. Maybe it should.
@@
 	\item One good idea / do too much
+		\begin{itemize}
+			\item You can have one good idea in a paper, the rest get lost
+			\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
+			\item If above I suggest to add a section, I think that the paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
+The analysis is also split between strict and weak commitments.
+Which scenario clearly develops the point/contribution you want to make?
+Shouldn't the other one occupy a smaller place (i.e.
+space) of the paper?)
+		\end{itemize}
 	\item BS2005 vs. MRC2007
+		\begin{itemize}
+			\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
+				\begin{itemize}
+					\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
+				\end{itemize}
+			\item I like the way you use Bagwell and Staiger (2005) as your benchmark.
+But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.
+I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.
+To keep things manageable, I wonder whether you could restrict attention to the case of ex post lobbying.
+This is where all the action is, right?
+In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.
+But your mechanisms seem quite different.
+They have capital mobility whereas your factors are specific to each sector.
+If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.
+What drives this feature in your model?
+Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.
+What is the relationship between lobby effort level and tariff policy?
+Is effort fixed first, after which the government sets its tariff?
+If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.
+It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.
+Having done that, you could bring in ex-ante lobbying and treat it the same way.
+			\item In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
+Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point (and in general it seems to me that you seem to "dependent" on Bagwell and Staiger...
+(e.g.
+beginning of Sections 4 and 6).
+		\end{itemize}
 	\item Structured differently from most papers
+		\begin{itemize}
+			\item How did Devashish structure his AER paper: the argument, the model?
+				\begin{itemize}
+					\item This paper is structured differently than the norm---is there a good reason to do it?
+					\item Don't want to challenge what people are used to unless I really need to
+					\item In general, how do AER papers \textit{look}?
+				\end{itemize}
+			\item If above I suggest to add a section, I think that the paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
+The analysis is also split between strict and weak commitments.
+Which scenario clearly develops the point/contribution you want to make?
+Shouldn't the other one occupy a smaller place (i.e.
+space) of the paper?)
+		\end{itemize}
 	\item Costly state verification
+		\begin{itemize}
+				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
+I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  Their idea is that if you want to use the escape clause then this comes at the cost of providing information to the GATT panel to persuade them that you really have been hit by a shock.
+This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
+How does your approach relate to costly state verification?
+		\end{itemize}
 	\item Define $\ga$ before using it
-	\item Take out repeats, especially of references to BS2005, rely to heavily on BS2005
+	\item Take out repeats, especially of references to BS2005, rely to heavily on BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
+Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point (and in general it seems to me that you seem to "dependent" on Bagwell and Staiger...
+(e.g.
+beginning of Sections 4 and 6).
 	\item Scrub unnecessary references to weak bindings (since strong are now mostly gone)
+	\item Perhaps make a bigger deal of the idea of $\ga$ decreasing function of $\tau$---move out of footnote
 \end{enumerate}

### 2016-03-30 to-do.tex
tighten up to-do to make decisions about revision

@@
 Again, your main point seems to get lost into the discussion instead of standing out.
-			\item In a sort of conclusion, I think that you need to think a bit more about the objective(s) of the paper and how best to achieve them. And I hope that some of the above is useful!
+			\item You need to think a bit more about the objective(s) of the paper and how best to achieve them.
 		\end{itemize}
+
+\newpage
 	\item More on escape clause / WEC
@@
 The paper does get to the heart of the issue, which is that `s' can be observed by the DSB but e cannot.
-But the paper then moves on to ex-ante lobbying before getting to the bottom of how to design a dispute settlement system that resolves the conflict created by `e vs s'.
+Get to the bottom of how to design a dispute settlement system that resolves the conflict created by `e vs s'.
 				\end{itemize}
@@
 					\item Zissimos (2007): a government `chooses the severity of its own punishment' by the extent of its initial deviation.
-In that paper, I took WEC for granted and then played out trade liberalization according to its rules.
-But this left open the question of why WEC made sense as an approach to punishment.
-Now I'm wondering whether your framework offers an answer.
+I'm wondering whether your framework offers an answer for why WEC made sense as an approach to punishment.
 					\item Under WEC, it is worth deviating from the agreement in proportion to $s$ because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
@@
 						\end{itemize}
-					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated: one that applies if the trade partner keeps to the agreement and one that applies if they deviate.
-As far as I know this latter approach does not mirror at all the institutional features actually set out by the GATT/WTO.
+					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated.
 				\end{itemize}
 			\item Possible puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round
+				\begin{itemize}
 					\item Read Chad's ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), at \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
-				\begin{itemize}
 					\item Need to understand better what changed, both with safeguard and with other policy instruments
-				\end{itemize}
-			\item There is this drop off in usage of escape clause (EC) after Uruguay Round (UR) reforms, even though retaliation prohibited for first three years. Why, when retaliation goes \textit{away}?
-				\begin{itemize}
-					\item \textbf{perhaps better abel to respond to $\ga(e)$ before dispute settlement?}
+					\item There is this drop off in usage of EC even though retaliation prohibited for first three years. Why?
+					\item \textbf{perhaps better able to respond to $\ga(e)$ before dispute settlement?}
 				\end{itemize}
@@
 			\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
-			\item If above I suggest to add a section, I think that the paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
-The analysis is also split between strict and weak commitments.
-Which scenario clearly develops the point/contribution you want to make?
-Shouldn't the other one occupy a smaller place (i.e.
-space) of the paper?)
+			\item The paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
 		\end{itemize}
+
+\newpage
 	\item BS2005 vs. MRC2007
@@
 I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.
-To keep things manageable, I wonder whether you could restrict attention to the case of ex post lobbying.
-This is where all the action is, right?
 In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.
@@
 Having done that, you could bring in ex-ante lobbying and treat it the same way.
-			\item In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
-Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point (and in general it seems to me that you seem to "dependent" on Bagwell and Staiger...
-(e.g.
-beginning of Sections 4 and 6).
+			\item In general it seems to me that you seem to "dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
 		\end{itemize}
@@
 				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
-I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  Their idea is that if you want to use the escape clause then this comes at the cost of providing information to the GATT panel to persuade them that you really have been hit by a shock.
-This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
+I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
 How does your approach relate to costly state verification?
@@
 	\item Define $\ga$ before using it
-	\item Take out repeats, especially of references to BS2005, rely to heavily on BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
-Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point (and in general it seems to me that you seem to "dependent" on Bagwell and Staiger...
-(e.g.
-beginning of Sections 4 and 6).
+	\item Take out repeats, especially of references to BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
+Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point.
 	\item Scrub unnecessary references to weak bindings (since strong are now mostly gone)

### 2016-03-31 to-do.tex
revise to-do outline regarding EC/WEC

@@
 		\begin{itemize}
-			\item I started to get really excited about the paper around page 19, especially with Result 3.
-The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.
+			\item The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.
 In fact I was itching to get to that part right from when you discussed it in the introduction.
-I also found the results in the static framework reasonably predictable.
-So I'm wondering whether you could dispense with most of the static analysis and get into the repeated game more quickly.
 				\begin{itemize}
-					\item First, the paper makes the extremely important point above almost in passing.
+					\item The paper makes the extremely important point almost in passing.
 I would be interested to see a more detailed examination of how this works.
@@
 I think it would be worth exploring this in detail.
-The paper does get to the heart of the issue, which is that `s' can be observed by the DSB but e cannot.
+The paper does get to the \textbf{heart of the issue, which is that `s' can be observed by the DSB but e cannot}.
 Get to the bottom of how to design a dispute settlement system that resolves the conflict created by `e vs s'.
@@
 				\begin{itemize}
-					\item Zissimos (2007): a government `chooses the severity of its own punishment' by the extent of its initial deviation.
+					\item Zissimos (2007): a government \textbf{`chooses the severity of its own punishment'} by the extent of its initial deviation.
 I'm wondering whether your framework offers an answer for why WEC made sense as an approach to punishment.
@@
 							\item \textbf{Think about how WEC may mitigate LOBBY'S incentives through reaction of gov't}
-							\item Ben doesn't have any need for escape, so have to shoehorn WEC into this framework
+							\item Ben doesn't have any need for escape; have to put WEC into my framework
+							\item Government feels $\ga$ the same whether it's elevated because of $s$ or $e$
+							\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
+							\item WEC should work against TOT, not PE shock?
 						\end{itemize}
@@
 				\end{itemize}
-			\item Possible puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round
+			\item Puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round, even though no retaliation for 3 yrs
 				\begin{itemize}
-					\item Read Chad's ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), at \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
+					\item ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
 					\item Need to understand better what changed, both with safeguard and with other policy instruments
-					\item There is this drop off in usage of EC even though retaliation prohibited for first three years. Why?
+					\item Note that safeguards agreement states that safeguard can only be at level necessary to remedy injury, and must liberalize as possible
 					\item \textbf{perhaps better able to respond to $\ga(e)$ before dispute settlement?}
+						\begin{itemize}
+							\item Lack of effective enforcement before WTO?
 						\end{itemize}
-			\item I'm not sure there's a way to make sense of Ben's point: in my set-up, government doesn't ``get'' something different from lobby than it does directly from the shock. Maybe it should.
+				\end{itemize}
+			\item What does WTO \textit{really} want? To discourage rent-seeking lobbying but allow governments to escape when there's a real shock.
 				\begin{itemize}
-					\item Government feels $\ga$ the same whether it's elevated because of $s$ or because of $e$
-					\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
+					\item There may be legitimate lobbying to communicate about the shock, so can't look at the presence of lobbying as a sufficient statistic
 				\end{itemize}
@@
 				\end{itemize}
+			\item After $s$ realized, $e$ choosen by lobby. Then government can choose $\tau$
+				\begin{enumerate}
+					\item consistent with $s$, no retaliation
+					\item consistent with $s$ and $e$ and be retaliated against according to WEC (need to work WEC into model--repeated game incentives necessary)
+						\begin{itemize}
+							\item Is responding to $e$ in some cases worth it?
+							\item Does WEC help or hurt vs. grim trigger, $T$ period Nash reversion?
 						\end{itemize}
-	\item One good idea / do too much
+				\end{enumerate}
+		\end{itemize}
+	\item Costly state verification
 		\begin{itemize}
-			\item You can have one good idea in a paper, the rest get lost
-			\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
-			\item The paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
+				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
+I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
+How does your approach relate to costly state verification?
 		\end{itemize}
@@
 		\end{itemize}
-	\item Costly state verification
+		\item One good idea / do too much
 		\begin{itemize}
-				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
-I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
-How does your approach relate to costly state verification?
+			\item You can have one good idea in a paper, the rest get lost
+			\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
+			\item The paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
+			\item So I'm wondering whether you could dispense with most of the static analysis and get into the repeated game more quickly.
 		\end{itemize}

### 2016-03-31 draft.tex
remove ref to weak bindings, change to biber and add path to bib file on wallis computer
move obj fcn section and shorten

@@
 
-%dell laptop
 \usepackage[backend=biber,block=space,style=authoryear-comp]{biblatex}
@@
 %dell laptop
-\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
+%\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
 %\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
+
+%wallis computer
+\addbibresource{C:/Users/wallis/Dropbox/Research/xBibs/tradeagreements.bib}
 
@@
 
+\subsection{Political Objective Functions and the Role of Trade Agreements}
+\label{sec:objfcn}
+When political pressure is determined endogenously, a government can predict the political pressure it will face and in some cases use a trade agreement to alter lobbying incentives.
+Since the government has some control through the trade agreement over the level of $\ga$ it will face, it becomes important to understand how the government's welfare varies in $\ga$.
+The perceived role that trade agreements can play is therefore influenced by the choice of government objective function.
+
+Let's begin with the \Textcite{baldwin}-style government objective function that is most common in the recent literature on trade agreements, specialized to the case of endogenous pressure only.
+\begin{equation}
+  W(\ga(e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
+  \tag{\ref{eq:wel}}
+\end{equation}
+We need to know at what $\ga(e)$ the government's welfare is maximized, given that it will choose $\tau$ according to either Equation~\ref{eq:nash} when setting tariffs unilaterally or Equation~\ref{eq:eff} when setting tariffs in the context of a trade agreement.
+
+Envelope-theorem style results imply that the first order condition in the same in the unilateral and joint problems: $\displaystyle \frac{\partial W_x}{\partial \ga} = 0$.
+See the \hyperlink{envelope}{Appendix} for details.
+
+An interior solution does not exist to this problem.
+By inspection, the partial derivative of government welfare with respect to $\ga$ is simply producer surplus in the import sector.
+This is always positive.
+Thus, under the government welfare function in Expression~\ref{eq:wel}, government welfare is maximized when $\ga$ takes on the highest value achievable.
+
+This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the partially weighted government welfare function from Expression~\ref{eq:wel} under the parameterization of \Textcite{bs2005} used throughout this paper.
+Graphed in terms of tariffs, one can see that from $\tau=0.05$---the tariff the home government optimally sets when making unilateral policy with no political pressure---government welfare steadily increases as the trade agreement tariff increases.
+This is because the governments can expect that ex-post political pressure will increase in the trade agreement tariff, and government welfare increases everywhere in political pressure.
+
+\begin{figure}
+\begin{center}
+\includegraphics[height=3in, width=4.25in]{unweight.jpg}
+\end{center}
+\caption{Government Welfare with Partially Weighted Objective Function \label{fig:unweight}}
+\end{figure}
+
+This result that government welfare is strictly increasing in the political pressure parameter would seem to be related to the fact that, although $\ga$ is often referred to as a ``political economy weight,'' Expression~\ref{eq:wel} is only partially weighted.
+That the extra importance given to profits of importers when $\ga$ increases is pure additional welfare with no concomitant reduction elsewhere may be an assumption worth weakening, particularly in the context of endogenous political activity.
+
+One alternative formulation for the government objective function is one in which $\ga$ is a true weight in the following straightforward sense:
+\begin{equation}
+  \frac{1}{4+\ga(e)}CS_X(\tau) + \frac{\ga(e)}{4+\ga(e)}PS_X(\tau) + \frac{1}{4+\ga(e)}CS_Y(\tau^*) + \frac{1}{4+\ga(e)}PS_Y(\tau^*) + \frac{1}{4+\ga(e)}TR(\tau)
+  \label{eq:weighted}
+\end{equation}
+
+See the \hyperlink{sol_weighted}{Appendix} for the mathematical derivations corresponding to Figure~\ref{fig:weight}.
+They show that, particularly for the trade agreement tariffs, government welfare first declines and then increases, with the minimum value occurring around $\tau = 0.057$ and $\ga = 1.05$.
+The maximum still unquestionably occurs at the maximum value of $\ga$, which leads to the prohibitive tariff of $\frac{1}{6}$.
+But if a government's objective function is closer to Expression~\ref{eq:weighted} than Expression~\ref{eq:wel}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
+
+\begin{figure}
+\begin{center}
+\includegraphics[height=3in, width=4.25in]{weight.jpg}
+\end{center}
+\caption{Government Welfare with Weighted Objective Function \label{fig:weight}}
+\end{figure}
+
+Whether or not the optimal trade agreement would set a tariff cap aimed at reducing lobbying depends on the lobby's incentives, which depends in turn on the shape of $\ga(e)$.
+Let's look at two parameterizations in a simple family of functions for $\ga(e)$ that demonstrate drastically different outcomes.
+
+I begin with $\ga(e) = 1 + e^{.1}$.
+With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0016$, which results in $\ga=1.525$ and a Nash tariff of $0.129$.
+Import-competing producer (net) profits are $0.1025$, while government welfare is $0.0862$.
+Because joint government welfare is convex in the tariff, the solution will be at a corner: either at $\tau^R_{W,e} =0$ or at the highest level for which the lobby will exert effort, i.e.
+the Nash tariff.
+Since joint welfare at the zero tariff is $0.0857$, the tariff cap is set at the trivial level of the non-cooperative Nash tariff.
+
+Here, given the way in which lobbying effort is translated into political support as represented by $\ga$, the government cannot improve its welfare by manipulating lobbying incentives.
+Discouraging lobbying would reduce welfare; in this parameter range, the government's interest is solely in encouraging lobbying effort, and there is nothing it can do in this regard.
+
+%This is does through the introduction of the trade agreement: by making protection more expensive, on the margin the lobby pays more in order to get even less protection than it did in the absence of the agreement.
+But there is nothing further the government can do to increase $\ga$.
+
+This example stands in contrast to a less concave (in the sense of Arrow Pratt) $\ga$.
+Take $\ga(e) = 1 + e^{.3}$.
+With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0010$, which results in $\ga=1.126$ and a unilateral tariff of $0.068$.
+Import-competing producer (net) profits are $0.0921$, while government welfare is $0.0855$.
+Since joint government welfare is higher at free trade, the trade agreement will set $\tau^R_{W,e}=0$, profits decrease to $0.0816$, and government welfare increases to $0.0857$.
+
+In this case, again the government does not have means to encourage any additional lobbying, but it can improve its situation by discouraging lobbying.
+By setting a tariff cap of zero---assuming the conditions for self enforcement outlined in Section~\ref{sec:selfweak} are met---the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.
+The stark prediction that the best the government can do is by setting the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has an interior global maximum in $\ga$.
+
+With this weighted welfare function, we see that the trade agreement is employed to different ends depending on $\ga(e)$.
+If the lobby's ability to make an impact on the political process is strong enough, the government will be chiefly concerned with the profits of the import-competing industry and will use the trade agreement solely to internalize the terms-of-trade externality.
+However, if the lobby's impact is below a given threshold, the distortions that protection create are weighed more heavily and the government will use the trade agreement to rein in lobbying activity.
+
+%In either case, we see that the government is able to provide reduced protection while extracting more effort from the lobby under a trade agreement: internalizing the terms of trade externality increases the marginal cost of protection.
+
+It is not claimed that the government welfare function in Expression~\ref{eq:weighted} perfectly represents the preferences of those who negotiate trade agreements.
+However, examining this alternative in combination with endogenous lobbying points out that the idea that trade agreements can function as domestic political commitment devices may be more general than previously demonstrated.
+It can also provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
+
 \subsection{Exogenous Political Pressure and Incentive Compatibility}
@@
 
-\subsubsection{Self Enforcement with Weak Bindings}
+\subsubsection{Self Enforcement}
 \label{sec:selfweak}
@@
 
-In the next section, I ask the question of whether this type of outcome is likely to be consistent with the ex-ante best interests of the governments themselves as a prelude to exploring the likely intentions behind the design of such escape clauses.
-
-\section{Political Objective Functions and the Role of Trade Agreements}
-\label{sec:objfcn}
-The previous sections analyze in general the effects of various forms of trade agreements when political pressure is endogenous.
-I now use the specific functional forms for the economy given in Section~\ref{sec:stage} and compare government objective functions to understand the impact of trade agreements on government welfare.
-
-To reiterate, in line with the dominant strain of the literature, I have assumed that the governments' choice of trade agreement tariff is optimized with respect to the political pressure it expects to experience after the trade agreement is signed.
-In the case of exogenous political pressure, the best one can do is take an expectation given the distribution of $\ga$, or possibly build some kind of flexibility into the agreement through tariff caps or an incentive compatible escape clause.
-
-With endogenously-determined political pressure with no shocks---the simple case examined so far here---each government can predict the political pressure it will face and in some cases use the trade agreement to alter the incentives of the lobby.
-Since the government has some control through the trade agreement over the level of $\ga$ it will face, it becomes important to understand how the government's welfare varies in $\ga$.
-
-Let's begin with the \Textcite{baldwin}-style government objective function that is most common in the recent literature on trade agreements and that has been in use throughout this paper so far, specialized to the case of endogenous pressure only.
-\begin{equation}
-  W(\ga(e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
-  \tag{\ref{eq:wel}}
-\end{equation}
-What we would like to know is at what $\ga(e)$ the government's welfare is maximized, given that it will choose $\tau$ according to either Equation~\ref{eq:nash} when setting tariffs unilaterally or Equation~\ref{eq:eff} when setting tariffs in the context of a trade agreement.
-
-The answer turns out to be the same for the unilateral and joint problems.
-Starting with the non-cooperative maximization problem with respect to $\ga$ (there is no loss in suppressing the dependence of $\ga$ on $e$ for the time being):
-\[
- \max_{\ga} W_x(\ga,\tau) + W_y(\tau^*)
-\]
-The first order condition is
-\[
-	\frac{\partial W_x}{\partial \ga} + \frac{\partial W_x}{\partial \tau}\frac{\partial \tau}{\partial \ga} = 0
-\]
-In an envelope-theorem style result, because the Nash tariff is the result of the government's optimization with respect to $\tau$, $\frac{\partial W_x}{\partial \tau}=0$ so that the first order condition reduces to $\frac{\partial W_x}{\partial \ga} = 0$.
-
-Turning to the maximization of joint welfare in the trade agreement, we have
-\[
-  \max_{\ga} W_x(\ga,\tau) + W_y(\tau^*) + W_x^*(\tau)+ W_y^*(\tau^*)
-\]
-The first order condition is 
-\[
-  \frac{\partial W_x}{\partial \ga} + \frac{\partial W_x}{\partial \tau}\frac{\partial \tau}{\partial \ga} + \frac{\partial W_x^*}{\partial \tau}\frac{\partial \tau}{\partial \ga} = \frac{\partial W_x}{\partial \ga} + \left[ \frac{\partial W_x}{\partial \tau} + \frac{\partial W_x^*}{\partial \tau}\right]\frac{\partial \tau}{\partial \ga} = 0
-\]
-Here, it is the efficient joint tariff that is chosen, and it is chosen precisely by setting the term in brackets equal to zero so that we have the first order condition simplifying to the same expression as in the unilateral case.
-
-An interior solution does not exist to this problem.
-By inspection, the partial derivative of government welfare with respect to $\ga$ is simply producer surplus in the import sector.
-This is always positive.
-Thus, under the government welfare function in Expression~\ref{eq:wel}, government welfare is maximized when $\ga$ takes on the highest value achievable.
-
-This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the partially weighted government welfare function from Expression~\ref{eq:wel} under the parameterization of \Textcite{bs2005} used throughout this paper.
-Graphed in terms of tariffs, one can see that from $\tau=0.05$---the tariff the home government optimally sets when making unilateral policy with no political pressure---government welfare steadily increases as the trade agreement tariff increases.
-This is because the governments can expect that ex-post political pressure will increase in the trade agreement tariff, and government welfare increases everywhere in political pressure.
-
-\begin{figure}
-\begin{center}
-\includegraphics[height=3.5in, width=5in]{unweight.jpg}
-\end{center}
-\caption{Government Welfare with Partially Weighted Objective Function \label{fig:unweight}}
-\end{figure}
-
-This result that government welfare is strictly increasing in the political pressure parameter would seem to be related to the fact that, although $\ga$ is often referred to as a ``political economy weight,'' Expression~\ref{eq:wel} is only partially weighted.
-That the extra importance given to profits of importers when $\ga$ increases is pure additional welfare with no concomitant reduction elsewhere may be an assumption worth weakening, particularly in the context of endogenous political activity.
-
-An alternative formulation for the government objective function is one in which $\ga$ is a true weight in the following straightforward sense:
-\begin{equation}
-  \frac{1}{4+\ga(e)}CS_X(\tau) + \frac{\ga(e)}{4+\ga(e)}PS_X(\tau) + \frac{1}{4+\ga(e)}CS_Y(\tau^*) + \frac{1}{4+\ga(e)}PS_Y(\tau^*) + \frac{1}{4+\ga(e)}TR(\tau)
-  \label{eq:weighted}
-\end{equation}
-
-See the \hyperlink{sol_weighted}{Appendix} for the mathematical derivations corresponding to Figure~\ref{fig:weight}.
-They show that, particularly for the trade agreement tariffs, government welfare first declines and then increases, with the minimum value occurring around $\tau = 0.057$ and $\ga = 1.05$.
-The maximum still unquestionably occurs at the maximum value of $\ga$, which leads to the prohibitive tariff of $\frac{1}{6}$.
-But if a government's objective function is closer to Expression~\ref{eq:weighted} than Expression~\ref{eq:wel}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
-
-\begin{figure}
-\begin{center}
-\includegraphics[height=3.5in, width=5in]{weight.jpg}
-\end{center}
-\caption{Government Welfare with Weighted Objective Function \label{fig:weight}}
-\end{figure}
-
-Whether or not the optimal trade agreement would set a tariff cap aimed at reducing lobbying depends on the lobby's incentives, which depends in turn on the shape of $\ga(e)$.
-Let's look at two parameterizations in a simple family of functions for $\ga(e)$ that demonstrate drastically different outcomes.
-
-I begin with $\ga(e) = 1 + e^{.1}$.
-With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0016$, which results in $\ga=1.525$ and a Nash tariff of $0.129$.
-Import-competing producer (net) profits are $0.1025$, while government welfare is $0.0862$.
-Because joint government welfare is convex in the tariff, the solution will be at a corner: either at $\tau^R_{W,e} =0$ or at the highest level for which the lobby will exert effort, i.e.
-the Nash tariff.
-Since joint welfare at the zero tariff is $0.0857$, the tariff cap is set at the trivial level of the non-cooperative Nash tariff.
-
-Here, given the way in which lobbying effort is translated into political support as represented by $\ga$, the government cannot improve its welfare by manipulating lobbying incentives.
-Discouraging lobbying would reduce welfare; in this parameter range, the government's interest is solely in encouraging lobbying effort, and there is nothing it can do in this regard.
-
-%This is does through the introduction of the trade agreement: by making protection more expensive, on the margin the lobby pays more in order to get even less protection than it did in the absence of the agreement.
-But there is nothing further the government can do to increase $\ga$.
-
-This example stands in contrast to a less concave (in the sense of Arrow Pratt) $\ga$.
-Take $\ga(e) = 1 + e^{.3}$.
-With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0010$, which results in $\ga=1.126$ and a unilateral tariff of $0.068$.
-Import-competing producer (net) profits are $0.0921$, while government welfare is $0.0855$.
-Since joint government welfare is higher at free trade, the trade agreement will set $\tau^R_{W,e}=0$, profits decrease to $0.0816$, and government welfare increases to $0.0857$.
-
-In this case, again the government does not have means to encourage any additional lobbying, but it can improve its situation by discouraging lobbying.
-By setting a tariff cap of zero---assuming the conditions for self enforcement outlined in Section~\ref{sec:selfweak} are met---the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.
-The stark prediction that the best the government can do is by setting the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has an interior global maximum in $\ga$.
-
-With this weighted welfare function, we see that the trade agreement is employed to different ends depending on $\ga(e)$.
-If the lobby's ability to make an impact on the political process is strong enough, the government will be chiefly concerned with the profits of the import-competing industry and will use the trade agreement solely to internalize the terms-of-trade externality.
-However, if the lobby's impact is below a given threshold, the distortions that protection create are weighed more heavily and the government will use the trade agreement to rein in lobbying activity.
-
-%In either case, we see that the government is able to provide reduced protection while extracting more effort from the lobby under a trade agreement: internalizing the terms of trade externality increases the marginal cost of protection.
-
-It is not claimed that the government welfare function in Expression~\ref{eq:weighted} perfectly represents the preferences of those who negotiate trade agreements.
-However, examining this alternative in combination with endogenous lobbying points out that the idea that trade agreements can function as domestic political commitment devices may be more general than previously demonstrated.
-It can also provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
-
 In the next section, I turn to the implications for the structure of the WTO procedures for administered protection, and the escape clause in particular, in a more realistic setting where both exogenous political shocks and endogenous lobbying are present.
@@
 \section{Appendix}
+\noindent \textbf{\hypertarget{envelope}{Unilateral and Joint Welfare Maximized at same $\ga$}}: \\
+Starting with the non-cooperative maximization problem with respect to $\ga$:
+\[
+ \max_{\ga} W_x(\ga,\tau) + W_y(\tau^*)
+\]
+The first order condition is
+\[
+	\frac{\partial W_x}{\partial \ga} + \frac{\partial W_x}{\partial \tau}\frac{\partial \tau}{\partial \ga} = 0
+\]
+In an envelope-theorem style result, because the Nash tariff is the result of the government's optimization with respect to $\tau$, $\frac{\partial W_x}{\partial \tau}=0$ so that the first order condition reduces to $\frac{\partial W_x}{\partial \ga} = 0$.
 
+Turning to the maximization of joint welfare in the trade agreement, we have
+\[
+  \max_{\ga} W_x(\ga,\tau) + W_y(\tau^*) + W_x^*(\tau)+ W_y^*(\tau^*)
+\]
+The first order condition is 
+\[
+  \frac{\partial W_x}{\partial \ga} + \frac{\partial W_x}{\partial \tau}\frac{\partial \tau}{\partial \ga} + \frac{\partial W_x^*}{\partial \tau}\frac{\partial \tau}{\partial \ga} = \frac{\partial W_x}{\partial \ga} + \left[ \frac{\partial W_x}{\partial \tau} + \frac{\partial W_x^*}{\partial \tau}\right]\frac{\partial \tau}{\partial \ga} = 0
+\]
+Here, it is the efficient joint tariff that is chosen, and it is chosen precisely by setting the term in brackets equal to zero so that we have the first order condition simplifying to the same expression as in the unilateral case.
+$\hfill\blacksquare$
+
+\vskip.3in
 \noindent \textbf{\hypertarget{sol_weighted}{Optimal $\ga$ for Weighted Government Welfare Function}}: \\

### 2016-04-01 draft.tex
add proposition on EC; begin new section on IC EC

@@
 
-It is not claimed that the government welfare function in Expression~\ref{eq:weighted} perfectly represents the preferences of those who negotiate trade agreements.
-However, examining this alternative in combination with endogenous lobbying points out that the idea that trade agreements can function as domestic political commitment devices may be more general than previously demonstrated.
+I do not argue that the government welfare function in Expression~\ref{eq:weighted} perfectly represents the preferences of those who negotiate trade agreements.
+Another reasonable candidate for the government objective function embodies diminishing returns to lobbying as tariffs increase as in \Textcite{ethier2012}.
+In the current model, one would take $\ga$ to be a decreasing function of $\tau$.
+This implies that government welfare must eventually decrease in $\ga$, providing for concavity in $\ga$ and a strict interior maximum.
+
+Examining these alternatives in combination with endogenous lobbying points out that the idea that trade agreements can function as domestic political commitment devices may be more general than previously demonstrated.
 It can also provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
@@
 But when the shock is low, the lobby will exert effort so that the sum of $\ga(s)$ and $\ga(e)$ equals the high realization.
+
+\begin{proposition}{Ineffectiveness of Political Criterion for Escape Clause}
+    Assume $\ga(s,e)$ is additively separable in $s$ and $e$ with $s$ and $e$ independent.
+If an escape clause conditions on $\ga(s,e)$ and the lobby's preferred tariff is higher than the escape clause binding, the lower ``normal'' tariff binding will never be applied.
+\end{proposition}
+
 From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section~\ref{sec:objfcn}, the governments' ex-ante welfare is reduced by excess lobbying.
@@
 Without this information, it thus can only certify the legitimacy of a Safeguard measure; it cannot determine the correct level at which the Safeguard measure should be applied.
+
+In this world that is far from first best, I next demonstrate how a WTO-styled escape clause can allow flexibility in the face of fundamental shocks $s$ while discouraging lobbyists from using the shock as cover for rent-seeking behavior.
+The timeline is as follows:
+\begin{enumerate}
+	\item The government sets the trade agreement tariff $\tau^F_{W,e}$ cooperatively.
+I assume this is a weak binding and add the subscript $F$ to denote the environment with flexibility to deal with shocks.
+	\item The exogenous shock $s$ is realized.
+	\item The lobby chooses its effort level $e_F$.
+	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau($s$)$ and $\tau^B(\ga(s,e_F))$.
+\end{enumerate}
+
+				\begin{enumerate}
+					\item consistent with $s$, no retaliation
+					\item consistent with $s$ and $e$ and be retaliated against according to WEC (need to work WEC into model--repeated game incentives necessary)
+						\begin{itemize}
+							\item Is responding to $e$ in some cases worth it?
+							\item Does WEC help or hurt vs. grim trigger, $T$ period Nash reversion?
+						\end{itemize}
+					\end{enumerate}
 

### 2016-04-01 draft.tex
create subsection for IC EC; explain this escape set up

@@
 
-In this world that is far from first best, I next demonstrate how a WTO-styled escape clause can allow flexibility in the face of fundamental shocks $s$ while discouraging lobbyists from using the shock as cover for rent-seeking behavior.
+\subsection{An Escape Clause For Endogenous Politics}
+\label{sec:ECendog}
+In this world that is far from first best, I next demonstrate how a WTO-style escape clause can allow flexibility in the face of fundamental shocks $s$ while discouraging lobbyists from using the shock as cover for rent-seeking behavior.
 The timeline is as follows:
@@
 	\item The lobby chooses its effort level $e_F$.
-	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau($s$)$ and $\tau^B(\ga(s,e_F))$.
+	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau(s)$ and $\tau^B(\ga(s,e_F))$.
 \end{enumerate}
+Here I take the higher, `escape' tariff to be continuous instead of set at a given binding.
+If the government applies $\tau^F_{W,e}$, implementation of the agreement goes on as before.
+Consistent with the WTO Agreement on Safeguards, the government can apply the tariff that is necessary to remedy the injury caused by the shock, $\tau(s)$, with no retaliation for three periods.
+If the government applies a tariff that is higher than either $\tau^F_{W,e}$, $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
+I label this tariff $\tau^B(\ga(s,e_F))$.
+This is a violation of the trade agreement even in the presence of the safeguard and incurs retaliation.
 
-				\begin{enumerate}
-					\item consistent with $s$, no retaliation
-					\item consistent with $s$ and $e$ and be retaliated against according to WEC (need to work WEC into model--repeated game incentives necessary)
 						\begin{itemize}
@@
 						\end{itemize}
-					\end{enumerate}
 

### 2016-04-01 draft.tex
finish most of new EC section

@@
 
-\section{The Purpose and Design of the Escape Clause}
+\subsection{The Purpose and Design of the Escape Clause}
 \label{sec:escape2}
@@
 If the government applies a tariff that is higher than either $\tau^F_{W,e}$, $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
-I label this tariff $\tau^B(\ga(s,e_F))$.
+I label this tariff $\tau^B(\ga(s,e_F))$, where $e^F$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.
+
 This is a violation of the trade agreement even in the presence of the safeguard and incurs retaliation.
+It is a violation of the agreement because the applied tariff level is higher than is needed to remedy the injury caused by the shock; the Agreement on Safeguards instructs that ``A Member shall apply safeguard measures only to the extent necessary to prevent or remedy serious injury and to facilitate adjustment'' (Article 5.1).\footnote{The paragraph goes on to give even more clarification: ``If a quantitative restriction is used, such a measure shall  not reduce the quantity of imports below the level of a recent period which shall be the average of imports in the last three representative years for which statistics are available, unless clear justification is given that a different level is necessary to prevent or remedy serious injury.''}
 
-						\begin{itemize}
-							\item Is responding to $e$ in some cases worth it?
-							\item Does WEC help or hurt vs. grim trigger, $T$ period Nash reversion?
-						\end{itemize}
+To see how such a scheme works in the context of repeated-game enforcement, we look to the incentive constraints of the government and lobby.
+I follow \Textcite{buzard2013a} by using limited reversions to the stage-game Nash equilibrium outcome.
+For ease of analysis, I use a punishment length of two periods.
+This simply implies that a country would return to the agreement either after applying a valid escape clause for three years or after violating the agreement.
+
+The government's incentive constraint is given by
+  \begin{multline}
+    \left[1+ \de +\de^2 \right] \left\{W_x(\ga(s,0),\tau(s)) + W_x^*(\tau^F_{W,e}) \right\} \geq \\ W_X(\ga(s,e^F),\tau^B(\ga(s,e^F)) + W_x^*(\tau^F_{W,e}) + \left[\de +\de^2 \right] \left\{W_X(\ga(s,e^N),\tau^N) + W_x^*(\tau^N) \right\}
+		\label{exp:govincentEC}
+  \end{multline}
+Label $\ov{e}$ the level of $e^F$ that is just high enough to cause the government to throw the trading relationship into dispute.
+We use $\ov{e}$ in the lobby's constraint, in which the lobby optimally exerts no effort if the government applies the safeguard tariff $\tau(s)$:
+\begin{equation}
+  \left[1+ \de +\de^2 \right] \pi(\tau(s)) \geq \left[1+ \de +\de^2 \right] \left\{\pi(\tau(\ga(s,\ov{e})) - \ov{e} \right\}
+	\label{exp:lobbyEC}
+\end{equation}
+
+Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may chooses to set the trade agreement tariff in the presence of an escape clause $\left(\tau^F_{W,e}\right)$ so that in some cases the lobby's incentive constraint is violated.
+This would lead to the legislature's constraint being violated as well, leading to a dispute.
+
+I plan to compare the Nash-reversion punishment to punishment via 'withdrawal of equivalent concessions,' as it appears on the face of it that this environment may provide a justification its use.
+The intuition is the following: \Textcite{buzard2013a} shows that the lobbying effort required to cause the government to enter into a dispute is higher than the Nash equilibrium effort levels.
+Then the tariffs during a dispute are also higher, and punishment at this higher level would be more severe than Nash---both for the government and the lobby.
 

### 2016-04-01 draft.tex
finish EC section

@@
 I assume this is a weak binding and add the subscript $F$ to denote the environment with flexibility to deal with shocks.
-	\item The exogenous shock $s$ is realized.
+	\item The exogenous shock $s$ is realized. It's level is assumed costlessly verifiable for simplicity.
 	\item The lobby chooses its effort level $e_F$.
@@
 
-I plan to compare the Nash-reversion punishment to punishment via 'withdrawal of equivalent concessions,' as it appears on the face of it that this environment may provide a justification its use.
+One important thing \textit{does not} happen in this environment: lobbies do not use the escape clause as cover for rent seeking.
+The government may use the escape clause to provide extra protection commensurate to the shock.
+This provides a looser incentive constraint than in the absence of the escape clause and improves self-enforceability of the agreement.
+Or the lobbies may still find it in their interest to exert effort to encourage the government to provide this higher level of protection.
+This will provoke a dispute, but the protection afforded by the higher tariff but this will not be hidden behind the escape clause.
+So an escape clause for the purposes of providing breathing room from exogenous shocks \textit{is} possible.
+
+It still does not seem feasible to provide this flexibility at precisely the level that the shock impacts the government's preferences $\ga(s)$.
+What's more, this may provide an answer to why Safeguard measures have fallen out of use.
+If the real purpose is to provide an outlet for political pressure, I have demonstrated above that the current WTO-style Safeguards rules seem designed explicitly to discourage rent-seeking behavior.
+Thus if other measures such as Anti-Dumping Duties are more susceptible to manipulation, it seems natural that lobbies' efforts would migrate there.
+
+I also plan to compare the Nash-reversion punishment to punishment via 'withdrawal of equivalent concessions,' as it appears on the face of it that this environment may provide a justification its use.
 The intuition is the following: \Textcite{buzard2013a} shows that the lobbying effort required to cause the government to enter into a dispute is higher than the Nash equilibrium effort levels.

### 2016-04-02 draft.tex
change header for dell laptop; revise abstract

@@
 %dell laptop
-%\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
-%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
+\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
 
 %wallis computer
-\addbibresource{C:/Users/wallis/Dropbox/Research/xBibs/tradeagreements.bib}
+%\addbibresource{C:/Users/wallis/Dropbox/Research/xBibs/tradeagreements.bib}
 
@@
 \begin{quote}
-{\small Political pressure is undoubtedly an important influence in the setting of trade policy and the formulation of trade agreements.
-Most of the literature in this area models the political pressure that governments face as resulting from an exogenous, stochastic process.
-This paper shows that when political pressure arises endogenously, important results can be overturned and new insights into the motivation for features of the trade agreements we observe and rules of organizations such as the WTO come to light.
+{\small Political pressure is undoubtedly an important influence in the formulation of trade agreements and the institutions that govern them.
+Much of the literature that speaks to the design of trade policy institutions takes the political pressure that governments face as resulting from an exogenous, stochastic process.
+This paper shows that when political pressure arises endogenously, important results may be overturned and new insights into the motivation for features of the trade agreements we observe and rules of organizations such as the WTO come to light.
 Developing a model that integrates both exogenous and endogenous political pressure and taking a general approach to the government objective function, I show that governments may want to use tariff caps both to force special interest groups to continue lobbying after a trade agreement is signed and to reduce the magnitude of that lobbying effort.
-The presence of endogenous politics can also destroy an escape clause's ability to provide flexibility in times of large negative political shocks when lobbies use the flexibility to seek rents.
-This can explain why use of WTO Safeguards are conditioned on measurable economic indicators as well as why Safeguard levels of protection are not regulated.}
+The presence of endogenous politics can also destroy the ability of an escape clause as traditionally defined in the literature to provide flexibility in times of large negative political shocks when lobbies use the flexibility to seek rents.
+This can explain why the use of WTO Safeguards are conditioned on measurable economic indicators.}
 \end{quote}

### 2016-04-02 draft.tex
revise intro up to lit review

@@
 
-\begin{proposition}{Ineffectiveness of Political Criterion for Escape Clause}
+\begin{proposition}[Ineffectiveness of Political Criterion for Escape Clause]
     Assume $\ga(s,e)$ is additively separable in $s$ and $e$ with $s$ and $e$ independent.
@@
 These conditions seem unlikely to be met in reality and we observe, again, that the DSB makes no attempt to determine the political parameter.
-Without this information, it thus can only certify the legitimacy of a Safeguard measure; it cannot determine the correct level at which the Safeguard measure should be applied.
+Without this information, it thus can only certify the legitimacy of a Safeguard measure and perhaps the level of tariff that is necessary to counteract $s$.
+It cannot determine the correct level at which the Safeguard measure should be applied for the purposes of matching the applied tariff to the political pressure felt by the government.
 
@@
 
-I also plan to compare the Nash-reversion punishment to punishment via 'withdrawal of equivalent concessions,' as it appears on the face of it that this environment may provide a justification its use.
+As in \Textcite{zissimos}, I also plan to compare the Nash-reversion punishment to punishment via 'withdrawal of equivalent concessions,' as it appears on the face of it that this environment may provide a justification its use.
 The intuition is the following: \Textcite{buzard2013a} shows that the lobbying effort required to cause the government to enter into a dispute is higher than the Nash equilibrium effort levels.

### 2016-04-02 intro.tex
revise intro up to lit review

@@
 Much of the work on the political economy of trade agreements focuses on questions of the optimal design of trade agreements, trade agreement negotiations, and trade dispute settlement that arise in the presence of asymmetric information about shocks to an exogenous political economy parameter.
+The assumption that political economy forces are entirely driven by exogenous forces is a rather drastic simplifying assumption.
+Do the predictions of our models hold up if this simplifying assumption is lifted?
 
-One of the basic ideas that emerges from this literature is that, in the presence of asymmetric information about the strength of the ex-post political economy shocks, it is often advantageous to grant governments a period of relief from trade commitments.
+One of the basic ideas that emerges from this literature is that in the presence of asymmetric information about the strength of the ex-post political economy shocks, it is often advantageous to grant governments a period of relief from trade commitments.
 That is, one would rather allow a short period of ``escape'' from the agreement rather than have the agreement abandoned forever because domestic political opposition is temporarily too strong to be resisted.
 
-Various trade agreement and institutional design features have been proposed in the literature that can help to make an escape clause work, since it is not incentive compatible to allow governments to take advantage of an escape clause whenever they wish.
-If some cost can be associated with the use of the escape clause or a dispute settlement body (DSB) such as that at the World Trade Organization (WTO) can procure a signal about the strength of $\ga$, it is often advantageous to grant this period of relief from trade commitments if the signal is strong enough.
-
-%In this vein, \Textcite{bs2005} find that governments prefer tariff caps to strong bindings when there is no escape clause, but that impatient governments may need an escape clause in order to make cooperation sustainable.
-
-This is an intuitively appealing story, but the logic can break down in the presence of endogenous political pressure.
+This is an intuitively appealing story, but it turns out that the logic can break down in the presence of endogenous political pressure.
 An escape clause allows a government to apply a higher tariff barrier when it experiences intense political pressure.
-But if a government gets a free pass from the WTO whenever it feels sufficient political pressure from domestic interest groups, those interest groups have a strong incentive to exert the required level of pressure, eviscerating the escape clause.
+But if a government gets a free pass at the WTO whenever it feels sufficient political pressure from domestic interest groups, those interest groups have a strong incentive to exert the required level of pressure regardless of the underlying state of the world, eviscerating the escape clause.
 
 This is one example of a design question whose answers are sensitive to assumptions about the endogeneity of political pressure.
-In order to examine it and others, I build a model that closely follows \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights.
-I find that the presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significantly ways compared with political pressure that is assumed to be exogenous.
+In order to examine it and others, I build a model that is comparable to \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights in a repeated-game setting.
 
+I first establish the baseline case with tariff caps and no escape clause.
+Here, I find that the presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significant ways compared with the case where political pressure is assumed to be exogenous.
 As in \Textcite{mrc2007}, I show that one of the uses of tariff caps is to incentivize the lobby to engage in the political process after the trade agreement is in place.
@@
 Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.\footnote{Both points extend results from \Textcite{mrc2007} to the case of perfectly immobile specific factors considered in this paper.
-For this, it must be shown that the results hold when lobbying is possible ex ante---that is, when special interests are able to influence the formation of the trade agreement.
+To truly extend those results, it must be shown that the results hold when lobbying is possible ex ante---that is, when special interests are able to influence the formation of the trade agreement.
 The results with ex-ante lobbying are available from the author upon request.}
 
-In order to see this, I compare welfare under the standard \Textcite{baldwin}-style government objective function and a similar, fully-weighted version of the government's objective function.
+In order to make this point clear, I compare welfare under the standard \Textcite{baldwin}-style government objective function and a similar, fully-weighted version of the government's objective function.
 At issue is the fact that the standard government objective function is everywhere increasing in political pressure.
 In an environment with endogenous political pressure, this leads inexorably to the conclusion that governments never want to discourage special interest groups from exerting pressure.
-The slight modification to the government objective function demonstrates that governments may indeed want to use trade agreements to reduce lobbying.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.}
+The slight modification to the government objective function demonstrates that governments may indeed want to use trade agreements to reduce lobbying.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
 
-Thus, examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
+In the main example of this paper, I show that the same combination---endogenously-determined political pressure and a general approach to the government's objective function---can explain why the conditions for invoking the WTO Safeguards measure rely on purely observable economic variables, and why the level of protection governments can choose when invoking a Safeguard is related to the injury imposed by the shock and not its political impact.
+The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role for which it was intended.
 
-This example shows that it may be important to consider how the features of the government objective function interact with the question being asked.
+Through the lens of these examples, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective function have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.
+It is important to consider how these features interact with the question being asked.
 For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the shape of the objective function is not crucial, there may be no loss in employing the simple form.
@@
 This cannot microfound a flexible model as in \Textcite{longvousden} in which weights depend on lobbying activity or in which weights are the result of shocks.}
-
-In another example, I show that the same combination---endogenously-determined political pressure and a general approach to the government's objective function---can explain why the conditions for invoking the WTO Safeguards measure rely on purely observable economic variables, and why the level of protection governments can choose when invoking a Safeguard is not restricted.
-The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role that is intended for it.
-
-Through the lens of these examples, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective function have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.
 

### 2016-04-02 intro.tex
adjust road map paragraph

@@
 
-In the next section, I present the model and some preliminary results.
-Section~\ref{sec:rigid} contains the analysis of rigid tariffs while Section~\ref{sec:escape} explores trade agreements with escape clauses.
-The political welfare function of the government is examined in depth in Section~\ref{sec:objfcn} before returning to examine the escape clause in an environment with both endogenous and exogenous political pressure in Section~\ref{sec:escape2}.
+In the next section, I present the model and an in-depth discussion of the political welfare function of the government.
+Section~\ref{sec:rigid} contains an analysis of tariff caps without the flexibility of escape and introduces the repeated-game environment.
+Section~\ref{sec:escape} explores questions regarding the design of the escape clause.
 Section~\ref{sec:concl} concludes.

### 2016-04-04 draft.tex
proof sec:stage

@@
 %dell laptop
-\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
+%\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
 
 %wallis computer
-%\addbibresource{C:/Users/wallis/Dropbox/Research/xBibs/tradeagreements.bib}
+\addbibresource{C:/Users/wallis/Dropbox/Research/xBibs/tradeagreements.bib}
 
@@
 
-As is standard in the literature, it is assumed that the production of each good requires the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factor is tied to the price of the good in whose production their factor is used.
+As is standard in the literature, I assume that the production of each good requires the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factor is tied to the price of the good in whose production their factor is used.
 
 Note that $P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective importing country's tariff.
-As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff.
+As profits and producer surplus (identical in this model) in a sector increase in the price of its good, profits in the import-competing sector also increase in the domestic tariff.
 This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.
@@
 The stage-game timing is as follows.
-First, the governments cooperatively form a trade agreement.\footnote{See Section~\ref{sec:rigid} for a discussion of why lobbies are assumed not to be involved at the trade agreement formation stage.} After the trade agreement is concluded, the special interest group representing the import-competing industry in each country lobbies the government for protection in the form of tariffs and any exogenous shock is realized.
-Finally, given the trade agreement, enforcement conditions, and political pressure it experiences, each government chooses the applied tariff level for its import-competing good.
+First, the governments cooperatively form a trade agreement.\footnote{See Section~\ref{sec:rigid} for a discussion of why lobbies are assumed not to be involved at the trade agreement formation stage.} After the trade agreement is concluded, any exogenous shock is realized.
+Next the special interest group representing the import-competing industry in each country lobbies the government for protection in the form of tariffs.
+Finally, given the trade agreement, enforcement conditions, and political pressure it experiences, each government chooses the applied tariff on its import-competing good.
 
@@
 where $\mathit{CS}$ is consumer surplus, $\pi$ represents profits, $\ga(s,e)$ is the weight placed on profits (producer surplus) in the import-competing industry, $s$ is an exogenous state variable, $e$ is lobbying effort, and $\mathit{TR}$ is tariff revenue.
-Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, is affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{The standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them.
+Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, is affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{While the standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them.
 An isomorphism can be made between the two forms as discussed in \Textcite{buzard2013b}.}
@@
 \begin{equation}
-  U_\text{L} = \pi(\tau(\ga(e))) - e
+  U_\text{L} = \pi(\tau(\ga(s,e))) - e
   \label{eq:lv3}
@@
 The implication is that the lobby can directly influence only the home government, and so the influence of one country's lobby on the other country's government occurs only through the tariffs selected.\footnote{cfr.
-\Textcite{gh95}, page 685.} $\ga(s)$ is assumed to be private information of the government, with specifics about its distributions given below.
+\Textcite{gh95}, page 685.} $\ga(s,e)$ is assumed to be private information of the government, with specifics about its distributions given below.
 This makes the environment one of asymmetric information, with asymmetries potentially coming from two sources: the exogenous shocks and the endogenous behavior of the special interest group.
@@
 $\tau^N$ is increasing in $\ga$ and the second order condition is satisfied for all values of $\ga < 17/2$.
-Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$ it seems reasonable to assume that this condition will be satisfied in equilibrium.
+Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$ I assume that this condition is satisfied in equilibrium.
 
-In order to derive the jointly efficient tariff choice---that is, the tariff that maximizes the joint welfare of the home and foreign government, we require the welfare of the foreign government, $W^*(\ga^*(s^*,e^*),\tau,\tau^*)$ = $W_Y^*(\ga^*(s^*,e^*),\tau^*) + W_X^*(\tau)$, where
+In order to derive the jointly efficient tariff choice---that is, the tariff that maximizes the joint welfare of the home and foreign government, we also need the welfare of the foreign government.
+We can write $W^*(\ga^*(s^*,e^*),\tau,\tau^*)$ = $W_Y^*(\ga^*(s^*,e^*),\tau^*) + W_X^*(\tau)$, where
 \[
@@
 \[
-  W_X^*(\tau) = \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2 
+  W_X^*(\tau) = \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2 .
 \]

### 2016-04-04 draft.tex
revise sec:objfcn

@@
 \label{sec:objfcn}
-When political pressure is determined endogenously, a government can predict the political pressure it will face and in some cases use a trade agreement to alter lobbying incentives.
-Since the government has some control through the trade agreement over the level of $\ga$ it will face, it becomes important to understand how the government's welfare varies in $\ga$.
-The perceived role that trade agreements can play is therefore influenced by the choice of government objective function.
+When political pressure is determined endogenously, a government can predict the political pressure it will face.
+I argue below that in some cases a government may want to use a trade agreement to alter lobbying incentives.
+If this is true, then the perceived role that trade agreements can play is influenced by the choice of government objective function.
 
@@
 \end{equation}
-We need to know at what $\ga(e)$ the government's welfare is maximized, given that it will choose $\tau$ according to either Equation~\ref{eq:nash} when setting tariffs unilaterally or Equation~\ref{eq:eff} when setting tariffs in the context of a trade agreement.
+The government chooses $\tau$ according to Equation~\ref{eq:eff} when setting tariffs in the context of a trade agreement.
+The government knows that this trade agreement tariff will impact lobbying incentives and therefore the realization of $\ga(e)$ at the applied tariff-setting phase.
+We want to know how $\ga(e)$ affects the government's welfare so that we can see how the desire to manipulate lobbying incentives may influence the government's choice of trade agreement tariff.
 
@@
 
-This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the partially weighted government welfare function from Expression~\ref{eq:wel} under the parameterization of \Textcite{bs2005} used throughout this paper.
-Graphed in terms of tariffs, one can see that from $\tau=0.05$---the tariff the home government optimally sets when making unilateral policy with no political pressure---government welfare steadily increases as the trade agreement tariff increases.
+This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the partially weighted government welfare function from Expression~\ref{eq:wel}.
+Note that the graph is in terms of tariffs, which are monotonically increasing in $\ga$.
+You can see that from $\tau=0.05$---the tariff the home government optimally sets when making unilateral policy with no political pressure---government welfare steadily increases as the trade agreement tariff increases.
 This is because the governments can expect that ex-post political pressure will increase in the trade agreement tariff, and government welfare increases everywhere in political pressure.
@@
 
-This example stands in contrast to a less concave (in the sense of Arrow Pratt) $\ga$.
-Take $\ga(e) = 1 + e^{.3}$.
+This example stands in contrast to one in which $\ga(e) = 1 + e^{.3}$.
 With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0010$, which results in $\ga=1.126$ and a unilateral tariff of $0.068$.
@@
 By setting a tariff cap of zero---assuming the conditions for self enforcement outlined in Section~\ref{sec:selfweak} are met---the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.
-The stark prediction that the best the government can do is by setting the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has an interior global maximum in $\ga$.
+The stark prediction that the best the government can do is to set the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has an interior global maximum in $\ga$.
+A candidate for the government objective function in which the maximum is strictly interior is one in which returns to lobbying decrease as tariffs increase as in \Textcite{ethier2012}.
+In the current model, one would take $\ga$ to be a decreasing function of $\tau$.
+This implies that government welfare must eventually decrease in $\ga$, providing for the required concavity in $\ga$.
 
-With this weighted welfare function, we see that the trade agreement is employed to different ends depending on $\ga(e)$.
+With the fully-weighted welfare function, we see that the trade agreement is employed to different ends depending on $\ga(e)$.
 If the lobby's ability to make an impact on the political process is strong enough, the government will be chiefly concerned with the profits of the import-competing industry and will use the trade agreement solely to internalize the terms-of-trade externality.
@@
 I do not argue that the government welfare function in Expression~\ref{eq:weighted} perfectly represents the preferences of those who negotiate trade agreements.
-Another reasonable candidate for the government objective function embodies diminishing returns to lobbying as tariffs increase as in \Textcite{ethier2012}.
-In the current model, one would take $\ga$ to be a decreasing function of $\tau$.
-This implies that government welfare must eventually decrease in $\ga$, providing for concavity in $\ga$ and a strict interior maximum.
+I only want to make the point that assumptions on the features of the government objective function can be important drivers of results on the role that trade agreements can play.
 

### 2016-04-04 draft.tex
move defn of \ga, \ga(s), \ga(e) to first place I use \ga

@@
 where $\mathit{CS}$ is consumer surplus, $\pi$ represents profits, $\ga(s,e)$ is the weight placed on profits (producer surplus) in the import-competing industry, $s$ is an exogenous state variable, $e$ is lobbying effort, and $\mathit{TR}$ is tariff revenue.
-Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, is affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{While the standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them.
-An isomorphism can be made between the two forms as discussed in \Textcite{buzard2013b}.}
+Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, may be affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{While the standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them.
+An isomorphism can be made between the two forms as discussed in \Textcite{buzard2013b}.} I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to the case where political pressure  results from the rent-seeking behavior of lobbies only.
 
@@
 In Section~\ref{sec:escape2} I will examine the more realistic case where endogenous and exogenous political forces coexist.
-Until then, I will consider the cases separately to establish clearly the different dynamics that arise from the two distinct types of influence.
-I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to political pressure that results from the rent-seeking behavior of lobbies.\footnote{See Section~\ref{sec:escape2} for a detailed discussion of the relationship and interactions between these two extreme cases.} I begin with the case of rigid tariffs---that is, trade agreements that have no provision for flexibility.
-Rigid tariffs cannot be made to depend on the realized level of $\ga$ and so incentive compatibility considerations will not come into play.
+Until then, I will consider the cases separately to clearly establish the different dynamics that arise from the two distinct types of influence.
+I begin with the case of rigid tariffs---that is, trade agreements that have no provision for flexibility.
+Rigid tariffs cannot be made to depend on the realized level of $\ga$ and so incentive compatibility considerations will do come into play.
 

### 2016-04-04 draft.tex
proof sec:perfect

@@
 The former, which is undoubtedly an important possibility in many institutional settings, is left as an important extension.\footnote{Note that the endogenous trade policy literature that involves trade agreements (e.g.
-\Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure while answering a different set of questions.}
+\Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure while answering a different set of questions.
+Results for the case of ex-ante lobbying are available from the author upon request.}
 
@@
 
-In the model under examination here with endogenous political effort,\footnote{As above, in this section I consider endogenous political pressure alone in place of exogenously-determined $\ga$.} in place of taking expectations over a range of probabilistically-determined levels of exogenous political pressure, we can backward induct to determine the lobby's effort decision and the government's optimal choice of tariff caps in anticipation of the lobby's behavior.
+When political pressure is endogenous, in place of taking expectations over a range of probabilistically-determined levels of exogenous political pressure, we can backward induct to determine the lobby's effort decision and the government's optimal choice of tariff caps in anticipation of the lobby's behavior.
 
 In the final stage, the government will seek to maximize its welfare by choice of $\tau$ given the political pressure that it faces and the enforcement of the tariff cap, here denoted $\tau_{W,e}^R$ for the case of endogenous political pressure, as distinct from $\tau_W^R$ in the case of exogenous political pressure.
-Thus, the home government unilaterally maximizes Expression~\ref{eq:wel}, so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as  $\tau^N(\ga(e)) \leq \tau_{W,e}^R$.
+The home government unilaterally maximizes Expression~\ref{eq:wel} so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as  $\tau^N(\ga(e)) \leq \tau_{W,e}^R$.
 Otherwise, it must set $\tau = \tau_{W,e}^R$ since the weak binding $\tau_{W,e}^R$ is externally enforced.
@@
 
-We know that the lobby will either choose its unconstrained optimal effort $e^L$, or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding.
+We know that the lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding.
 The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$.
-The government is unable to use a weak binding to encourage extra lobbying---at least in the case with only one import-competing lobby under consideration here.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}
+The government is unable to use a weak binding to encourage extra lobbying in the case with only one import-competing lobby under consideration here.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}
 

### 2016-04-04 draft.tex
proof sec:selfweak

@@
 This implies that $\tau^R_{W,e}$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments do not want as much lobbying as possible.
-To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.
-See Section~\ref{sec:objfcn} for further discussion.}
+To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.}
 
@@
 In fact, any strong binding accomplishes this very feat, but is also \textit{unable} to generate lobbying activity because of its effect on lobbying incentives.
-Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is a related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.}
+Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.}
 
@@
 \label{sec:self}
-Here I relax the assumption of perfect external enforcement, since this is not widely available in the context of international trade relations.
+Here I drop the assumption of perfect external enforcement since external enforcement is not widely available in the context of international trade relations.
 External enforcement is replaced with promises of future cooperation and punishment is modeled via a repeated game.
-I will first describe this repeated game and then present the analysis for the case of (rigid) weak bindings.
+I begin by describing the repeated-game set-up.
 
 \subsubsection{Repeated Game}
-For the setting of exogenous political pressure, Bagwell and Staiger assume that $\ga$ is independently and identically distributed across governments and time.
-Because the governments are faced with asymmetric information, perfect public equilibrium (PPE) is the appropriate solution concept.
-Here attention is restricted to symmetric, stationary PPE.
+Because the governments are faced with asymmetric information, perfect public equilibrium (PPE) is the appropriate solution concept. Here attention is restricted to symmetric, stationary PPE.
 
@@
 We know that the former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here.
-Following \Textcite{bs2005} and most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternatives that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.}
+Following most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternatives that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.}
 	%see KRW, Buzard 2014 for alternatives
@@
 We will see that moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem.
-However, it does not alter the within-period costs and benefits and so in the case of endogenous political pressure we will see the same result that the optimal trade agreement tariff is invariant to the enforcement conditions.
+Moving to the repeated-game environment with self enforcement does not alter the governments costs and benefits so that the government's optimal trade agreement tariff is invariant to the enforcement conditions.
+This creates the possibility that the politically-optimal trade agreement tariff may not be self-enforcing.
 
@@
 
-If there is no $\de$ such that Inequalities~\ref{exp:govincent} and \ref{exp:lobby} hold simultaneously, the governments must raise the trade agreement tariff, which loosens both incentives constraints simultaneously.
+If there is no $\de$ such that Inequalities~\ref{exp:govincent} and \ref{exp:lobby} hold simultaneously, the governments must raise the trade agreement tariff, which loosens both incentive constraints simultaneously.
 

### 2016-04-04 draft.tex
revise entire EC section

@@
 \label{sec:escape}
-I follow \Textcite{bs2005} in examining a trade agreement with an escape clause by adding a second negotiated binding to the agreement so that there is one (weak) binding for when political pressure is low / normal and a higher (weak) binding for exceptional circumstances of high levels of political pressure.
+I begin by following \Textcite{bs2005} in examining a trade agreement with an escape clause by adding a second negotiated binding to the agreement so that there is one (weak) binding for when political pressure is low and a higher, weak binding for exceptional circumstances of high levels of political pressure.
 Escape clauses are common features of trade agreements in practice, and when one models political pressure as exogenous, an escape clause seems attractive: in the presence of a particularly large negative political shock, being bound to a tariff designed for normal times would cause significant welfare losses to the politician.
@@
 The government would actually be \textit{truthfully} reporting the higher level of $\ga(e)$ as long as it is worthwhile for the lobby to increase its effort to this level.
-The appeal of an escape clause is missing in this case, but the potential for governments to be forced ex-post to exploit the escape clause in a way that damages their ex-ante welfare remains.\footnote{Notice that the ex-post rent-seeking that creates this ex-ante loss is different in the two cases.
+The appeal of an escape clause is missing in this case.
+What remains is the potential for governments to be forced ex-post to exploit the escape clause in a way that damages their ex-ante welfare.\footnote{Notice that the ex-post rent-seeking that creates this ex-ante loss is different in the two cases.
 With exogenous shocks, the government increases its welfare through the terms-of-trade channel.
@@
 
-\subsection{Escape Clause with Side Payments}
+\subsection{Escape Clauses with Side Payments}
 Suppose instead that the use of the escape clause must be accompanied by a side payment when the higher tariff is applied.
@@
 Here the government is truthfully revealing the level of political pressure it faces, but that pressure is created by endogenous forces that are intensified by the existence of the escape clause.
+The escape clause is incentive compatible, but as will be clarified in the next section, the spirit of the escape clause is not necessarily upheld.
 
@@
 
-\subsection{The Purpose and Design of the Escape Clause}
+\subsection{Escape Clauses with both Exogenous and Endogenous Political Pressure}
 \label{sec:escape2}
-The title of this section pays direct homage to \Textcite{bs2005} for their above-referenced, foundational work on the escape clause in a setting with exogenous political pressure.
-Here, I ask how the addition of endogenous political pressure alters the optimal design of an escape clause.
-
 When $\ga$ is exogenous, the purpose of the escape clause is clear: it improves welfare by allowing the government to apply a higher tariff when the realization of $\ga$ is particularly high.
@@
 
-In Section~\ref{sec:escape}, I show that there is no such benefit when political pressure is purely endogenous in nature.
+There is no such benefit when political pressure is purely endogenous.
 All political pressure can be anticipated by the government and accounted for in the normal binding, so any higher tariff provided for escape would either not be used or would reduce government welfare by encouraging excess political pressure in the case where the trade agreement tariff is being employed to reduce lobbying.
@@
 In reality, it is likely that both exogenous shocks and endogenous forces contribute to the pressure to which policy makers are subject.
-Thus, the desire for flexibility that derives from the purely exogenous case remains and we desire to see if it is possible to implement an escape clause in the face of the additional endogenous source of political pressure.
+Thus, the need for flexibility that derives from the purely exogenous case remains and we desire to see if it is possible to implement an escape clause in the face of the additional endogenous source of political pressure.
 
-So we go back to the general formulation where $\ga$ is a function of both lobbying effort $e$ and exogenously-determined events that create political pressure as in \Textcite{ms2011}, $s$.
+So we go back to the general formulation where $\ga$ is a function of both lobbying effort $e$ and exogenously-determined events $s$ that create political pressure as in \Textcite{ms2011}.
 Let us take a simple case where the political pressure from the two sources are additively separable so that $\ga(s,e) = \ga(s) + \ga(e)$.
@@
 
-Given that the intent of the escape clause is to provide flexibility for high realizations of $\ga(s)$, we see that the presence of endogenous lobbying can easily destroy its efficacy for this purpose as the lower binding will never be used.
+Given that the intent of the escape clause is to provide flexibility for high realizations of $\ga(s)$, we see that the presence of endogenous lobbying can easily destroy its efficacy as the lower binding will never be used.
 The water in the escape clause will be filled in by endogenous political pressure.
@@
 
+In the next section, I turn to the kind of escape clause that is both implementable and able---at least in some cases---to address shocks without encouraging rent-seeking.
+
 \subsection{An Escape Clause For Endogenous Politics}
@@
 I assume this is a weak binding and add the subscript $F$ to denote the environment with flexibility to deal with shocks.
-	\item The exogenous shock $s$ is realized. It's level is assumed costlessly verifiable for simplicity.
-	\item The lobby chooses its effort level $e_F$.
-	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau(s)$ and $\tau^B(\ga(s,e_F))$.
+	\item The exogenous shock $s$ is realized. Its level is assumed costlessly verifiable for simplicity.
+	\item The lobby chooses its effort level $e^F$.
+	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau(s)$ and $\tau^B(\ga(s,e^F))$.
 \end{enumerate}
@@
 If the government applies a tariff that is higher than either $\tau^F_{W,e}$, $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
-I label this tariff $\tau^B(\ga(s,e_F))$, where $e^F$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.
+I label this tariff $\tau^B(\ga(s,e^F))$, where $e^F$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.
 
 This is a violation of the trade agreement even in the presence of the safeguard and incurs retaliation.
-It is a violation of the agreement because the applied tariff level is higher than is needed to remedy the injury caused by the shock; the Agreement on Safeguards instructs that ``A Member shall apply safeguard measures only to the extent necessary to prevent or remedy serious injury and to facilitate adjustment'' (Article 5.1).\footnote{The paragraph goes on to give even more clarification: ``If a quantitative restriction is used, such a measure shall  not reduce the quantity of imports below the level of a recent period which shall be the average of imports in the last three representative years for which statistics are available, unless clear justification is given that a different level is necessary to prevent or remedy serious injury.''}
+It is a violation of the agreement because the applied tariff level is higher than is needed to remedy the injury caused by the shock; the WTO Agreement on Safeguards instructs that ``A Member shall apply safeguard measures only to the extent necessary to prevent or remedy serious injury and to facilitate adjustment'' (Article 5.1).\footnote{The paragraph goes on to give even more clarification: ``If a quantitative restriction is used, such a measure shall  not reduce the quantity of imports below the level of a recent period which shall be the average of imports in the last three representative years for which statistics are available, unless clear justification is given that a different level is necessary to prevent or remedy serious injury.''}
 
@@
 For ease of analysis, I use a punishment length of two periods.
-This simply implies that a country would return to the agreement either after applying a valid escape clause for three years or after violating the agreement.
+This implies that a country would return to the agreement either after applying a valid escape clause for three years or after violating the agreement.
 
@@
 Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may chooses to set the trade agreement tariff in the presence of an escape clause $\left(\tau^F_{W,e}\right)$ so that in some cases the lobby's incentive constraint is violated.
-This would lead to the legislature's constraint being violated as well, leading to a dispute.
+This would lead to the legislature's constraint being violated as well, creating a dispute.
 
-One important thing \textit{does not} happen in this environment: lobbies do not use the escape clause as cover for rent seeking.
-The government may use the escape clause to provide extra protection commensurate to the shock.
+One important outcome \textit{does not} occur in this environment: lobbies do not use the escape clause as cover for rent seeking.
+The government may use the escape clause to provide extra protection to compensate for a shock.
 This provides a looser incentive constraint than in the absence of the escape clause and improves self-enforceability of the agreement.
 Or the lobbies may still find it in their interest to exert effort to encourage the government to provide this higher level of protection.
-This will provoke a dispute, but the protection afforded by the higher tariff but this will not be hidden behind the escape clause.
+This will provoke a dispute, but the protection afforded by the higher tariff will not be hidden behind the escape clause.
 So an escape clause for the purposes of providing breathing room from exogenous shocks \textit{is} possible.

### 2016-04-04 intro.tex
clarify contribution on tractability; move repeated games cites

@@
 This is one example of a design question whose answers are sensitive to assumptions about the endogeneity of political pressure.
-In order to examine it and others, I build a model that is comparable to \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights in a repeated-game setting.
+In order to examine this question and others, I use a model that is comparable to \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights in a repeated-game setting.
+Because of its tractability, this model can address a wide range of institutional design questions while also allowing endogenous politics to be taken into account.\footnote{Repeated non-cooperative game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.}
 
@@
 
+This work is also related to the literature on the endogenous political economy of trade.
+The foundational work is \Textcite{gh94}; the insights are applied to trade agreements in \Textcite{gh95}.
+\Textcite{mrc2007} advanced the literature by demonstrating that there is a domestic commitment role for trade agreements.
+\Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government has a separation-of-powers structure, while \Textcite{coatesludema} demonstrate that, in the presence of opposition to a trade agreement from foreign lobbies, it may be optimal to liberalize trade unilaterally.
+
 A rich literature has been developed to address questions concerning the design and enforcement of trade agreements.
-Repeated non-cooperative game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.
 
@@
 
-This work is also related to the literature on the endogenous political economy of trade.
-The foundational work is \Textcite{gh94}; the insights are applied to trade agreements in \Textcite{gh95}.
-\Textcite{mrc2007} advanced the literature by demonstrating that there is a domestic commitment role for trade agreements.
-\Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government has a separation-of-powers structure, while \Textcite{coatesludema} demonstrate that, in the presence of opposition to a trade agreement from foreign lobbies, it may be optimal to liberalize trade unilaterally.
-
 In the next section, I present the model and an in-depth discussion of the political welfare function of the government.

### 2016-04-04 draft.tex
remove cites to irrelevant papers; move punishment papers to WEC paragraph in sec:ECendog

@@
 The intuition is the following: \Textcite{buzard2013a} shows that the lobbying effort required to cause the government to enter into a dispute is higher than the Nash equilibrium effort levels.
-Then the tariffs during a dispute are also higher, and punishment at this higher level would be more severe than Nash---both for the government and the lobby.
+Then the tariffs during a dispute are also higher, and punishment at this higher level would be more severe than Nash---both for the government and the lobby.\footnote{\Textcite{beshkar2010a} shows that when one assumes that utility is not transferable between countries as has become common in the literature, the optimal mechanism involves less-than-proportional retaliation against parties who have defected from the agreement.
+\Textcite{martinvergote} demonstrate that future punishment provides for higher welfare than contemporaneous punishment when governments are sufficiently patient.
+Indeed, they show that retaliation is a necessary feature of any efficient equilibrium in this environment.
+\Textcite{hungerford} and \Textcite{riezman1991} also consider the impact of different assumptions about reactions and timing of punishments for deviations from agreements.}
 

### 2016-04-04 intro.tex
remove cites to irrelevant papers; move punishment papers to WEC paragraph in sec:ECendog

@@
 
-Maggi and Staiger have a series of papers that employ an exogenous political economy force to study questions about the design of trade agreements and trade dispute settlement.
+%Maggi and Staiger have a series of papers that employ an exogenous political economy force to study questions about the design of trade agreements and trade dispute settlement.
 \Textcite{ms2012a} study the conditions under which property versus liability rules will be optimal when renegotiation of agreements is possible.
@@
 
-\Textcite{beshkar2010a} shows that when one assumes that utility is not transferable between countries as has become common in the literature, the optimal mechanism involves less-than-proportional retaliation against parties who have defected from the agreement.
 \Textcite{beshkar2010b} compares the GATT escape clause to the WTO Safeguards agreement and shows that the DSB as a non-binding arbitrator can assist governments in self-enforcing their trade agreements.
-\Textcite{martinvergote} demonstrate that future punishment provides for higher welfare than contemporaneous punishment when governments are sufficiently patient.
-Indeed, they show that retaliation is a necessary feature of any efficient equilibrium in this environment.
-\Textcite{hungerford} and \Textcite{riezman1991} also consider the impact of different assumptions about reactions and timing of punishments for deviations from agreements.
 

### 2016-04-04 intro.tex
revise paragraphs on political commitment device in intro

@@
 In order to examine this question and others, I use a model that is comparable to \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights in a repeated-game setting.
-Because of its tractability, this model can address a wide range of institutional design questions while also allowing endogenous politics to be taken into account.\footnote{Repeated non-cooperative game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.}
+Because of its tractability, this model can address a wide range of institutional design questions while also allowing endogenous politics to be taken into account.\footnote{Repeated non-cooperative game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{coatesludema}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.
+\Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government has a separation-of-powers structure.}
 
@@
 At issue is the fact that the standard government objective function is everywhere increasing in political pressure.
-In an environment with endogenous political pressure, this leads inexorably to the conclusion that governments never want to discourage special interest groups from exerting pressure.
-The slight modification to the government objective function demonstrates that governments may indeed want to use trade agreements to reduce lobbying.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
+The objective function use in the foundational work in the endogenous political economy of trade, \Textcite{gh94} and its extension to trade agreements in \Textcite{gh95} is also everywhere increase in political pressure.
+This leads inexorably to the conclusion that governments do not want to discourage special interest groups from exerting pressure except in special cases such as those demonstrated by \Textcite{mrc2007} and \Textcite{mitra}.
+
+The modification to the government objective function I propose leads to the result that governments may indeed want to use trade agreements to reduce lobbying in many environments.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
 
@@
 
-This work is also related to the literature on the endogenous political economy of trade.
-The foundational work is \Textcite{gh94}; the insights are applied to trade agreements in \Textcite{gh95}.
-\Textcite{mrc2007} advanced the literature by demonstrating that there is a domestic commitment role for trade agreements.
-\Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government has a separation-of-powers structure, while \Textcite{coatesludema} demonstrate that, in the presence of opposition to a trade agreement from foreign lobbies, it may be optimal to liberalize trade unilaterally.
-
-A rich literature has been developed to address questions concerning the design and enforcement of trade agreements.
+%A rich literature has been developed to address questions concerning the design and enforcement of trade agreements.
 

### 2016-04-04 intro.tex
finish revisions to intro

@@
 The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role for which it was intended.
+If the purpose of contingent protection measures is actually to provide a political safety valve, this could provide an explanation for the infrequency with which safeguards are invoked under the WTO, as I argue that the WTO's escape clause has been designed to provide relief only from economic shocks.
 
@@
 
+\textbf{A few additional comments on the literature}.
 Tariff caps have received particular attention in the literature that takes political pressure to be exogenous.
@@
 
+On escape clauses, \Textcite{bs1990} is similar in spirit to \Textcite{bs2005}.
+Here, there is no private information and trade volume shocks in the presence of self-enforcement constraints are behind the need for an escape clause.
+\Textcite{bc} find empirical support for this theory.
 \Textcite{beshkar2010b} compares the GATT escape clause to the WTO Safeguards agreement and shows that the DSB as a non-binding arbitrator can assist governments in self-enforcing their trade agreements.

### 2016-04-06 feedback.tex
rename results to propositions
add notes from rochester seminar to feedback

@@
 	\item Page 1:  Should define gamma before using it, especially given how critical it is to the paper.
-	\item Page 17: `is invariant'
-	\item Page 26: `an interior'
-	\item I'm wondering why are the results referred to as `Results' and not `Propositions'?
+	\item \sout{Page 17: `is invariant'} DONE
+	\item \sout{Page 26: `an interior'} DONE
+	\item \sout{I'm wondering why are the results referred to as `Results' and not `Propositions'?
 Odd question when I put it like that!
 But since we're used to `Propositions', saying `Results' is a bit distracting.
-Or maybe I'm just being old fashioned ;-)
+Or maybe I'm just being old fashioned ;-)} DONE
 \end{enumerate}
@@
 
+\newpage
+Rochester Seminar, April 6, 2016 (Mostly Mark Bils)
+\begin{itemize}
+	\item Need to say that I'm ruling out side payments so it's clear that there can't just be a menu of punishments that exactly keeps each government in line
+		\begin{itemize}
+			\item Risk (average) vs. incentives (marginal) impossibility result --- but with no money transfers, even this breaks down
+			\item Can I think of this as an insurance problem between the governments?
+		\end{itemize}
+	\item Show what the optimum is, then why they can't get it, how they have to try to mimic it with these instruments (tariff caps, escape clause, these types of punishments, etc.)
+	\item Going from $\ga(s)$ to $\ga(e)$ --- essentially taking away the structure of the $s$'s was confusing b/c the policy space doesn't make any sense
+\end{itemize}
+
 \end{document}

### 2016-06-20 presentation.tex
copy rochester presentation from dropbox as presentation.tex and compile

%\documentclass{beamer} 
\documentclass[handout]{beamer} 
\usetheme{Ilmenau}
\usepackage{graphicx,verbatim,hyperref}
\usepackage{textpos}

\usecolortheme{beaver}
\useinnertheme{default}
\setbeamertemplate{itemize item}[triangle]
\setbeamertemplate{itemize subitem}[triangle]
\setbeamertemplate{itemize subsubitem}[circle]
\setbeamertemplate{enumerate items}[default]
\setbeamertemplate{blocks}[upper=block head,rounded]
\setbeamercolor{item}{fg=black}
\usefonttheme{serif} %should allow ccfonts to take effect

\usepackage{cite}
\usepackage{times, verbatim,xcolor,bm}
%\usepackage[usenames,dvipsnames]{color}
\usepackage{amsbsy,amssymb, amsmath, amsthm}
\usepackage{booktabs}
%David miller's fonts
	\usepackage[T1]{fontenc}
	\usepackage[boldsans]{ccfonts}
	%\usepackage[boldsans]{concmath}
	\usepackage[euler-hat-accent]{eulervm}

\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\btn}{\bm{\tau^{tw}}}
\newcommand{\ga}{\gamma}
\newcommand{\ve}{\varepsilon}
\newcommand{\ta}{\theta}
\newcommand{\de}{\delta}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}

\newenvironment{changemargin}[2]{% 
  \begin{list}{}{% 
    \setlength{\topsep}{0pt}% 
    \setlength{\leftmargin}{#1}% 
    \setlength{\rightmargin}{#2}% 
    \setlength{\listparindent}{\parindent}% 
    \setlength{\itemindent}{\parindent}% 
    \setlength{\parsep}{\parskip}% 
  }% 
  \item[]}{\end{list}}

	\let\Tiny=\tiny

\title[Endogenous Politics and the Design of Trade Institutions\hspace{2.05in}\insertframenumber/\inserttotalframenumber]{Endogenous Politics and \\ the Design of Trade Institutions}
%\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{April 6, 2016}
\begin{document}
\maketitle
%\insertpresentationendpage removed b/c of appendix

\section{Overview}
\subsection{Preview}
\begin{frame}{The Questions}

\pause
\begin{enumerate}[<+->]
\item When is endogenizing political pressure important for answering optimal design questions?
	\begin{itemize}
		\item Exogenous vs. endogenous politics
	\end{itemize}
	\vskip.1in
\item Can trade agreements be used to manipulate domestic lobbying incentives?
	\begin{itemize}
		\item Government objective function
	\end{itemize}
\end{enumerate}

\end{frame}

\begin{frame}{Preview}
\frametitle{Political Economy of Trade Institutions}
\pause
With a few exceptions, TA literature has taken political economy forces to be exogenous. I:
\pause
\begin{itemize}[<+->]
	\item endogenize politics into a standard model for studying TA design questions
	\item carefully distinguish between dynamics induced by exogenous and endogenous politics for
		\begin{itemize}[<+->]
			\item base case with tariff caps
			\item tariff caps with escape clause
		\end{itemize}
	\item examine escape clause design when both exogenous and endogenous forces are present
\end{itemize}
\end{frame}

\begin{frame}{Preview}
\frametitle{Results}

\pause
\begin{itemize}[<+->]
	\item Show that TAs may be used to manipulate domestic political actors (no long-run distortions)
	\item For both tariff caps and escape clauses, outcomes are very different with endogenous politics
	\item Demonstrate that (standard, theoretical) escape clause can't work in the presence of endogenous political pressure
		\begin{itemize}
			\item Points to real-world design of WTO Agreement on Safeguards
			\item May explain why escape clause has fallen out of use
		\end{itemize}
\end{itemize}
\end{frame}

\begin{comment}
\begin{frame}{Outline of Talk}
\pause
\begin{enumerate}[<+->]
	\item Model
	\item Government Objective Function
	\item Base Model with Tariff Caps
	\item Escape Clause
	\item Conclusion
\end{enumerate}
\end{frame}
\end{comment}

\subsection{Role and Design of TAs}
\begin{frame}{Role of Trade Agreements: TOT Externality}
\pause
Bagwell and Staiger (2002)
\pause
\begin{itemize}[<+->]
	\item Joint social welfare maximized at free trade
	\item Trade war (i.e. no agreement)
		\begin{itemize}
			\item Maximize with respect to home country welfare only
			\item Terms of trade (TOT) externality $\Rightarrow$ positive tariffs  
		\end{itemize}
	\item Trade agreements
    \begin{itemize}
			\item Now take into account impact on foreign welfare
			\item Internalize TOT externality $\Rightarrow$ free trade
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Role of Trade Agreements: TOT Externality}
\pause
Grossman and Helpman (1995)
\pause
\begin{itemize}[<+->]
	\item Add endogenous politics
	\item Now in ``Trade War'': two reasons for positive tariff
		\begin{itemize}
			\item TOT externality + pressure from import competing lobby
		\end{itemize}

	\item Trade agreement: only internalizes TOT externality
\end{itemize}
\end{frame}

\begin{frame}{Role of Trade Agreements: New Trade Theory}
\pause
``New Trade Theory'' externalities
\pause
\begin{itemize}[<+->]
	\item Source: imperfect competition (firm level)
		\begin{itemize}
			\item Delocation: Ossa 2011, Bagwell and Staiger 2012
			\item Profit Shifting: Mrazova 2011, Ossa 2012  
		\end{itemize}
	\item Trade agreements
    \begin{itemize}
			\item Now take into account impact on foreign welfare
			\item Internalize these new externalities
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Role of Trade Agreements: Domestic Commitment}
\pause
\begin{itemize}[<+->]
	\item Maggi and Rodriguez-Clare (1998, 2007)
		\begin{itemize}
			\item Allow for (imperfect) capital mobility
			\item Domestic investment decisions depend on level of protection
			\item Inability to commit $\Rightarrow$ investment too high b/c importers know protection will respond
			\item Trade agreements provide commitment device
    \end{itemize}
	\item Mitra (2002)
		\begin{itemize}
			\item Here distortion is wasted resources in lobby formation
		\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Design of Trade Agreements}
\pause
\begin{itemize}[<+->]
	\item \textbf{Tariff caps}: Bagwell and Staiger 2005, Horn et al 2010, Amador and Bagwell 2012; Beshkar and Bond 2012
	\item \textbf{Escape clause}: Bagwell and Staiger 2005, Horn et al 2010, 
	\item \textbf{Shallow vs. deep integration}: Bagwell and Staiger 2001, DeRemer 2014
	\item \textbf{Dispute settlement}: Maggi 1999, Ludema 2001, Maggi and Staiger 2011/2013, Klimenko et al 2008
	%\item \textbf{Property vs. liability rules}: Pauwelyn 2008, Beshkar 2010, Maggi and Staiger 2014
	\item \textbf{Retaliation}: Bown 2002/2004, Beshkar 2010
\end{itemize}
\end{frame}

\section{Model}
\subsection{Economic and Political Structure}
\begin{frame}{Economy}
Two countries: home and foreign (${}^*$)
\pause
\begin{itemize}[<+->]
	\item Separable in two goods: $X$ and $Y$
			\begin{itemize}
				\item $P_i$: home price of good $i$
				\item $P_i^*$: foreign price of good $i$
			\end{itemize}
	\item Demand identical for both goods in both countries
		\begin{itemize}
			\item $D(P_i) = 1 - P_i$
		\end{itemize}
	\item Supply: $Q_X^*(P_X) > Q_X(P_X)$ $\forall P_X$; symmetric for $Y$ 
		\begin{itemize}
			\item $Q_X(P_X) = \frac{P_X}{2}$; $Q_Y(P_Y) = P_Y$
			\item Home net importer of $X$, net exporter of $Y$
		\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{Policy and Politics}

\pause
Home levies $\tau$ on $X$, Foreign levies $\tau^*$ on $Y$
\pause
\begin{itemize}
	\item $P_X=P_X^W + \tau$ increasing in $\tau$
	\pause
	\item $\pi_X(P_X)$ increasing in $P_X$, therefore also $\tau$
\end{itemize}

\pause
\vskip.2in
Non-tradable specific factors motivate political activity
\end{frame}

\begin{frame}{Timeline}
\pause
%\vskip.1in
Each period:
\pause
\begin{enumerate}[<+->]
	\item {\bfseries Trade Agreement Formed}
		\begin{enumerate}[i.]
			%\pause
			\item Governments set trade policy in international agreement
		\end{enumerate}
	%\pause
	\item \textbf{Domestic Politics Played Out}
		\begin{enumerate}[i.]
			%\pause
			\item Exogenous shocks are realized AND/OR
			%\pause
			\item Import-competing industry lobbies government for protection 
		\end{enumerate}
	%\pause
	\item \textbf{Tariffs are Applied}
	%\pause
		\begin{enumerate}[i.]
			\item Given political pressure, governments choose applied tariff levels
		\end{enumerate}
\end{enumerate}
\end{frame}

\subsection{The Players}

\begin{frame}{Applied Tariff Decision}
\pause
  Baldwin-style government objective function:
\pause
\[
  W = \mathit{CS}_X(\tau) + \ga(s,e) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]
\vskip-.1in
\pause
\begin{itemize}[<+->]
	\item Standard \textit{except} weight on import-competing industry profits $\ga(s,e)$: 
		\begin{itemize}
			\item $s$: exogenous shock
			\item $e$: lobbying effort
		\end{itemize}
	\item Optimal applied tariff is a function of $\ga(s,e)$
		\begin{itemize}
			\item Ignores foreign welfare
			\item Takes into account trade agreement enforcement
		\end{itemize}
	\item Assume $\ga$, $\ga^*$ is private info of each government
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Domestic Political Pressure}
Two potential sources
\pause
\begin{enumerate}[<+->]
	\item Exogenous shocks
		\begin{itemize}[<+->]
			\item Shock directly to $\ga$ as in Bagwell $\&$ Staiger (2005): $\ga$, $\ga^*$ with CDF $H(\ga)$ on support $\left[\un{\ga},\ov{\ga}\right]$; or
			\item Can take $\ga$ as a function of $s$: $\ga(s)$
		\end{itemize}

	\item Endogenous effort choice of lobby, $e$
		\begin{itemize}[<+->]
			\item Lobby chooses effort to maximize profits, $\pi(\cdot)$, net of lobbying effort, $e$
			\item Call lobby's optimal effort choice $e^L$
						\[
						  e^L = \max_e \pi(\tau(\ga(e))) - e
						\]
		\end{itemize}
\end{enumerate}

\end{frame}

\begin{frame}{Trade Agreement Negotiation}
	\pause
Model as Nash bargain between the two countries' governments

	\pause
\begin{itemize}[<+->]
	\item Maximize joint political welfare
	\item Disagreement point: non-cooperative outcome
\end{itemize}

\vskip.2in
\pause
Once agreement is set, cooperation enforced by repeated-game punishments conditioned on history, history + DSB signal

\end{frame}

\section{Objective Fcn}
\subsection{Objective Function}
\begin{frame}{Restraining Political Pressure through TAs}
\pause
%It's easy to see that weak bindings keep lobbying alive relative to strong bindings
\begin{itemize}[<+->]
	\item Will TA be used to discourage lobbying? Depends on how gov't welfare varies in $\ga$
	\item With standard Baldwin-style objective function, welfare always increases with $\ga$:
\[
  W = \mathit{CS}_X(\tau) + \ga \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]
	\begin{itemize}
		\item Note `Protection for Sale' objective function is isomorphic
	\end{itemize}

	\item If we instead make weights sum to 1, welfare is no longer monotonic in $\ga$
\end{itemize}
\end{frame}

\begin{frame}
\includegraphics[height=2.75in, width=4.25in]{weight.jpg}
\end{frame}

\section{Tariff Caps}
\subsection{Tariff Caps}
\begin{comment}
\begin{frame}{Tariff Caps: Exogenous vs. Endogenous $\ga$}
\pause
Strong binding: an exact tariff commitment
\begin{itemize}
	\pause
	\item Must apply the precise tariff level specified in TA
\end{itemize}

\pause
\vskip.2in
Weak binding: a tariff cap
\begin{itemize}
\pause
	\item Must set tariff at or below specified level 
\end{itemize}

\pause
\vskip.4in
Perfect external enforcement vs. self-enforcement (repeated-game)

\end{frame}

\begin{frame}{Strong Bindings}
Must set tariff at specified level
\pause
\begin{itemize}[<+->]
	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): optimal (rigid) strong binding is the tariff that is politically optimal for the expected realization of $\ga$
  \item $\ga$ endogenous: if governments have no inherent bias toward protection, the lobbies exert no effort and are afforded no protection 
\end{itemize}
\end{frame}
\end{comment}

{\begin{frame}{Tariff Caps: Exogenous vs. Endogenous $\ga$}
\pause
Must set tariff at or below specified level (aka tariff cap)
\pause
\begin{itemize}[<+->]
	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): Negotiated weak bindings (a) are higher than those gov'ts would choose if they instead negotiated strong bindings and (b) imply that governments with low realizations of $\ga$ set their applied tariffs strictly below the bound level.

	\item $\ga$ endogenous: Governments will not set applied tariffs strictly below the bound level. They may use the weak tariff binding either to encourage and/or restrain endogenous political pressure.

\end{itemize}
\end{frame}

\begin{frame}{Tariff Caps with Self Enforcement}
\pause
\begin{itemize}[<+->]
	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): if governments patient enough ($\de$ high enough), optimal externally-enforced weak binding can be self-enforced
  \item $\ga$ endogenous: not necessarily true
		\begin{itemize}
			\item Problem: lobby is an additional repeated-game player
			\item Lobby's incentive constraint is harder to satisfy as $\de$ increases
		\end{itemize}

\end{itemize}

\end{frame}

\begin{comment}
\begin{frame}{Highlight: One-Shot Intuition}
\pause
\textbf{Legislature}
\pause
\begin{itemize}
	\item Breaks agreement if median legislator prefers $\btn$ to $\bta$
\end{itemize}

\pause
\vskip.1in
\textbf{Lobby}
\pause
\begin{itemize}[<+->]
	\item Given the $\bta$ it faces, lobby knows what $e_b$ is required to break the agreement
 	\item Lobby pays if $\bm{\pi}(\btn)$ - lobbying costs $> \pi(\bta)$
\end{itemize}

\pause
\vskip.13in
\textbf{Executives}
\pause
\begin{itemize}[<+->]
	\item Set $\bta$ to make paying $e_b$ unprofitable
		\pause
		\begin{itemize}
			\item[$\Rightarrow$] $e_b=0$, agreement remains in force
		\end{itemize}
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\end{frame}
\end{comment}

\begin{frame}{Repeated Game Intuition}
\pause
Legislature: break agreement if punishment not strong enough
\pause
\begin{itemize}
	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
\end{itemize}

\vskip.1in
\pause
Lobby: solve for lowest effort ($\ov{e}_b$) that breaks this constraint
\pause
\begin{itemize}
	\item pay $\ov{e}_b$ if it's less than gain from $T$ periods of trade-war profits
\end{itemize}

\pause
\vskip.1in
Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
\pause
\begin{itemize}
	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
	\pause
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\end{frame}

\section{Escape Clause}
\subsection{Escape Clause}
\begin{frame}{Escape Clause with Exogenous Politics}

\pause
When $\ga$ is \textit{only} exogenous (Bagwell $\&$ Staiger 2005):

\pause
\begin{itemize}[<+->]
	\item Simple escape clause: add a second (higher) negotiated weak binding
		\begin{itemize}
			\item Escape clause is designed to allow higher applied tariff when realization of $\ga$ is high
		\end{itemize}
	\item Improves political efficiency
	\item Can improve self-enforcement
	\item Incentive compatibility becomes an issue
\end{itemize}
\end{frame}

\begin{frame}{Incentive compatibility}
\pause
Escape clause is meant to allow higher applied tariff when realized $\ga$ is high
\pause
\begin{itemize}[<+->]
	\item $\ga$ is private information
	\item We want truthful revelation, but truth-telling must be in the best interest of each gov't
	\item Gov't can exploit TOT externality by reporting high $\ga$ even when $\ga$ is low
		\begin{itemize}
			\item Only way to prevent this is with some cost of using escape clause
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Escape Clause with Endogenous Politics}
When $\ga$ is \textit{only} endogenous:
\pause
\begin{itemize}[<+->]
	\item Benefit of escape clause from exogenous case is gone
	\item Assuming lower binding is set to maximize political welfare, escape clause encourages inefficiently high lobbying effort / protection
	\item Incentive compatibility still an issue, but often not the central one
		\begin{itemize}[<+->]
			\item If lobby's preferred tariff $\geq$ escape clause binding, gov't experiences high $\ga$, no need to lie
		\end{itemize}
  \end{itemize}

\pause
\vskip.2in

If $\ga$ is only endogenous, escape clause causes problems, provides no benefits
\end{frame}

\begin{frame}{When the world is more complicated...}
Now suppose political pressure is a result of both endogenous and exogenous forces (i.e. $\ga(s,e)$):
\pause
\begin{itemize}[<+->]
	\item Want escape clause to deal with exogenous shock
	\item But endogenous part $\Rightarrow$ lobbying incentives make it hard to implement escape clause
\end{itemize}

\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Ineffectiveness of Political Criterion for Escape Clause}
  Assume $\ga(s,e) = \ga(s) + \ga(e)$. If an escape clause conditions on $\ga(s,e)$ and $\ga(s^L) < \ga(s^H) < \ga(e^L)$, the lower ``normal'' tariff binding will never be applied.
\end{beamerboxesrounded}
\end{frame}

\begin{frame}{When the world is more complicated... (2)}
\begin{itemize}[<+->]
	\item To make escape clause work, can't use $\ga$
	\begin{itemize}[<+->]
		\item Need signal of shock that is not influenced by endogenous pressure
	\end{itemize}
	\item Can condition directly on $s$
		\begin{itemize}
			\item This seems to be what the WTO actually \textit{does}
		\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{An Escape Clause for Endogenous Politics}
\pause
Assume a WTO-like set up: gov't can choose between $\tau^a$, `escape' tariff $\tau(s)$, or politically-optimal $\tau$ matched to $\ga(s,e)$
\pause
\begin{itemize}[<+->]
	\item Assume $s$ verifiable, so no punishment for $\tau(s)$
	\item Punishment for $\tau(\ga(s,e)) > \tau(s)$
\end{itemize}

\vskip.2in
\pause
Optimal $\tau^a$ may lead to government applying $\tau(\ga(s,e))$
\pause
\begin{itemize}[<+->]
	\item When this happens, it leads to dispute, not valid escape
	\item Otherwise, no extra rent-seeking is encouraged
\end{itemize}

\vskip.2in
\pause
May explain why escape clause has fallen out of use

\end{frame}

%\begin{frame}{Optimal Punishments}
%\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Optimal DSI}
%  In the case of political certainty, if non-trivial cooperation is possible in the presence of a lobby, the optimal punishment scheme is finite when the influence of the lobby on legislative preferences is sufficiently strong $\left(\frac{\partial \gamma}{\partial e} \text{ is sufficiently high} \right)$.
%\end{beamerboxesrounded}
%\end{frame}

\section{Conclusion}

\begin{comment}
\begin{frame}{Ex-Ante Lobbying}
Without ex-ante: gov't anticipates ex-post lobbying, chooses $\tau^{\text{TA}}$ that maximizes objective function given $e^{\text{EP}}$
  \[
	  \max_\tau W(\ga(e^{\text{EP}}),\tau) + W^*(\tau)
	\]
\pause
When ex-ante lobbying ($e^{\text{EA}}$) is possible, we could have
\pause
  \[
	  \max_\tau W(\ga(e^{\text{EA}}),\tau) + W^*(\tau)
	\]
	but the lobby would now have to pay \textit{twice}.
\end{frame}
\end{comment}

\begin{frame}{Future Work}
\pause
\begin{itemize}[<+->]
	\item Application of framework to other design questions
	\item Interactions between $\ga(s)$ and $\ga(e)$
	\item Choice between protective measures over time
\end{itemize}

\end{frame}

\begin{frame}{Conclusion}
Taking into account endogenous political forces alongside exogenous ones...
\pause
\begin{itemize}[<+->]
		\item provides additional general explanation for tariff caps
		\item demonstrates that TAs can be used to discourage lobbing activity in general
		\item helps explain the structure and enforcement of the WTO Safeguards measure
\end{itemize}

\end{frame}

\end{document}

### 2016-06-20 presentation.tex
remove slide on 'new trade theory' role of trade agreements
move 'design' lit review slide ahead of 'role' slides
fill out 'design' lit review slide
minor exposition improvements in tariff caps section

@@
 \author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
-\date{April 6, 2016}
+\date{June 30, 2016}
 \begin{document}
@@
 \pause
-With a few exceptions, TA literature has taken political economy forces to be exogenous. I:
+With a few exceptions, TA design literature has taken political economy forces to be exogenous. I:
 \pause
@@
 \subsection{Role and Design of TAs}
+\begin{frame}{Design of Trade Agreements}
+\pause
+\begin{itemize}[<+->]
+	\item \textbf{Tariff caps}: Bagwell and Staiger 2005, Horn et al 2010, Amador and Bagwell 2012; Beshkar and Bond 2012
+	\item \textbf{Escape clause}: Bagwell and Staiger 2005, Horn et al 2010, 
+	\item \textbf{Shallow vs. deep integration}: Bagwell and Staiger 2001, DeRemer 2014
+	\item \textbf{Dispute settlement}: Maggi 1999, Ludema 2001, Maggi and Staiger 2011/2013, Klimenko et al 2008
+	\item \textbf{Property vs. liability rules}: Pauwelyn 2008, Beshkar 2010, Maggi and Staiger 2014
+	\item \textbf{Retaliation}: Bown 2002/2004, Beshkar 2010
+\end{itemize}
+\end{frame}
+
 \begin{frame}{Role of Trade Agreements: TOT Externality}
@@
 
+\begin{comment} 
 \begin{frame}{Role of Trade Agreements: New Trade Theory}
@@
 \end{frame} 
+\end{comment}
 
@@
 		\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Design of Trade Agreements}
-\pause
-\begin{itemize}[<+->]
-	\item \textbf{Tariff caps}: Bagwell and Staiger 2005, Horn et al 2010, Amador and Bagwell 2012; Beshkar and Bond 2012
-	\item \textbf{Escape clause}: Bagwell and Staiger 2005, Horn et al 2010, 
-	\item \textbf{Shallow vs. deep integration}: Bagwell and Staiger 2001, DeRemer 2014
-	\item \textbf{Dispute settlement}: Maggi 1999, Ludema 2001, Maggi and Staiger 2011/2013, Klimenko et al 2008
-	%\item \textbf{Property vs. liability rules}: Pauwelyn 2008, Beshkar 2010, Maggi and Staiger 2014
-	\item \textbf{Retaliation}: Bown 2002/2004, Beshkar 2010
 \end{itemize}
@@
 	\item Will TA be used to discourage lobbying? Depends on how gov't welfare varies in $\ga$
-	\item With standard Baldwin-style objective function, welfare always increases with $\ga$:
+	\item With standard Baldwin-style objective function, welfare always increases with $\ga$\spaceskip1pt
 \[
-  W = \mathit{CS}_X(\tau) + \ga \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
+  W = \mathit{CS}_X(\tau) + \ga \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
 \]
+
 	\begin{itemize}
@@
 
-	\item If we instead make weights sum to 1, welfare is no longer monotonic in $\ga$
+	\item If weights must sum to 1, welfare no longer monotonic in $\ga$
 \end{itemize}
@@
 	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): if governments patient enough ($\de$ high enough), optimal externally-enforced weak binding can be self-enforced
-  \item $\ga$ endogenous: not necessarily true
+  \item $\ga$ endogenous: optimal externally-enforced weak binding may not be self-enforcing
 		\begin{itemize}

### 2016-06-20 presentation.tex
more exposition improvements

@@
 	\item Assuming lower binding is set to maximize political welfare, escape clause encourages inefficiently high lobbying effort / protection
-	\item Incentive compatibility still an issue, but often not the central one
+	\item (Incentive compatibility still an issue, but often not the central one)
 		\begin{itemize}[<+->]
-			\item If lobby's preferred tariff $\geq$ escape clause binding, gov't experiences high $\ga$, no need to lie
+			\item (If lobby's preferred tariff $\geq$ escape clause binding, gov't experiences high $\ga$, no need to lie)
 		\end{itemize}
@@
 
-\begin{frame}{When the world is more complicated... (2)}
+\begin{frame}{When the world is more complicated... (con't)}
 \begin{itemize}[<+->]
@@
 \pause
-Optimal $\tau^a$ may lead to government applying $\tau(\ga(s,e))$
+Optimal $\tau^a$ may lead government to apply $\tau(\ga(s,e))$
 \pause

### 2016-07-18 feedback.tex
add comments from SITE presentation

@@
 \begin{document}
+\begin{center}
+Feedback on Apr. 4 draft
+\end{center}
+
+From SITE presentation
+\begin{itemize}
+	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with contributions 'felonious'
+	\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
+	\item $\ga(e)$
+		\begin{itemize}
+			\item Giovanni: think of expost $\ga(e)$ as effort to elect/choose ex-post decision-maker
+			\item At worst, I can use leg / exec split to deal with this criticism
+			\item Kyle (???): advertising paper (???)
+			\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
+				\begin{itemize}
+					\item Giovanni: if I can show properties survive  monotonic transformation, I'm fine
+					\item They don't like gov't choosing its own preferences. May just need to change wording
+				\end{itemize}
+			\item Bob: what is fundamental structure of preferences?
+			\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
+		\end{itemize}
+	\item Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies
+		\begin{itemize}
+			\item Maybe I should just get rid of tariff cap result altogether
+		\end{itemize}
+	\item David DeRemer: Safeguards have dropped off in use around the world
+		\begin{itemize}
+			\item Look at U.S. vs. ROW (he sent papers)
+		\end{itemize}
+	\item I think that I need to bring people along more carefully from GH/MRC to what I'm doing
+	\item Bob and Kyle: take my time, make it as good as possible. Better to have quality.
+\end{itemize}
+
+\newpage
 \begin{center}

### 2016-07-19 feedback.tex
add SITE presentation notes and section headings/bookmarks

@@
 
+\usepackage[pdftex,
+bookmarks=true,
+bookmarksnumbered=false,
+pdfview=fitH,
+bookmarksopen=true,hyperfootnotes=false]{hyperref}
+
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 \begin{document}
-\begin{center}
-Feedback on Apr. 4 draft
-\end{center}
+\section{Feedback on Apr. 4, 2016 draft}
 
-From SITE presentation
+\subsection{Rochester Seminar, April 6, 2016 (mostly Mark Bils)}
+\begin{itemize}
+	\item Need to say that I'm ruling out side payments so it's clear that there can't just be a menu of punishments that exactly keeps each government in line
+		\begin{itemize}
+			\item Risk (average) vs. incentives (marginal) impossibility result --- but with no money transfers, even this breaks down
+			\item Can I think of this as an insurance problem between the governments?
+		\end{itemize}
+	\item Show what the optimum is, then why they can't get it, how they have to try to mimic it with these instruments (tariff caps, escape clause, these types of punishments, etc.)
+	\item Going from $\ga(s)$ to $\ga(e)$ --- essentially taking away the structure of the $s$'s was confusing b/c the policy space doesn't make any sense
+\end{itemize}
+
+\subsection{From SITE 2016 presentation}
 \begin{itemize}
@@
 \newpage
-\begin{center}
-Feedback on Jan. 10 draft
-\end{center}
+\section{Feedback on Jan. 10, 2015 draft}
 
-Chad Bown, Jan. 22, 2015
+\subsection{Chad Bown, Jan. 22, 2015}
 \begin{itemize}
@@
 \newpage
-Ben Zissimos, May 11, 2015
+\subsection{Ben Zissimos, May 11, 2015}
 
@@
 \newpage
-Maurizio Zanardi, June 22, 2015
+\subsection{Maurizio Zanardi, June 22, 2015}
 \begin{itemize}
@@
 \newpage
-Rochester Seminar, April 6, 2016 (Mostly Mark Bils)
+\section{Presentation Notes}
+For SITE 2016
 \begin{itemize}
-	\item Need to say that I'm ruling out side payments so it's clear that there can't just be a menu of punishments that exactly keeps each government in line
-		\begin{itemize}
-			\item Risk (average) vs. incentives (marginal) impossibility result --- but with no money transfers, even this breaks down
-			\item Can I think of this as an insurance problem between the governments?
-		\end{itemize}
-	\item Show what the optimum is, then why they can't get it, how they have to try to mimic it with these instruments (tariff caps, escape clause, these types of punishments, etc.)
-	\item Going from $\ga(s)$ to $\ga(e)$ --- essentially taking away the structure of the $s$'s was confusing b/c the policy space doesn't make any sense
+	\item Ethier 2012 for protection decreasing function of $\tau$ [NEED TO REVISIT IN TEXT]. ``The Political-Support Approach to Protection,'' Global Journal of Economics
+	\item $\ga(s)$ comes from MS 2011 QJE, ``The Role of Dispute Settlement Procedures in Trade Agreements''
+	\item BS approach from 2002 book originates from 1999 AER ``An Economic Theory of GATT''
+	\item Add title, axis to graph
+	\item Review DSB literature (slide 13)
+	\item Review my ex-ante results
 \end{itemize}

### 2016-07-20 feedback.tex
add notes from 2014 spring Midwest mtgs

@@
 	\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
+		\begin{itemize}
+			\item Rohan: $(1-a)\cdot C + a \cdot W \equiv C + a \cdot (W-C)$
+			\item When weights on $\pi_X$ are not linear, what are the implications for resource waste? Is it the same as $+ C / - C$?
+		\end{itemize}
 	\item $\ga(e)$
@@
 \newpage
+\section{Feedback on 2014 draft}
+\subsection{May 30, 2014, Spring Midwest meetings}
+\begin{itemize}
+	\item Rick Bond, Ian Wooton, James Lake, Mostafa, David Kuenzel at the talk
+	\item James: need to be more precise about when $\ga$ is evaluated
+\end{itemize}
+
+\newpage
 \section{Presentation Notes}
@@
 	\item Ethier 2012 for protection decreasing function of $\tau$ [NEED TO REVISIT IN TEXT]. ``The Political-Support Approach to Protection,'' Global Journal of Economics
+		\begin{itemize}
+			\item Making $W$ quadratic in $\ga$ such as $\ga - \ga^2$ gives a nice interior max
+		\end{itemize}
 	\item $\ga(s)$ comes from MS 2011 QJE, ``The Role of Dispute Settlement Procedures in Trade Agreements''

### 2016-07-20 feedback.tex
add feedback from 2016 PEIO presentation

@@
 \newpage
+\subsection{January 2016, PEIO Salt Lake City}
+\begin{itemize}
+	\item Elias Bengtsson: more on policy implications, who's doing what to whom
+	\item Cristiane Carneiro
+		\begin{itemize}
+			\item Flexible instruments
+			\item Agreement duration, exit (legal scholarship)
+			\item Look at Summer 2001 special edition of IO
+			\item Korgmenos 2013 JCR
+			\item Korgmenos and Nav 2011
+			\item Coil
+			\item Helper 2005, Virginia Law Review
+		\end{itemize}
+	\item Renee Bowen: what's the contribution relative to GH; her paper, Peter's, McLaren (but she was a bit confused about what I was doing we figured out later)
+	\item Competing lobbies--exporters
+	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover
+\end{itemize}
+
+\newpage
 \section{Feedback on 2014 draft}

### 2016-07-21 feedback.tex
add feedback from 2015 CEAs
add devashish's feedback after SITE

@@
 			\item When weights on $\pi_X$ are not linear, what are the implications for resource waste? Is it the same as $+ C / - C$?
+			\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
+				\begin{itemize}
+					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's
+					\item Find something in persuasion literature
+				\end{itemize}
 		\end{itemize}
@@
 		\end{itemize}
-	\item Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies
+	\item Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc.
 		\begin{itemize}
@@
 \newpage
+\subsection {Summer 2015 CEA}
+\begin{itemize}
+	\item Richard: what happens if $\ga(e^L) < \ga(s^H)$?
+	\item Dev: Reputation: do they use safeguards less b/c worried about reputation?
+	\item Chris: What is an exogenous shock to political pressure? What signal would be correlated?
+\end{itemize}
+
+\vskip.5in
 \subsection{January 2016, PEIO Salt Lake City}

### 2016-07-21 feedback.tex
add feedback from talks in 2014

@@
 	\item Rick Bond, Ian Wooton, James Lake, Mostafa, David Kuenzel at the talk
+	\item Rick: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$
 	\item James: need to be more precise about when $\ga$ is evaluated
+		\begin{itemize}
+			\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
+		\end{itemize}
+	\item Ex-ante lobbying: need to understand MRC07 result that tariff-cap = strong binding
+\end{itemize}
+
+\subsection{Sept. 12, 2014, ETSG}
+\begin{itemize}
+	\item Paola: How do I want to sell the paper?
+		\begin{itemize}
+			\item In relation to what literature?
+				\begin{itemize}
+					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
+					\item What literature \textit{doesn't}?
+				\end{itemize}
+			\item Baldwin-style welfare function works for a set of questions, but not for this one
+			\item GH is isomorphic
+		\end{itemize}
+	\item Coates Ludema 2001
+		\begin{itemize}
+			\item What if WTO actually \textit{did} what literature says it does? It wouldn't work!
+			\item Be more explicit about the relationship between $\ga(e)$ and $\ga(s)$
+		\end{itemize}
+	\item I need to be able to say which questions are okay with
+		\begin{enumerate}
+			\item unweighted/standard welfare function
+			\item $\ga(s)$ only (i.e. predictions/design predictions won't change lobbying incentives) 
+		\end{enumerate}
+\end{itemize}
+
+\subsection{Oct. 17, 2014, McGill}
+\begin{itemize}
+	\item Hanson and Sargent, model uncertainty
+	\item Why Nash bargain for tariff cap?
+\end{itemize}
+
+\subsection{Nov. 7, 2014, Ryerson}
+\begin{itemize}
+	\item Export lobby at ex-post b/c wants to reduce chance of trade war
+	\item $e$ in terms of numeraire
+	\item Could escape clause be made to work with some fancier set up, e.g. dynamic use constraint?
 \end{itemize}

### 2016-07-21 to-do.tex
add big picture table to to-do list
add dynamic preferences references

@@
 	\item Define $\ga$ before using it
+	\item Deal with ``dynamic preferences'' issue
+		\begin{itemize}
+			\item References from Krishna and Sadowski (2014) ECTA
+				\begin{itemize}
+					\item DILLENBERGER, D., J. S. LLERAS, P. SADOWSKI, AND N. TAKEOKA (2013): “A Theory of Subjective Learning,” Working Paper. 
+					\item GUL, F., AND W. PESENDORFER (2004): “Self-Control and the Theory of Consumption,” Econometrica, 72 (1), 119–158
+					\item JONES, R. A., AND J. M. OSTROY (1984): “Flexibility and Uncertainty,” Review of Economic Studies, 51 (1), 13–32
+					\item KOOPMANS, T. C. (1964): “On the Flexibility of Future Preferences,” in Human Judgments and Optimality, ed. by M. W. Shelly and G. L. Bryan. New York: Wiley.
+					\item KREPS, D. M., AND E. L. PORTEUS (1978): “Temporal Resolution of Uncertainty and Dynamic Choice Theory,” Econometrica, 46 (1), 185–200.
+				\end{itemize}
+		\end{itemize}
 	\item Take out repeats, especially of references to BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
@@
 	\item Perhaps make a bigger deal of the idea of $\ga$ decreasing function of $\tau$---move out of footnote
+	\item Big picture of which changes do what \\
+		\begin{tabular}{|l|c|c|}
+			 & rigid & escape \\
+			BS2005  & on-schedule (truthtelling): trivial; & need cost  \\
+			 & off (repeated): optimal static if patient enough & \\
+			$\ga(e)$ & on: trivial& no need, but revisit side payments\\
+			 & off: mabye not b/c of lobby & \\
+			$\ga(e,s)$ & ignore for now & is $s$ verifiable? def not w/$\ga(s,e)$
+		\end{tabular}
 \end{enumerate}

### 2016-07-22 to-do.tex
turn to-do items into sections

--- 2016-03-31 to-do.tex
@@
 March 30, 2016
-\begin{enumerate}
-	\item Motivation
+
+\section{Motivation}
 		\begin{itemize}
@@
 \newpage
-	\item More on escape clause / WEC
+\section{More on escape clause / WEC}
 		\begin{itemize}
@@
 				\begin{itemize}
-					\item ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
+					\item ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), %\url{C:\Users\Kristy\Dropbox\Research\PEshocks\Literature\2016}
 					\item Need to understand better what changed, both with safeguard and with other policy instruments
@@
 		\end{itemize}
-	\item Costly state verification
+\section{Costly state verification}
 		\begin{itemize}
@@
 \newpage
-	\item BS2005 vs. MRC2007
+\section{BS2005 vs. MRC2007}
 		\begin{itemize}
@@
 		\end{itemize}
-	\item Structured differently from most papers
+\section{Structured differently from most papers}
 		\begin{itemize}
@@
 		\end{itemize}
-		\item One good idea / do too much
+\section{One good idea / do too much}
 		\begin{itemize}
@@
 		\end{itemize}
-	\item Define $\ga$ before using it
-	\item Take out repeats, especially of references to BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
+\section{Define $\ga$ before using it}
+\section{Deal with ``dynamic preferences'' issue}
+		\begin{itemize}
+			\item References from Krishna and Sadowski (2014) ECTA
+				\begin{itemize}
+					\item DILLENBERGER, D., J. S. LLERAS, P. SADOWSKI, AND N. TAKEOKA (2013): ``A Theory of Subjective Learning,'' Working Paper. 
+					\item GUL, F., AND W. PESENDORFER (2004): ``Self-Control and the Theory of Consumption,'' Econometrica, 72 (1), 119-158
+					\item JONES, R. A., AND J. M. OSTROY (1984): ``Flexibility and Uncertainty,'' Review of Economic Studies, 51 (1), 13-32
+					\item KOOPMANS, T. C. (1964): ``On the Flexibility of Future Preferences,'' in Human Judgments and Optimality, ed. by M. W. Shelly and G. L. Bryan. New York: Wiley.
+					\item KREPS, D. M., AND E. L. PORTEUS (1978): ``Temporal Resolution of Uncertainty and Dynamic Choice Theory,'' Econometrica, 46 (1), 185-200.
+				\end{itemize}
+		\end{itemize}
+\section{Take out repeats}
+	Especially of references to BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
 Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point.
-	\item Scrub unnecessary references to weak bindings (since strong are now mostly gone)
-	\item Perhaps make a bigger deal of the idea of $\ga$ decreasing function of $\tau$---move out of footnote
-\end{enumerate}
+
+\section{Make a bigger deal of the idea of $\ga$ decreasing function of $\tau$---move out of footnote}
+\section{Big picture of which changes do what}
+		\begin{tabular}{|l|c|c|}
+			 & rigid & escape \\
+			BS2005  & on-schedule (truthtelling): trivial; & need cost  \\
+			 & off (repeated): optimal static if patient enough & \\
+			$\ga(e)$ & on: trivial& no need, but revisit side payments\\
+			 & off: mabye not b/c of lobby & \\
+			$\ga(e,s)$ & ignore for now & is $s$ verifiable? def not w/$\ga(s,e)$
+		\end{tabular}
 

### 2016-07-22 to-do.tex
integrate old written to-do list into electronic one

@@
 			\item You need to think a bit more about the objective(s) of the paper and how best to achieve them.
+			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics
 		\end{itemize}
@@
 		\end{itemize}
-\section{Define $\ga$ before using it}
+
+\section{Misc.}
+\begin{itemize}
+	\item Define $\ga$ before using it
+	\item If add back in assumption 1, add Ethier footnote
+	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
+\end{itemize}
+
 \section{Deal with ``dynamic preferences'' issue}
@@
 \section{Big picture of which changes do what}
+\vskip.2in
 		\begin{tabular}{|l|c|c|}
@@
 		\end{tabular}
+\vskip.2in
+\begin{itemize}
+	\item Make sure this story is clear in text: I establish baseline tariff cap case, then add escape clause. We can already see a story of $\ldots$ emerge...
+	\item Need to be clear when it is flexibility, enforcement, endogenous $\ga$ that screws things up. THEN need to bmake sure it comes out in paper
+	\item May want to ditch repeated game part...
+\end{itemize}
+
+\section{Lit review}
+\begin{itemize}
+	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
+		\begin{itemize}
+			\item Bagwell and Staiger 1999 AER
+		\end{itemize}
+	\item Have to make connection to endogenous papers clearer
+	\item Review Bown, Bagwell $\&$ Staiger
+	\item Do I include papers like BS2001 (non-repeated) in lit review?
+\end{itemize}
+
+\section{Clarify information structure}
+
+\section{Add empirics / evidence?}
+
+\section{Existence proofs?}
+
+\section{Lobby's optimal effort}
+Need to be clear about when lobby's optimal $\tau$ is higher than gov'ts: that is, when TA would want to rein in lobby and escape clause would ruin that
+
+\section{Dynamic use constraint}
+When would lobby exert effort to top up?
+\begin{itemize}
+	\item Do I want correlation between endogenous/exogenous parts? 
+\end{itemize}
 

### 2016-07-23 to-do.tex
add questions re:Limao & Tovar to to-do

@@
 	\item Do I include papers like BS2001 (non-repeated) in lit review?
+	\item Limao and Tovar 2011
+		\begin{enumerate}
+			\item Their 2nd (and final) stage is Nash bargaining. Is there an isomorphism to my setup?
+			\item It IS clear to take their first, limited case of commitment to tariff cap only (ignore NTB) as analogy for gov't in my model choosing commitment to different tariff level.
+Need to answer Q1 above in order to answer points 3 and 4:
+			\item Their gains from commitment derive from improvement in bargaining power vis-a-vis the lobby (Schelling conjecture invoked). How does this relate to the gain from commitment in my model?
+			\item Top of page 6: ``large enough bargaining power.'' Can I make connection to $\ga(e)$?
+			\item Note: in their equation 3, $C$ enters expression for NBS's ($\tau$) in equilibrium, it doesn't disappear as in GH
+		\end{enumerate}
 \end{itemize}

### 2016-07-25 to-do.tex
fix math in bookmarks issue in to-do.tex

@@
 
-\section{Make a bigger deal of the idea of $\ga$ decreasing function of $\tau$---move out of footnote}
+\section{Make a bigger deal of the idea of \texorpdfstring{$\gamma$}{gamma} decreasing function of tariffs---move out of footnote}
+Ethier 2012 for protection decreasing function of $\tau$ [NEED TO REVISIT IN TEXT]. ``The Political-Support Approach to Protection,'' Global Journal of Economics
+	\begin{itemize}
+		\item Making $W$ quadratic in $\ga$ such as $\ga - \ga^2$ gives a nice interior max
+	\end{itemize}
+
 \section{Big picture of which changes do what}

### 2016-07-25 feedback.tex
integrate Ryerson feedback into to-do
add ETSG 2014 and McGill feedback to to-do

@@
 \begin{itemize}
-	\item Paola: How do I want to sell the paper?
+	\item Paola: How do I want to sell the paper? [on to-do list]
 		\begin{itemize}
@@
 				\end{itemize}
-			\item Baldwin-style welfare function works for a set of questions, but not for this one
-			\item GH is isomorphic
+			\item Baldwin-style welfare function works for a set of questions, but not for this one [DONE]
+			\item GH is isomorphic [on to-do list]
 		\end{itemize}
-	\item Coates Ludema 2001
+	\item Coates Ludema 2001 [on to-do list]
 		\begin{itemize}
-			\item What if WTO actually \textit{did} what literature says it does? It wouldn't work!
-			\item Be more explicit about the relationship between $\ga(e)$ and $\ga(s)$
+			\item What if WTO actually \textit{did} what literature says it does? It wouldn't work! [done]
+			\item Be more explicit about the relationship between $\ga(e)$ and $\ga(s)$ [done]
 		\end{itemize}
-	\item I need to be able to say which questions are okay with
+	\item I need to be able to say which questions are okay with [done]
 		\begin{enumerate}
@@
 \begin{itemize}
-	\item Hanson and Sargent, model uncertainty
-	\item Why Nash bargain for tariff cap?
+	\item Hanson and Sargent, model uncertainty [doesn't seem that relevant]
+	\item Why Nash bargain for tariff cap? [why not?]
 \end{itemize}
@@
 \begin{itemize}
-	\item Export lobby at ex-post b/c wants to reduce chance of trade war
-	\item $e$ in terms of numeraire
-	\item Could escape clause be made to work with some fancier set up, e.g. dynamic use constraint?
+	\item Export lobby at ex-post b/c wants to reduce chance of trade war [outside the scope of this paper]
+	\item $e$ in terms of numeraire [on to-do list]
+	\item Could escape clause be made to work with some fancier set up, e.g. dynamic use constraint? [on to-do list]
 \end{itemize}

### 2016-07-25 feedback.tex
add PEIO, CEA and Midwest 2014 feedback to to-do

@@
 \begin{itemize}
-	\item Richard: what happens if $\ga(e^L) < \ga(s^H)$?
-	\item Dev: Reputation: do they use safeguards less b/c worried about reputation?
-	\item Chris: What is an exogenous shock to political pressure? What signal would be correlated?
+	\item Richard: what happens if $\ga(e^L) < \ga(s^H)$? [done]
+	\item Dev: Reputation: do they use safeguards less b/c worried about reputation? [outside the scope]
+	\item Chris: What is an exogenous shock to political pressure? What signal would be  correlated? [don't think I need to go there]
 \end{itemize}
@@
 \begin{itemize}
-	\item Elias Bengtsson: more on policy implications, who's doing what to whom
+	\item Elias Bengtsson: more on policy implications, who's doing what to whom [done]
 	\item Cristiane Carneiro
@@
 		\end{itemize}
-	\item Renee Bowen: what's the contribution relative to GH; her paper, Peter's, McLaren (but she was a bit confused about what I was doing we figured out later)
-	\item Competing lobbies--exporters
-	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover
+	\item Renee Bowen: what's the contribution relative to GH; her paper, Peter's, McLaren (but she was a bit confused about what I was doing we figured out later) [done / don't need to do]
+	\item Competing lobbies--exporters [outside scope]
+	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover [on to-do list]
 \end{itemize}
@@
 	\item Rick Bond, Ian Wooton, James Lake, Mostafa, David Kuenzel at the talk
-	\item Rick: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$
-	\item James: need to be more precise about when $\ga$ is evaluated
+	\item Rick: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$ [on to-do list]
+	\item James: need to be more precise about when $\ga$ is evaluated [on to-do list]
 		\begin{itemize}
@@
 		\end{itemize}
-	\item Ex-ante lobbying: need to understand MRC07 result that tariff-cap = strong binding
+	\item Ex-ante lobbying: need to understand MRC07 result that tariff-cap = strong binding [been there, done that, got the t-shirt, gave it back]
 \end{itemize}

### 2016-07-25 to-do.tex
integrate Ryerson feedback into to-do
add ETSG 2014 and McGill feedback to to-do
add PEIO, CEA and Midwest 2014 feedback to to-do

@@
 	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
+	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
 \end{itemize}
@@
 	\begin{itemize}
+		\item James: need to be more precise about when $\ga$ is evaluated
+			\begin{itemize}
+				\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
+			\end{itemize}
 		\item References from Krishna and Sadowski (2014) ECTA
@@
 	\item Have to make connection to endogenous papers clearer
+		\begin{itemize}
+			\item and make clear that GH is isomorphic / provides microfoundations for Baldwin objective function
+			\item More on Coates $\&$ Ludema (2001)
+		\end{itemize}
+	\item Paola: How do I want to sell the paper?
+		\begin{itemize}
+			\item In relation to what literature?
+				\begin{itemize}
+					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
+					\item What literature \textit{doesn't}?
+				\end{itemize}
+			\end{itemize}
 	\item Review Bown, Bagwell $\&$ Staiger
 	\item Do I include papers like BS2001 (non-repeated) in lit review?
+	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
 	\item Limao and Tovar 2011
@@
 
+\section{Numeraire}
+Do I need to add a numeraire so I can put $e$ in terms of it?
+
+\section{Lobbying cost included in gov't welfare}
+	\begin{itemize}
+		\item Rick Bond: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$
+	\end{itemize}
 \end{document}

### 2016-07-25 feedback.tex
complete integration of feedback into to-do

@@
 \begin{itemize}
-	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with contributions 'felonious'
-	\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
+	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with contributions 'felonious' [in to-do list]
+	\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not [in to-do list]
 		\begin{itemize}
@@
 		\end{itemize}
-	\item $\ga(e)$
+	\item $\ga(e)$ [in to-do list]
 		\begin{itemize}
@@
 		\end{itemize}
-	\item Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc.
+	\item Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc. [in to-do list]
 		\begin{itemize}
@@
 		\end{itemize}
-	\item David DeRemer: Safeguards have dropped off in use around the world
+	\item David DeRemer: Safeguards have dropped off in use around the world [in to-do list]
 		\begin{itemize}

### 2016-07-25 to-do.tex
move one point from Ben's feedback into to-do
complete integration of feedback into to-do

@@
 	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
+	\item Ben: I'm wondering how you view the time-frame over which your model is set.
+On one hand a specific-factors set-up is normally associated with the short-run.
+But on the other hand you have an infinite time horizon which suggests very long run.
+	\item Clarify information structure
+
 \end{itemize}
@@
 				\end{itemize}
+			\item Giovanni: think of expost $\ga(e)$ as effort to elect/choose ex-post decision-maker
+			\item At worst, I can use leg / exec split to deal with this criticism
+			\item Kyle (???): advertising paper (???)
+			\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
+				\begin{itemize}
+					\item Giovanni: if I can show properties survive  monotonic transformation, I'm fine
+					\item They don't like gov't choosing its own preferences. May just need to change wording
+				\end{itemize}
+			\item Bob: what is fundamental structure of preferences?
+			\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
+		\end{itemize}
+	Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc.
+	\begin{itemize}
+			\item Maybe I should just get rid of tariff cap result altogether [me]
 		\end{itemize}
@@
 		\end{itemize}
+	\item I think that I need to bring people along more carefully from GH/MRC to what I'm doing
 	\item Paola: How do I want to sell the paper?
@@
 	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
+	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
+	 contributions 'felonious'
+	\item David DeRemer: Safeguards have dropped off in use around the world
+		\begin{itemize}
+			\item Look at U.S. vs. ROW (he sent papers)
+		\end{itemize}
 	\item Limao and Tovar 2011
@@
 \end{itemize}
-
-\section{Clarify information structure}
 
@@
 		\item Rick Bond: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$
+		\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
+		\begin{itemize}
+			\item Rohan: $(1-a)\cdot C + a \cdot W \equiv C + a \cdot (W-C)$
+			\item When weights on $\pi_X$ are not linear, what are the implications for resource waste? Is it the same as $+ C / - C$?
+			\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
+				\begin{itemize}
+					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's
+					\item Find something in persuasion literature
+				\end{itemize}
+		\end{itemize}
 	\end{itemize}

### 2016-07-27 to-do.tex
reorganize to-do list
start work on dynamic preferences

@@
 
-March 30, 2016
-
-\section{Motivation}
+\section{``Dynamic preferences''}
 	\begin{itemize}
-			\item What it's missing: intersection with reality
+		\item James: need to be more precise about when $\ga$ is evaluated
 			\begin{itemize}
-					\item Anecdotes
-					\item How policy is used
+				\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
+			\end{itemize}
+		\item References from Krishna and Sadowski (2014) ECTA
 				\begin{itemize}
-							\item I thought that escape clause implications \textit{were} this
+					\item DILLENBERGER, D., J. S. LLERAS, P. SADOWSKI, AND N. TAKEOKA (2013): ``A Theory of Subjective Learning,'' Working Paper. 
+					\item GUL, F., AND W. PESENDORFER (2004): ``Self-Control and the Theory of Consumption,'' Econometrica, 72 (1), 119-158
+					\item \sout{JONES, R. A., AND J. M. OSTROY (1984): ``Flexibility and Uncertainty,'' Review of Economic Studies, 51 (1), 13-32}
+					\item (couldn't find) KOOPMANS, T. C. (1964): ``On the Flexibility of Future Preferences,'' in Human Judgments and Optimality, ed. by M. W. Shelly and G. L. Bryan. New York: Wiley.
+					\item \sout{KREPS, D. M., AND E. L. PORTEUS (1978): ``Temporal Resolution of Uncertainty and Dynamic Choice Theory,'' Econometrica, 46 (1), 185-200.}
 				\end{itemize}
-					\item Something like how GH predicts export taxes but it doesn't work that way in the real world
-					\item Well articulated question / paradox
+			\item Giovanni: think of expost $\ga(e)$ as effort to elect/choose ex-post decision-maker
+			\item At worst, I can use leg / exec split to deal with this criticism
+			\item Kyle (???): advertising paper (???)
+			\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
+				\begin{itemize}
+					\item Giovanni: if I can show properties survive  monotonic transformation, I'm fine
+					\item They don't like gov't choosing its own preferences. May just need to change wording
 				\end{itemize}
-			\item Look at ``Theory of Managed Trade'' model motivation (BS)
-			\item Intro doesn't reflect body of paper
-			\item What kinds of empirical questions might we ask with this model?
-			\item What would be different because we use this model?
-			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper.
-However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government.
-Is this aspect crucial for the point you want to make?
-If yes, it should be integrate with the previous object from the beginning.
-If not, it should be postponed a bit.
-			\item I also find the introduction too long.
-You could/should split the literature review that begins on page 4 to a separate section.
-And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it.
-Again, your main point seems to get lost into the discussion instead of standing out.
-			\item You need to think a bit more about the objective(s) of the paper and how best to achieve them.
-			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics
+			\item Bob: what is fundamental structure of preferences?
+			\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
+			\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
+				\begin{itemize}
+					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's
+					\item Find something in persuasion literature
+				\end{itemize}
+		\end{itemize}
+	Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc.
+	\begin{itemize}
+			\item Maybe I should just get rid of tariff cap result altogether [me]
 		\end{itemize}
 
-\newpage
+\section{Lobbying cost included in gov't welfare}
+	\begin{itemize}
+		\item Rick Bond: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$
+		\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
+		\begin{itemize}
+			\item Rohan: $(1-a)\cdot C + a \cdot W \equiv C + a \cdot (W-C)$
+			\item When weights on $\pi_X$ are not linear, what are the implications for resource waste? Is it the same as $+ C / - C$?
+		\end{itemize}
+	\end{itemize}
+
+\section{Numeraire}
+Do I need to add a numeraire so I can put $e$ in terms of it?
+
+\section{Lit review}
+\begin{itemize}
+	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
+		\begin{itemize}
+			\item Bagwell and Staiger 1999 AER
+		\end{itemize}
+	\item Have to make connection to endogenous papers clearer
+		\begin{itemize}
+			\item and make clear that GH is isomorphic / provides microfoundations for Baldwin objective function
+			\item More on Coates $\&$ Ludema (2001)
+		\end{itemize}
+	\item I think that I need to bring people along more carefully from GH/MRC to what I'm doing
+	\item Paola: How do I want to sell the paper?
+		\begin{itemize}
+			\item In relation to what literature?
+				\begin{itemize}
+					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
+					\item What literature \textit{doesn't}?
+				\end{itemize}
+			\end{itemize}
+	\item Review Bown, Bagwell $\&$ Staiger
+	\item Do I include papers like BS2001 (non-repeated) in lit review?
+	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
+	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
+	 contributions 'felonious'
+	\item David DeRemer: Safeguards have dropped off in use around the world
+		\begin{itemize}
+			\item Look at U.S. vs. ROW (he sent papers)
+		\end{itemize}
+	\item Limao and Tovar 2011
+		\begin{enumerate}
+			\item Their 2nd (and final) stage is Nash bargaining. Is there an isomorphism to my setup?
+			\item It IS clear to take their first, limited case of commitment to tariff cap only (ignore NTB) as analogy for gov't in my model choosing commitment to different tariff level.
+Need to answer Q1 above in order to answer points 3 and 4:
+			\item Their gains from commitment derive from improvement in bargaining power vis-a-vis the lobby (Schelling conjecture invoked). How does this relate to the gain from commitment in my model?
+			\item Top of page 6: ``large enough bargaining power.'' Can I make connection to $\ga(e)$?
+			\item Note: in their equation 3, $C$ enters expression for NBS's ($\tau$) in equilibrium, it doesn't disappear as in GH
+		\end{enumerate}
+\end{itemize}
+
 \section{More on escape clause / WEC}
@@
 		\end{itemize}
+
 \section{Costly state verification}
@@
 
-\newpage
+\section{Dynamic use constraint}
+When would lobby exert effort to top up?
+\begin{itemize}
+	\item Do I want correlation between endogenous/exogenous parts? 
+\end{itemize}
+
+\section{Motivation}
+		\begin{itemize}
+			\item What it's missing: intersection with reality
+				\begin{itemize}
+					\item Anecdotes
+					\item How policy is used
+						\begin{itemize}
+							\item I thought that escape clause implications \textit{were} this
+						\end{itemize}
+					\item Something like how GH predicts export taxes but it doesn't work that way in the real world
+					\item Well articulated question / paradox
+				\end{itemize}
+			\item Look at ``Theory of Managed Trade'' model motivation (BS)
+			\item Intro doesn't reflect body of paper
+			\item What kinds of empirical questions might we ask with this model?
+			\item What would be different because we use this model?
+			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper.
+However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government.
+Is this aspect crucial for the point you want to make?
+If yes, it should be integrate with the previous object from the beginning.
+If not, it should be postponed a bit.
+			\item I also find the introduction too long.
+You could/should split the literature review that begins on page 4 to a separate section.
+And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it.
+Again, your main point seems to get lost into the discussion instead of standing out.
+			\item You need to think a bit more about the objective(s) of the paper and how best to achieve them.
+			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics
+		\end{itemize}
+
 \section{BS2005 vs. MRC2007}
@@
 		\end{itemize}
+
 \section{Structured differently from most papers}
@@
 		\end{itemize}
+
 \section{One good idea / do too much}
@@
 
-\section{Misc.}
-\begin{itemize}
-	\item Define $\ga$ before using it
-	\item If add back in assumption 1, add Ethier footnote
-	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
-	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
-	\item Ben: I'm wondering how you view the time-frame over which your model is set.
-On one hand a specific-factors set-up is normally associated with the short-run.
-But on the other hand you have an infinite time horizon which suggests very long run.
-	\item Clarify information structure
-
-\end{itemize}
-
-\section{Deal with ``dynamic preferences'' issue}
-	\begin{itemize}
-		\item James: need to be more precise about when $\ga$ is evaluated
-			\begin{itemize}
-				\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
-			\end{itemize}
-		\item References from Krishna and Sadowski (2014) ECTA
-				\begin{itemize}
-					\item DILLENBERGER, D., J. S. LLERAS, P. SADOWSKI, AND N. TAKEOKA (2013): ``A Theory of Subjective Learning,'' Working Paper. 
-					\item GUL, F., AND W. PESENDORFER (2004): ``Self-Control and the Theory of Consumption,'' Econometrica, 72 (1), 119-158
-					\item JONES, R. A., AND J. M. OSTROY (1984): ``Flexibility and Uncertainty,'' Review of Economic Studies, 51 (1), 13-32
-					\item KOOPMANS, T. C. (1964): ``On the Flexibility of Future Preferences,'' in Human Judgments and Optimality, ed. by M. W. Shelly and G. L. Bryan. New York: Wiley.
-					\item KREPS, D. M., AND E. L. PORTEUS (1978): ``Temporal Resolution of Uncertainty and Dynamic Choice Theory,'' Econometrica, 46 (1), 185-200.
-				\end{itemize}
-			\item Giovanni: think of expost $\ga(e)$ as effort to elect/choose ex-post decision-maker
-			\item At worst, I can use leg / exec split to deal with this criticism
-			\item Kyle (???): advertising paper (???)
-			\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
-				\begin{itemize}
-					\item Giovanni: if I can show properties survive  monotonic transformation, I'm fine
-					\item They don't like gov't choosing its own preferences. May just need to change wording
-				\end{itemize}
-			\item Bob: what is fundamental structure of preferences?
-			\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
-		\end{itemize}
-	Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc.
-	\begin{itemize}
-			\item Maybe I should just get rid of tariff cap result altogether [me]
-		\end{itemize}
 \section{Take out repeats}
@@
 Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point.
-
-\section{Make a bigger deal of the idea of \texorpdfstring{$\gamma$}{gamma} decreasing function of tariffs---move out of footnote}
-Ethier 2012 for protection decreasing function of $\tau$ [NEED TO REVISIT IN TEXT]. ``The Political-Support Approach to Protection,'' Global Journal of Economics
-	\begin{itemize}
-		\item Making $W$ quadratic in $\ga$ such as $\ga - \ga^2$ gives a nice interior max
-	\end{itemize}
 
@@
 
-\section{Lit review}
+\section{Misc.}
 \begin{itemize}
-	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
+	\item Define $\ga$ before using it
+	\item If add back in assumption 1, add Ethier footnote
+	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
+	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
+	\item Ben: I'm wondering how you view the time-frame over which your model is set.
+On one hand a specific-factors set-up is normally associated with the short-run.
+But on the other hand you have an infinite time horizon which suggests very long run.
+	\item Clarify information structure
+\end{itemize}
+
+\section{Make a bigger deal of the idea of \texorpdfstring{$\gamma$}{gamma} decreasing function of tariffs---move out of footnote}
+Ethier 2012 for protection decreasing function of $\tau$ [NEED TO REVISIT IN TEXT]. ``The Political-Support Approach to Protection,'' Global Journal of Economics
 	\begin{itemize}
-			\item Bagwell and Staiger 1999 AER
+		\item Making $W$ quadratic in $\ga$ such as $\ga - \ga^2$ gives a nice interior max
 	\end{itemize}
-	\item Have to make connection to endogenous papers clearer
-		\begin{itemize}
-			\item and make clear that GH is isomorphic / provides microfoundations for Baldwin objective function
-			\item More on Coates $\&$ Ludema (2001)
-		\end{itemize}
-	\item I think that I need to bring people along more carefully from GH/MRC to what I'm doing
-	\item Paola: How do I want to sell the paper?
-		\begin{itemize}
-			\item In relation to what literature?
-				\begin{itemize}
-					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
-					\item What literature \textit{doesn't}?
-				\end{itemize}
-			\end{itemize}
-	\item Review Bown, Bagwell $\&$ Staiger
-	\item Do I include papers like BS2001 (non-repeated) in lit review?
-	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
-	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
-	 contributions 'felonious'
-	\item David DeRemer: Safeguards have dropped off in use around the world
-		\begin{itemize}
-			\item Look at U.S. vs. ROW (he sent papers)
-		\end{itemize}
-	\item Limao and Tovar 2011
-		\begin{enumerate}
-			\item Their 2nd (and final) stage is Nash bargaining. Is there an isomorphism to my setup?
-			\item It IS clear to take their first, limited case of commitment to tariff cap only (ignore NTB) as analogy for gov't in my model choosing commitment to different tariff level.
-Need to answer Q1 above in order to answer points 3 and 4:
-			\item Their gains from commitment derive from improvement in bargaining power vis-a-vis the lobby (Schelling conjecture invoked). How does this relate to the gain from commitment in my model?
-			\item Top of page 6: ``large enough bargaining power.'' Can I make connection to $\ga(e)$?
-			\item Note: in their equation 3, $C$ enters expression for NBS's ($\tau$) in equilibrium, it doesn't disappear as in GH
-		\end{enumerate}
-\end{itemize}
+
+\section{Lobby's optimal effort}
+Need to be clear about when lobby's optimal $\tau$ is higher than gov'ts: that is, when TA would want to rein in lobby and escape clause would ruin that
 
@@
 
-\section{Lobby's optimal effort}
-Need to be clear about when lobby's optimal $\tau$ is higher than gov'ts: that is, when TA would want to rein in lobby and escape clause would ruin that
-
-\section{Dynamic use constraint}
-When would lobby exert effort to top up?
-\begin{itemize}
-	\item Do I want correlation between endogenous/exogenous parts? 
-\end{itemize}
-
-\section{Numeraire}
-Do I need to add a numeraire so I can put $e$ in terms of it?
-
-\section{Lobbying cost included in gov't welfare}
-	\begin{itemize}
-		\item Rick Bond: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$
-		\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
-		\begin{itemize}
-			\item Rohan: $(1-a)\cdot C + a \cdot W \equiv C + a \cdot (W-C)$
-			\item When weights on $\pi_X$ are not linear, what are the implications for resource waste? Is it the same as $+ C / - C$?
-			\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
-				\begin{itemize}
-					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's
-					\item Find something in persuasion literature
-				\end{itemize}
-		\end{itemize}
-	\end{itemize}
 \end{document}

### 2016-08-13 to-do.tex
continue reading on dynamic inconsistency

@@
 				\begin{itemize}
-					\item DILLENBERGER, D., J. S. LLERAS, P. SADOWSKI, AND N. TAKEOKA (2013): ``A Theory of Subjective Learning,'' Working Paper. 
-					\item GUL, F., AND W. PESENDORFER (2004): ``Self-Control and the Theory of Consumption,'' Econometrica, 72 (1), 119-158
+					\item \sout{DILLENBERGER, D., J. S. LLERAS, P. SADOWSKI, AND N. TAKEOKA (2013): ``A Theory of Subjective Learning,'' Working Paper.}
+						\begin{itemize}
+							\item This is about figuring out the subjective learning structure by looking at a DM's preference for flexibility.
+Not directly relevant to my question b/c the information structure is objective.
+						\end{itemize}
+					\item GUL, F., AND W. PESENDORFER (2004): ``Self-Control and the Theory of Consumption,'' Econometrica, 72 (1), 119-158 (infinite horizon model)
+					\item Gul and Pesendorfer's (2001) Econometrica is a two-period model. No dynamic inconsistency; derive temptation and desire for commitment in a model with dynamically consistent preferences.
 					\item \sout{JONES, R. A., AND J. M. OSTROY (1984): ``Flexibility and Uncertainty,'' Review of Economic Studies, 51 (1), 13-32}
@@
 					\item \sout{KREPS, D. M., AND E. L. PORTEUS (1978): ``Temporal Resolution of Uncertainty and Dynamic Choice Theory,'' Econometrica, 46 (1), 185-200.}
+					\item Follow up on highlighted references in Kreps (1979): Hammond, Stigler $\&$ Becker, Strotz, Yaari
+					\item Promising: Hammond, Peter J., ``Changing Tastes and Coherent Dynamic Choice,'' The Review of Economic Studies, Vol. 43, No. 1 (Feb., 1976)
 				\end{itemize}
@@
 
+\section{Big picture of which changes do what}
+\vskip.2in
+		\begin{tabular}{|l|c|c|}
+			 & rigid & escape \\
+			BS2005  & on-schedule (truthtelling): trivial; & need cost  \\
+			 & off (repeated): optimal static if patient enough & \\
+			$\ga(e)$ & on: trivial& no need, but revisit side payments\\
+			 & off: mabye not b/c of lobby & \\
+			$\ga(e,s)$ & ignore for now & is $s$ verifiable? def not w/$\ga(s,e)$
+		\end{tabular}
+\vskip.2in
+\begin{itemize}
+	\item Make sure this story is clear in text: I establish baseline tariff cap case, then add escape clause. We can already see a story of $\ldots$ emerge...
+	\item Need to be clear when it is flexibility, enforcement, endogenous $\ga$ that screws things up. THEN need to bmake sure it comes out in paper
+	\item May want to ditch repeated game part...
+\end{itemize}
+
 \section{BS2005 vs. MRC2007}
@@
 
-\section{Big picture of which changes do what}
-\vskip.2in
-		\begin{tabular}{|l|c|c|}
-			 & rigid & escape \\
-			BS2005  & on-schedule (truthtelling): trivial; & need cost  \\
-			 & off (repeated): optimal static if patient enough & \\
-			$\ga(e)$ & on: trivial& no need, but revisit side payments\\
-			 & off: mabye not b/c of lobby & \\
-			$\ga(e,s)$ & ignore for now & is $s$ verifiable? def not w/$\ga(s,e)$
-		\end{tabular}
-\vskip.2in
-\begin{itemize}
-	\item Make sure this story is clear in text: I establish baseline tariff cap case, then add escape clause. We can already see a story of $\ldots$ emerge...
-	\item Need to be clear when it is flexibility, enforcement, endogenous $\ga$ that screws things up. THEN need to bmake sure it comes out in paper
-	\item May want to ditch repeated game part...
-\end{itemize}
-
 \section{Misc.}

### 2016-08-16 to-do.tex
finish prelim lit review on changing preferences

@@
 					\item \sout{KREPS, D. M., AND E. L. PORTEUS (1978): ``Temporal Resolution of Uncertainty and Dynamic Choice Theory,'' Econometrica, 46 (1), 185-200.}
-					\item Follow up on highlighted references in Kreps (1979): Hammond, Stigler $\&$ Becker, Strotz, Yaari
+					\item Follow up on highlighted references in Kreps (1979): Strotz, which is about inconsistencies in discount function (non-exponential discounting).
 					\item Promising: Hammond, Peter J., ``Changing Tastes and Coherent Dynamic Choice,'' The Review of Economic Studies, Vol. 43, No. 1 (Feb., 1976)
+					\item Even better, especially advertising example: Becker, Gary.
+S.
+and George J.
+Stigler, ``De Gustibus Non Est Disputandum,'' The American Economic Review, Vol.
+67, No.
+2 (Mar., 1977), pp.
+76-90
+						\begin{itemize}
+							\item I can make my argument almost exactly in analogy to this advertising example. My $\ga(e)$ function is like their $g(A)$.
+							\item Check Kyle's Discussion paper on advertising to get context
+						\end{itemize}
 				\end{itemize}

### 2016-08-17 to-do.tex
added Becker & Murphy (1993), removed unnecessary lit from dynamic preferences'

@@
 				\begin{itemize}
-					\item \sout{DILLENBERGER, D., J. S. LLERAS, P. SADOWSKI, AND N. TAKEOKA (2013): ``A Theory of Subjective Learning,'' Working Paper.}
-						\begin{itemize}
-							\item This is about figuring out the subjective learning structure by looking at a DM's preference for flexibility.
-Not directly relevant to my question b/c the information structure is objective.
-						\end{itemize}
-					\item GUL, F., AND W. PESENDORFER (2004): ``Self-Control and the Theory of Consumption,'' Econometrica, 72 (1), 119-158 (infinite horizon model)
 					\item Gul and Pesendorfer's (2001) Econometrica is a two-period model. No dynamic inconsistency; derive temptation and desire for commitment in a model with dynamically consistent preferences.
-					\item \sout{JONES, R. A., AND J. M. OSTROY (1984): ``Flexibility and Uncertainty,'' Review of Economic Studies, 51 (1), 13-32}
-					\item (couldn't find) KOOPMANS, T. C. (1964): ``On the Flexibility of Future Preferences,'' in Human Judgments and Optimality, ed. by M. W. Shelly and G. L. Bryan. New York: Wiley.
-					\item \sout{KREPS, D. M., AND E. L. PORTEUS (1978): ``Temporal Resolution of Uncertainty and Dynamic Choice Theory,'' Econometrica, 46 (1), 185-200.}
-					\item Follow up on highlighted references in Kreps (1979): Strotz, which is about inconsistencies in discount function (non-exponential discounting).
 					\item Promising: Hammond, Peter J., ``Changing Tastes and Coherent Dynamic Choice,'' The Review of Economic Studies, Vol. 43, No. 1 (Feb., 1976)
-					\item Even better, especially advertising example: Becker, Gary.
+					\item Even better, especially advertising example (credit Kyle): Becker, Gary.
 S.
@@
 							\item I can make my argument almost exactly in analogy to this advertising example. My $\ga(e)$ function is like their $g(A)$.
-							\item Check Kyle's Discussion paper on advertising to get context
 						\end{itemize}
+					\item Simpler: Becker, G. S. and K. M. Murphy (1993), “A Simple Theory of Advertising as a Good or Bad,” Quarterly Journal of Economics, 942-64.
 				\end{itemize}
-			\item Giovanni: think of expost $\ga(e)$ as effort to elect/choose ex-post decision-maker
 			\item At worst, I can use leg / exec split to deal with this criticism
-			\item Kyle (???): advertising paper (???)
 			\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
@@
 				\begin{itemize}
-					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's
-					\item Find something in persuasion literature
+					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's ``expost $\ga(e)$ as effort to elect/choose ex-post decision-maker''
 				\end{itemize}
@@
 	Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc.
-	\begin{itemize}
-			\item Maybe I should just get rid of tariff cap result altogether [me]
-		\end{itemize}
 

### 2016-08-17 game-tree.tex
add game-tree.tex from sop-rie and begin mods for this project

\begin{comment}
\documentclass[10pt]{article}

\addtolength{\textwidth}{1.3in}
\addtolength{\oddsidemargin}{-.65in} %left margin
\addtolength{\evensidemargin}{-.65in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
\usepackage{setspace}
\doublespacing

\usepackage{titling}

\usepackage[pdftex]{graphicx}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont,pdfsync}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{tikz}

% disables chapter, section and subsection numbering
\setcounter{secnumdepth}{-1}

\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
\usepackage[hang,flushmargin]{footmisc}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}

\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\bte}{\bm{\tau^E}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}

\begin{document}
\end{comment}

\tikzset{
% Two node styles for game trees: solid and hollow
solid node/.style={circle,draw,inner sep=1.5,fill=black},
hollow node/.style={circle,draw,inner sep=1.5,fill=white}
%non node/.style={circle,draw,inner sep=.1,fill=black}
}
\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
% Specify spacing for each level of the tree
\tikzstyle{level 1}=[level distance=10mm,sibling distance=8mm]
\tikzstyle{level 2}=[level distance=8mm,sibling distance=8mm]
% The Tree
\node(0)[solid node,label=above:{$E,E^*$}]{} 
child{node(1)[solid node,label=below:{$0$}]{}}
child{[white] node(2)[solid node,xshift=12,label=above:{L}]{}  
	child{[black] node(4)[solid node,label=below:{$0$}]{}}
	child{[white] node(5)[hollow node,xshift=12]{} 
	 child[sibling distance=11mm]{[black] node(22)[solid node]{}  }
	 child{[white] node(24)[solid node,xshift=7]{}
		child[sibling distance=8mm]{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
		child[sibling distance=8mm]{[white] node(8)[solid node,xshift=12]{}  
			child[sibling distance=10mm]{[black] node(10)[solid node]{}  }
			child{[white] node(11)[solid node,xshift=7]{}
				child[sibling distance=5mm]{[black] node(13)[solid node,label=below:{$0$}]{}}
				child{[white] node(15)[solid node,xshift=7]{}	}
				child[sibling distance=5mm]{[black] node(14)[solid node,label=above:{$\infty$}]{} }
				}
			child[sibling distance=10mm]{[black] node(12)[solid node]{}  }
			%edge from parent node[black, xshift=20]{$\ta$} 
			}
		child[sibling distance=8mm]{[black] node(9)[solid node,label=above:{$\infty$}]{} }
		}
	 child[sibling distance=11mm]{[black] node(23)[solid node]{}  }
	 edge from parent node[black, xshift=20]{$e_b$} 	
	 }
	child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
	edge from parent node[black, xshift=23]{$\bta$} 
	}
child{node(3)[solid node,label=above:{$\infty$}]{}
}
;
% information set
    \draw[solid,bend right](1)to(3);
		\draw[solid,bend right](4)to(6);
		\draw[solid,bend right](7)to(9);
		\draw[solid,bend right](13)to(14);
		\draw[solid,bend right](10)to(12);
		\draw[solid,bend right](22)to(23);
% labels
		\node[above,xshift=3,yshift=5]at(2){$L$}; 
		%\node[above,xshift=1,yshift=5]at(23){$N$}; 
		\node[above,xshift=3,yshift=5]at(5){$N$}; 
		\node[above,xshift=6,yshift=5]at(8){$Leg$}; 
		\node[above,xshift=5,yshift=5]at(12){$L$};
		\node[above,xshift=7,yshift=5]at(15){$N$};
		\draw[solid](5)circle(.05cm);
		\draw[solid](0)circle(.15cm);
		%\node[right]at (18) {$W_E\left(\btn\right),$};
		%\node[right,yshift=-10]at (18) {$\pi\left(\tau^n\right)-e_b -e_n,$};
		%\node[right,yshift=-20]at (18) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
\end{tikzpicture}

%\end{document}

### 2016-08-18 game-tree.tex
more mods to game-tree.tex, include so it can compile standalone for now

@@
-\begin{comment}
+%\begin{comment}
 \documentclass[10pt]{article}
@@
 \begin{document}
-\end{comment}
+%\end{comment}
 
@@
 % The Tree
-\node(0)[solid node,label=above:{$E,E^*$}]{} 
+\node(0)[solid node,label=above:{$G,G^*$}]{} 
 child{node(1)[solid node,label=below:{$0$}]{}}
@@
 	child{[white] node(5)[hollow node,xshift=12]{} 
-	 child[sibling distance=11mm]{[black] node(22)[solid node]{}  }
-	 child{[white] node(24)[solid node,xshift=7]{}
-		child[sibling distance=8mm]{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
-		child[sibling distance=8mm]{[white] node(8)[solid node,xshift=12]{}  
-			child[sibling distance=10mm]{[black] node(10)[solid node]{}  }
-			child{[white] node(11)[solid node,xshift=7]{}
-				child[sibling distance=5mm]{[black] node(13)[solid node,label=below:{$0$}]{}}
-				child{[white] node(15)[solid node,xshift=7]{}	}
-				child[sibling distance=5mm]{[black] node(14)[solid node,label=above:{$\infty$}]{} }
-				}
-			child[sibling distance=10mm]{[black] node(12)[solid node]{}  }
+	 %child[sibling distance=11mm]{[black] node(22)[solid node]{}  }
+	 child{[black] node(22)[solid node]{}  }
+	 child{[white] node(24)[solid node,xshift=12]{}
+		child{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
+		child{[white] node(8)[solid node,xshift=12]{}  
+			child{[black] node(10)[solid node]{}  }
+			child{[white] node(11)[solid node,xshift=14]{} }
+			child{[black] node(12)[solid node]{}  }
 			%edge from parent node[black, xshift=20]{$\ta$} 
 			}
-		child[sibling distance=8mm]{[black] node(9)[solid node,label=above:{$\infty$}]{} }
+		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
 		}
-	 child[sibling distance=11mm]{[black] node(23)[solid node]{}  }
+	 child{[black] node(23)[solid node]{}  }
 	 edge from parent node[black, xshift=20]{$e_b$} 	
@@
 		\draw[solid,bend right](7)to(9);
-		\draw[solid,bend right](13)to(14);
 		\draw[solid,bend right](10)to(12);
@@
 		\node[above,xshift=5,yshift=5]at(12){$L$};
-		\node[above,xshift=7,yshift=5]at(15){$N$};
 		\draw[solid](5)circle(.05cm);
 		\draw[solid](0)circle(.15cm);
-		%\node[right]at (18) {$W_E\left(\btn\right),$};
-		%\node[right,yshift=-10]at (18) {$\pi\left(\tau^n\right)-e_b -e_n,$};
-		%\node[right,yshift=-20]at (18) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
+		\node[right]at (11) {$W_E\left(\btn\right),$};
+		\node[right,yshift=-10]at (11) {$\pi\left(\tau^n\right)-e,$};
+		\node[right,yshift=-20]at (11) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
 \end{tikzpicture}
 
-%\end{document}
+\end{document}

### 2016-08-22 game-tree.tex
compile with game tree, remove pdftex option on graphicx package (fully functional on dell)
labels on game-tree

@@
-%\begin{comment}
+\begin{comment}
 \documentclass[10pt]{article}
@@
 \begin{document}
-%\end{comment}
+\end{comment}
 
@@
 	 %child[sibling distance=11mm]{[black] node(22)[solid node]{}  }
-	 child{[black] node(22)[solid node]{}  }
+	 child{[black] node(22)[solid node,,label=below:{$-\infty$}]{}  }
 	 child{[white] node(24)[solid node,xshift=12]{}
-		child{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
+		child{[black] node(7)[solid node,label=below:{$0$}]{}}
 		child{[white] node(8)[solid node,xshift=12]{}  
-			child{[black] node(10)[solid node]{}  }
+			child{[black] node(10)[solid node,label=below:{$0$}]{}  }
 			child{[white] node(11)[solid node,xshift=14]{} }
-			child{[black] node(12)[solid node]{}  }
+			child{[black] node(12)[solid node,label=above:{$\infty$}]{}  }
 			%edge from parent node[black, xshift=20]{$\ta$} 
@@
 		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
+		edge from parent node[black, xshift=20]{$s$} 	
 		}
-	 child{[black] node(23)[solid node]{}  }
-	 edge from parent node[black, xshift=20]{$e_b$} 	
+	 child{[black] node(23)[solid node,label=above:{$\infty$}]{}  }
+	 edge from parent node[black, xshift=20]{$e$} 	
 	 }
@@
 		\node[above,xshift=3,yshift=5]at(2){$L$}; 
-		%\node[above,xshift=1,yshift=5]at(23){$N$}; 
+		\node[above,xshift=3,yshift=5]at(24){$G$}; 
 		\node[above,xshift=3,yshift=5]at(5){$N$}; 
-		\node[above,xshift=6,yshift=5]at(8){$Leg$}; 
-		\node[above,xshift=5,yshift=5]at(12){$L$};
+		\node[above,xshift=6,yshift=5]at(8){$G^*$}; 
+		%\node[above,xshift=5,yshift=5]at(12){$L$};
 		\draw[solid](5)circle(.05cm);
@@
 
-\end{document}
+%\end{document}

### 2016-08-23 game-tree.tex
fix rest of game-tree labels; will examine notation later

@@
 			child{[black] node(12)[solid node,label=above:{$\infty$}]{}  }
-			%edge from parent node[black, xshift=20]{$\ta$} 
+			edge from parent node[black, xshift=20]{$\tau$} 
 			}
@@
 		\draw[solid](0)circle(.15cm);
-		\node[right]at (11) {$W_E\left(\btn\right),$};
-		\node[right,yshift=-10]at (11) {$\pi\left(\tau^n\right)-e,$};
-		\node[right,yshift=-20]at (11) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
+		\node[left]at (11) {$\tau^*$};
+		\node[right]at (11) {$W\left(\ga(e,s),\bm{\tau}\right),W^*\left(\ga(e^*,s^*),\bm{\tau}\right)$};
+		\node[right,yshift=-10]at (11) {$\pi\left(\tau \right)-e$};
 \end{tikzpicture}

### 2016-08-25 to-do.tex
incorporate notes and streamline 'dynamic preferences' section

@@
 	\begin{itemize}
-		\item James: need to be more precise about when $\ga$ is evaluated
-			\begin{itemize}
-				\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
-			\end{itemize}
-		\item References from Krishna and Sadowski (2014) ECTA
-				\begin{itemize}
+
 					\item Gul and Pesendorfer's (2001) Econometrica is a two-period model. No dynamic inconsistency; derive temptation and desire for commitment in a model with dynamically consistent preferences.
-					\item Promising: Hammond, Peter J., ``Changing Tastes and Coherent Dynamic Choice,'' The Review of Economic Studies, Vol. 43, No. 1 (Feb., 1976)
 					\item Even better, especially advertising example (credit Kyle): Becker, Gary.
@@
 76-90
+					\item Simpler: Becker, G. S. and K. M. Murphy (1993), Simple Theory of Advertising as a Good or Bad, Quarterly Journal of Economics, 942-64.
 						\begin{itemize}
-							\item I can make my argument almost exactly in analogy to this advertising example. My $\ga(e)$ function is like their $g(A)$.
+							\item Their firms are my firm. Their consumers are my government
+							\item Kyle's lit review talkes about `persuasive,' `informative,' and `complementary' views of advertisting. This is complementary.
+							\item $e$ and $\tau$ enter objective function.
+Gov't is sophisticated, can make commitment that alters $e$ (i.e.
+behavior of another actor).
+But $CS + \ga(e)PS + \ldots$, just like $u(X,Y,A)$; says consumer could work to pass law about number of adds on TV each hour.
+							\item Use newly drawn extensive form to explain this, lays out when welfare is realized
+							\item Addresses James' concern about needing to be more precise about when $\ga$ is evaluated)
+			\begin{itemize}
+				\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
 			\end{itemize}
-					\item Simpler: Becker, G. S. and K. M. Murphy (1993), “A Simple Theory of Advertising as a Good or Bad,” Quarterly Journal of Economics, 942-64.
+
 				\end{itemize}
-			\item At worst, I can use leg / exec split to deal with this criticism
 			\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
@@
 		\end{itemize}
-	Bob: maybe gov't has some other reason for setting non-zero tariff that opens the door for lobbies. 2nd best remedy for market failure, etc.
 

### 2016-08-25 to-do.tex
incorporate some notes into gamma decreasing in tau section of to-do
add outline of paper w/propositions to to-do list

@@
 	\item Making $W$ quadratic in $\ga$ such as $\ga - \ga^2$ gives a nice interior max
+	\item I think what I have already has returns to $e$ decreasing in $\tau$. Don't I show it? If not in this paper, in RoIE paper.
+	\item What embodies returns to lobbying?
+		\begin{itemize}
+			\item $e \rightarrow \ga(e)$ is how lobbying is translated into weight in political process. It's not `returns.'
+			\item $\tau$ isn't even returns. $\pi(\tau(e_2)) - \pi(\tau(e_1))$ is returns.
+		\end{itemize}
+	\item To get returns decreasing in $\tau$, maybe weighting can't be confined to $PS$
 \end{itemize}
@@
 
+\newpage
+Current structure of paper (August 25, 2016)
+\begin{enumerate}
+	\item Introduction
+	\item Model
+	\item Rigid Tariffs with Endogenous Political Pressure
+		\begin{itemize}
+			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
+			\item[3.2] $\cdot$
+				\begin{itemize}
+					\item[3.2.1] $\cdot$
+					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
+				\end{itemize}
+		\end{itemize}
+	\item Endogenous Political Pressure and the Escape Clause
+		\begin{itemize}
+			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
+			\item[4.2] Side payments: spirit not upheld, but IC
+			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
+			\item[4.4] EC for endogenous politics
+		\end{itemize}
+	\item Conclusion
+\end{enumerate}
+
 \end{document}

### 2016-12-06 figures.R
add pe_gov_utility_exante from ex-ante folder and begin to convert to r

@@ -1,0 +1,132 @@
+g = [1:.001:7/4];
+g = seq(1,7/4,0.001)
+
+W = 9/98 + 8*g/98 + ((8*g - 8)/49)*(8*g - 5)/(4*(17-2*g)) + (2*(2*g-17) + 9)/49 * (8*g - 5)/(4*(17-2*g))^2;
+
+plot(g,W);
+
+%unilateral
+t = (8*g - 5)/(4*(17-2*g));
+gtn = (68*t + 5)/(8*(1+t));
+
+PSx = ((2 +2*t)^2)/49;
+CSx = 5*((3 -4*t)^2)/49;
+TR = (t - 6*t^2)/7;
+CSy = ((3 +3*t)^2)/98;
+PSy = ((4 -3*t)^2)/98;
+
+w = 1./(4+g);
+w = 1;
+
+W2 = w*CSx + g*w*PSx + w*CSy + w*PSy + w*TR;
+plot(g,W);
+
+%efficient tariff
+te = (4*g - 4)./(25-4*g);
+gte = (25*t + 4)./(4+4*t);
+
+PSxe = ((2 +2*te).^2)/49;
+CSxe = .5*((3 -4*te).^2)/49;
+TRe = (te - 6*te.^2)/7;
+CSye = ((3 +3*te).^2)/98;
+PSye = ((4 -3*te).^2)/98;
+
+w = 1./(4+g);
+
+We = w.*CSxe + g.*w.*PSxe + w.*CSye + w.*PSye + w.*TRe;
+plot(g,We);
+
+
+                          
+                          
+##########################################################
+#          Do everything as a function of t              #
+##########################################################
+t = seq(0,1/6,0.001)
+                          
+PSx = ((2 +2*t)^2)/49;
+CSx = .5*((3 -4*t)^2)/49;
+TR = (t - 6*t^2)/7;
+CSy = ((3 +3*t)^2)/98;
+PSy = ((4 -3*t)^2)/98;
+                          
+gtn = (68*t + 5)/(8*(1+t));  #gamma as fcn of tau, unilateral maximization
+
+
+#some values of tau are so low that gamma would be less than one
+gtn2 = zeros(size(gtn));
+gtn2 = matrix(0,length(gtn),1)
+
+for (j in 1:length(gtn)) {
+  x = gtn[j]
+  gtn2[j] = max(x,1)
+}  
+
+  
+
+w = 1./(4+gtn2);
[…]

### 2016-12-06 figures.R
add '-e' to code for graphs in rough test run

@@ -1,13 +1,112 @@
-g = [1:.001:7/4];
+##########################################################
+#          Do everything as a function of t              #
+##########################################################
+t = seq(0,1/6,0.001)           #vector of non-prohibitive tariff values
+                          
+PSx = ((2 +2*t)^2)/49
+CSx = .5*((3 -4*t)^2)/49
+TR = (t - 6*t^2)/7
+CSy = ((3 +3*t)^2)/98
+PSy = ((4 -3*t)^2)/98
+                          
+gtn = (68*t + 5)/(8*(1+t))    #gamma as fcn of tau, unilateral maximization
+e = (gtn - 1)^4
+
+
+#some values of tau are so low that gamma would be less than one
+gtn2 = matrix(0,length(gtn),1)
+
+for (j in 1:length(gtn)) {
+  x = gtn[j]
+  gtn2[j] = max(x,1)
+}  
+
+w = 1/(4+gtn2)                 #create weights as function of gamma
+                          
+W = w*CSx + gtn2*w*PSx + w*CSy + w*PSy + w*TR  #fully weighted welfare
+plot(t,W)
+                          
+W2 = CSx + gtn2*PSx + CSy + PSy + TR  #partially weighted welfare
+plot(t,W2)
+
+W3 = CSx + gtn2*PSx + CSy + PSy + TR - e  #partially weighted welfare - e
+plot(t,W3)
+
+                          
+##########################################################
+#               Unconverted Matlab code                  #
+##########################################################
+# Create figure for weighted function
+figure1 = figure('Color',[1 1 1]);
+                          
+# Create axes
+axes1 = axes('Parent',figure1);
+                          
+# set X-limits of the axes
+xlim(axes1,[0 1/6]);
+# set Y-limits of the axes
+ylim(axes1,[0.0854 0.087]);
+box(axes1,'on');
+hold(axes1,'all');
+                          
+# Create multiple lines using matrix input to plot
+plot1 = plot(t,[W],'Parent',axes1,'LineWidth',1);
+set(plot1(1),'DisplayName','Trade Agreement Tariffs',...
+'Color',[0 0 0]);
+                          
+# Create title
+#title('Government Welfare with Weighted Objective Function',...
+'FontWeight','demi',...
+'FontSize',12);
+                          
[…]

### 2016-12-08 to-do.tex
update to-do after JIE_revision (broad strokes)
add language from JIE_revision and letter to to_do

@@
 	\begin{itemize}
-
 					\item Gul and Pesendorfer's (2001) Econometrica is a two-period model. No dynamic inconsistency; derive temptation and desire for commitment in a model with dynamically consistent preferences.
@@
 			\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
+				\begin{itemize}
+					\item ``How the assumption of a non-unitary legislature translates into this incentive constraint.
+In the non-unitary legislature, if the current period incentive constraint includes any future value of lobbying effort, this would mean the current-period median legislator evaluating her current-period incentive constraint using a mixture of her own political-economy weight with those of the legislators who are median in the future in the trade agreement and trade-war scenarios.
+The first major step I have taken to clarify the incentive constraint is the analysis in Appendix B.1 of the unitary model (described above), which I would guess is the model you had in mind.
+I discuss the differences between the unitary and non-unitary models in the beginning of that Appendix.''
+					\item ``Note that the median legislator, whose identity is determined by the lobby's effort level $e_b$, evaluates future payoffs according to her own political economy weight, $\ga(e_b)$.
+Of course, depending on legislator $e_b$'s choice, either legislator $e_a$ or legislator $e_{tw}$ will be the decision maker in those future periods.
+But legislator $e_b$, who is the decision maker in the current period, maximizes her own welfare given the predicted behavior of future decision makers.''
+
+				\end{itemize}
 			\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
@@
 				\item When weights on $\pi_X$ are not linear, what are the implications for resource waste? Is it the same as $+ C / - C$?
+			\end{itemize}
+		\item Try subtracting $e$ and see what it does to the graphs. Does it give me the kind of non-monotonicity I'd like? YES
+			\begin{itemize}
+				\item Does it do anything to the analysis?
+			\end{itemize}
+		\item JIE$\_$revision: I have also verified that \emph{none} of the results of the model change if lobbying effort is subtracted from the legislative/executive welfare function (except, of course, the evaluated level of welfare).
+I have refrained from adding the ``$-e$'' to the welfare function out of a desire to be consistent with the existing literature.
+	\end{itemize}
+
+\section{BS2005 vs. MRC2007}
+	\begin{itemize}
+		\item Review appendix of JIE$\_$revision to see what connections I want to make
+		\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
+			\begin{itemize}
+				\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
+			\end{itemize}
+		\item I like the way you use Bagwell and Staiger (2005) as your benchmark.
+But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.
+I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.
+In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.
+But your mechanisms seem quite different.
+They have capital mobility whereas your factors are specific to each sector.
+If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.
+What drives this feature in your model?
+Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.
+What is the relationship between lobby effort level and tariff policy?
+Is effort fixed first, after which the government sets its tariff?
+If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.
+It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.
+Having done that, you could bring in ex-ante lobbying and treat it the same way.
+		\item In general it seems to me that you seem to "dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
+		\item DGH / MRC analogy from JIE$\_$revision
+			\begin{itemize}
+				\item I've supplied the suggested results to connect my model to that of Dixit, Grossman and Helpman (1997) in Appendix B.2, along with some text pointing to it at the end of Section 2.2 on page 10.
+The results from the DGH-style model are qualitatively different in a minor way.
+I discuss this difference in Appendix B.1 where I present a unitary government/legislature version of the model that most closely matches Maggi and Rodriguez-Clare 2007.
+				\item The model does admit an interpretation in which the same branch of government both negotiates the trade agreement and decides on the applied tariff ex-post.
+In this sense, the structure of the one-shot game shares much in common with Maggi and Rodriguez-Clare 2007.
+				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007.
+Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored.
+In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps.
+The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff.
+There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff.
+Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement.
+				\item Consider the following, alternative interpretation of the model.
+Assume there is only one decision-making body but the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated.
+For ease of exposition, take this single decision-making body to be the legislature so that the single decision-making body has the preferences in Expression 2.
+At the ex-ante stage, the welfare function would reduce to that in Expression 4.
+This interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.\footnote{To match the assumption of a social-wefare maximing executive, this requires the additional assumption that $\ga(0)=1.$ The model is qualitatively unchanged for other values of $\ga(0)$ as long as the analogue of Assumption 2 below holds.} Note that this remains a non-unitary model.
+Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g.
+during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$).
+A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results.
+I discuss the unitary version of the model in Appendix B.1.
+Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
+				\item The model admits this `one-branch' interpretation only if the one branch is non-unitary
+				\item the model analyzed in the body of this paper can be given an interpretation in line with Maggi and Rodriguez-Clare 2007 if capital is completely mobile in the long run and the single decision-making body is non-unitary.
+In Appendix B.1, I compare the results of this paper to those from a model with a unitary government, as this would seem to bring the model most closely into alignment with Maggi and Rodriguez-Clare 2007 and a large part of the literature.
+Then, in Appendix B.2, since the government welfare function used in this paper can be interpreted as a special case of the one proposed by DGH97, I will show that a model patterned after DGH97 and specialized to this environment---which is also a unitary model---provides results that are qualitatively similar to the unitary model.
+				\item I have added Appendix B.1 with analysis of the unitary model
+					\begin{itemize}
+						\item In the model with a non-unitary government, the lobby's effort $e$ determines the identity of the decision maker.
+Thus the decision maker at the time the lobby is pushing for the agreement to be broken ($e=e_b>e_a$ in order to have a median legislator who will break the agreement) is different from the decision maker during the trade war phase $(e = e_{tw})$ and the decision maker during a trade agreement phase ($e=e_a(\tau^a)$).
+The weight on the lobbying industry's profits changes with lobbying effort because lobbying effort determines a different median voter in the legislature or government more generally.
+However, when the median voter during a `break' phase is evaluating her incentive constraint, she values future tariff choices --- which will be determined by whoever is median in the future --- with her own preferences.
+Although she will not be the decision maker in the future, there is no reason for her to evaluate future welfare according to some other legislator's preference.
+						\item If the model is interpreted as having a unitary government, there is only one decision maker.
+There is a fixed mapping from lobbying effort to the weight the decision maker places on the lobby's profits, and the realization of this weight changes with the realization of lobbying effort.
+Thus, predicting future lobbying effort, the decision maker knows what the realization of her weight on the lobby's profits will be and evaluates the incentive constraint accordingly.
+Expression (10) would be modified while the lobbying constraint is unchanged.
+					\end{itemize}
+				\item Appendix B.2 with analysis of DGH model
+					\begin{itemize}
+						\item I claim that the government welfare function presented in this paper can be interpreted as a special case of the general welfare function proposed in DGH97.
+They specify government welfare as $G(a,c)$ where $a$ is the policy vector $\left(\left(\tau,\tau^*\right)\right)$ and $c$ is a vector of payments from each lobby ($e$).
+						\item Here I examine a version of the DGH97 model specialized to this environment.
+Because the DGH97 model is fundamentally a unitary government model, the relevant benchmark is the unitary version of the model.
+The results of the two unitary models are qualitatively the same.
+						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
+In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort.
+This distinction is mainly important in the context of the repeated game.
+Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
+						\item In the auction-menu set-up of the DGH97 model, there does not seem to be a natural alternative to the unitary government interpretation.
+The lobby offers a contribution schedule of effort levels and tariffs and the government chooses among these pairs.
+In the non-unitary model, each different effort level determines a different government, so the government cannot choose between elements of a schedule.
+To analyze this model, I make the standard assumptions of a unitary government and that the lobby has all the bargaining power.
+						\item Denote the government's welfare function as $W_G(\tau,\tau^*,e(\tau)) = W(\tau,\tau^*) + g(e(\tau))$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare. Then do analysis...
+					\end{itemize}
+				\item Appendix B.3 with analysis of the strong binding case. 
+					\begin{itemize}
+						\item If the trade agreement involves strong bindings instead of tariff caps (i.e.
+weak bindings), the legislature must deliver $\tau^a$ and the lobby's optimal effort during a period in which the trade agreement holds is $e_a = 0$.
+That the lobby pays less---here, nothing---for the protection it receives under the trade agreement is the only change from the base model.
+This means that nothing changes in the trade war.
+Likewise the legislature's repeated-game incentive constraint is unchanged.
+But the lobby's incentive constraint IS changed...
+						\item As in MRC, tariff caps in this model serve to keep the lobby `in the game.' A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement.
+From the ex-ante point of view, strong bindings are therefore preferable.
+Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
+					\end{itemize}
 			\end{itemize}
@@
 
-\section{BS2005 vs. MRC2007}
-		\begin{itemize}
-			\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
-				\begin{itemize}
-					\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
-				\end{itemize}
-			\item I like the way you use Bagwell and Staiger (2005) as your benchmark.
-But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.
-I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.
-In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.
-But your mechanisms seem quite different.
-They have capital mobility whereas your factors are specific to each sector.
-If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.
-What drives this feature in your model?
-Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.
-What is the relationship between lobby effort level and tariff policy?
-Is effort fixed first, after which the government sets its tariff?
-If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.
-It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.
-Having done that, you could bring in ex-ante lobbying and treat it the same way.
-			\item In general it seems to me that you seem to "dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
-		\end{itemize}
-
 \section{Structured differently from most papers}
@@
 
-\newpage
+\vskip.5in
 Current structure of paper (August 25, 2016)

### 2016-12-09 draft.tex
compile draft for dell laptop
add game-tree.tex from sop-rie and begin mods for this project
take out game-tree for now, switch to office location of bib
compile with game tree, remove pdftex option on graphicx package (fully functional on dell)
fix rest of game-tree labels; will examine notation later
update bibtex file with info on my paper revisions

@@
 %dell laptop
-%\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
-
-%wallis computer
-\addbibresource{C:/Users/wallis/Dropbox/Research/xBibs/tradeagreements.bib}
+\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
 
 %campus computer
-%\addbibresource{C:/Apps-SU/Dropbox/Research/xBibs/tradeagreements.bib}
+\addbibresource{C:/Apps-SU/Dropbox/Research/xBibs/tradeagreements.bib}
 \renewcommand{\newunitpunct}{,}
@@
 \usepackage[american]{babel}
+\usepackage[usenames,dvipsnames]{color}
+\usepackage{tikz}
+\tikzset{
+% Two node styles for game trees: solid and hollow
+solid node/.style={circle,draw,inner sep=1.5,fill=black},
+hollow node/.style={circle,draw,inner sep=1.5,fill=white}
+%non node/.style={circle,draw,inner sep=.1,fill=black}
+}
 
@@
 bookmarksopen=true,hyperfootnotes=false]{hyperref}
-\usepackage[pdftex]{graphicx}
+\usepackage{graphicx}
 
@@
 \usepackage[hang,flushmargin]{footmisc}
-\usepackage[usenames,dvipsnames]{color}
+
 %\usepackage{cite}
@@
 Finally, given the trade agreement, enforcement conditions, and political pressure it experiences, each government chooses the applied tariff on its import-competing good.
+
+\begin{figure}
+	\begin{center}
+		\input{game-tree.tex}
+	\end{center}
+	\caption{Extensive Form\label{fig:ext}}
+\end{figure}
 

### 2016-12-09 to-do.tex
update bibtex file with info on my paper revisions

@@
 	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
-	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dyanmic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
+	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
 	\item Ben: I'm wondering how you view the time-frame over which your model is set.
@@
 
-\vskip.5in
+\newpage
 Current structure of paper (August 25, 2016)
@@
 			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
-			\item[3.2] $\cdot$
+			\item[3.2] Self-Enforcing Trade Agreements
 				\begin{itemize}
-					\item[3.2.1] $\cdot$
+					\item[3.2.1] Repeated Game
 					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement

### 2016-12-10 draft.tex
make minor changes from read through sec:rigid and sec:escape

@@
 
-When political pressure is endogenous, in place of taking expectations over a range of probabilistically-determined levels of exogenous political pressure, we can backward induct to determine the lobby's effort decision and the government's optimal choice of tariff caps in anticipation of the lobby's behavior.
+When political pressure is endogenous, in place of taking expectations over a range of probabilistically-determined levels of exogenous political pressure, we backward induct to determine the lobby's effort decision and the government's optimal choice of tariff caps in anticipation of the lobby's behavior.
 
@@
 
-We know that the lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding.
+The lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding.
 The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$.
-The government is unable to use a weak binding to encourage extra lobbying in the case with only one import-competing lobby under consideration here.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}
+The government is unable to use a weak binding to encourage extra lobbying when only import-competing interests are represented by a lobby.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}
 
@@
 
-Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage according to the chosen tariff cap $\tau^R_{W,e}$.
+Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage.
 The governments get to choose the $\tau^R_{W,e}$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.
@@
 Notice that the maximand in Expression~\ref{exp:1} is exactly the same as that in this third-stage maximization problem except it adds  foreign government welfare.
-This implies that $\tau^R_{W,e}$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments do not want as much lobbying as possible.
-To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.}
+This implies that $\tau^R_{W,e}$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments want as much lobbying as possible.
+To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.\label{fn:tot}}
 
@@
 In fact, any strong binding accomplishes this very feat, but is also \textit{unable} to generate lobbying activity because of its effect on lobbying incentives.
-Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.}
+Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.\label{fn:mrc1}}
 
@@
 We know that the former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here.
-Following most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternatives that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.}
+Following most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternative that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.\label{fn:krw}}
 	%see KRW, Buzard 2014 for alternatives
@@
 We will see that moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem.
-Moving to the repeated-game environment with self enforcement does not alter the governments costs and benefits so that the government's optimal trade agreement tariff is invariant to the enforcement conditions.
+Moving to the repeated-game environment with self enforcement does not alter the government's costs and benefits so that the government's optimal trade agreement tariff is invariant to the enforcement conditions.
 This creates the possibility that the politically-optimal trade agreement tariff may not be self-enforcing.
@@
 
-To see this, start by noticing that the lefthand side of Expression~\ref{exp:govincent} in increasing in $\de$, while the righthand side is constant in $\de$.
+To see this, start by noticing that the lefthand side of Expression~\ref{exp:govincent} is increasing in $\de$, while the righthand side is constant in $\de$.
 That is, as the government becomes more patient, it is easier to satisfy the government's constraint.
@@
   The presence of the lobby as a repeated-game player may imply that the politically optimal self-enforcing trade agreement tariff is strictly greater than $\tau^R_{W,e}$.
+	\label{res:repeated}
 \end{proposition}
@@
 From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section~\ref{sec:objfcn}, the governments' ex-ante welfare is reduced by excess lobbying.
+However, the problem is not interesting if government welfare is everywhere increasing in lobbying effort because the lower binding will be set to the prohibitive tariff in this environment with no counter-lobbying by export groups.
 
@@
 If the government applies $\tau^F_{W,e}$, implementation of the agreement goes on as before.
-Consistent with the WTO Agreement on Safeguards, the government can apply the tariff that is necessary to remedy the injury caused by the shock, $\tau(s)$, with no retaliation for three periods.
-If the government applies a tariff that is higher than either $\tau^F_{W,e}$, $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
-I label this tariff $\tau^B(\ga(s,e^F))$, where $e^F$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.
+Consistent with the WTO Agreement on Safeguards, the government can apply the tariff that is necessary to remedy the economic injury---not necessarily equal to the political injury---caused by the shock, $\tau(s)$, with no retaliation for three periods.
+If the government applies a tariff that is higher than either $\tau^F_{W,e}$ or $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
+I label this tariff $\tau^B(\ga(s,e^F))$, where $e^F$ must be high enough to give the government incentive to break the trade agreement in the absence of external enforcement.
 
@@
 
-Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may chooses to set the trade agreement tariff in the presence of an escape clause $\left(\tau^F_{W,e}\right)$ so that in some cases the lobby's incentive constraint is violated.
+Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau^F_{W,e}\right)$ so that in some cases the lobby's incentive constraint is violated.
 This would lead to the legislature's constraint being violated as well, creating a dispute.
@@
 
-It still does not seem feasible to provide this flexibility at precisely the level that the shock impacts the government's preferences $\ga(s)$.
+It still does not seem feasible to provide this flexibility at precisely the level that the shock impacts the government's preferences $\ga(s)$ if the WTO does not know the mapping from $s$ to $\ga$.
 What's more, this may provide an answer to why Safeguard measures have fallen out of use.

### 2016-12-10 to-do.tex
make minor changes from read through sec:rigid and sec:escape

@@
 	\begin{itemize}
-		\item Review appendix of JIE$\_$revision to see what connections I want to make
 		\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
@@
 			\begin{itemize}
+				\item Footnote fn:mrc1 is good place to start to put references to MRC 2007
 				\item I've supplied the suggested results to connect my model to that of Dixit, Grossman and Helpman (1997) in Appendix B.2, along with some text pointing to it at the end of Section 2.2 on page 10.
@@
 	\item Clarify information structure
+	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together better
+	\item Add ref to JIE R$\&$R to fn:krw ?
+	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
+	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
 \end{itemize}

### 2016-12-10 draft.tex
add thanks to draft; integrate to-dos from read through

@@
 \begin{document}
-\title{\vskip-0.6in Endogenous Politics and the Design of Trade Institutions}
+\title{Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Chad Bown, Nuno Limao, Giovanni Maggi, Anson Soderbery, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at SITE for very helpful comments and discussions.}}
 \author{Kristy Buzard\thanks{Syracuse University, Economics Department. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard.}}
@@
 \subsection{Escape Clauses with Strong Bindings}
+\label{sec:strong}
 Suppose that political pressure derives only from endogenous sources but an escape clause is to be implemented anyway.

### 2016-12-10 to-do.tex
add thanks to draft; integrate to-dos from read through

@@
 	\begin{itemize}
-		\item Rick Bond: need lobbying to use up resources: if it changes net profits, it should also change $\pi$ in $W(\tau,\tau^*)$
 		\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
@@
 			\end{itemize}
-		\item Try subtracting $e$ and see what it does to the graphs. Does it give me the kind of non-monotonicity I'd like? YES
+		\item Subtracting $e$ gives me the kind of non-monotonicity I'd like in the graphs.
 			\begin{itemize}
-				\item Does it do anything to the analysis?
+				\item Does it do anything to the analysis? Just shifts govt's ideal point
 			\end{itemize}
@@
 I have refrained from adding the ``$-e$'' to the welfare function out of a desire to be consistent with the existing literature.
+		\item I think I'd like to just change to the ``$-e$'' version.
+			\begin{itemize}
+				\item Worries: won't match literature. Want to make analogies to GH, DGH, MRC, LT.
+			\end{itemize}
 	\end{itemize}
@@
 					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated.
+					\item I say I'm going to do this on Page 27. Need to decide and either do it or take this out.
 				\end{itemize}
@@
 How does your approach relate to costly state verification?
+				\item I say that I'm going to provide a result along the lines of Beshkar 2010b. Same or different from costly state verification? Which is more helpful?
 		\end{itemize}
@@
 	\item Clarify information structure
-	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together better
+	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together
 	\item Add ref to JIE R$\&$R to fn:krw ?
@@
 	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
+	\item Streamline notation. If not doing strong binding anywhere (except in sec:strong), don't need $R$ subscript on $\tau$ and $e$.
 \end{itemize}
@@
 
+\newpage
+\begin{center}
+\textbf{\large{For Later}}
+\end{center}
+\begin{enumerate}
+	\item Add export lobby. 
+		\begin{itemize}
+			\item Extension or appendix b/c want to make comparison to existing literature.
+		\end{itemize}
+\end{enumerate}
 \end{document}

### 2016-12-31 to-do.tex
add new work on GH94/MRC07 to to-do

@@
 					\end{itemize}
+			\end{itemize}
+		\item Comparison to GH94
+			\begin{itemize}
+				\item It's the truthful contribution schedule in GH94 that implies the constant weighting of $1+a$ for sectors represented by a lobby.
+Won't get this clean, constant weighting with other schedules.
+				\item Gov't doesn't care about \textit{anything} because it gets the same payoff no matter what
+					\begin{itemize}
+						\item With one lobby, this lobby-has-all-the-bargaining-power feature is an equilibrium implication of the menu-auction set-up where the lobby makes a TIOLI offer
+						\item Gov't gets all the surplus when there are two or more lobbies
+						\item This result holds in DGH97, p. 767. Footnote 10 explicitly says that this is because lobby makes TIOLI offer.
+						\item So go to MRC07 where bargaining power can be manipulated
+					\end{itemize}
+				\item Menu auction eqm requirement: $(C(\tau),\tau)$ must maximize joint surplus
+					\begin{itemize}
+						\item NB: $\tau$ that maximizes gov't welfare doesn't generically maximize joint surplus (unless it also maximizes lobby payoff)
+						\item Careful of transferable vs. non-transferable utility
+					\end{itemize}
+				\item Not sure if contingent vs. non-contingent payments are a big deal
+			\end{itemize}
+		\item MRC07
+			\begin{itemize}
+				\item Need case of fixed capital ($z=0$)
+				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
+				\item Shift all bargaining power to gov't. They do this in WP version. Read to see what their result is for $z=0$.
+					\begin{itemize}
+						\item Should have gov't make lobby indifferent and maximize joint surplus. But again, this doesn't necessarily maximize gov't welfare
+					\end{itemize}
+				\item Then re-do MRC07 with $W(\tau) + g(e)$, and version where I subtract $e$
 			\end{itemize}

### 2016-12-31 game-tree.tex
add dashed line in game-tree for simultaneous choice
add language to game-tree so it will switch to draft.tex to compile
fix order of moves in stage game extensive form

@@
+% !TEX root = ../draft.tex
 \begin{comment}
@@
 child{[white] node(2)[solid node,xshift=12,label=above:{L}]{}  
-	child{[black] node(4)[solid node,label=below:{$0$}]{}}
+	child{[black] node(4)[solid node,label=below:{$-\infty$}]{}}
 	child{[white] node(5)[hollow node,xshift=12]{} 
 	 %child[sibling distance=11mm]{[black] node(22)[solid node]{}  }
-	 child{[black] node(22)[solid node,,label=below:{$-\infty$}]{}  }
+	 child{[black] node(22)[solid node,,label=below:{$0$}]{}  }
 	 child{[white] node(24)[solid node,xshift=12]{}
@@
 		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
-		edge from parent node[black, xshift=20]{$s$} 	
+		edge from parent node[black, xshift=20]{$e$} 	
 		}
 	 child{[black] node(23)[solid node,label=above:{$\infty$}]{}  }
-	 edge from parent node[black, xshift=20]{$e$} 	
+	 edge from parent node[black, xshift=20]{$s$} 	
 	 }
@@
 		\draw[solid,bend right](4)to(6);
+		\draw[dashed,bend right]([xshift=2]7.center)to([xshift=2]9.center);
 		\draw[solid,bend right](7)to(9);
@@
 % labels
-		\node[above,xshift=3,yshift=5]at(2){$L$}; 
+		\node[above,xshift=3,yshift=5]at(2){$N$}; 
 		\node[above,xshift=3,yshift=5]at(24){$G$}; 
-		\node[above,xshift=3,yshift=5]at(5){$N$}; 
+		\node[above,xshift=3,yshift=5]at(5){$L$}; 
 		\node[above,xshift=6,yshift=5]at(8){$G^*$}; 

### 2016-12-31 draft.tex
fix order of moves in stage game extensive form
remove ga decr in tau sec from to-do; update 'thanks' in draft

@@
 \begin{document}
-\title{Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Chad Bown, Nuno Limao, Giovanni Maggi, Anson Soderbery, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at SITE for very helpful comments and discussions.}}
+\title{Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Renee Bowen, Chad Bown, Paola Conconi, David DeRemer, James Lake, Nuno Limao, Giovanni Maggi, Devashish Mitra, Peter Rosendorff, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at McGill, Rochester, Ryerson, Wesleyan, ETSG, InsTED, the Midwest International Trade Meetings, PEIO and SITE for very helpful comments and discussions.}}
 \author{Kristy Buzard\thanks{Syracuse University, Economics Department. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard.}}
@@
 
-The stage-game timing is as follows.
+The stage-game timing is shown in Figure~\ref{fig:ext}.
 First, the governments cooperatively form a trade agreement.\footnote{See Section~\ref{sec:rigid} for a discussion of why lobbies are assumed not to be involved at the trade agreement formation stage.} After the trade agreement is concluded, any exogenous shock is realized.
@@
 	\end{center}
-	\caption{Extensive Form\label{fig:ext}}
+	\caption{Extensive Form of the Stage Game\label{fig:ext}}
 \end{figure}
@@
 The stark prediction that the best the government can do is to set the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has an interior global maximum in $\ga$.
-A candidate for the government objective function in which the maximum is strictly interior is one in which returns to lobbying decrease as tariffs increase as in \Textcite{ethier2012}.
-In the current model, one would take $\ga$ to be a decreasing function of $\tau$.
-This implies that government welfare must eventually decrease in $\ga$, providing for the required concavity in $\ga$.
+There are many candidate for the government objective function in which the maximum is strictly interior.
+For example, in \Textcite{ethier2012}, returns to lobbying decrease as tariffs increase.
+In the current model, one could take government welfare to be quadratic in $\ga$ or to simply subtract lobbying effort from Equation~\ref{eq:wel}.
 

### 2016-12-31 to-do.tex
remove ga decr in tau sec from to-do; update 'thanks' in draft

@@
 
-\section{Make a bigger deal of the idea of \texorpdfstring{$\gamma$}{gamma} decreasing function of tariffs---move out of footnote}
-Ethier 2012 for protection decreasing function of $\tau$ [NEED TO REVISIT IN TEXT]. ``The Political-Support Approach to Protection,'' Global Journal of Economics
-\begin{itemize}
-	\item Making $W$ quadratic in $\ga$ such as $\ga - \ga^2$ gives a nice interior max
-	\item I think what I have already has returns to $e$ decreasing in $\tau$. Don't I show it? If not in this paper, in RoIE paper.
-	\item What embodies returns to lobbying?
-		\begin{itemize}
-			\item $e \rightarrow \ga(e)$ is how lobbying is translated into weight in political process. It's not `returns.'
-			\item $\tau$ isn't even returns. $\pi(\tau(e_2)) - \pi(\tau(e_1))$ is returns.
-		\end{itemize}
-	\item To get returns decreasing in $\tau$, maybe weighting can't be confined to $PS$
-\end{itemize}
-
 \section{Lobby's optimal effort}

### 2017-01-01 presentation.tex
move repeated game intuition slide to appendix
reverse order of conclusion slides
move lit review to obj fcn section
first compile of presentation for ASSAs, recompile to-do

@@
 
+\usepackage{appendixnumberbeamer}
+
 \title[Endogenous Politics and the Design of Trade Institutions\hspace{2.05in}\insertframenumber/\inserttotalframenumber]{Endogenous Politics and \\ the Design of Trade Institutions}
 %\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
-\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
-\date{June 30, 2016}
+\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
+\date{January 6, 2017}
 \begin{document}
@@
 \end{comment}
-
-\subsection{Role and Design of TAs}
-\begin{frame}{Design of Trade Agreements}
-\pause
-\begin{itemize}[<+->]
-	\item \textbf{Tariff caps}: Bagwell and Staiger 2005, Horn et al 2010, Amador and Bagwell 2012; Beshkar and Bond 2012
-	\item \textbf{Escape clause}: Bagwell and Staiger 2005, Horn et al 2010, 
-	\item \textbf{Shallow vs. deep integration}: Bagwell and Staiger 2001, DeRemer 2014
-	\item \textbf{Dispute settlement}: Maggi 1999, Ludema 2001, Maggi and Staiger 2011/2013, Klimenko et al 2008
-	\item \textbf{Property vs. liability rules}: Pauwelyn 2008, Beshkar 2010, Maggi and Staiger 2014
-	\item \textbf{Retaliation}: Bown 2002/2004, Beshkar 2010
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Role of Trade Agreements: TOT Externality}
-\pause
-Bagwell and Staiger (2002)
-\pause
-\begin{itemize}[<+->]
-	\item Joint social welfare maximized at free trade
-	\item Trade war (i.e. no agreement)
-		\begin{itemize}
-			\item Maximize with respect to home country welfare only
-			\item Terms of trade (TOT) externality $\Rightarrow$ positive tariffs  
-		\end{itemize}
-	\item Trade agreements
-    \begin{itemize}
-			\item Now take into account impact on foreign welfare
-			\item Internalize TOT externality $\Rightarrow$ free trade
-    \end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Role of Trade Agreements: TOT Externality}
-\pause
-Grossman and Helpman (1995)
-\pause
-\begin{itemize}[<+->]
-	\item Add endogenous politics
-	\item Now in ``Trade War'': two reasons for positive tariff
-		\begin{itemize}
-			\item TOT externality + pressure from import competing lobby
-		\end{itemize}
-
-	\item Trade agreement: only internalizes TOT externality
-\end{itemize}
-\end{frame}
-
-\begin{comment} 
-\begin{frame}{Role of Trade Agreements: New Trade Theory}
-\pause
-``New Trade Theory'' externalities
-\pause
-\begin{itemize}[<+->]
-	\item Source: imperfect competition (firm level)
-		\begin{itemize}
-			\item Delocation: Ossa 2011, Bagwell and Staiger 2012
-			\item Profit Shifting: Mrazova 2011, Ossa 2012  
-		\end{itemize}
-	\item Trade agreements
-    \begin{itemize}
-			\item Now take into account impact on foreign welfare
-			\item Internalize these new externalities
-    \end{itemize}
-\end{itemize}
-\end{frame} 
-\end{comment}
-
-\begin{frame}{Role of Trade Agreements: Domestic Commitment}
-\pause
-\begin{itemize}[<+->]
-	\item Maggi and Rodriguez-Clare (1998, 2007)
-		\begin{itemize}
-			\item Allow for (imperfect) capital mobility
-			\item Domestic investment decisions depend on level of protection
-			\item Inability to commit $\Rightarrow$ investment too high b/c importers know protection will respond
-			\item Trade agreements provide commitment device
-    \end{itemize}
-	\item Mitra (2002)
-		\begin{itemize}
-			\item Here distortion is wasted resources in lobby formation
-		\end{itemize}
-\end{itemize}
-\end{frame}
 
@@
 \section{Objective Fcn}
+
+\subsection{Role and Design of TAs}
+\begin{frame}{Design of Trade Agreements}
+\pause
+\begin{itemize}[<+->]
+	\item \textbf{Tariff caps}: Bagwell and Staiger 2005, Horn et al 2010, Amador and Bagwell 2012; Beshkar and Bond 2012
+	\item \textbf{Escape clause}: Bagwell and Staiger 2005, Horn et al 2010, 
+	\item \textbf{Shallow vs. deep integration}: Bagwell and Staiger 2001, DeRemer 2014
+	\item \textbf{Dispute settlement}: Maggi 1999, Ludema 2001, Maggi and Staiger 2011/2013, Klimenko et al 2008
+	\item \textbf{Property vs. liability rules}: Pauwelyn 2008, Beshkar 2010, Maggi and Staiger 2014
+	\item \textbf{Retaliation}: Bown 2002/2004, Beshkar 2010
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Role of Trade Agreements: TOT Externality}
+\pause
+Bagwell and Staiger (2002)
+\pause
+\begin{itemize}[<+->]
+	\item Joint social welfare maximized at free trade
+	\item Trade war (i.e. no agreement)
+		\begin{itemize}
+			\item Maximize with respect to home country welfare only
+			\item Terms of trade (TOT) externality $\Rightarrow$ positive tariffs  
+		\end{itemize}
+	\item Trade agreements
+    \begin{itemize}
+			\item Now take into account impact on foreign welfare
+			\item Internalize TOT externality $\Rightarrow$ free trade
+    \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Role of Trade Agreements: TOT Externality}
+\pause
+Grossman and Helpman (1995)
+\pause
+\begin{itemize}[<+->]
+	\item Add endogenous politics
+	\item Now in ``Trade War'': two reasons for positive tariff
+		\begin{itemize}
+			\item TOT externality + pressure from import competing lobby
+		\end{itemize}
+
+	\item Trade agreement: only internalizes TOT externality
+\end{itemize}
+\end{frame}
+
+\begin{comment} 
+\begin{frame}{Role of Trade Agreements: New Trade Theory}
+\pause
+``New Trade Theory'' externalities
+\pause
+\begin{itemize}[<+->]
+	\item Source: imperfect competition (firm level)
+		\begin{itemize}
+			\item Delocation: Ossa 2011, Bagwell and Staiger 2012
+			\item Profit Shifting: Mrazova 2011, Ossa 2012  
+		\end{itemize}
+	\item Trade agreements
+    \begin{itemize}
+			\item Now take into account impact on foreign welfare
+			\item Internalize these new externalities
+    \end{itemize}
+\end{itemize}
+\end{frame} 
+\end{comment}
+
+\begin{frame}{Role of Trade Agreements: Domestic Commitment}
+\pause
+\begin{itemize}[<+->]
+	\item Maggi and Rodriguez-Clare (1998, 2007)
+		\begin{itemize}
+			\item Allow for (imperfect) capital mobility
+			\item Domestic investment decisions depend on level of protection
+			\item Inability to commit $\Rightarrow$ investment too high b/c importers know protection will respond
+			\item Trade agreements provide commitment device
+    \end{itemize}
+	\item Mitra (2002)
+		\begin{itemize}
+			\item Here distortion is wasted resources in lobby formation
+		\end{itemize}
+\end{itemize}
+\end{frame}
+
 \subsection{Objective Function}
@@
 	\begin{itemize}
-		\item Note `Protection for Sale' objective function is isomorphic
+		\item Isomorphic to `Protection for Sale' objective function
 	\end{itemize}
@@
 
-\begin{frame}{Tariff Caps with Self Enforcement}
+\begin{frame}[label=self]
+\frametitle{Tariff Caps with Self Enforcement}
 \pause
@@
 \end{itemize}
-
+\pause
+\hyperlink{repeated<1>}{\beamergotobutton{Repeated Game Intuition}}
 \end{frame}
@@
 \end{comment}
-
-\begin{frame}{Repeated Game Intuition}
-\pause
-Legislature: break agreement if punishment not strong enough
-\pause
-\begin{itemize}
-	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
-\end{itemize}
-
-\vskip.1in
-\pause
-Lobby: solve for lowest effort ($\ov{e}_b$) that breaks this constraint
-\pause
-\begin{itemize}
-	\item pay $\ov{e}_b$ if it's less than gain from $T$ periods of trade-war profits
-\end{itemize}
-
-\pause
-\vskip.1in
-Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
-\pause
-\begin{itemize}
-	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
-	\pause
-	\item High tariffs, no lobbying, no trade disruptions
-\end{itemize}
-\end{frame}
 
@@
 
+\begin{frame}{Conclusion}
+Taking into account endogenous political forces alongside exogenous ones...
+\pause
+\begin{itemize}[<+->]
+		\item helps explain the structure and enforcement of the WTO Safeguards measure
+		\item can help us think about optimal design of trading insitutions
+		\item demonstrates that TAs can be used to discourage lobbing activity in general
+		\item provides additional general explanation for tariff caps
+\end{itemize}
+
+\end{frame}
+
 \begin{frame}{Future Work}
@@
 
-\begin{frame}{Conclusion}
-Taking into account endogenous political forces alongside exogenous ones...
+\appendix
+\begin{frame}[label=repeated]
+\frametitle{Repeated Game Intuition}
 \pause
-\begin{itemize}[<+->]
-		\item provides additional general explanation for tariff caps
-		\item demonstrates that TAs can be used to discourage lobbing activity in general
-		\item helps explain the structure and enforcement of the WTO Safeguards measure
+Legislature: break agreement if punishment not strong enough
+\pause
+\begin{itemize}
+	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
 \end{itemize}
 
+\vskip.1in
+\pause
+Lobby: solve for lowest effort ($\ov{e}_b$) that breaks this constraint
+\pause
+\begin{itemize}
+	\item pay $\ov{e}_b$ if it's less than gain from $T$ periods of trade-war profits
+\end{itemize}
+
+\pause
+\vskip.1in
+Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
+\pause
+\begin{itemize}
+	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
+	\pause
+	\item High tariffs, no lobbying, no trade disruptions
+\end{itemize}
+\hyperlink{self<1>}{\beamergotobutton{Go Back}}
 \end{frame}

### 2017-01-02 to-do.tex
add notes on joint surplus in my model

@@
 						\item NB: $\tau$ that maximizes gov't welfare doesn't generically maximize joint surplus (unless it also maximizes lobby payoff)
+						\item In my model, optimum for lobby would be prohibitive tariff, $e=0$.
+Joint optimum determined by $\frac{\partial \ga}{\partial e}\pi_X = 2$ and $\frac{\partial CS_X}{\partial \tau} + \left[ 1 + \ga(e)\right]\frac{\partial \pi_X}{\partial \tau} + \frac{\partial TR}{\partial \tau} = 0$.
+Not clear how this is related to lobby's problem, but likely to have higher $\tau$ and lower $e$ than government's optimum (unless $\ga$ changes very steeply between the two points $e_J$ and $e_G$)
 						\item Careful of transferable vs. non-transferable utility
@@
 				\item Not sure if contingent vs. non-contingent payments are a big deal
+				\item If I change nothing else but subtract $C$ from gov't welfare function, no qualitative change b/c of 0 bargaining weight
 			\end{itemize}

### 2017-01-02 to-do.tex
add old misc notes; MRC WP result and next steps

@@
 In the non-unitary legislature, if the current period incentive constraint includes any future value of lobbying effort, this would mean the current-period median legislator evaluating her current-period incentive constraint using a mixture of her own political-economy weight with those of the legislators who are median in the future in the trade agreement and trade-war scenarios.
-The first major step I have taken to clarify the incentive constraint is the analysis in Appendix B.1 of the unitary model (described above), which I would guess is the model you had in mind.
+The first major step I have taken to clarify the incentive constraint is the analysis in Appendix B.1 of the unitary model.
 I discuss the differences between the unitary and non-unitary models in the beginning of that Appendix.''
@@
 It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.
-Having done that, you could bring in ex-ante lobbying and treat it the same way.
-		\item In general it seems to me that you seem to "dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
+		\item In general it seems to me that you seem to ``dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
 		\item DGH / MRC analogy from JIE$\_$revision
@@
 			\begin{itemize}
+				\item Contingent payments are hard to reconcile with uncertainty
 				\item It's the truthful contribution schedule in GH94 that implies the constant weighting of $1+a$ for sectors represented by a lobby.
@@
 				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
-				\item Shift all bargaining power to gov't. They do this in WP version. Read to see what their result is for $z=0$.
+					\begin{itemize}
+						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period).
+Here, could be free trade but could be interior to get some rents.
+					\end{itemize}
+				\item Then re-do MRC07 with $W(\tau) + g(e)$, and version where I subtract $e$
 					\begin{itemize}
 						\item Should have gov't make lobby indifferent and maximize joint surplus. But again, this doesn't necessarily maximize gov't welfare
+						\item Or rather, should I follow Limao and Tovar?
 					\end{itemize}
-				\item Then re-do MRC07 with $W(\tau) + g(e)$, and version where I subtract $e$
 			\end{itemize}
@@
 Do I need to add a numeraire so I can put $e$ in terms of it?
+\begin{itemize}
+	\item Does DGH97 or similar have it?
+\end{itemize}
 
@@
 I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
-					\item Then the paper begins to address the equally important question of what can be done about this problem, but again stops short of providing a really compelling answer.
-I think it would be worth exploring this in detail.
-The paper does get to the \textbf{heart of the issue, which is that `s' can be observed by the DSB but e cannot}.
-Get to the bottom of how to design a dispute settlement system that resolves the conflict created by `e vs s'.
 				\end{itemize}
@@
 				\end{itemize}
-			\item Need to go to continuous value of EC tariff
-				\begin{itemize}
-					\item There will be optimal tariff for whatever value of $s$ is realized
-						\begin{itemize}
-							\item Not sure how fine I want to go with $s$
-						\end{itemize}
-					\item Lobby will choose its optimal level of $e$. Will this be constrained in any way? Probably not, if there isn't a fixed EC-binding.
-				\end{itemize}
-			\item After $s$ realized, $e$ choosen by lobby. Then government can choose $\tau$
-				\begin{enumerate}
-					\item consistent with $s$, no retaliation
-					\item consistent with $s$ and $e$ and be retaliated against according to WEC (need to work WEC into model--repeated game incentives necessary)
-						\begin{itemize}
-							\item Is responding to $e$ in some cases worth it?
 			\item Does WEC help or hurt vs. grim trigger, $T$ period Nash reversion?
-						\end{itemize}
-				\end{enumerate}
 		\end{itemize}
@@
 				\item I say that I'm going to provide a result along the lines of Beshkar 2010b. Same or different from costly state verification? Which is more helpful?
+				\item Need to check this for sure in new section. I don't have cost to verification.
 		\end{itemize}

### 2017-01-04 presentation.tex
remove lit review from presentation for ASSAs
remove tariff cap slides and references for ASSAs
remove superfluous notes and combine others in to-do

@@
 	\vskip.1in
-\item Can trade agreements be used to manipulate domestic lobbying incentives?
+\item When do governments want to use trade agreements to manipulate domestic lobbying incentives?
 	\begin{itemize}
@@
 	\item endogenize politics into a standard model for studying TA design questions
-	\item carefully distinguish between dynamics induced by exogenous and endogenous politics for
-		\begin{itemize}[<+->]
-			\item base case with tariff caps
-			\item tariff caps with escape clause
-		\end{itemize}
+	\item carefully distinguish between dynamics induced by exogenous and endogenous politics for tariff caps with escape clause
+		%\begin{itemize}[<+->]
+			%\item base case with tariff caps
+			%\item 
+		%\end{itemize}
 	\item examine escape clause design when both exogenous and endogenous forces are present
@@
 	\item Show that TAs may be used to manipulate domestic political actors (no long-run distortions)
-	\item For both tariff caps and escape clauses, outcomes are very different with endogenous politics
+	\item Escape clause outcomes are very different with endogenous politics
 	\item Demonstrate that (standard, theoretical) escape clause can't work in the presence of endogenous political pressure
@@
 Two countries: home and foreign (${}^*$)
-\pause
-\begin{itemize}[<+->]
+%\pause
+\begin{itemize}
 	\item Separable in two goods: $X$ and $Y$
@@
 \[
-  W = \mathit{CS}_X(\tau) + \ga(s,e) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
+  W = \mathit{CS}_X(\tau) + \ga(s,e) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau) - e
 \]
@@
 \begin{itemize}[<+->]
-	\item Standard \textit{except} weight on import-competing industry profits $\ga(s,e)$: 
+	\item Standard \textit{except} for $s$ and $e$: 
 		\begin{itemize}
@@
 		\end{itemize}
-	\item Assume $\ga$, $\ga^*$ is private info of each government
+	%\item Assume $\ga$, $\ga^*$ is private info of each government
 \end{itemize}
@@
 
+\begin{comment}
 \subsection{Role and Design of TAs}
@@
 \end{frame}
-\end{comment}
 
@@
 
-	\item If weights must sum to 1, welfare no longer monotonic in $\ga$
+	\item When subtracting lobbying effort, welfare no longer monotonic in $\ga$
 \end{itemize}
@@
 
-\begin{frame}
-\includegraphics[height=2.75in, width=4.25in]{weight.jpg}
-\end{frame}
+%\begin{frame}
+%\includegraphics[height=2.75in, width=4.25in]{weight.jpg}
+%\end{frame}
 
+\begin{comment}
 \section{Tariff Caps}
 \subsection{Tariff Caps}
-\begin{comment}
+
 \begin{frame}{Tariff Caps: Exogenous vs. Endogenous $\ga$}
@@
 
+\begin{comment}
 {\begin{frame}{Tariff Caps: Exogenous vs. Endogenous $\ga$}
@@
 \end{frame}
+\end{comment}
 
@@
 	\item Can improve self-enforcement
-	\item Incentive compatibility becomes an issue
+	%\item Incentive compatibility becomes an issue
 \end{itemize}
@@
 
+\begin{comment}
 \begin{frame}{Incentive compatibility}
@@
 	\item Assuming lower binding is set to maximize political welfare, escape clause encourages inefficiently high lobbying effort / protection
-	\item (Incentive compatibility still an issue, but often not the central one)
-		\begin{itemize}[<+->]
-			\item (If lobby's preferred tariff $\geq$ escape clause binding, gov't experiences high $\ga$, no need to lie)
-		\end{itemize}
+	%\item (Incentive compatibility still an issue, but often not the central one)
+		%\begin{itemize}[<+->]
+			%\item (If lobby's preferred tariff $\geq$ escape clause binding, gov't experiences high $\ga$, no need to lie)
+		%\end{itemize}
   \end{itemize}
@@
 		\item can help us think about optimal design of trading insitutions
-		\item demonstrates that TAs can be used to discourage lobbing activity in general
-		\item provides additional general explanation for tariff caps
+		%\item demonstrates that TAs can be used to discourage lobbing activity in general
+		%\item provides additional general explanation for tariff caps
 \end{itemize}
@@
 
+\begin{comment}
 \appendix

### 2017-01-04 to-do.tex
remove superfluous notes and combine others in to-do

@@
 	\begin{itemize}
-					\item Gul and Pesendorfer's (2001) Econometrica is a two-period model. No dynamic inconsistency; derive temptation and desire for commitment in a model with dynamically consistent preferences.
-					\item Even better, especially advertising example (credit Kyle): Becker, Gary.
-S.
-and George J.
-Stigler, ``De Gustibus Non Est Disputandum,'' The American Economic Review, Vol.
-67, No.
-2 (Mar., 1977), pp.
-76-90
-					\item Simpler: Becker, G. S. and K. M. Murphy (1993), Simple Theory of Advertising as a Good or Bad, Quarterly Journal of Economics, 942-64.
+		\item Becker, G. S. and K. M. Murphy (1993), Simple Theory of Advertising as a Good or Bad, Quarterly Journal of Economics, 942-64.
 			\begin{itemize}
@@
 behavior of another actor).
-But $CS + \ga(e)PS + \ldots$, just like $u(X,Y,A)$; says consumer could work to pass law about number of adds on TV each hour.
+But $CS + \ga(e)PS + \ldots$, just like $u(X,Y,A)$; says consumer could work to pass law about number of ads on TV each hour.
 				\item Use newly drawn extensive form to explain this, lays out when welfare is realized
-							\item Addresses James' concern about needing to be more precise about when $\ga$ is evaluated)
+				\item Addresses James' concern about needing to be more precise about when $\ga$ is evaluated
 					\begin{itemize}
@@
 					\end{itemize}
-
 			\end{itemize}
@@
 But legislator $e_b$, who is the decision maker in the current period, maximizes her own welfare given the predicted behavior of future decision makers.''
-
 			\end{itemize}
@@
 			\end{itemize}
-		\item JIE$\_$revision: I have also verified that \emph{none} of the results of the model change if lobbying effort is subtracted from the legislative/executive welfare function (except, of course, the evaluated level of welfare).
-I have refrained from adding the ``$-e$'' to the welfare function out of a desire to be consistent with the existing literature.
 		\item I think I'd like to just change to the ``$-e$'' version.
@@
 				\item Footnote fn:mrc1 is good place to start to put references to MRC 2007
-				\item I've supplied the suggested results to connect my model to that of Dixit, Grossman and Helpman (1997) in Appendix B.2, along with some text pointing to it at the end of Section 2.2 on page 10.
-The results from the DGH-style model are qualitatively different in a minor way.
-I discuss this difference in Appendix B.1 where I present a unitary government/legislature version of the model that most closely matches Maggi and Rodriguez-Clare 2007.
-				\item The model does admit an interpretation in which the same branch of government both negotiates the trade agreement and decides on the applied tariff ex-post.
-In this sense, the structure of the one-shot game shares much in common with Maggi and Rodriguez-Clare 2007.
-				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007.
-Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored.
-In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps.
-The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff.
-There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff.
-Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement.
+				\item The model does admit an interpretation in which the same branch of government both negotiates the trade agreement and decides on the applied tariff ex-post, as long as the one branch is non-unitary.
+The model analyzed in the body of this paper can be given an interpretation in line with Maggi and Rodriguez-Clare 2007 if capital is completely mobile in the long run.
+In Appendix B.1, I compare the results of this paper to those from a model with a unitary government, as this would seem to bring the model most closely into alignment with Maggi and Rodriguez-Clare 2007 and a large part of the literature.
 				\item Consider the following, alternative interpretation of the model.
@@
 Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
-				\item The model admits this `one-branch' interpretation only if the one branch is non-unitary
-				\item the model analyzed in the body of this paper can be given an interpretation in line with Maggi and Rodriguez-Clare 2007 if capital is completely mobile in the long run and the single decision-making body is non-unitary.
-In Appendix B.1, I compare the results of this paper to those from a model with a unitary government, as this would seem to bring the model most closely into alignment with Maggi and Rodriguez-Clare 2007 and a large part of the literature.
-Then, in Appendix B.2, since the government welfare function used in this paper can be interpreted as a special case of the one proposed by DGH97, I will show that a model patterned after DGH97 and specialized to this environment---which is also a unitary model---provides results that are qualitatively similar to the unitary model.
+				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007.
+Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored.
+In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps.
+The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff.
+There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff.
+Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement.
+A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement.
+From the ex-ante point of view, strong bindings are therefore preferable.
+Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
+
 				\item I have added Appendix B.1 with analysis of the unitary model
@@
 They specify government welfare as $G(a,c)$ where $a$ is the policy vector $\left(\left(\tau,\tau^*\right)\right)$ and $c$ is a vector of payments from each lobby ($e$).
-						\item Here I examine a version of the DGH97 model specialized to this environment.
-Because the DGH97 model is fundamentally a unitary government model, the relevant benchmark is the unitary version of the model.
-The results of the two unitary models are qualitatively the same.
+
 						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
@@
 Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
-						\item In the auction-menu set-up of the DGH97 model, there does not seem to be a natural alternative to the unitary government interpretation.
-The lobby offers a contribution schedule of effort levels and tariffs and the government chooses among these pairs.
-In the non-unitary model, each different effort level determines a different government, so the government cannot choose between elements of a schedule.
-To analyze this model, I make the standard assumptions of a unitary government and that the lobby has all the bargaining power.
+
 						\item Denote the government's welfare function as $W_G(\tau,\tau^*,e(\tau)) = W(\tau,\tau^*) + g(e(\tau))$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare. Then do analysis...
-					\end{itemize}
-				\item Appendix B.3 with analysis of the strong binding case. 
-					\begin{itemize}
-						\item If the trade agreement involves strong bindings instead of tariff caps (i.e.
-weak bindings), the legislature must deliver $\tau^a$ and the lobby's optimal effort during a period in which the trade agreement holds is $e_a = 0$.
-That the lobby pays less---here, nothing---for the protection it receives under the trade agreement is the only change from the base model.
-This means that nothing changes in the trade war.
-Likewise the legislature's repeated-game incentive constraint is unchanged.
-But the lobby's incentive constraint IS changed...
-						\item As in MRC, tariff caps in this model serve to keep the lobby `in the game.' A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement.
-From the ex-ante point of view, strong bindings are therefore preferable.
-Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
 					\end{itemize}
@@
 	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
-	 contributions 'felonious'
+	 contributions `felonious'
 	\item David DeRemer: Safeguards have dropped off in use around the world
@@
 		\begin{itemize}
-			\item What it's missing: intersection with reality
-				\begin{itemize}
-					\item Anecdotes
-					\item How policy is used
-						\begin{itemize}
-							\item I thought that escape clause implications \textit{were} this
-						\end{itemize}
-					\item Something like how GH predicts export taxes but it doesn't work that way in the real world
-					\item Well articulated question / paradox
-				\end{itemize}
-			\item Look at ``Theory of Managed Trade'' model motivation (BS)
+			\item What it's missing: Well articulated question / paradox (I think it's just not well-articulated)
 			\item Intro doesn't reflect body of paper
@@
 			\item You need to think a bit more about the objective(s) of the paper and how best to achieve them.
-			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics
+			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics.
 		\end{itemize}

### 2017-01-09 to-do.tex
add notes from Wouters & Zissimos (2016)
outline approach to introducing my obj fcn in end of sec 3

@@
 It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.
-		\item In general it seems to me that you seem to ``dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
 		\item DGH / MRC analogy from JIE$\_$revision
@@
 Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
-
-						\item Denote the government's welfare function as $W_G(\tau,\tau^*,e(\tau)) = W(\tau,\tau^*) + g(e(\tau))$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare. Then do analysis...
 					\end{itemize}
@@
 			\end{itemize}
+		\item Approach: denote the government's welfare function as $W_G(\tau,\tau^*,e) = W(\tau,\tau^*) + g(e)$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare.
+			\begin{itemize}
+				\item This is like DGH97 (they write as $G(a,C)$, so no problem putting an $e$ in there), BS1999 `Economic Theory of GATT' (I think), $W(p,\tilde{p}^w)$
+					\begin{itemize}
+						\item I've already written $W(\ga(e,s),\tau,\tau^*)$
+						\item I think people are most concerned with underlying economic model. I need a numeraire anyway. DGH97 doesn't have numeraire b/c small country
+					\end{itemize}
+				\item Then specialize to BS2005 when necessary (tariff cap example, picture for interior max of gov't obj fcn)
+				\item Menu auction framework doesn't work with one lobby because gov't doesn't get any surplus. Switch to MRC07 (bargaining). But want decreasing returns (LT).
+				\item Make argument that want this version like Baldwin-style literature for comparison
+					\begin{itemize}
+						\item Want to make argument that things go wrong in standard model when you insert $e$ into $\ga$ (also, when $e$ goes into $\ga$ need to subtract $e$ too, and this often results in interior optimum)
+						\item This is an easier way to get the result from MRC07 / LT, can make comparisons, develop design results more easily
+					\end{itemize}
+				\item Want non-contingent set-up for different assumptions on timing / meaning of $s$; contingent only makes sense if $e$ chosen after $s$ and $s$ is completely verifiable to lobby.
+If $s$ is political uncertainty, it probably realizes after/contemporaneously with $e$
+					\begin{itemize}
+						\item I can leave general, then specify for escape clause. Or I can just go with my order of $s$ then $e$, which assumes $s$ is an economic shock.
+					\end{itemize}
+				\item Bargaining framework is not appealing when gov't is non-unitary
+				\item Limao $\&$ Tovar: lobby objective the same as mine.
+$G = \Psi(C) + aW(p)$.
+Gov't decides whether to commit in int'l agreement, then Nash bargains with lobby over $C$ and policies $(t,\tau)$ where $\tau$ is a NTB
+					\begin{itemize}
+						\item Section 3.3: They show exactly what I need to show.
+If gov't can commit with this obj fcn, it will as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying.
+						\item Gov't does not have an interest in maximizing joint surplus, because it doesn't appropriate all the gains
+					\end{itemize}
+				\item Only difference between Limao / Tovar and mine is $\Psi(e)$ vs. $[\ga(e) - 1]\pi_x(\tau)$
+					\begin{itemize}
+						\item This means there's a positive cross partial between $e$ and $\tau$ in my formulation, so that an incraese in $\tau$ increases $W$ more the higher is $e$
+					\end{itemize}
+			\end{itemize}
 	\end{itemize}
@@
 			\item More on Coates $\&$ Ludema (2001)
+			\item Check out Ben and James new paper
 		\end{itemize}
@@
 			\item Look at U.S. vs. ROW (he sent papers)
+			\item Wouters and Zissimos (2016 WP): Appellate Body failed to give clear direction in cases in early 2000s as to when Safeguards can be applied. See Sykes (2003).
 		\end{itemize}
@@
 
-\section{Add empirics / evidence?}
-
 \section{Existence proofs?}

### 2017-01-09 draft.tex
clarify when gov't prefers to reduce e is sec:perfect

@@
 
-A second motive is operative here that is absent in the case of exogenous political pressure: the trade agreement can act as a domestic political commitment device, allowing the home government to manipulate ex-post lobbying incentives.
-In fact, whenever the cap is strictly below the non-cooperative level, that is when $\tau^R_{W,e} < \tau^N$, the weak binding serves to restrain political pressure whether or not the governments desire it.
+A second motive may be operative here that is absent in the case of exogenous political pressure.
+When the government's optimum in terms of $\ga(e)$ is lower than the lobby's, the government will use the trade agreement as a domestic political commitment device to manipulate ex-post lobbying incentives.
+Whenever the cap is strictly below the non-cooperative level, that is when $\tau^R_{W,e} < \tau^N$, the weak binding serves to restrain political pressure.
 

### 2017-01-11 to-do.tex
remove to-do item about when gov't prefers to reduce e
add notes on changes from subtracting e to to-do

@@
 				\item Worries: won't match literature. Want to make analogies to GH, DGH, MRC, LT.
+			\end{itemize}
+		\item Note: there's no ``$e$'' to subtract when $\ga$ is exogenous
+		\item Changes from version without $-e$:
+			\begin{itemize}
+				\item 3.1: $\tau_{W,e}^R$ reduced because gov't optimum is lower. $\tau^N(\ga(e))$ doesn't change because $-e$ is just a welfare shifter, isn't a function of $\tau$ is it is in menu auction setup
+				\item 3.2.2: $\tau_{W,e}^R$ changes, but qualitative result doesn't change.
+Conjecture: more likely that this lower $\tau_{W,e}^R$ is not self-enforcing because it's lower.
+Should be same proof as strong vs.
+weak binding in JIE R$\&$R: $\pi(\tau_{W,e}^R) - e < \pi(\tau_{W,e}^R)$ is smaller, so it's harder to self-enforce.
+Creates a bigger gap with trade war net profit
+				\item 4.1: no change other than level of $\tau_{S,e}^F$
+				\item 4.2: Side payments: need to check how transfers would change, but no obvious reasons
+				\item 4.3: I was implicitly assuming something like this anyway. But I think I want to try a simple version with an export lobby.
+				\item 4.4: Should write (10) more carefully because different $e$'s in three places
 			\end{itemize}
@@
 
-\section{Lobby's optimal effort}
-Need to be clear about when lobby's optimal $\tau$ is higher than gov'ts: that is, when TA would want to rein in lobby and escape clause would ruin that
-
 \section{Existence proofs?}

### 2017-01-11 draft.tex
remove _W from notation since not doing strong bindings anymore

@@
 Import-competing producer (net) profits are $0.0921$, while government welfare is $0.0855$.
-Since joint government welfare is higher at free trade, the trade agreement will set $\tau^R_{W,e}=0$, profits decrease to $0.0816$, and government welfare increases to $0.0857$.
+Since joint government welfare is higher at free trade, the trade agreement will set the tariff cap at zero, profits decrease to $0.0816$, and government welfare increases to $0.0857$.
 
@@
 
-In the final stage, the government will seek to maximize its welfare by choice of $\tau$ given the political pressure that it faces and the enforcement of the tariff cap, here denoted $\tau_{W,e}^R$ for the case of endogenous political pressure, as distinct from $\tau_W^R$ in the case of exogenous political pressure.
-The home government unilaterally maximizes Expression~\ref{eq:wel} so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as  $\tau^N(\ga(e)) \leq \tau_{W,e}^R$.
-Otherwise, it must set $\tau = \tau_{W,e}^R$ since the weak binding $\tau_{W,e}^R$ is externally enforced.
+In the final stage, the government will seek to maximize its welfare by choice of $\tau$ given the political pressure that it faces and the enforcement of the tariff cap, here denoted $\tau_e^R$ for the case of endogenous political pressure with rigid bindings, as distinct from $\tau^R$ in the case of exogenous political pressure with rigid bindings.
+The home government unilaterally maximizes Expression~\ref{eq:wel} so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as $\tau^N(\ga(e)) \leq \tau_e^R$.
+Otherwise, it must set $\tau = \tau_e^R$ since the weak binding $\tau_e^R$ is externally enforced.
 
 Knowing how the government will set the applied tariff, in the second stage the lobby makes its effort decision to maximize net profits according to Expression~\ref{eq:lv3}.
-If the lobby's optimal effort level in the unconstrained problem (label it $e^L$) would lead to $\tau^N(\ga(e)) > \tau_{W,e}^R$, the lobby will reduce its choice so that no effort is wasted.
-I label the lobbying effort choice that leads to the weak binding level as $e^R_{W,e}$ so that $\tau^N(\ga(e^R_{W,e})) = \tau_{W,e}^R$.
+If the lobby's optimal effort level in the unconstrained problem (label it $e^L$) would lead to $\tau^N(\ga(e)) > \tau_e^R$, the lobby will reduce its choice so that no effort is wasted.
+I label the lobbying effort choice that leads to the weak binding level as $e^R_{W,e}$ so that $\tau^N(\ga(e^R_{W,e})) = \tau_e^R$.
 
@@
 
-I assume for simplicity of exposition that the government does not set the tariff binding $\tau^R_{W,e}$ higher than that which would result from the lobby's optimal effort $e^L$, as setting a higher tariff cap would not change lobbying behavior or outcomes in any way.
+I assume for simplicity of exposition that the government does not set the tariff binding $\tau_e^R$ higher than that which would result from the lobby's optimal effort $e^L$, as setting a higher tariff cap would not change lobbying behavior or outcomes in any way.
 This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.
@@
 
-To understand how the optimal home tariff $\tau^R_{W,e}$ under the trade agreement is set, notice that it must maximize joint government welfare.
+To understand how the optimal home tariff $\tau_e^R$ under the trade agreement is set, notice that it must maximize joint government welfare.
 As argued in Section~\ref{sec:nasheff}, joint government welfare (with variables that do not affect the $X$ sector suppressed) is
@@
 	\end{equation}
-where the lobby will exert no more effort than that which produces the capped $\tau^R_{W,e}=\tau^N(\ga(e))$.
+where the lobby will exert no more effort than that which produces the capped $\tau_e^R=\tau^N(\ga(e))$.
 Recall that the Nash tariff, $\tau^N(\ga(e))$ is the solution to the unilateral optimization problem the home government faces when choosing the applied tariff level in the final stage of the game.
 It is given by Expression~\ref{eq:nash}, which contains the mapping between the political pressure the lobby exerts and the government's unilateral tariff response.
-We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e))$ equal to the tariff cap $\tau^R_{W,e}$.
+We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e))$ equal to the tariff cap $\tau_e^R$.
 This $e$ is that which is labeled $e^R_{W,e}$ above.
@@
 Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage.
-The governments get to choose the $\tau^R_{W,e}$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.
+The governments get to choose the $\tau_e^R$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.
 
 Notice that the maximand in Expression~\ref{exp:1} is exactly the same as that in this third-stage maximization problem except it adds  foreign government welfare.
-This implies that $\tau^R_{W,e}$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments want as much lobbying as possible.
+This implies that $\tau_e^R$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments want as much lobbying as possible.
 To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.\label{fn:tot}}
@@
 When the government's optimum in terms of $\ga(e)$ is lower than the lobby's, the government will use the trade agreement as a domestic political commitment device to manipulate ex-post lobbying incentives.
-Whenever the cap is strictly below the non-cooperative level, that is when $\tau^R_{W,e} < \tau^N$, the weak binding serves to restrain political pressure.
+Whenever the cap is strictly below the non-cooperative level, that is when $\tau_e^R < \tau^N$, the weak binding serves to restrain political pressure.
 
-On the other hand, if the level of the optimal weak binding $\tau^R_{W,e}$ is strictly positive, the weak binding is at least in part serving to generate political involvement---most often thought of as campaign contributions---that the government finds beneficial from its politically-motivated point of view.
+On the other hand, if the level of the optimal weak binding $\tau_e^R$ is strictly positive, the weak binding is at least in part serving to generate political involvement---most often thought of as campaign contributions---that the government finds beneficial from its politically-motivated point of view.
 The government could have chosen a cap of zero and eliminated lobbying activity altogether.
@@
   \begin{multline}
-    \frac{\de}{1-\de} \left\{W_x(\ga(e^R),\tau^R_{W,e}) + W_x^*(\tau^R_{W,e}) - \left[W_x(\ga(e^N),\tau^N(\ga(e^N)) + W_x^*(\tau^N(\ga(e^N)) \right] \right\} \\ \geq W_X(\ga(e^B),\tau^B(\ga(e^B)) - W_X(\ga(e^R),\tau^R_{W,e})
+    \frac{\de}{1-\de} \left\{W_x(\ga(e^R),\tau_e^R) + W_x^*(\tau_e^R) - \left[W_x(\ga(e^N),\tau^N(\ga(e^N)) + W_x^*(\tau^N(\ga(e^N)) \right] \right\} \\ \geq W_X(\ga(e^B),\tau^B(\ga(e^B)) - W_X(\ga(e^R),\tau_e^R)
 		\label{exp:govincent}
   \end{multline}
-where $\de$ is the discount factor assumed common to the government and lobby, $\tau^R_{W,e}$ is the same tariff cap from Section~\ref{sec:perfect}, and $\tau^B$ is the ``break'' tariff.
-Any $\tau > \tau^R_{W,e}$ is sufficient to breach the trade agreement and trigger Nash reversion; $\tau^B(\ga(e^B))$ is defined as the tariff the government would choose as unilaterally optimal if faced with $e^B$, where $e^B$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.
+where $\de$ is the discount factor assumed common to the government and lobby, $\tau_e^R$ is the same tariff cap from Section~\ref{sec:perfect}, and $\tau^B$ is the ``break'' tariff.
+Any $\tau > \tau_e^R$ is sufficient to breach the trade agreement and trigger Nash reversion; $\tau^B(\ga(e^B))$ is defined as the tariff the government would choose as unilaterally optimal if faced with $e^B$, where $e^B$ is the minimum lobbying effort that would give the government incentive to break the trade agreement in the absence of external enforcement.
 
@@
 \begin{equation}
-  \frac{1}{1-\de}\left[\pi(\tau^R_{W,e}) - e^R\right] \geq \pi(\tau^B) - e^B + \frac{\de}{1-\de}\left[\pi(\tau^N) - e^N\right]
+  \frac{1}{1-\de}\left[\pi(\tau_e^R) - e^R\right] \geq \pi(\tau^B) - e^B + \frac{\de}{1-\de}\left[\pi(\tau^N) - e^N\right]
 	\label{exp:lobby}
@@
 \begin{proposition}
-  The presence of the lobby as a repeated-game player may imply that the politically optimal self-enforcing trade agreement tariff is strictly greater than $\tau^R_{W,e}$.
+  The presence of the lobby as a repeated-game player may imply that the politically optimal self-enforcing trade agreement tariff is strictly greater than $\tau_e^R$.
 	\label{res:repeated}

### 2017-01-11 draft.tex
minor textual changes in sec:rigid and perfect; remove W from flexible tariff notation

@@
 \label{sec:rigid}
-In this section and the next, I demonstrate the implications of taking account of endogenous political pressure on some central design features of trade agreements.
+We are now ready to demonstrate the implications of taking account of endogenous political pressure on some central design features of trade agreements.
 I begin by developing some basic results about the design of trade agreements with tariff caps when political pressure is endogenous.
-I then compare these results to those from from \Textcite{bs2005} for the design of trade agreements when political considerations are exogenous.
+I then compare and contrast these results to the case where political considerations are exogenous.
 
@@
 I begin with the case of rigid tariffs---that is, trade agreements that have no provision for flexibility.
-Rigid tariffs cannot be made to depend on the realized level of $\ga$ and so incentive compatibility considerations will do come into play.
+Rigid tariffs cannot be made to depend on the realized level of $\ga$ and so incentive compatibility considerations do not come into play.
 
@@
 I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement.
-Although it is widely agreed that this is unrealistic, this is an important baseline case in which there is no incentive compatibility problem because the contractually-allowed tariff may not vary with $\ga$.
+Although it is widely agreed that this is unrealistic, this is an important baseline case.
 Section~\ref{sec:self} will relax this assumption and require that trade agreements be self-enforcing.
@@
 This implies that $\tau_e^R$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments want as much lobbying as possible.
-To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.\label{fn:tot}}
+To illustrate: if joint welfare is everywhere increasing in $\ga$, although reducing tariffs would improve welfare when $\ga$ is \textit{exogenously given} because each country would not be imposing the terms-of-trade externality on the other, when $\ga$ is endogenous, reducing the tariff reduces $\ga$ and therefore joint welfare.\label{fn:tot}}
 
@@
 \begin{enumerate}
-	\item The government sets the trade agreement tariff $\tau^F_{W,e}$ cooperatively.
+	\item The government sets the trade agreement tariff $\tau_e^R$ cooperatively.
 I assume this is a weak binding and add the subscript $F$ to denote the environment with flexibility to deal with shocks.
@@
 	\item The lobby chooses its effort level $e^F$.
-	\item The government chooses the applied tariff level from among $\tau^F_{W,e}$, $\tau(s)$ and $\tau^B(\ga(s,e^F))$.
+	\item The government chooses the applied tariff level from among $\tau_e^R$, $\tau(s)$ and $\tau^B(\ga(s,e^F))$.
 \end{enumerate}
 Here I take the higher, `escape' tariff to be continuous instead of set at a given binding.
-If the government applies $\tau^F_{W,e}$, implementation of the agreement goes on as before.
+If the government applies $\tau_e^R$, implementation of the agreement goes on as before.
 Consistent with the WTO Agreement on Safeguards, the government can apply the tariff that is necessary to remedy the economic injury---not necessarily equal to the political injury---caused by the shock, $\tau(s)$, with no retaliation for three periods.
-If the government applies a tariff that is higher than either $\tau^F_{W,e}$ or $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
+If the government applies a tariff that is higher than either $\tau_e^R$ or $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
 I label this tariff $\tau^B(\ga(s,e^F))$, where $e^F$ must be high enough to give the government incentive to break the trade agreement in the absence of external enforcement.
@@
   \begin{multline}
-    \left[1+ \de +\de^2 \right] \left\{W_x(\ga(s,0),\tau(s)) + W_x^*(\tau^F_{W,e}) \right\} \geq \\ W_X(\ga(s,e^F),\tau^B(\ga(s,e^F)) + W_x^*(\tau^F_{W,e}) + \left[\de +\de^2 \right] \left\{W_X(\ga(s,e^N),\tau^N) + W_x^*(\tau^N) \right\}
+    \left[1+ \de +\de^2 \right] \left\{W_x(\ga(s,0),\tau(s)) + W_x^*(\tau_e^R) \right\} \geq \\ W_X(\ga(s,e^F),\tau^B(\ga(s,e^F)) + W_x^*(\tau_e^R) + \left[\de +\de^2 \right] \left\{W_X(\ga(s,e^N),\tau^N) + W_x^*(\tau^N) \right\}
 		\label{exp:govincentEC}
@@
 
-Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau^F_{W,e}\right)$ so that in some cases the lobby's incentive constraint is violated.
+Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau_e^R\right)$ so that in some cases the lobby's incentive constraint is violated.
 This would lead to the legislature's constraint being violated as well, creating a dispute.

### 2017-01-11 draft.tex
simplify notation for e^R as well

@@
 I begin by assuming that perfect external enforcement is available for ensuring that the trading partners abide by the terms of the trade agreement.
-Although it is widely agreed that this is unrealistic, this is an important baseline case.
-Section~\ref{sec:self} will relax this assumption and require that trade agreements be self-enforcing.
+Although this is an unrealistic assumption, it establishes an important baseline case.
+Section~\ref{sec:self} adds the requirement that trade agreements be self-enforcing.
 
@@
 The home government unilaterally maximizes Expression~\ref{eq:wel} so that the applied tariff is the Nash tariff $\tau^N(\ga(e))$ as long as $\tau^N(\ga(e)) \leq \tau_e^R$.
-Otherwise, it must set $\tau = \tau_e^R$ since the weak binding $\tau_e^R$ is externally enforced.
+Otherwise, the government must set $\tau = \tau_e^R$ since the weak binding $\tau_e^R$ is externally enforced.
 
@@
 If the lobby's optimal effort level in the unconstrained problem (label it $e^L$) would lead to $\tau^N(\ga(e)) > \tau_e^R$, the lobby will reduce its choice so that no effort is wasted.
-I label the lobbying effort choice that leads to the weak binding level as $e^R_{W,e}$ so that $\tau^N(\ga(e^R_{W,e})) = \tau_e^R$.
+I label the lobbying effort choice that leads to the weak binding level as $e^R$ so that $\tau^N(\ga(e^R)) = \tau_e^R$.
 
@@
 
-The lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R_{W,e}$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding.
+The lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding.
 The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$.
@@
 We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e))$ equal to the tariff cap $\tau_e^R$.
-This $e$ is that which is labeled $e^R_{W,e}$ above.
+This $e$ is that which is labeled $e^R$ above.
 
-Ex-post lobbying will constrain the governments in that the $\ga\left(e^R_{W,e}\right)$ they experience will be determined through the unilateral maximization problem in the last stage.
+Ex-post lobbying will constrain the governments in that the $\ga\left(e^R\right)$ they experience will be determined through the unilateral maximization problem in the last stage.
 The governments get to choose the $\tau_e^R$ that maximizes their joint welfare, but they cannot change this future lobbying process; they can only reduce the level at which it takes place through the tariff cap.

### 2017-01-12 draft.tex
revise sec 3 w/view toward generalizing model

@@
 We are now ready to demonstrate the implications of taking account of endogenous political pressure on some central design features of trade agreements.
-I begin by developing some basic results about the design of trade agreements with tariff caps when political pressure is endogenous.
-I then compare and contrast these results to the case where political considerations are exogenous.
+I develop some basic results about the design of trade agreements with tariff caps when political pressure is endogenous and compare and contrast these results to the case where political considerations are exogenous.
 
@@
 Rigid tariffs cannot be made to depend on the realized level of $\ga$ and so incentive compatibility considerations do not come into play.
-
-%Importantly, both cases suffer from asymmetric information, but we will see that the nature of the informational lacuna differs.
-In the  exogenous case, there is a difference between the ex-ante and ex-post information of each government that has very important implications; this difference does not exist in the endogenous case.
-The implications for the features of trade agreements that are desirable---such as escape clauses---result directly from this difference.
 
@@
 I label the lobbying effort choice that leads to the weak binding level as $e^R$ so that $\tau^N(\ga(e^R)) = \tau_e^R$.
+The lobby will choose $e^R$ when its optimal effort would lead to a tariff level higher than that allowed by the weak binding; otherwise it will choose its unconstrained optimal effort $e^L$.
 
@@
 
-The lobby will choose one of two effort levels: its unconstrained optimal effort $e^L$ or $e^R$ in the case that its optimal effort would lead to a tariff level higher than that allowed by the weak binding.
-The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$.
-The government is unable to use a weak binding to encourage extra lobbying when only import-competing interests are represented by a lobby.\footnote{Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}
+I assume for simplicity of exposition that the government does not set the tariff binding $\tau_e^R$ higher than that which would result from the lobby's optimal effort $e^L$, as setting a higher tariff cap would not change lobbying behavior or outcomes in any way.\footnote{The government's tariff cap choice can reduce lobbying effort; it cannot, however, increase lobbying effort above $e^L$.
+The government is unable to use a weak binding to encourage extra lobbying when only import-competing interests are represented by a lobby.
+Introducing lobbies for the export industry could reverse this: lobbying effort by exporters could be encouraged in support of capping the import tariffs of the partner country.}
 
-I assume for simplicity of exposition that the government does not set the tariff binding $\tau_e^R$ higher than that which would result from the lobby's optimal effort $e^L$, as setting a higher tariff cap would not change lobbying behavior or outcomes in any way.
-This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.
-In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.
+This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.\footnote{Note that in their treatment, Bagwell and Staiger (2005) assume linear supply and demand.} In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.
 
@@
 	\end{equation}
-where the lobby will exert no more effort than that which produces the capped $\tau_e^R=\tau^N(\ga(e))$.
+where the lobby will exert no more effort than that which produces the capped $\tau_e^R=\tau^N(\ga(e^R))$.
 Recall that the Nash tariff, $\tau^N(\ga(e))$ is the solution to the unilateral optimization problem the home government faces when choosing the applied tariff level in the final stage of the game.
 It is given by Expression~\ref{eq:nash}, which contains the mapping between the political pressure the lobby exerts and the government's unilateral tariff response.
-We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e))$ equal to the tariff cap $\tau_e^R$.
-This $e$ is that which is labeled $e^R$ above.
+We can therefore use Expression~\ref{eq:nash} to determine the level of $e$ the lobby must exert in order to receive a particular $\tau^N(\ga(e^R))$ equal to the tariff cap $\tau_e^R$.
 
@@
 
-Notice that the maximand in Expression~\ref{exp:1} is exactly the same as that in this third-stage maximization problem except it adds  foreign government welfare.
+Notice that the maximand in Expression~\ref{exp:1} is exactly the same as that in the third-stage maximization problem except it adds  foreign government welfare.
 This implies that $\tau_e^R$ is (weakly) lower than the unilateral tariff as the terms-of-trade externality is internalized.\footnote{The terms-of-trade motivation for the trade agreement may not strictly lower the tariff if the solution to the government's joint maximization problem is not interior in the sense that the governments want as much lobbying as possible.
@@
 When the government's optimum in terms of $\ga(e)$ is lower than the lobby's, the government will use the trade agreement as a domestic political commitment device to manipulate ex-post lobbying incentives.
-Whenever the cap is strictly below the non-cooperative level, that is when $\tau_e^R < \tau^N$, the weak binding serves to restrain political pressure.
+Whenever the cap is strictly below the non-cooperative level, that is when $\tau_e^R < \tau^N$, the weak binding restrains political pressure.
 
-On the other hand, if the level of the optimal weak binding $\tau_e^R$ is strictly positive, the weak binding is at least in part serving to generate political involvement---most often thought of as campaign contributions---that the government finds beneficial from its politically-motivated point of view.
+On the other hand, if the level of the optimal weak binding $\tau_e^R$ is strictly positive, the weak binding is at least in part serving to generate political effort---most often thought of as campaign contributions---that the government finds beneficial from its politically-motivated point of view.
 The government could have chosen a cap of zero and eliminated lobbying activity altogether.
 In fact, any strong binding accomplishes this very feat, but is also \textit{unable} to generate lobbying activity because of its effect on lobbying incentives.
-Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.\footnote{Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.\label{fn:mrc1}}
+Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.
 
@@
 Thus we see that the governments can use a tariff cap to manipulate lobbying incentives so that political effort is neither too low or too high.
+Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.
 
@@
 \label{sec:self}
-Here I drop the assumption of perfect external enforcement since external enforcement is not widely available in the context of international trade relations.
+Here I remove the assumption of perfect external enforcement since external enforcement is not widely available in the context of international trade relations.
 External enforcement is replaced with promises of future cooperation and punishment is modeled via a repeated game.
@@
 In order to establish an equilibrium, we must ensure that both the on-schedule (static) incentive constraint and the off-schedule (repeated-game) incentive constraint is satisfied.
-We know that the former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here.
+The former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here.
 Following most of the literature, I assume that any deviation triggers a reversion to the static Nash equilibrium---what is known as `grim trigger.'\footnote{See \Textcite{krw} for an alternative that takes seriously the threat of governments renegotiating out of punishments that are not themselves incentive compatible.\label{fn:krw}}
@@
 When $\ga$ is endogenously-determined, repeated-game enforcement is altered relative to the case of exogenously-given political pressure.
-This is because, in place of a stochastic process that determines $\ga$, $\ga$ is determined by a new repeated-game player who has incentives relative to whether the trade agreement is followed or reversion to the non-cooperative Nash outcome takes place.
-
-When $\ga$ is exogenous, the optimal trade agreement tariff is the same under perfect external enforcement and self-enforcement because the change in enforcement conditions cannot affect the realization of $\ga$.
-We will see that moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem.
-Moving to the repeated-game environment with self enforcement does not alter the government's costs and benefits so that the government's optimal trade agreement tariff is invariant to the enforcement conditions.
-This creates the possibility that the politically-optimal trade agreement tariff may not be self-enforcing.
+This is because, in place of a stochastic process that determines $\ga$, $\ga$ is determined by a new repeated-game player who cares whether the trade agreement is followed or reversion to the non-cooperative Nash outcome takes place.
 
@@
 Enforcement considerations do not change the protection demands when $\ga$ is exogenous.
-Similarly, because the government's objective function does not change and it is able to use the tariff cap to control the lobby's effort level, as long as it is sufficiently patient, it will set the same tariff cap.
+Similarly, because the government's objective function does not change and it is able to use the tariff cap to control the lobby's effort level, as long as the government is sufficiently patient, it would like to set the same tariff cap in the face of self-enforcement constraints as when the agreement is externally enforced.
+
+When $\ga$ is exogenous, the optimal trade agreement tariff is the same under perfect external enforcement and self-enforcement because the change in enforcement conditions cannot affect the realization of $\ga$.
+In contrast, moving from a static problem with external enforcement to a repeated-game problem with an enforcement constraint changes the lobby's decision problem.
+Moving to the repeated-game environment with self enforcement does not alter the government's ideal trade agreement tariff, but the introduction of the lobby creates the possibility that the politically-optimal trade agreement tariff may not be self-enforcing.
 
@@
   \begin{multline}
-    \frac{\de}{1-\de} \left\{W_x(\ga(e^R),\tau_e^R) + W_x^*(\tau_e^R) - \left[W_x(\ga(e^N),\tau^N(\ga(e^N)) + W_x^*(\tau^N(\ga(e^N)) \right] \right\} \\ \geq W_X(\ga(e^B),\tau^B(\ga(e^B)) - W_X(\ga(e^R),\tau_e^R)
+    \frac{\de}{1-\de} \left\{W_X(\ga(e^R),\tau_e^R) + W_X^*(\tau_e^R) - \left[W_X(\ga(e^N),\tau^N(\ga(e^N)) + W_X^*(\tau^N(\ga(e^N)) \right] \right\} \\ \geq W_X(\ga(e^B),\tau^B(\ga(e^B)) - W_X(\ga(e^R),\tau_e^R)
 		\label{exp:govincent}
@@
 
-On the left side of the inequality is the discounted per-period gain from maintaining the trade agreement relative to Nash reversion.
+On the left side of the inequality is the discounted gain from maintaining the trade agreement relative to Nash reversion.
 In order for the government to have the incentive to abide by the agreement, this must be at least as large as the benefit to cheating.
@@
 
-Here we have the possibility of a starkly different result from that under exogenous political pressure, where \Textcite{bs2005} show that sufficiently patient governments can always sustain the optimal trade agreement tariffs.
+Here we have the possibility of a starkly different result from that under exogenous political pressure.
 Instead, the addition of a second incentive constraint can make it impossible to sustain the politically efficient tariffs.
@@
 
-Shifting attention to the lobby's incentive constraint in Expression~\ref{exp:lobby}, this becomes harder to satisfy as $\de$ increases.
-That is, the lefthand side is decreasing in $\de$ and the righthand side is increasing in $\de$, and this relationship holds even taking into account the influence from the government's constraint on $e^B$ as long as $e^B$ is to the left of the lobby's optimal choice of effort.
-Once $e^B$ passes $e^L$, $\pi(\tau^B) - e^B$ begins to decrease and it's possible for the effect of $\de$ on the lobby's constraint to eventually be overturned.
+Shifting attention to the lobby's incentive constraint in Expression~\ref{exp:lobby}, this becomes harder to satisfy as $\de$ increases for most levels of the trade agreement tariff.
+That is, the lefthand side is decreasing and the righthand side is increasing as a direct function of $\de$.
+When taking into account the influence of $\de$ from the government's constraint on $e^B$, for a small part of the parameter space where $e^B$ must be very large, it's possible for the lobby's constraint to eventually loosen as $\de$ increases since $\pi(\tau^B) - e^B$ decreases to the right of $\tau^N$.
 
@@
 At issue is the behavior of the lobby, which is now a formal player in its own right and must be incentivized to keep its behavior on the equilibrium path.
-More patient lobbies are willing to work harder to encourage the government to break the trade agreement so that they can enjoy infinite periods of Nash tariffs, so increasing the discount factor makes it more difficult rather than easier to satisfy the lobby's constraint.\footnote{The essential point underlying this result was first made by \Textcite{buzard2013a}.}
+More patient lobbies are willing to work harder to encourage the government to break the trade agreement so that they can enjoy infinite periods of Nash tariffs.
+Increasing the discount factor makes it more difficult rather than easier to satisfy the lobby's constraint.\footnote{The essential point underlying this result was first made by \Textcite{buzard2013a}.}
 
@@
 \begin{enumerate}
-	\item The government sets the trade agreement tariff $\tau_e^R$ cooperatively.
-I assume this is a weak binding and add the subscript $F$ to denote the environment with flexibility to deal with shocks.
+	\item The government sets the trade agreement tariff $\tau_e^F$ cooperatively. I assume this is a weak binding. The subscript $F$ denotes the environment with flexibility to deal with shocks.
 	\item The exogenous shock $s$ is realized. Its level is assumed costlessly verifiable for simplicity.
 	\item The lobby chooses its effort level $e^F$.
-	\item The government chooses the applied tariff level from among $\tau_e^R$, $\tau(s)$ and $\tau^B(\ga(s,e^F))$.
+	\item The government chooses the applied tariff level from among $\tau_e^F$, $\tau(s)$ and $\tau^B(\ga(s,e^F))$.
 \end{enumerate}
-Here I take the higher, `escape' tariff to be continuous instead of set at a given binding.
-If the government applies $\tau_e^R$, implementation of the agreement goes on as before.
+I take the higher, `escape' tariff $\tau(s)$ to be continuous.
+If the government applies $\tau_e^F$, implementation of the agreement goes on as before.
 Consistent with the WTO Agreement on Safeguards, the government can apply the tariff that is necessary to remedy the economic injury---not necessarily equal to the political injury---caused by the shock, $\tau(s)$, with no retaliation for three periods.
-If the government applies a tariff that is higher than either $\tau_e^R$ or $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
+If the government applies a tariff that is higher than either $\tau_e^F$ or $\tau(s)$, it will choose its myopic best response to the pressure it experiences.
 I label this tariff $\tau^B(\ga(s,e^F))$, where $e^F$ must be high enough to give the government incentive to break the trade agreement in the absence of external enforcement.
@@
   \begin{multline}
-    \left[1+ \de +\de^2 \right] \left\{W_x(\ga(s,0),\tau(s)) + W_x^*(\tau_e^R) \right\} \geq \\ W_X(\ga(s,e^F),\tau^B(\ga(s,e^F)) + W_x^*(\tau_e^R) + \left[\de +\de^2 \right] \left\{W_X(\ga(s,e^N),\tau^N) + W_x^*(\tau^N) \right\}
+    \left[1+ \de +\de^2 \right] \left\{W_x(\ga(s,0),\tau(s)) + W_x^*(\tau_e^F) \right\} \geq \\ W_X(\ga(s,e^F),\tau^B(\ga(s,e^F)) + W_x^*(\tau_e^F) + \left[\de +\de^2 \right] \left\{W_X(\ga(s,e^N),\tau^N) + W_x^*(\tau^N) \right\}
 		\label{exp:govincentEC}
@@
 This provides a looser incentive constraint than in the absence of the escape clause and improves self-enforceability of the agreement.
-Or the lobbies may still find it in their interest to exert effort to encourage the government to provide this higher level of protection.
+Or the lobbies may still find it in their interest to exert effort to encourage the government to provide this higher level of protection $\tau(\ga(s,e^F))$.
 This will provoke a dispute, but the protection afforded by the higher tariff will not be hidden behind the escape clause.
@@
 What's more, this may provide an answer to why Safeguard measures have fallen out of use.
-If the real purpose is to provide an outlet for political pressure, I have demonstrated above that the current WTO-style Safeguards rules seem designed explicitly to discourage rent-seeking behavior.
+If the real purpose is to provide an outlet for political pressure, I have demonstrated above that the current WTO-style Safeguards rules seem designed explicitly to discourage such rent-seeking behavior.
 Thus if other measures such as Anti-Dumping Duties are more susceptible to manipulation, it seems natural that lobbies' efforts would migrate there.
-
-As in \Textcite{zissimos}, I also plan to compare the Nash-reversion punishment to punishment via 'withdrawal of equivalent concessions,' as it appears on the face of it that this environment may provide a justification its use.
-The intuition is the following: \Textcite{buzard2013a} shows that the lobbying effort required to cause the government to enter into a dispute is higher than the Nash equilibrium effort levels.
-Then the tariffs during a dispute are also higher, and punishment at this higher level would be more severe than Nash---both for the government and the lobby.\footnote{\Textcite{beshkar2010a} shows that when one assumes that utility is not transferable between countries as has become common in the literature, the optimal mechanism involves less-than-proportional retaliation against parties who have defected from the agreement.
-\Textcite{martinvergote} demonstrate that future punishment provides for higher welfare than contemporaneous punishment when governments are sufficiently patient.
-Indeed, they show that retaliation is a necessary feature of any efficient equilibrium in this environment.
-\Textcite{hungerford} and \Textcite{riezman1991} also consider the impact of different assumptions about reactions and timing of punishments for deviations from agreements.}
 
@@
 There is also strong potential to gain insights concerning very realistic dynamics between political and economics shocks and lobbying, especially insofar as they impact questions of enforcement and administered protection.
+
+As in \Textcite{zissimos}, it would also be interesting to compare the Nash-reversion punishment to punishment via 'withdrawal of equivalent concessions,' as it appears on the face of it that this environment may provide a justification its use.
+The intuition is the following: \Textcite{buzard2013a} shows that the lobbying effort required to cause the government to enter into a dispute is higher than the Nash equilibrium effort levels.
+Then the tariffs during a dispute are also higher, and punishment at this higher level would be more severe than Nash---both for the government and the lobby.\footnote{\Textcite{beshkar2010a} shows that when one assumes that utility is not transferable between countries as has become common in the literature, the optimal mechanism involves less-than-proportional retaliation against parties who have defected from the agreement.
+\Textcite{martinvergote} demonstrate that future punishment provides for higher welfare than contemporaneous punishment when governments are sufficiently patient.
+Indeed, they show that retaliation is a necessary feature of any efficient equilibrium in this environment.
+\Textcite{hungerford} and \Textcite{riezman1991} also consider the impact of different assumptions about reactions and timing of punishments for deviations from agreements.}
 

### 2017-01-12 draft.tex
revise sec 4 w/view toward generalizing model

@@
 \label{sec:escape}
-I begin by following \Textcite{bs2005} in examining a trade agreement with an escape clause by adding a second negotiated binding to the agreement so that there is one (weak) binding for when political pressure is low and a higher, weak binding for exceptional circumstances of high levels of political pressure.
+I begin by following the literature in examining a trade agreement with an escape clause by adding a second negotiated binding to the agreement so that there is one (weak) binding for when political pressure is low and a higher, weak binding for exceptional circumstances of high levels of political pressure.
 Escape clauses are common features of trade agreements in practice, and when one models political pressure as exogenous, an escape clause seems attractive: in the presence of a particularly large negative political shock, being bound to a tariff designed for normal times would cause significant welfare losses to the politician.
@@
 But in the stark case under examination here with purely endogenous politics, an exceptionally high $\ga$ can only derive from higher effort exerted by the lobby.
-If one chooses an optimal tariff cap, a second, higher cap can only encourage excess, sub-optimal lobbying.
+If one chooses an optimal tariff cap, a second, higher cap can only encourage excess lobbying that is sub-optimal from the point of view of the government.
 
@@
 
-Thus, in this case the strong binding does not really create an incentive-compatibility-inducing cost, but again the central problem is that of lobbying incentives.
+Thus, in this case the strong binding does not create an incentive-compatibility-inducing cost, but again the central problem is that of lobbying incentives.
 This could explain, as Bagwell and Staiger point out, why the WTO does not incorporate strong bindings for the escape clause.
@@
 Here the government is truthfully revealing the level of political pressure it faces, but that pressure is created by endogenous forces that are intensified by the existence of the escape clause.
-The escape clause is incentive compatible, but as will be clarified in the next section, the spirit of the escape clause is not necessarily upheld.
+The escape clause is incentive compatible, but as will be clarified in the next section, the spirit of the escape clause is not upheld.
 
@@
 When the realization of the shock is high, the government will report the high shock and apply the escape clause tariff.
-But when the shock is low, the lobby will exert effort so that the sum of $\ga(s)$ and $\ga(e)$ equals the high realization.
+But when the shock is low, the lobby will exert effort so that the sum of $\ga(s)$ and $\ga(e)$ equals $\ga(\ov{s})$.
 
@@
 From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section~\ref{sec:objfcn}, the governments' ex-ante welfare is reduced by excess lobbying.
-However, the problem is not interesting if government welfare is everywhere increasing in lobbying effort because the lower binding will be set to the prohibitive tariff in this environment with no counter-lobbying by export groups.
 
@@
 This is not a problem of incentive compatibility.
-Similar to \Textcite{beshkar2010b}, it can be shown [formal result coming soon] that a dispute settlement body that provides an independent signal about the value of $\ga(s,e)$ on which the governments condition punishments for misreporting can make truth-telling incentive compatible.
+Similar to \Textcite{beshkar2010b}, it can be shown that a dispute settlement body that provides an independent signal about the value of $\ga(s,e)$ on which the governments condition punishments for misreporting can make truth-telling incentive compatible.
 That is, a government will not over-report its political-economy pressure in order to impose a terms-of-trade externality on its trading partner.

### 2017-01-12 draft.tex
integrate sec:nasheff into sec:stage

@@
 
+Expression~\ref{eq:wel} can be written $W(\ga(s,e),\tau,\tau^*) = W_X(\ga(s,e),\tau) +W_Y(\tau^*)$, where $W_X(\ga(s,e),\tau)$ is the utility derived from consumer surplus, producer surplus and tariff revenues in the import-competing industry and $W_Y(\tau^*)$ is the utility derived from consumer and producer surplus in the exporting industry.
+As there are no interactions between $\tau$ and $\tau^*$, when setting the tariff unilaterally, the government simply maximizes $W(\ga(s,e),\tau, \tau^*)$ by choice of $\tau$ given $\ga(s,e)$ and $\tau^*$ and sets the non-cooperative tariff
+\begin{equation}
+  \tau^N(\ga(s,e)) = \max_\tau W_X(\ga(s,e),\tau)
+  \label{eq:nash}
+\end{equation}
+I refer to this as the Nash tariff because it is the result of Nash equilibrium in the non-cooperative game between the governments.
+$\tau^N$ is increasing in $\ga$ and I assume the realization of $\ga$ is below the threshold need to ensure the second order condition is satisfied.
+This assumption does not in general rule out that the solution to Expression~\ref{eq:nash} could be the prohibitive tariff.
+
 Given the government's preferences, the home lobby chooses its lobbying effort $e$ to maximize profits net of lobbying effort:
@@
 \end{equation}
-Note that in this symmetric environment, this is the Nash bargaining solution where the disagreement point is the welfare resulting from the Nash equilibrium in the non-cooperative game.
 
+Government welfare in both countries is additively separable in $\tau$ and we can write the welfare of the foreign government as $W^*(\ga^*(s^*,e^*),\tau,\tau^*)$ = $W_Y^*(\ga^*(s^*,e^*),\tau^*) + W_X^*(\tau)$.
+Therefore maximizing Expression~\ref{eq:jv} in $\tau$ is equivalent to maximizing $W_X(\ga(s,e),\tau) + W_X^*(\tau)$.
+The trade agreement tariff is then given by
+\begin{equation}
+  \tau^E(\ga(s,e)) = \max_\tau W_X(\ga(s,e),\tau) + W_X^*(\tau)
+  \label{eq:eff}
+\end{equation}
+As in the applied tariff problem, I assume the realization of $\ga$ is below the threshold need to ensure the second order condition is satisfied.
+
+Note that in this symmetric environment, this problem can be solved by applying the Nash bargaining solution where the disagreement point is the welfare resulting from the Nash equilibrium in the non-cooperative game.
+
+\begin{comment}
 \subsection{Nash Tariffs and Internationally Efficient Tariffs}
 \label{sec:nasheff}
+
 The home government's welfare is $W(\ga(s,e),\tau,\tau^*) = W_X(\ga(s,e),\tau) +W_Y(\tau^*)$, where
@@
 The second order condition is satisfied for $\ga < \frac{25}{4}$, so again for all values of $\ga$ that lead to a non-prohibitive tariff.
+\end{comment}
 
@@
 To understand how the optimal home tariff $\tau_e^R$ under the trade agreement is set, notice that it must maximize joint government welfare.
-As argued in Section~\ref{sec:nasheff}, joint government welfare (with variables that do not affect the $X$ sector suppressed) is
+As argued in Section~\ref{sec:stage}, joint government welfare (with variables that do not affect the $X$ sector suppressed) is
   \begin{equation}

### 2017-01-12 draft.tex
change \tau^E to \tau^a

@@
 \begin{equation}
-  \tau^E(\ga(s,e)) = \max_\tau W_X(\ga(s,e),\tau) + W_X^*(\tau)
+  \tau^a(\ga(s,e)) = \max_\tau W_X(\ga(s,e),\tau) + W_X^*(\tau)
   \label{eq:eff}
@@
 \begin{equation}
-  \tau^E(\ga(s,e)) = \frac{4\left[\ga(s,e)-1\right]}{25-4\ga(s,e)}
+  \tau^a(\ga(s,e)) = \frac{4\left[\ga(s,e)-1\right]}{25-4\ga(s,e)}
   \label{eq:eff}
@@
 
-Multiplying Equation (\ref{eq:ea}) by $\frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^E}{\partial \ga}}$, we have
+Multiplying Equation (\ref{eq:ea}) by $\frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^a}{\partial \ga}}$, we have
 	\begin{equation}
-	  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^N}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = \frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^E}{\partial \ga}} + 1.
+	  \frac{\partial \pi}{\partial \tau} \frac{\partial \tau^N}{\partial \ga} \frac{\partial \ga}{\partial e^{EA}} = \frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^a}{\partial \ga}} + 1.
 			\label{eq:ea2}
 		\end{equation}
-Because $\frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^E}{\partial \ga}} > 0$, the solution to the ex-ante weak binding first order condition as given in Equation~(\ref{eq:ea2}) is smaller than the solution to Equation~(\ref{eq:nta}), so that the lobby exerts less effort in the ex-ante stage of trade agreement negotiations than when there is no trade agreement.
-Since $\tau^E(\cdot) < \tau^N(\cdot)$ for all $e$, $\tau^{EA}\left(e^{EA}\right) < \tau^N\left(e^N\right)$ when bindings are weak.
+Because $\frac{\frac{\partial \tau^N}{\partial \ga}}{\frac{\partial \tau^a}{\partial \ga}} > 0$, the solution to the ex-ante weak binding first order condition as given in Equation~(\ref{eq:ea2}) is smaller than the solution to Equation~(\ref{eq:nta}), so that the lobby exerts less effort in the ex-ante stage of trade agreement negotiations than when there is no trade agreement.
+Since $\tau^a(\cdot) < \tau^N(\cdot)$ for all $e$, $\tau^{EA}\left(e^{EA}\right) < \tau^N\left(e^N\right)$ when bindings are weak.
 

### 2017-01-13 draft.tex
re-do figures in R for -e welfare fcn

@@
 
-Envelope-theorem style results imply that the first order condition in the same in the unilateral and joint problems: $\displaystyle \frac{\partial W_x}{\partial \ga} = 0$.
+Envelope-theorem style results imply that the first order condition is the same in the unilateral and joint problems: $\displaystyle \frac{\partial W_x}{\partial \ga} = 0$.
 See the \hyperlink{envelope}{Appendix} for details.
@@
 \begin{center}
-\includegraphics[height=3in, width=4.25in]{unweight.jpg}
+\includegraphics[height=3in, width=5.25in]{unweight.jpeg}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=3in, width=4.25in]{weight.jpg}
+\includegraphics[height=3in, width=5.25in]{weight.jpeg}
 \end{center}

### 2017-01-13 figures.R

@@ -10,8 +10,6 @@ CSy = ((3 +3*t)^2)/98
 PSy = ((4 -3*t)^2)/98
                           
 gtn = (68*t + 5)/(8*(1+t))    #gamma as fcn of tau, unilateral maximization
-e = (gtn - 1)^4
-
 
 #some values of tau are so low that gamma would be less than one
 gtn2 = matrix(0,length(gtn),1)
@@ -21,16 +19,12 @@ for (j in 1:length(gtn)) {
   gtn2[j] = max(x,1)
 }  
 
-w = 1/(4+gtn2)                 #create weights as function of gamma
-                          
-W = w*CSx + gtn2*w*PSx + w*CSy + w*PSy + w*TR  #fully weighted welfare
-plot(t,W)
-                          
-W2 = CSx + gtn2*PSx + CSy + PSy + TR  #partially weighted welfare
-plot(t,W2)
+W = CSx + gtn2*PSx + CSy + PSy + TR  #partially weighted welfare
+plot(t,W, type = "l",xlab = "Tariff",ylab = "Gov't Welfare")
 
-W3 = CSx + gtn2*PSx + CSy + PSy + TR - e  #partially weighted welfare - e
-plot(t,W3)
+e = (gtn2 - 1)^10
+W2 = CSx + gtn2*PSx + CSy + PSy + TR - e  #partially weighted welfare - e
+plot(t,W2, type = "l",xlab = "Tariff",ylab = "Gov't Welfare")
 
                           
 ##########################################################

### 2017-01-14 draft.tex
more changes to figures.R

@@
 \begin{equation}
-  W(\ga(s,e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(s,e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
+  W(\ga(s,e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(s,e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau) - e
   \label{eq:wel}
@@
   W(\ga(e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
-  \tag{\ref{eq:wel}}
+  \tag{\ref{eq:baldwin}}
 \end{equation}
@@
 
-This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the partially weighted government welfare function from Expression~\ref{eq:wel}.
+This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the government welfare function from Expression~\ref{eq:baldwin}.
 Note that the graph is in terms of tariffs, which are monotonically increasing in $\ga$.
@@
 
-This result that government welfare is strictly increasing in the political pressure parameter would seem to be related to the fact that, although $\ga$ is often referred to as a ``political economy weight,'' Expression~\ref{eq:wel} is only partially weighted.
-That the extra importance given to profits of importers when $\ga$ increases is pure additional welfare with no concomitant reduction elsewhere may be an assumption worth weakening, particularly in the context of endogenous political activity.
-
-One alternative formulation for the government objective function is one in which $\ga$ is a true weight in the following straightforward sense:
-\begin{equation}
-  \frac{1}{4+\ga(e)}CS_X(\tau) + \frac{\ga(e)}{4+\ga(e)}PS_X(\tau) + \frac{1}{4+\ga(e)}CS_Y(\tau^*) + \frac{1}{4+\ga(e)}PS_Y(\tau^*) + \frac{1}{4+\ga(e)}TR(\tau)
-  \label{eq:weighted}
-\end{equation}
-
-See the \hyperlink{sol_weighted}{Appendix} for the mathematical derivations corresponding to Figure~\ref{fig:weight}.
-They show that, particularly for the trade agreement tariffs, government welfare first declines and then increases, with the minimum value occurring around $\tau = 0.057$ and $\ga = 1.05$.
-The maximum still unquestionably occurs at the maximum value of $\ga$, which leads to the prohibitive tariff of $\frac{1}{6}$.
-But if a government's objective function is closer to Expression~\ref{eq:weighted} than Expression~\ref{eq:wel}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
+When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, this monotonicity often breaks down.
+For most specifications of the $\ga(e)$ function, there is an interior optimum.
+Figure~\ref{fig:weight} illustrates for the case of $\ga(e) = 1 + e^{.2}$.
+Here, the optimum occurs at $\tau = 0.105$ and $e = 0.0072$, which is about $7\%$ of profits in the import-competing industry.
+Here $ga(0.0065) = 1.37$.
+The lobby's optimum is at $\tau = But if a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
 
@@
 
-I begin with $\ga(e) = 1 + e^{.1}$.
-With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0016$, which results in $\ga=1.525$ and a Nash tariff of $0.129$.
-Import-competing producer (net) profits are $0.1025$, while government welfare is $0.0862$.
-Because joint government welfare is convex in the tariff, the solution will be at a corner: either at $\tau^R_{W,e} =0$ or at the highest level for which the lobby will exert effort, i.e.
-the Nash tariff.
-Since joint welfare at the zero tariff is $0.0857$, the tariff cap is set at the trivial level of the non-cooperative Nash tariff.
-
-Here, given the way in which lobbying effort is translated into political support as represented by $\ga$, the government cannot improve its welfare by manipulating lobbying incentives.
-Discouraging lobbying would reduce welfare; in this parameter range, the government's interest is solely in encouraging lobbying effort, and there is nothing it can do in this regard.
-
-%This is does through the introduction of the trade agreement: by making protection more expensive, on the margin the lobby pays more in order to get even less protection than it did in the absence of the agreement.
-But there is nothing further the government can do to increase $\ga$.
-
-This example stands in contrast to one in which $\ga(e) = 1 + e^{.3}$.
-With no trade agreement, the lobby maximizes its net profits by choosing $e=0.0010$, which results in $\ga=1.126$ and a unilateral tariff of $0.068$.
-Import-competing producer (net) profits are $0.0921$, while government welfare is $0.0855$.
-Since joint government welfare is higher at free trade, the trade agreement will set the tariff cap at zero, profits decrease to $0.0816$, and government welfare increases to $0.0857$.
+Here, the optimum occurs at $\tau = 0.142$ and $e = 0.0065$, which is about $6\%$ of profits in the import-competing industry.
+Here $ga(0.0065) = 1.60$.
+But if a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
 
@@
 
-With the fully-weighted welfare function, we see that the trade agreement is employed to different ends depending on $\ga(e)$.
-If the lobby's ability to make an impact on the political process is strong enough, the government will be chiefly concerned with the profits of the import-competing industry and will use the trade agreement solely to internalize the terms-of-trade externality.
-However, if the lobby's impact is below a given threshold, the distortions that protection create are weighed more heavily and the government will use the trade agreement to rein in lobbying activity.
-
-%In either case, we see that the government is able to provide reduced protection while extracting more effort from the lobby under a trade agreement: internalizing the terms of trade externality increases the marginal cost of protection.
-
-I do not argue that the government welfare function in Expression~\ref{eq:weighted} perfectly represents the preferences of those who negotiate trade agreements.
+I do not argue that the government welfare function in Expression~\ref{eq:wel} perfectly represents the preferences of those who negotiate trade agreements.
 I only want to make the point that assumptions on the features of the government objective function can be important drivers of results on the role that trade agreements can play.
@@
 
+\begin{comment}
 \vskip.3in
@@
 
-\begin{comment}
 \vskip.2in

### 2017-01-14 figures.R

@@ -22,9 +22,14 @@ for (j in 1:length(gtn)) {
 W = CSx + gtn2*PSx + CSy + PSy + TR  #partially weighted welfare
 plot(t,W, type = "l",xlab = "Tariff",ylab = "Gov't Welfare")
 
-e = (gtn2 - 1)^10
-W2 = CSx + gtn2*PSx + CSy + PSy + TR - e  #partially weighted welfare - e
+e = (gtn2 - 1)^5
+e = -(log(2 - gtn2))/30
+W2 = CSx + gtn2*PSx + CSy + PSy + TR - gtn2*e  #partially weighted welfare - e
 plot(t,W2, type = "l",xlab = "Tariff",ylab = "Gov't Welfare")
+g = which.max(W2)
+
+P = PSx - e
+l = which.max(P) #lobby's optimum
 
                           
 ##########################################################

### 2017-01-15 draft.tex
remove examples from sec:objfcn

@@
   W(\ga(e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
-  \tag{\ref{eq:baldwin}}
+  \label{eq:baldwin}
 \end{equation}
@@
 Thus, under the government welfare function in Expression~\ref{eq:wel}, government welfare is maximized when $\ga$ takes on the highest value achievable.
+This is because the governments can expect that ex-post political pressure will increase in the trade agreement tariff, and government welfare increases everywhere in political pressure.
 
+\begin{comment}
 This can be seen in Figure~\ref{fig:unweight}, which displays a graph of the government welfare function from Expression~\ref{eq:baldwin}.
@@
 You can see that from $\tau=0.05$---the tariff the home government optimally sets when making unilateral policy with no political pressure---government welfare steadily increases as the trade agreement tariff increases.
-This is because the governments can expect that ex-post political pressure will increase in the trade agreement tariff, and government welfare increases everywhere in political pressure.
 
@@
 \end{figure}
+\end{comment}
 
@@
 For most specifications of the $\ga(e)$ function, there is an interior optimum.
+If a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
+By setting a tariff cap below the lobby's optimal tariff, the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.
+There are many candidates for the government objective function that are not monotonic in political pressure.
+For example, in \Textcite{ethier2012}, returns to lobbying decrease as tariffs increase.
+In the current model, one could take government welfare to be quadratic in $\ga$ or make the weights on the terms in the objective function sum to one.
+
+\begin{comment}
 Figure~\ref{fig:weight} illustrates for the case of $\ga(e) = 1 + e^{.2}$.
@@
 Here $ga(0.0065) = 1.37$.
-The lobby's optimum is at $\tau = But if a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
 
@@
 
-Here, the optimum occurs at $\tau = 0.142$ and $e = 0.0065$, which is about $6\%$ of profits in the import-competing industry.
-Here $ga(0.0065) = 1.60$.
-But if a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
+Here, the optimum occurs at $\tau = 0.142$ and $e = 0.0065$, which is about $6\%$ of profits in the import-competing industry. Here $ga(0.0065) = 1.60$.
 
 In this case, again the government does not have means to encourage any additional lobbying, but it can improve its situation by discouraging lobbying. 
-By setting a tariff cap of zero---assuming the conditions for self enforcement outlined in Section~\ref{sec:selfweak} are met---the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.
-The stark prediction that the best the government can do is to set the trade agreement to zero would be softened in an environment such as that in Section~\ref{sec:escape2} where there are both endogenous and exogenous sources of political pressure or with a government welfare function that has an interior global maximum in $\ga$.
-There are many candidate for the government objective function in which the maximum is strictly interior.
-For example, in \Textcite{ethier2012}, returns to lobbying decrease as tariffs increase.
-In the current model, one could take government welfare to be quadratic in $\ga$ or to simply subtract lobbying effort from Equation~\ref{eq:wel}.
+\end{comment}
 

### 2017-01-15 draft.tex
generalize economic model to GH95

@@
 \label{sec:stage}
+\begin{comment}
 I employ a two good partial equilibrium model with two countries: home (no asterisk) and foreign (asterisk).
@@
 Market clearing implies that world and home prices of $X$ are $P_X^W = \frac{4-3\tau}{7}$ and $P_X = \frac{4+4\tau}{7}$, symmetric for $Y$.
+\end{comment}
 
-As is standard in the literature, I assume that the production of each good requires the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factor is tied to the price of the good in whose production their factor is used.
+\subsection{The Economic Setting}
+\label{sec:economic}
+I begin by describing the economic setting within which trade occurs.\footnote{Note this is the economic environment of \Textcite{gh95} specialized to two tradable goods with symmetry assumed for simplicity of exposition.} It is a three-good model with two countries: home (no asterisk) and foreign (asterisk).
+In the home country, the representative consumer maximizes
+\[
+  u = c_N + u_X(c_X) + u_Y(c_Y).
+\]
+where $c_i$ for $i \in \left\{i,X,Y\right\}$ denotes consumption of the numeraire, X and Y goods respectively.
+The functions $u_X(\cdot)$ and $u_Y(\cdot)$ are assumed to be increasing, strictly concave and differentiable.
+The demand functions for $X$ and $Y$ are written $D(P_i)$ where $D(\cdot)$ is the inverse of $u'_i(\cdot)$ and $P_i$ ($P_i^*$) denotes the home (foreign) price of good $i \in \left\{N,X,Y\right\}$.
+Demand is assumed symmetric in foreign.
+
+Good $N$ is produced with labor alone so that $Q_N = l_N$.
+I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is positive.
+The production of goods $X$ and $Y$ requires labor and a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
+Production of these goods is also assumed to exhibit constant returns to scale.
+The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
+For any such $P_X$, I assume $Q_X^*(P_X) > Q_X(P_X)$ so that the home country is a net importer of good $X$.
+The production structure for good $Y$ is symmetric.
+
+For simplicity, I assume each government's only trade policy instrument is a specific tariff on its import-competing good: the home country levies a tariff $\tau$ on good $X$ while the foreign country applies a tariff $\tau^*$ to good $Y$.
+Local prices are then $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y = P_Y^W$ and $P_Y^* = P_Y^W + \tau^*$ where a $W$ superscript indicates world prices.
+Equilibrium prices are determined by the market clearing conditions
+$$M_X(P_X)= D(P_X)-Q_X(P_X) = Q_X^*(P_X^*) - D(P_X^*) = E_X^*(P_X^*)$$
+$$E_Y(P_Y)=Q_Y(P_Y)-D(P_Y) = D(P_Y^*)-Q_Y^*(P_Y) = M_Y^*(P_Y^*)$$
+where e.g. $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$. The price of the numeraire is equal to one in both countries and on the world market.
+
+$P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff.
+This gives rise to a terms-of-trade externality.
+Profits in a sector are increasing in the price of its good and also in the domestic tariff.
+This fact, combined with the assumptions on specific factor ownership, motivates political activity by import-competing lobbies.
 
@@
 They assume that $\ga$ and $\ga^*$ are each drawn independently from probability distributions with cumulative distribution functions $H(\ga)$ with $h(\ga)=H'(\ga)$.
-The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga} < \frac{7}{4}$.
+The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga}$ is small enough that the government's optimal tariff will be non-prohibitive.
 They also assume $\ga$ and $\ga^*$ are private information.

### 2017-01-15 draft.tex
add X superscript to lobby's profits

@@
 \begin{equation}
-  U_\text{L} = \pi(\tau(\ga(s,e))) - e
+  U_\text{L} = \pi_X(\tau(\ga(s,e))) - e
   \label{eq:lv3}
@@
 \begin{equation}
-  \frac{1}{1-\de}\left[\pi(\tau_e^R) - e^R\right] \geq \pi(\tau^B) - e^B + \frac{\de}{1-\de}\left[\pi(\tau^N) - e^N\right]
+  \frac{1}{1-\de}\left[\pi_X(\tau_e^R) - e^R\right] \geq \pi_X(\tau^B) - e^B + \frac{\de}{1-\de}\left[\pi_X(\tau^N) - e^N\right]
 	\label{exp:lobby}
@@
 That is, the lefthand side is decreasing and the righthand side is increasing as a direct function of $\de$.
-When taking into account the influence of $\de$ from the government's constraint on $e^B$, for a small part of the parameter space where $e^B$ must be very large, it's possible for the lobby's constraint to eventually loosen as $\de$ increases since $\pi(\tau^B) - e^B$ decreases to the right of $\tau^N$.
+When taking into account the influence of $\de$ from the government's constraint on $e^B$, for a small part of the parameter space where $e^B$ must be very large, it's possible for the lobby's constraint to eventually loosen as $\de$ increases since $\pi_X(\tau^B) - e^B$ decreases to the right of $\tau^N$.
 
@@
 \begin{equation}
-  \left[1+ \de +\de^2 \right] \pi(\tau(s)) \geq \left[1+ \de +\de^2 \right] \left\{\pi(\tau(\ga(s,\ov{e})) - \ov{e} \right\}
+  \left[1+ \de +\de^2 \right] \pi_X(\tau(s)) \geq \left[1+ \de +\de^2 \right] \left\{\pi_X(\tau(\ga(s,\ov{e})) - \ov{e} \right\}
 	\label{exp:lobbyEC}

### 2017-01-15 draft.tex
finish adding numeraire

@@
 \end{equation}
-where $\mathit{CS}$ is consumer surplus, $\pi$ represents profits, $\ga(s,e)$ is the weight placed on profits (producer surplus) in the import-competing industry, $s$ is an exogenous state variable, $e$ is lobbying effort, and $\mathit{TR}$ is tariff revenue.
+where $\mathit{CS}$ is consumer surplus, $\pi$ represents profits, $\ga(s,e)$ is the weight placed on profits (producer surplus) in the import-competing industry, $s$ is an exogenous state variable, $e$ is lobbying effort in terms of the numeraire, and $\mathit{TR}$ is tariff revenue.
 Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, may be affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{While the standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them.

### 2017-01-15 to-do.tex
remove _W from notation since not doing strong bindings anymore
revise sec 3 w/view toward generalizing model
finish adding numeraire

@@
 			\begin{itemize}
-				\item Footnote fn:mrc1 is good place to start to put references to MRC 2007
+				\item After Prop1 is a good place to start to put references to MRC 2007
 				\item The model does admit an interpretation in which the same branch of government both negotiates the trade agreement and decides on the applied tariff ex-post, as long as the one branch is non-unitary.
@@
 			\end{itemize}
-	\end{itemize}
-
-\section{Numeraire}
-Do I need to add a numeraire so I can put $e$ in terms of it?
-\begin{itemize}
-	\item Does DGH97 or similar have it?
 	\end{itemize}
@@
 	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
-	\item Streamline notation. If not doing strong binding anywhere (except in sec:strong), don't need $R$ subscript on $\tau$ and $e$.
 \end{itemize}

### 2017-01-15 draft.tex
move outline of obj fcn explanation into draft.tex

@@
 
-\subsection{The Economic Setting}
+\subsection{Setup}
 \label{sec:economic}
@@
 \label{sec:objfcn}
+Approach: denote the government's welfare function as $W_G(\tau,\tau^*,e) = W(\tau,\tau^*) + g(e)$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare.
+			\begin{itemize}
+				\item This is like DGH97 (they write as $G(a,C)$, so no problem putting an $e$ in there), BS1999 `Economic Theory of GATT' (I think), $W(p,\tilde{p}^w)$
+				\item Careful of transferable vs. non-transferable utility
+					\begin{itemize}
+						\item I've already written $W(\ga(e,s),\tau,\tau^*)$
+						\item It's the truthful contribution schedule in GH94 that implies the constant weighting of $1+a$ for sectors represented by a lobby.
+Won't get this clean, constant weighting with other schedules.
+						\item I claim that the government welfare function presented in this paper can be interpreted as a special case of the general welfare function proposed in DGH97.
+They specify government welfare as $G(a,c)$ where $a$ is the policy vector $\left(\left(\tau,\tau^*\right)\right)$ and $c$ is a vector of payments from each lobby ($e$).
+						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
+In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort.
+This distinction is mainly important in the context of the repeated game.
+Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
+					\end{itemize}
+				\item Then specialize to BS2005 when necessary (tariff cap example, picture for interior max of gov't obj fcn)
+				\item Menu auction framework doesn't work with one lobby because gov't doesn't get any surplus. Switch to MRC07 (bargaining). But want decreasing returns (LT).
+				\begin{itemize}
+					\item The model does admit an interpretation in which the same branch of government both negotiates the trade agreement and decides on the applied tariff ex-post, as long as the one branch is non-unitary.
+The model analyzed in the body of this paper can be given an interpretation in line with Maggi and Rodriguez-Clare 2007 if capital is completely mobile in the long run.
+In Appendix B.1, I compare the results of this paper to those from a model with a unitary government, as this would seem to bring the model most closely into alignment with Maggi and Rodriguez-Clare 2007 and a large part of the literature.
+					\item Consider the following, alternative interpretation of the model.
+Assume there is only one decision-making body but the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated.
+For ease of exposition, take this single decision-making body to be the legislature so that the single decision-making body has the preferences in Expression 2.
+At the ex-ante stage, the welfare function would reduce to that in Expression 4.
+This interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.\footnote{To match the assumption of a social-wefare maximing executive, this requires the additional assumption that $\ga(0)=1.$ The model is qualitatively unchanged for other values of $\ga(0)$ as long as the analogue of Assumption 2 below holds.} Note that this remains a non-unitary model.
+Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g.
+during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$).
+A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results.
+I discuss the unitary version of the model in Appendix B.1.
+Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
+				\end{itemize}
+				\item Make argument that want this version like Baldwin-style literature for comparison
+					\begin{itemize}
+						\item Want to make argument that things go wrong in standard model when you insert $e$ into $\ga$ (also, when $e$ goes into $\ga$ need to subtract $e$ too, and this often results in interior optimum)
+						\item This is an easier way to get the result from MRC07 / LT, can make comparisons, develop design results more easily
+					\end{itemize}
+				\item Want non-contingent set-up for different assumptions on timing / meaning of $s$; contingent only makes sense if $e$ chosen after $s$ and $s$ is completely verifiable to lobby.
+If $s$ is political uncertainty, it probably realizes after/contemporaneously with $e$
+					\begin{itemize}
+						\item I can leave general, then specify for escape clause. Or I can just go with my order of $s$ then $e$, which assumes $s$ is an economic shock.
+						\item Contingent payments are hard to reconcile with uncertainty
+					\end{itemize}
+				\item Bargaining framework is not appealing when gov't is non-unitary
+				\item Limao $\&$ Tovar: lobby objective the same as mine.
+$G = \Psi(C) + aW(p)$.
+Gov't decides whether to commit in int'l agreement, then Nash bargains with lobby over $C$ and policies $(t,\tau)$ where $\tau$ is a NTB
+					\begin{itemize}
+						\item Careful of transferable vs. non-transferable utility
+						\item Gov't doesn't care about \textit{anything} because it gets the same payoff no matter what
+							\begin{itemize}
+								\item Menu auction eqm requirement: $(C(\tau),\tau)$ must maximize joint surplus
+								\item With one lobby, this lobby-has-all-the-bargaining-power feature is an equilibrium implication of the menu-auction set-up where the lobby makes a TIOLI offer
+								\item Gov't gets all the surplus when there are two or more lobbies
+								\item This result holds in DGH97, p. 767. Footnote 10 explicitly says that this is because lobby makes TIOLI offer.
+								\item So go to MRC07 where bargaining power can be manipulated
+							\end{itemize}
+						\item Section 3.3: They show exactly what I need to show.
+If gov't can commit with this obj fcn, it will as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying.
+						\item Gov't does not have an interest in maximizing joint surplus, because it doesn't appropriate all the gains
+					\end{itemize}
+				\item Only difference between Limao / Tovar and mine is $\Psi(e)$ vs. $[\ga(e) - 1]\pi_x(\tau)$
+					\begin{itemize}
+						\item This means there's a positive cross partial between $e$ and $\tau$ in my formulation, so that an increase in $\tau$ increases $W$ more the higher is $e$
+					\end{itemize}
+			\end{itemize}
+
 When political pressure is determined endogenously, a government can predict the political pressure it will face.

### 2017-01-15 to-do.tex
move outline of obj fcn explanation into draft.tex

@@
 	\begin{itemize}
-		\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
-			\begin{itemize}
-				\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
-			\end{itemize}
 		\item I like the way you use Bagwell and Staiger (2005) as your benchmark.
@@
 				\item After Prop1 is a good place to start to put references to MRC 2007
-				\item The model does admit an interpretation in which the same branch of government both negotiates the trade agreement and decides on the applied tariff ex-post, as long as the one branch is non-unitary.
-The model analyzed in the body of this paper can be given an interpretation in line with Maggi and Rodriguez-Clare 2007 if capital is completely mobile in the long run.
-In Appendix B.1, I compare the results of this paper to those from a model with a unitary government, as this would seem to bring the model most closely into alignment with Maggi and Rodriguez-Clare 2007 and a large part of the literature.
-				\item Consider the following, alternative interpretation of the model.
-Assume there is only one decision-making body but the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated.
-For ease of exposition, take this single decision-making body to be the legislature so that the single decision-making body has the preferences in Expression 2.
-At the ex-ante stage, the welfare function would reduce to that in Expression 4.
-This interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.\footnote{To match the assumption of a social-wefare maximing executive, this requires the additional assumption that $\ga(0)=1.$ The model is qualitatively unchanged for other values of $\ga(0)$ as long as the analogue of Assumption 2 below holds.} Note that this remains a non-unitary model.
-Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g.
-during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$).
-A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results.
-I discuss the unitary version of the model in Appendix B.1.
-Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
+
 				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007.
@@
 					\end{itemize}
-				\item Appendix B.2 with analysis of DGH model
-					\begin{itemize}
-						\item I claim that the government welfare function presented in this paper can be interpreted as a special case of the general welfare function proposed in DGH97.
-They specify government welfare as $G(a,c)$ where $a$ is the policy vector $\left(\left(\tau,\tau^*\right)\right)$ and $c$ is a vector of payments from each lobby ($e$).
-
-						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
-In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort.
-This distinction is mainly important in the context of the repeated game.
-Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
-					\end{itemize}
 			\end{itemize}
@@
 			\begin{itemize}
-				\item Contingent payments are hard to reconcile with uncertainty
-				\item It's the truthful contribution schedule in GH94 that implies the constant weighting of $1+a$ for sectors represented by a lobby.
-Won't get this clean, constant weighting with other schedules.
-				\item Gov't doesn't care about \textit{anything} because it gets the same payoff no matter what
-					\begin{itemize}
-						\item With one lobby, this lobby-has-all-the-bargaining-power feature is an equilibrium implication of the menu-auction set-up where the lobby makes a TIOLI offer
-						\item Gov't gets all the surplus when there are two or more lobbies
-						\item This result holds in DGH97, p. 767. Footnote 10 explicitly says that this is because lobby makes TIOLI offer.
-						\item So go to MRC07 where bargaining power can be manipulated
-					\end{itemize}
 				\item Menu auction eqm requirement: $(C(\tau),\tau)$ must maximize joint surplus
@@
 Not clear how this is related to lobby's problem, but likely to have higher $\tau$ and lower $e$ than government's optimum (unless $\ga$ changes very steeply between the two points $e_J$ and $e_G$)
-						\item Careful of transferable vs. non-transferable utility
 					\end{itemize}
@@
 					\begin{itemize}
-						\item This means there's a positive cross partial between $e$ and $\tau$ in my formulation, so that an incraese in $\tau$ increases $W$ more the higher is $e$
+						\item This means there's a positive cross partial between $e$ and $\tau$ in my formulation, so that an increase in $\tau$ increases $W$ more the higher is $e$
 					\end{itemize}
@@
 			\item The paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
-			\item So I'm wondering whether you could dispense with most of the static analysis and get into the repeated game more quickly.
 		\end{itemize}

### 2017-01-15 draft.tex
finish revising sec:objfcn w/connection to literature

--- 2017-01-15 draft.tex
@@
 Here, the weight the government places on the profits of the import-competing industry, $\ga(s,e)$, may be affected both by a shock (due to \Textcite{ms2011}) and by the level of lobbying effort.\footnote{While the standard `Protection for Sale' modeling of \Textcite{gh94} would specify $W = C + aW$, this form allows the incorporation of endogenous political pressure into the government objective function that is most often used to examine questions concerning the design of international trade agreements and the institutions that facilitate them.
-An isomorphism can be made between the two forms as discussed in \Textcite{buzard2013b}.} I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to the case where political pressure  results from the rent-seeking behavior of lobbies only.
+See Section~\ref{sec:objfcn} for a detailed discussion.} I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to the case where political pressure  results from the rent-seeking behavior of lobbies only.
 
@@
 \label{sec:objfcn}
-When political pressure is determined endogenously, a government can predict the political pressure it will face.
-I argue below that in some cases a government may want to use a trade agreement to alter lobbying incentives.
-If this is true, then the perceived role that trade agreements can play is influenced by the choice of government objective function.
+There are several advantages to using the government objective function in Expression~\ref{eq:wel}, first of all tractability for analyzing institutional design questions.
+I do not propose this alternative lightly, and it is closely related to objective functions that are well-studied in the literature.
 
-Let's begin with the \Textcite{baldwin}-style government objective function that is most common in the recent literature on trade agreements, specialized to the case of endogenous pressure only.
+\Textcite{gh94}'s well-known government objective function both restricts the returns to lobbying to be constant and assumes that utility is transferable between the government and lobby, preventing analysis of distributional questions.
+\Textcite{dgh97} point this out and proposes a more general form where the government objective is a function of a policy vector payments from each interest group (here $\left(\tau,\tau^*\right)$) and $e$, respectively).
+When there are no exogenous shocks, Expression~\ref{eq:wel} is therefore a special case of the \Textcite{dgh97} objective function.
+
+Part of the appeal of the \Textcite{gh94} model is that it provides theoretical foundations for the standard \Textcite{baldwin}-style political objective function that is often used in the literature on the design of trade institutions:
 \begin{equation}
-  W(\ga(e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga(e) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
+  W(\ga(e),\tau,\tau^*) = \mathit{CS}_X(\tau) + \ga \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
   \label{eq:baldwin}
 \end{equation}
-The government chooses $\tau$ according to Equation~\ref{eq:eff} when setting tariffs in the context of a trade agreement.
-The government knows that this trade agreement tariff will impact lobbying incentives and therefore the realization of $\ga(e)$ at the applied tariff-setting phase.
-We want to know how $\ga(e)$ affects the government's welfare so that we can see how the desire to manipulate lobbying incentives may influence the government's choice of trade agreement tariff.
 
-Envelope-theorem style results imply that the first order condition is the same in the unilateral and joint problems: $\displaystyle \frac{\partial W_x}{\partial \ga} = 0$.
+The \Textcite{gh94} result is that $\ga$ must equal $1+a$, where $a$ is the weight the government places on social welfare.
+The literature that uses both the endogenous and the exogenous versions of this welfare function has typically concluded that without some kind of long-run inefficiency, governments will not want to use the trade agreement to reduce lobbying.
+That is, trade agreements internalize the terms of trade externality but nothing more.
+Careful inspection of Expression~\ref{eq:baldwin} helps explain why.
+If we ask where the maximum of the objective function is in terms of $\ga$, an envelope-theorem style result implies that the first order condition is the same in the unilateral and joint problems: $\displaystyle \frac{\partial W_x}{\partial \ga} = 0$.
 See the \hyperlink{envelope}{Appendix} for details.
@@
 This is always positive.
-Thus, under the government welfare function in Expression~\ref{eq:wel}, government welfare is maximized when $\ga$ takes on the highest value achievable.
-This is because the governments can expect that ex-post political pressure will increase in the trade agreement tariff, and government welfare increases everywhere in political pressure.
+Thus, under the government welfare function in Expression~\ref{eq:baldwin}, government welfare is maximized when $\ga$ takes on the highest value achievable.
+
+Thus the objective function in \Textcite{gh94} and Expression~\ref{eq:baldwin} are not well-suited to answering the question of whether governments would like to use trade agreements as a political commitment device because the functional forms rule this out by assumption.
+For the exogenous version to be well micro-founded, it also requires the assumption that lobbies present truthful contribution schedules that are contingent, take-it-or-leave-it offers.
+With one lobby, this means the government never receives any surplus.
+Even if the government did not prefer as much political pressure as possible, in the menu-auction set-up the government is made indifferent between all outcomes.
+This feature remains in the more general \Textcite{dgh97} model because it's a result of the take-it-or-leave-it offer of the lobby.
+
+We would like a framework in which the government's payoff is not constant.
+An appealing candidate is \Textcite{mrc2007}, which is in the spirit of \Textcite{gh94} but uses a bargaining framework.
+Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights.
+
+In \Textcite{lt}'s setup, the government, with objective function $G = \Psi(e) + aW(\tau,\tau^*)$, decides whether to commit in an international agreement and then Nash bargains with the lobby over $e$ and the tariff.\footnote{I have changed notation to match that which I use here and restricted attention to their Section 3.3.
+with the tariff only; their main results address the choice between tariffs and non-tariff barriers.} They show that the government will use the trade agreement to commit vis-\`{a}-vis the lobby as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying.
+Here, the government does not have an interest in maximizing joint surplus because it is unable to appropriate all the gains.\footnote{In the working paper version of \Textcite{mrc2007}, a similar result is demonstrated for the case of immobile capital when the government has all the bargaining power.}
+
+The setup that I have assumed has very similar dynamics.
+When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, the monotonicity in $\ga$ inherent in \ref{eq:baldwin} typically breaks down.
+For most specifications of the $\ga(e)$ function, there is an interior optimum.
+If a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin} with $e$ endogenized, the government will likely be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
+By setting a tariff cap below the lobby's optimal tariff, the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.\footnote{There are many candidates for the government objective function that are not monotonic in political pressure.
+For example, in \Textcite{ethier2012}, returns to lobbying decrease as tariffs increase.
+In the current model, one could take government welfare to be quadratic in $\ga$ or make the weights on the terms in the objective function sum to one.}
+
+I do not argue that the government welfare function in Expression~\ref{eq:wel} perfectly represents the preferences of those who negotiate trade agreements.
+I only want to make the point that assumptions on the features of the government objective function can be important drivers of results on the role that trade agreements can play.
+And this can potentially provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
+
+There are several additional advantages to the setup assumed here.
+First, it allows me to take lobbying incentives into account in the same framework typically used in the literature to examine institutional design questions so that comparisons to the existing literature are direct.
+Second, it is very tractable, making it easily adopted to ask how endogenous politics changes the answers to a wide range of policy questions.
+Third, if one prefers a non-unitary interpretation of the government, this framework in which the lobby exerts effort that determines the identity of the pivotal decision-maker within the government is more appealing than bargaining.
+Finally, in an environment where exogenous shocks are present, contingent lobbying effort / payments are hard to reconcile with many types of uncertainty.
+
+If one takes $a=1$, the only difference between the \Textcite{lt} government objective function and Expression~\ref{eq:wel} is that in place of their $\Psi(e)$, I have $[\ga(e) - 1]\pi_x(\tau) - e$.\footnote{Note that this is still consistent with the \Textcite{dgh97} form, although of course I have the lobby maximizing net profits instead of making a take-it-or-leave-it offer.}	This means there's a positive cross derivative between lobbying effort and tariffs in my formulation, so that an increase in tariffs increases government welfare more the higher is lobbying effort.
 
@@
 \end{comment}
-
-When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, this monotonicity often breaks down.
-For most specifications of the $\ga(e)$ function, there is an interior optimum.
-If a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin}, the government may be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
-By setting a tariff cap below the lobby's optimal tariff, the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.
-There are many candidates for the government objective function that are not monotonic in political pressure.
-For example, in \Textcite{ethier2012}, returns to lobbying decrease as tariffs increase.
-In the current model, one could take government welfare to be quadratic in $\ga$ or make the weights on the terms in the objective function sum to one.
 
@@
 \end{comment}
-
-I do not argue that the government welfare function in Expression~\ref{eq:wel} perfectly represents the preferences of those who negotiate trade agreements.
-I only want to make the point that assumptions on the features of the government objective function can be important drivers of results on the role that trade agreements can play.
-
-Examining these alternatives in combination with endogenous lobbying points out that the idea that trade agreements can function as domestic political commitment devices may be more general than previously demonstrated.
-It can also provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
 

### 2017-01-15 to-do.tex
finish revising sec:objfcn w/connection to literature

@@
 
-				\item I have added Appendix B.1 with analysis of the unitary model
+				\item Unitary vs. non-unitary
 					\begin{itemize}
+						\item Assume the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated.
+At the ex-ante stage, $e=0$.
+If this government is also unitary, then this interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
+						\item Note that this remains a non-unitary model.
+Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g.
+during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$).
+A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results.
+Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
+						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
+In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort.
+This distinction is mainly important in the context of the repeated game.
+Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
 						\item In the model with a non-unitary government, the lobby's effort $e$ determines the identity of the decision maker.

### 2017-01-16 to-do.tex
cut down to-do to account for what I've finished

@@
 				\end{itemize}
+								\item Unitary vs. non-unitary
+					\begin{itemize}
+						\item Assume the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated.
+At the ex-ante stage, $e=0$.
+If this government is also unitary, then this interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
+						\item Note that this remains a non-unitary model.
+Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g.
+during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$).
+A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results.
+Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
+						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
+In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort.
+This distinction is mainly important in the context of the repeated game.
+Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
+						\item In the model with a non-unitary government, the lobby's effort $e$ determines the identity of the decision maker.
+Thus the decision maker at the time the lobby is pushing for the agreement to be broken ($e=e_b>e_a$ in order to have a median legislator who will break the agreement) is different from the decision maker during the trade war phase $(e = e_{tw})$ and the decision maker during a trade agreement phase ($e=e_a(\tau^a)$).
+The weight on the lobbying industry's profits changes with lobbying effort because lobbying effort determines a different median voter in the legislature or government more generally.
+However, when the median voter during a `break' phase is evaluating her incentive constraint, she values future tariff choices --- which will be determined by whoever is median in the future --- with her own preferences.
+Although she will not be the decision maker in the future, there is no reason for her to evaluate future welfare according to some other legislator's preference.
+						\item If the model is interpreted as having a unitary government, there is only one decision maker.
+There is a fixed mapping from lobbying effort to the weight the decision maker places on the lobby's profits, and the realization of this weight changes with the realization of lobbying effort.
+Thus, predicting future lobbying effort, the decision maker knows what the realization of her weight on the lobby's profits will be and evaluates the incentive constraint accordingly.
+Expression (10) would be modified while the lobbying constraint is unchanged.
+					\end{itemize}
 		\end{itemize}
@@
 Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
-
-				\item Unitary vs. non-unitary
-					\begin{itemize}
-						\item Assume the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated.
-At the ex-ante stage, $e=0$.
-If this government is also unitary, then this interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
-						\item Note that this remains a non-unitary model.
-Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g.
-during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$).
-A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results.
-Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
-						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
-In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort.
-This distinction is mainly important in the context of the repeated game.
-Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
-						\item In the model with a non-unitary government, the lobby's effort $e$ determines the identity of the decision maker.
-Thus the decision maker at the time the lobby is pushing for the agreement to be broken ($e=e_b>e_a$ in order to have a median legislator who will break the agreement) is different from the decision maker during the trade war phase $(e = e_{tw})$ and the decision maker during a trade agreement phase ($e=e_a(\tau^a)$).
-The weight on the lobbying industry's profits changes with lobbying effort because lobbying effort determines a different median voter in the legislature or government more generally.
-However, when the median voter during a `break' phase is evaluating her incentive constraint, she values future tariff choices --- which will be determined by whoever is median in the future --- with her own preferences.
-Although she will not be the decision maker in the future, there is no reason for her to evaluate future welfare according to some other legislator's preference.
-						\item If the model is interpreted as having a unitary government, there is only one decision maker.
-There is a fixed mapping from lobbying effort to the weight the decision maker places on the lobby's profits, and the realization of this weight changes with the realization of lobbying effort.
-Thus, predicting future lobbying effort, the decision maker knows what the realization of her weight on the lobby's profits will be and evaluates the incentive constraint accordingly.
-Expression (10) would be modified while the lobbying constraint is unchanged.
-					\end{itemize}
-			\end{itemize}
-		\item Comparison to GH94
-			\begin{itemize}
-				\item Menu auction eqm requirement: $(C(\tau),\tau)$ must maximize joint surplus
-					\begin{itemize}
-						\item NB: $\tau$ that maximizes gov't welfare doesn't generically maximize joint surplus (unless it also maximizes lobby payoff)
-						\item In my model, optimum for lobby would be prohibitive tariff, $e=0$.
-Joint optimum determined by $\frac{\partial \ga}{\partial e}\pi_X = 2$ and $\frac{\partial CS_X}{\partial \tau} + \left[ 1 + \ga(e)\right]\frac{\partial \pi_X}{\partial \tau} + \frac{\partial TR}{\partial \tau} = 0$.
-Not clear how this is related to lobby's problem, but likely to have higher $\tau$ and lower $e$ than government's optimum (unless $\ga$ changes very steeply between the two points $e_J$ and $e_G$)
-					\end{itemize}
-				\item Not sure if contingent vs. non-contingent payments are a big deal
-				\item If I change nothing else but subtract $C$ from gov't welfare function, no qualitative change b/c of 0 bargaining weight
 			\end{itemize}
@@
 					\end{itemize}		
-				\item Then re-do MRC07 with $W(\tau) + g(e)$, and version where I subtract $e$
-					\begin{itemize}
-						\item Should have gov't make lobby indifferent and maximize joint surplus. But again, this doesn't necessarily maximize gov't welfare
-						\item Or rather, should I follow Limao and Tovar?
 			\end{itemize}
-			\end{itemize}
-		\item Approach: denote the government's welfare function as $W_G(\tau,\tau^*,e) = W(\tau,\tau^*) + g(e)$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare.
-			\begin{itemize}
-				\item This is like DGH97 (they write as $G(a,C)$, so no problem putting an $e$ in there), BS1999 `Economic Theory of GATT' (I think), $W(p,\tilde{p}^w)$
-					\begin{itemize}
-						\item I've already written $W(\ga(e,s),\tau,\tau^*)$
-						\item I think people are most concerned with underlying economic model. I need a numeraire anyway. DGH97 doesn't have numeraire b/c small country
-					\end{itemize}
-				\item Then specialize to BS2005 when necessary (tariff cap example, picture for interior max of gov't obj fcn)
-				\item Menu auction framework doesn't work with one lobby because gov't doesn't get any surplus. Switch to MRC07 (bargaining). But want decreasing returns (LT).
 		\item Make argument that want this version like Baldwin-style literature for comparison
@@
 				\item This is an easier way to get the result from MRC07 / LT, can make comparisons, develop design results more easily
-					\end{itemize}
-				\item Want non-contingent set-up for different assumptions on timing / meaning of $s$; contingent only makes sense if $e$ chosen after $s$ and $s$ is completely verifiable to lobby.
-If $s$ is political uncertainty, it probably realizes after/contemporaneously with $e$
-					\begin{itemize}
-						\item I can leave general, then specify for escape clause. Or I can just go with my order of $s$ then $e$, which assumes $s$ is an economic shock.
-					\end{itemize}
-				\item Bargaining framework is not appealing when gov't is non-unitary
-				\item Limao $\&$ Tovar: lobby objective the same as mine.
-$G = \Psi(C) + aW(p)$.
-Gov't decides whether to commit in int'l agreement, then Nash bargains with lobby over $C$ and policies $(t,\tau)$ where $\tau$ is a NTB
-					\begin{itemize}
-						\item Section 3.3: They show exactly what I need to show.
-If gov't can commit with this obj fcn, it will as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying.
-						\item Gov't does not have an interest in maximizing joint surplus, because it doesn't appropriate all the gains
-					\end{itemize}
-				\item Only difference between Limao / Tovar and mine is $\Psi(e)$ vs. $[\ga(e) - 1]\pi_x(\tau)$
-					\begin{itemize}
-						\item This means there's a positive cross partial between $e$ and $\tau$ in my formulation, so that an increase in $\tau$ increases $W$ more the higher is $e$
-					\end{itemize}
 			\end{itemize}
@@
 		\begin{itemize}
-			\item and make clear that GH is isomorphic / provides microfoundations for Baldwin objective function
 			\item More on Coates $\&$ Ludema (2001)
@@
 		\end{itemize}
-	\item I think that I need to bring people along more carefully from GH/MRC to what I'm doing
 	\item Paola: How do I want to sell the paper?
@@
 		\end{itemize}
-	\item Limao and Tovar 2011
-		\begin{enumerate}
-			\item Their 2nd (and final) stage is Nash bargaining. Is there an isomorphism to my setup?
-			\item It IS clear to take their first, limited case of commitment to tariff cap only (ignore NTB) as analogy for gov't in my model choosing commitment to different tariff level.
-Need to answer Q1 above in order to answer points 3 and 4:
-			\item Their gains from commitment derive from improvement in bargaining power vis-a-vis the lobby (Schelling conjecture invoked). How does this relate to the gain from commitment in my model?
-			\item Top of page 6: ``large enough bargaining power.'' Can I make connection to $\ga(e)$?
-			\item Note: in their equation 3, $C$ enters expression for NBS's ($\tau$) in equilibrium, it doesn't disappear as in GH
-		\end{enumerate}
 \end{itemize}
@@
 
-\section{Take out repeats}
-	Especially of references to BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3.
-Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point.
-
 \section{Misc.}
 \begin{itemize}
+	\item Specialize to BS2005 when necessary (tariff cap example)
 	\item Define $\ga$ before using it

### 2017-01-16 draft.tex
cannibalize incentive compatibility subsection under model

@@
 In the first stage, the governments choose the trade agreement tariffs via a negotiating process that I assume to be efficient.
-This process therefore maximizes the joint payoffs produced by the trade agreement:
+In this strand of the literature, it is typically assumed that trade agreements are negotiated to maximize the expected welfare of politically motivated governments given the exogenously-determined political pressure they expect to face at the time of the agreement's implementation.
+I will follow this convention throughout, making necessary adjustments for endogenous politics.
+The trade negotiation therefore is assumed to maximize the joint payoffs produced by the trade agreement:
 \begin{equation}
@@
 \end{equation}
-As in the applied tariff problem, I assume the realization of $\ga$ is below the threshold need to ensure the second order condition is satisfied.
+As in the applied tariff problem, I assume the realization of $\ga$ is below the threshold needed to ensure the second order condition is satisfied.
 
@@
 
-\subsection{Exogenous Political Pressure and Incentive Compatibility}
-In order to compare the results for the case of endogenous political pressure to \Textcite{bs2005} results for exogenous political shocks, I provide the essentials of their setup.
-
-They assume that $\ga$ and $\ga^*$ are each drawn independently from probability distributions with cumulative distribution functions $H(\ga)$ with $h(\ga)=H'(\ga)$.
-The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga}$ is small enough that the government's optimal tariff will be non-prohibitive.
-They also assume $\ga$ and $\ga^*$ are private information.
-
-This leads to a concern about incentive compatibility.
-It must be in a government's best interest to truthfully reveal its $\ga$ or else it will misrepresent its private information about $\ga$ in order to raise tariffs and create a terms-of-trade gain, thereby improving its unilateral welfare.
-They provide conditions that guarantee incentive compatibility on page 481.
-
-On a separate note, it is important to make clear that in this strand of the literature, it is assumed that trade agreements are negotiated to maximize the expected welfare of politically motivated governments given the exogenously-determined political pressure they expect to face at the time of the agreement's implementation.
-I will follow this convention throughout.
-
 \section{Rigid Tariffs with Endogenous Political Pressure}
@@
 
-This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.\footnote{Note that in their treatment, Bagwell and Staiger (2005) assume linear supply and demand.} In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.
+This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.\footnote{\Textcite{bs2005} assume that $\ga$ and $\ga^*$ are each drawn independently from probability distributions with cumulative distribution functions $H(\ga)$ with $h(\ga)=H'(\ga)$.
+The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga}$ is small enough that the government's optimal tariff will be non-prohibitive.
+They also assume $\ga$ and $\ga^*$ are private information.
+They also assume linear supply and demand.} In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.
 
@@
 
-However, because the tariff allowed under the trade agreement with an escape clause varies with the announced level of political pressure, the incentive compatibility constraint now comes into play: a government may gain an advantage by misreporting the level of $\ga$ it experiences.
-In their Proposition 4, Bagwell and Staiger show that a trade agreement with a costless escape clause cannot improve welfare because it cannot be made incentive compatible.
+However, because the tariff allowed under the trade agreement with an escape clause varies with the announced level of political pressure, introducing an escape clause leads to a concern about incentive compatibility.
+It must be in a government's best interest to truthfully reveal its $\ga$ or else it will misrepresent its private information about $\ga$ in order to raise tariffs and create a terms-of-trade gain, thereby improving its unilateral welfare.
+\Textcite{bs2005} provide conditions that guarantee incentive compatibility, but they also show that a trade agreement with a costless escape clause cannot improve welfare because it cannot be made incentive compatible.
 No matter the realization of the stochastic $\ga$, it is always in the government's interest to announce that $\ga$ is high, which allows it to apply any tariff up to the higher weak binding.

### 2017-01-16 draft.tex
rewrite end of sec:objfcn

@@
 
-The setup that I have assumed has very similar dynamics.
-When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, the monotonicity in $\ga$ inherent in \ref{eq:baldwin} typically breaks down.
+The setup that I have assumed is similar to \Textcite{lt} and produces very similar dynamics.\footnote{If one takes $a=1$, the only difference between the \Textcite{lt} government objective function and Expression~\ref{eq:wel} is that Expresion~\ref{eq:wel} replaces $\Psi(e)$ with $[\ga(e) - 1]\pi_x(\tau) - e$.
+This means there's a positive cross derivative between lobbying effort and tariffs in my formulation, so that an increase in tariffs increases government welfare more the higher is lobbying effort.
+Note that this is still consistent with the \Textcite{dgh97} form, although I have the lobby maximizing net profits instead of making a take-it-or-leave-it offer.
+} When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, the monotonicity in $\ga$ inherent in \ref{eq:baldwin} typically breaks down.
 For most specifications of the $\ga(e)$ function, there is an interior optimum.
@@
 There are several additional advantages to the setup assumed here.
-First, it allows me to take lobbying incentives into account in the same framework typically used in the literature to examine institutional design questions so that comparisons to the existing literature are direct.
-Second, it is very tractable, making it easily adopted to ask how endogenous politics changes the answers to a wide range of policy questions.
+First, it is very tractable, making it easily adopted to ask how endogenous politics might change the answers to a wide range of policy questions.
+Second, it allows me to take lobbying incentives into account in the same framework typically used in the literature to examine institutional design questions so that comparisons to the existing literature are direct.
 Third, if one prefers a non-unitary interpretation of the government, this framework in which the lobby exerts effort that determines the identity of the pivotal decision-maker within the government is more appealing than bargaining.
 Finally, in an environment where exogenous shocks are present, contingent lobbying effort / payments are hard to reconcile with many types of uncertainty.
-
-If one takes $a=1$, the only difference between the \Textcite{lt} government objective function and Expression~\ref{eq:wel} is that in place of their $\Psi(e)$, I have $[\ga(e) - 1]\pi_x(\tau) - e$.\footnote{Note that this is still consistent with the \Textcite{dgh97} form, although of course I have the lobby maximizing net profits instead of making a take-it-or-leave-it offer.}	This means there's a positive cross derivative between lobbying effort and tariffs in my formulation, so that an increase in tariffs increases government welfare more the higher is lobbying effort.
 

### 2017-01-16 to-do.tex
rewrite end of sec:objfcn

@@
 					\end{itemize}		
-			\end{itemize}
-		\item Make argument that want this version like Baldwin-style literature for comparison
-			\begin{itemize}
-				\item Want to make argument that things go wrong in standard model when you insert $e$ into $\ga$ (also, when $e$ goes into $\ga$ need to subtract $e$ too, and this often results in interior optimum)
-				\item This is an easier way to get the result from MRC07 / LT, can make comparisons, develop design results more easily
 			\end{itemize}

### 2017-01-16 draft.tex
finish introducing and justifying -e

@@
 See Section~\ref{sec:objfcn} for a detailed discussion.} I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to the case where political pressure  results from the rent-seeking behavior of lobbies only.
+I subtract the cost of lobbying effort from the government's objective function because if stands to reason that if the government cares about lobbying gross profits, it should also care about the costs.\footnote{Aside from making the government's optimum interior in $\ga$, taking account of the cost of lobbying changes the results that follow quantitatively but not qualitatively because there is no marginal impact on the government's decision-making regarding the optimal applied tariff.} This is often abstracted from in the closely-related literature, but it seems important to take into account in this environment of non-transferable utility and decreasing returns to lobbying.
+Note that there is no cost of lobbying effort in models that assume political pressure is exogenous only.
 

### 2017-01-16 to-do.tex
finish introducing and justifying -e

@@
 Expression (10) would be modified while the lobbying constraint is unchanged.
-					\end{itemize}
-		\end{itemize}
-
-\section{Lobbying cost included in gov't welfare}
-	\begin{itemize}
-		\item Kyle: why doesn't government (dis)value expenditure by lobby? Needs to be counted in government welfare function or justify why not
-			\begin{itemize}
-				\item Rohan: $(1-a)\cdot C + a \cdot W \equiv C + a \cdot (W-C)$
-				\item When weights on $\pi_X$ are not linear, what are the implications for resource waste? Is it the same as $+ C / - C$?
-			\end{itemize}
-		\item Subtracting $e$ gives me the kind of non-monotonicity I'd like in the graphs.
-			\begin{itemize}
-				\item Does it do anything to the analysis? Just shifts govt's ideal point
-			\end{itemize}
-		\item I think I'd like to just change to the ``$-e$'' version.
-			\begin{itemize}
-				\item Worries: won't match literature. Want to make analogies to GH, DGH, MRC, LT.
-			\end{itemize}
-		\item Note: there's no ``$e$'' to subtract when $\ga$ is exogenous
-		\item Changes from version without $-e$:
-			\begin{itemize}
-				\item 3.1: $\tau_{W,e}^R$ reduced because gov't optimum is lower. $\tau^N(\ga(e))$ doesn't change because $-e$ is just a welfare shifter, isn't a function of $\tau$ is it is in menu auction setup
-				\item 3.2.2: $\tau_{W,e}^R$ changes, but qualitative result doesn't change.
-Conjecture: more likely that this lower $\tau_{W,e}^R$ is not self-enforcing because it's lower.
-Should be same proof as strong vs.
-weak binding in JIE R$\&$R: $\pi(\tau_{W,e}^R) - e < \pi(\tau_{W,e}^R)$ is smaller, so it's harder to self-enforce.
-Creates a bigger gap with trade war net profit
-				\item 4.1: no change other than level of $\tau_{S,e}^F$
-				\item 4.2: Side payments: need to check how transfers would change, but no obvious reasons
-				\item 4.3: I was implicitly assuming something like this anyway. But I think I want to try a simple version with an export lobby.
-				\item 4.4: Should write (10) more carefully because different $e$'s in three places
 					\end{itemize}
@@
 
-\section{Structured differently from most papers}
-		\begin{itemize}
-			\item How did Devashish structure his AER paper: the argument, the model?
-				\begin{itemize}
-					\item This paper is structured differently than the norm---is there a good reason to do it?
-					\item Don't want to challenge what people are used to unless I really need to
-					\item In general, how do AER papers \textit{look}?
-				\end{itemize}
-			\item If above I suggest to add a section, I think that the paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
-The analysis is also split between strict and weak commitments.
-Which scenario clearly develops the point/contribution you want to make?
-Shouldn't the other one occupy a smaller place (i.e.
-space) of the paper?)
-		\end{itemize}
-
 \section{One good idea / do too much}

### 2017-01-16 draft.tex
integrate explanation of \ga(e) into sec:economic

@@
 
-%\begin{assumption}
- % $\ga(s,e)$ is continuously differentiable, strictly increasing and concave in $e$.
-  %\label{as:ga_c3}
-%\end{assumption}
+\begin{assumption}
+  $\ga(s,e)$ is continuously differentiable, strictly increasing and concave in $e$.
+  \label{as:ga_c3}
+\end{assumption}
 
-%Assumption~\ref{as:ga_c3} formalizes the intuition that the government favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.
+Assumption~\ref{as:ga_c3} represents the idea that the government favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.
+
+This objective function for the government is consistent with \Textcite{bm93}, which explores advertising and the goods being advertised as complementary in a `stable metautility function.' Their consumers are sophisticated and can advocate for policy to change the level of advertising with which they will be confronted.
+Similarly, the government here has a stable mapping from the amount of lobbying effort it experiences to its optimal choice of tariffs and it can use the trade agreement to change the level of lobbying effort with which it will be confronted.
+The government's preferences don't change with $e$, as the preferences related to lobbying effort are represented by the mapping $\ga(\cdot)$.
+When $e$ changes, the metautility function is evaluated at a different value.
+
+As Figure~\ref{fig:ext} makes clear, lobbying, as well as the realization of any shocks, occurs after the trade agreement is negotiated and before the applied tariff decision.
+$\ga(s,e)$ is thus evaluated at the ex-post stage, while the trade agreement is negotiated in the shadow of the realizations of $s$ and $e$.
+We can thus interpret the fundamental structure of preferences in a couple of ways.
+We can think of lobbying effort as a kind of persuasion than affects decision making of a single actor or we can think of lobbying effort as influencing the identify of the ex-post decision-maker.
+In either case, $\ga(s,e)$ is the weight the decision-maker places on the lobbying industry's profits ex-post, and the realization of this weight changes with the realization of the shock and lobbying effort even while the mapping---the preferences---do not.\footnote{The model also admits an interpretation in which the decision-making body is non-unitary, such as a large legislature.
+Here, $s$ and $e$ determine the identity of the median legislator.
+This interpretation results in minor, qualitative changes to the repeated-game results because the non-unitary decision-maker knows that different decision-makers will be in power in the future and evaluates future welfare with her own preferences.
+In contrast, the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future.}
 
@@
 An appealing candidate is \Textcite{mrc2007}, which is in the spirit of \Textcite{gh94} but uses a bargaining framework.
+In my setup, the lobby is not active during the trade agreement negotiation; in order to to fit into the framework of \Textcite{mrc2007}, we can assume that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
+
 Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights.
@@
 Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau_e^R\right)$ so that in some cases the lobby's incentive constraint is violated.
-This would lead to the legislature's constraint being violated as well, creating a dispute.
+This would lead to the government's constraint being violated as well, creating a dispute.
 

### 2017-01-16 to-do.tex
integrate explanation of \ga(e) into sec:economic

@@
 \end{center}
-
-\section{``Dynamic preferences''}
-	\begin{itemize}
-		\item Becker, G. S. and K. M. Murphy (1993), Simple Theory of Advertising as a Good or Bad, Quarterly Journal of Economics, 942-64.
-			\begin{itemize}
-				\item Their firms are my firm. Their consumers are my government
-				\item Kyle's lit review talkes about `persuasive,' `informative,' and `complementary' views of advertisting. This is complementary.
-				\item $e$ and $\tau$ enter objective function.
-Gov't is sophisticated, can make commitment that alters $e$ (i.e.
-behavior of another actor).
-But $CS + \ga(e)PS + \ldots$, just like $u(X,Y,A)$; says consumer could work to pass law about number of ads on TV each hour.
-				\item Use newly drawn extensive form to explain this, lays out when welfare is realized
-				\item Addresses James' concern about needing to be more precise about when $\ga$ is evaluated
-					\begin{itemize}
-						\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
-					\end{itemize}
-			\end{itemize}
-		\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
-				\begin{itemize}
-					\item Giovanni: if I can show properties survive  monotonic transformation, I'm fine
-					\item They don't like gov't choosing its own preferences. May just need to change wording
-				\end{itemize}
-		\item Bob: what is fundamental structure of preferences?
-		\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
-			\begin{itemize}
-				\item ``How the assumption of a non-unitary legislature translates into this incentive constraint.
-In the non-unitary legislature, if the current period incentive constraint includes any future value of lobbying effort, this would mean the current-period median legislator evaluating her current-period incentive constraint using a mixture of her own political-economy weight with those of the legislators who are median in the future in the trade agreement and trade-war scenarios.
-The first major step I have taken to clarify the incentive constraint is the analysis in Appendix B.1 of the unitary model.
-I discuss the differences between the unitary and non-unitary models in the beginning of that Appendix.''
-				\item ``Note that the median legislator, whose identity is determined by the lobby's effort level $e_b$, evaluates future payoffs according to her own political economy weight, $\ga(e_b)$.
-Of course, depending on legislator $e_b$'s choice, either legislator $e_a$ or legislator $e_{tw}$ will be the decision maker in those future periods.
-But legislator $e_b$, who is the decision maker in the current period, maximizes her own welfare given the predicted behavior of future decision makers.''
-			\end{itemize}
-		\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
-				\begin{itemize}
-					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's ``expost $\ga(e)$ as effort to elect/choose ex-post decision-maker''
-				\end{itemize}
-								\item Unitary vs. non-unitary
-					\begin{itemize}
-						\item Assume the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated.
-At the ex-ante stage, $e=0$.
-If this government is also unitary, then this interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
-						\item Note that this remains a non-unitary model.
-Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g.
-during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$).
-A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results.
-Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
-						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}.
-In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort.
-This distinction is mainly important in the context of the repeated game.
-Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
-						\item In the model with a non-unitary government, the lobby's effort $e$ determines the identity of the decision maker.
-Thus the decision maker at the time the lobby is pushing for the agreement to be broken ($e=e_b>e_a$ in order to have a median legislator who will break the agreement) is different from the decision maker during the trade war phase $(e = e_{tw})$ and the decision maker during a trade agreement phase ($e=e_a(\tau^a)$).
-The weight on the lobbying industry's profits changes with lobbying effort because lobbying effort determines a different median voter in the legislature or government more generally.
-However, when the median voter during a `break' phase is evaluating her incentive constraint, she values future tariff choices --- which will be determined by whoever is median in the future --- with her own preferences.
-Although she will not be the decision maker in the future, there is no reason for her to evaluate future welfare according to some other legislator's preference.
-						\item If the model is interpreted as having a unitary government, there is only one decision maker.
-There is a fixed mapping from lobbying effort to the weight the decision maker places on the lobby's profits, and the realization of this weight changes with the realization of lobbying effort.
-Thus, predicting future lobbying effort, the decision maker knows what the realization of her weight on the lobby's profits will be and evaluates the incentive constraint accordingly.
-Expression (10) would be modified while the lobbying constraint is unchanged.
-					\end{itemize}
-		\end{itemize}
 
@@
 	\item Specialize to BS2005 when necessary (tariff cap example)
+	\item They don't like gov't choosing its own preferences. May just need to change wording.
 	\item Define $\ga$ before using it

### 2017-01-17 draft.tex
integrate more sign-posting to MRC2007

@@
 In my setup, the lobby is not active during the trade agreement negotiation; in order to to fit into the framework of \Textcite{mrc2007}, we can assume that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
+Of course, I consider only the case of perfectly immobile capital in the short run, whereas \Textcite{mrc2007} examines all possible levels of capital mobility.
 
-Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights.
+Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights in an environment with no ex-ante lobbying.
 
@@
 with the tariff only; their main results address the choice between tariffs and non-tariff barriers.} They show that the government will use the trade agreement to commit vis-\`{a}-vis the lobby as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying.
-Here, the government does not have an interest in maximizing joint surplus because it is unable to appropriate all the gains.\footnote{In the working paper version of \Textcite{mrc2007}, a similar result is demonstrated for the case of immobile capital when the government has all the bargaining power.}
+Here, the government does not have an interest in maximizing joint surplus because it is unable to appropriate all the gains.\footnote{In the working paper version of \Textcite{mrc2007}, a similar result is demonstrated for the case of immobile capital when the government has all the bargaining power and there is no ex-ante lobbying.}
 
@@
 The former, which is undoubtedly an important possibility in many institutional settings, is left as an important extension.\footnote{Note that the endogenous trade policy literature that involves trade agreements (e.g.
-\Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure while answering a different set of questions.
+\Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure.
 Results for the case of ex-ante lobbying are available from the author upon request.}
@@
 Thus we see that the governments can use a tariff cap to manipulate lobbying incentives so that political effort is neither too low or too high.
-Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.
+The fact that the applied tariffs are equal to the negotiated binding is reminiscent of \Textcite{mrc2007}.
+Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored.
+At the same time, in both models the government typically wants to use the trade agreement tariff cap to reduce lobbying below the lobby's optimum.
+In the model of this paper, it is because the government gets disutility from very high lobbying expenditure; in \Textcite{mrc2007}, it is because high tariff rates encourage inefficient over-investment in the import-competing industry.
+
+%In their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.
 

### 2017-01-17 to-do.tex
integrate more sign-posting to MRC2007

@@
 \end{center}
-
-\section{BS2005 vs. MRC2007}
-	\begin{itemize}
-		\item I like the way you use Bagwell and Staiger (2005) as your benchmark.
-But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.
-I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.
-In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.
-But your mechanisms seem quite different.
-They have capital mobility whereas your factors are specific to each sector.
-If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.
-What drives this feature in your model?
-Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.
-What is the relationship between lobby effort level and tariff policy?
-Is effort fixed first, after which the government sets its tariff?
-If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.
-It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.
-		\item DGH / MRC analogy from JIE$\_$revision
-			\begin{itemize}
-				\item After Prop1 is a good place to start to put references to MRC 2007
-
-				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007.
-Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored.
-In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps.
-The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff.
-There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff.
-Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement.
-A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement.
-From the ex-ante point of view, strong bindings are therefore preferable.
-Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
-			\end{itemize}
-		\item MRC07
-			\begin{itemize}
-				\item Need case of fixed capital ($z=0$)
-				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
-					\begin{itemize}
-						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period).
-Here, could be free trade but could be interior to get some rents.
-					\end{itemize}		
-			\end{itemize}
-	\end{itemize}
 
@@
 				\item Need to check this for sure in new section. I don't have cost to verification.
-		\end{itemize}
-
-\section{Dynamic use constraint}
-When would lobby exert effort to top up?
-\begin{itemize}
-	\item Do I want correlation between endogenous/exogenous parts? 
 		\end{itemize}
@@
 \end{enumerate}
+
+\section{Dynamic use constraint}
+When would lobby exert effort to top up?
+\begin{itemize}
+	\item Do I want correlation between endogenous/exogenous parts? 
+\end{itemize}
+
 \end{document}

### 2017-01-17 to-do.tex
reorganize to-do for intro rewrite

@@
 
+\section{Motivation}
+		\begin{itemize}
+			\item Intro doesn't reflect body of paper
+			\item What kinds of empirical questions might we ask with this model?
+			\item Articulate the question / paradox well
+			\item What would be different because we use this model?
+				\begin{itemize}
+					\item Make sure the story is clear: I establish baseline tariff cap case, repeated-game implications. We can already see a story of $\ldots$ emerge... Then add escape clause.
+				\end{itemize}
+			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper.
+However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government.
+This should be integrated with the previous object from the beginning.
+			\item I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
+			\item I also find the introduction too long.
+You could/should split the literature review that begins on page 4 to a separate section.
+And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it.
+Again, your main point seems to get lost into the discussion instead of standing out.
+			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics.
+		\end{itemize}
+
 \section{Lit review}
@@
 	\item Do I include papers like BS2001 (non-repeated) in lit review?
-	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
 	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
@@
 \end{itemize}
+
+\section{Costly state verification}
+		\begin{itemize}
+				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
+This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
+How does your approach relate to costly state verification?
+				\item I say that I'm going to provide a result along the lines of Beshkar 2010b. Same or different from costly state verification? Which is more helpful?
+				\item Need to check this for sure in new section. I don't have cost to verification.
+		\end{itemize}
+
+\section{Misc.}
+\begin{itemize}
+	\item Specialize to BS2005 when necessary (tariff cap example)
+	\item They don't like gov't choosing its own preferences. May just need to change wording.
+	\item Define $\ga$ before using it
+	\item If add back in assumption 1, add Ethier footnote
+	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
+	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
+	\item Ben: I'm wondering how you view the time-frame over which your model is set.
+On one hand a specific-factors set-up is normally associated with the short-run.
+But on the other hand you have an infinite time horizon which suggests very long run.
+	\item Clarify information structure
+	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together
+	\item Add ref to JIE R$\&$R to fn:krw ?
+	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
+	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
+\end{itemize}
+
+\section{Existence proofs?}
 
@@
 					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated.
-					\item I say I'm going to do this on Page 27. Need to decide and either do it or take this out.
 				\end{itemize}
@@
 		\end{itemize}
-
-\section{Costly state verification}
-		\begin{itemize}
-				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
-I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
-How does your approach relate to costly state verification?
-				\item I say that I'm going to provide a result along the lines of Beshkar 2010b. Same or different from costly state verification? Which is more helpful?
-				\item Need to check this for sure in new section. I don't have cost to verification.
-		\end{itemize}
-
-\section{Motivation}
-		\begin{itemize}
-			\item What it's missing: Well articulated question / paradox (I think it's just not well-articulated)
-			\item Intro doesn't reflect body of paper
-			\item What kinds of empirical questions might we ask with this model?
-			\item What would be different because we use this model?
-			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper.
-However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government.
-Is this aspect crucial for the point you want to make?
-If yes, it should be integrate with the previous object from the beginning.
-If not, it should be postponed a bit.
-			\item I also find the introduction too long.
-You could/should split the literature review that begins on page 4 to a separate section.
-And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it.
-Again, your main point seems to get lost into the discussion instead of standing out.
-			\item You need to think a bit more about the objective(s) of the paper and how best to achieve them.
-			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics.
-		\end{itemize}
-
-\section{Big picture of which changes do what}
-\vskip.2in
-		\begin{tabular}{|l|c|c|}
-			 & rigid & escape \\
-			BS2005  & on-schedule (truthtelling): trivial; & need cost  \\
-			 & off (repeated): optimal static if patient enough & \\
-			$\ga(e)$ & on: trivial& no need, but revisit side payments\\
-			 & off: mabye not b/c of lobby & \\
-			$\ga(e,s)$ & ignore for now & is $s$ verifiable? def not w/$\ga(s,e)$
-		\end{tabular}
-\vskip.2in
-\begin{itemize}
-	\item Make sure this story is clear in text: I establish baseline tariff cap case, then add escape clause. We can already see a story of $\ldots$ emerge...
-	\item Need to be clear when it is flexibility, enforcement, endogenous $\ga$ that screws things up. THEN need to bmake sure it comes out in paper
-	\item May want to ditch repeated game part...
-\end{itemize}
-
-\section{One good idea / do too much}
-		\begin{itemize}
-			\item You can have one good idea in a paper, the rest get lost
-			\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
-			\item The paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
-		\end{itemize}
-
-\section{Misc.}
-\begin{itemize}
-	\item Specialize to BS2005 when necessary (tariff cap example)
-	\item They don't like gov't choosing its own preferences. May just need to change wording.
-	\item Define $\ga$ before using it
-	\item If add back in assumption 1, add Ethier footnote
-	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
-	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
-	\item Ben: I'm wondering how you view the time-frame over which your model is set.
-On one hand a specific-factors set-up is normally associated with the short-run.
-But on the other hand you have an infinite time horizon which suggests very long run.
-	\item Clarify information structure
-	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together
-	\item Add ref to JIE R$\&$R to fn:krw ?
-	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
-	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
-\end{itemize}
-
-\section{Existence proofs?}
 
@@
 \end{center}
-\begin{enumerate}
-	\item Add export lobby. 
-		\begin{itemize}
-			\item Extension or appendix b/c want to make comparison to existing literature.
-		\end{itemize}
-\end{enumerate}
 
@@
 
+\section{Add export lobby}
+		\begin{itemize}
+			\item Extension or appendix b/c want to make comparison to existing literature.
+		\end{itemize}
+
+Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
+
 \end{document}

### 2017-01-17 draft.tex
first pass at revising intro

@@
 Developing a model that integrates both exogenous and endogenous political pressure and taking a general approach to the government objective function, I show that governments may want to use tariff caps both to force special interest groups to continue lobbying after a trade agreement is signed and to reduce the magnitude of that lobbying effort.
-The presence of endogenous politics can also destroy the ability of an escape clause as traditionally defined in the literature to provide flexibility in times of large negative political shocks when lobbies use the flexibility to seek rents.
+The presence of endogenous politics can also destroy the ability of a traditionally-defined escape clause to provide flexibility in times of large negative political shocks when lobbies use the flexibility to seek rents.
 This can explain why the use of WTO Safeguards are conditioned on measurable economic indicators.}
@@
 
+Note that \Textcite{mrc2007} do not have this result for the case of perfectly immobile capital in the published version of the paper, so this would seem to be an extension of their result.
+It is, but in a limited sense.
+In the working paper version of their paper, this result goes through for fixed capital when the government has all the bargaining power and there is no ex-ante lobbying.
+To truly extend those results, it must be shown that the results hold when lobbying is possible ex ante---that is, when special interests
+ are able to influence the formation of the trade agreement. This is indeed the case; the results with ex-ante lobbying are available from the author upon request.
+
 %In their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.

### 2017-01-17 intro.tex
first pass at revising intro

@@
 The assumption that political economy forces are entirely driven by exogenous forces is a rather drastic simplifying assumption.
-Do the predictions of our models hold up if this simplifying assumption is lifted?
+Do the predictions of our models hold up if this simplifying assumption is relaxed?
 
 One of the basic ideas that emerges from this literature is that in the presence of asymmetric information about the strength of the ex-post political economy shocks, it is often advantageous to grant governments a period of relief from trade commitments.
-That is, one would rather allow a short period of ``escape'' from the agreement rather than have the agreement abandoned forever because domestic political opposition is temporarily too strong to be resisted.
+That is, one would rather allow a short period of ``escape'' from the agreement rather than have the agreement violated because domestic political opposition is temporarily too strong to be resisted.
 
-This is an intuitively appealing story, but it turns out that the logic can break down in the presence of endogenous political pressure.
+This is an intuitively appealing story, but the logic can break down in the presence of endogenous political pressure.
 An escape clause allows a government to apply a higher tariff barrier when it experiences intense political pressure.
@@
 This is one example of a design question whose answers are sensitive to assumptions about the endogeneity of political pressure.
-In order to examine this question and others, I use a model that is comparable to \Textcite{bs2005}, adding an endogenously-determined element to their exogenously-determined political economy weights in a repeated-game setting.
-Because of its tractability, this model can address a wide range of institutional design questions while also allowing endogenous politics to be taken into account.\footnote{Repeated non-cooperative game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{coatesludema}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.
-\Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government has a separation-of-powers structure.}
+In order to examine this question and others, I use a simple model that is standard in the literature (e.g.
+\Textcite{bown2002}, \Textcite{bown2004}, \Textcite{bs2005}, \Textcite{martinvergote}, \Textcite{bagwell2009}, \Textcite{beshkar2010a}, \Textcite{ms2012a}, and \Textcite{ms2013}) and add an endogenously-determined element to the usual exogenously-determined political economy weights.
+Because of its tractability, this model can address a wide range of institutional design questions while also allowing endogenous politics to be taken into account in a repeated-game setting.\footnote{Repeated non-cooperative-game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{coatesludema}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.
+\Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government is non-unitary.}
 
@@
 This, perhaps, provides an explanation for the ubiquity of tariff caps.
-Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.\footnote{Both points extend results from \Textcite{mrc2007} to the case of perfectly immobile specific factors considered in this paper.
-To truly extend those results, it must be shown that the results hold when lobbying is possible ex ante---that is, when special interests are able to influence the formation of the trade agreement.
-The results with ex-ante lobbying are available from the author upon request.}
+Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.
 
-In order to make this point clear, I compare welfare under the standard \Textcite{baldwin}-style government objective function and a similar, fully-weighted version of the government's objective function.
-At issue is the fact that the standard government objective function is everywhere increasing in political pressure.
-The objective function use in the foundational work in the endogenous political economy of trade, \Textcite{gh94} and its extension to trade agreements in \Textcite{gh95} is also everywhere increase in political pressure.
-This leads inexorably to the conclusion that governments do not want to discourage special interest groups from exerting pressure except in special cases such as those demonstrated by \Textcite{mrc2007} and \Textcite{mitra}.
+In order to make this point clear, I compare political welfare under the standard \Textcite{baldwin}-style government objective function to political welfare under the government objective function I introduce that takes account of endogenous politics.
+In my formulation, the government is not necessarily better off as political pressure climbs higher and higher, while the standard government objective function is everywhere increasing in political pressure.
+\Textcite{gh94} and its extension to trade agreements in \Textcite{gh95} provide microfoundations for this government objective function that is everywhere increasing in political pressure, where a fixed weight of $1+a$ is attached to the surplus of groups who lobby, and $a$ to the surplus of those who don't lobby.\footnote{Note that $a$ is the weight the government places on social welfare.} This cannot microfound a flexible model as in \Textcite{longvousden} or in which weights are the result of shocks to any parameter other than the weight the government places on social welfare.
 
+More importantly, this functional form leads inexorably to the conclusion that governments do not want to discourage special interest groups from exerting pressure except in special cases such as those demonstrated by \Textcite{mitra} and \Textcite{mrc2007}.
 The modification to the government objective function I propose leads to the result that governments may indeed want to use trade agreements to reduce lobbying in many environments.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
@@
 The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role for which it was intended.
-If the purpose of contingent protection measures is actually to provide a political safety valve, this could provide an explanation for the infrequency with which safeguards are invoked under the WTO, as I argue that the WTO's escape clause has been designed to provide relief only from economic shocks.
+We miss important dynamics like this one when we assume political pressure is exogenous.
+Moreover, if the purpose of contingent protection measures is actually to provide a political safety valve, this could help explain the infrequency with which safeguards are invoked under the WTO, as I argue that the WTO's escape clause has been designed to provide relief only from economic shocks.
 
@@
 It is important to consider how these features interact with the question being asked.
-For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the shape of the objective function is not crucial, there may be no loss in employing the simple form.
-However, for other questions that directly involve the government maximizing with respect to political pressure, a more general treatment may be prudent.\footnote{Note that, although \Textcite{gh94} provide microfoundations for one particular \Textcite{baldwin}-style objective function, it is a very specific form with fixed weights in which $1+a$ is attached to the surplus of groups who lobby, and $a$ to the surplus of those who don't lobby.
-This cannot microfound a flexible model as in \Textcite{longvousden} in which weights depend on lobbying activity or in which weights are the result of shocks.}
+For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the shape of the objective function is not crucial, there may be no loss in employing a simple form with exogenous political pressure.
+However, for other questions in which lobbying incentives might be impacted or the government's desire for a commitment mechanism is central, a more general treatment may be prudent.
 

### 2017-01-17 intro.tex
finish revising inro

@@
 I first establish the baseline case with tariff caps and no escape clause.
-Here, I find that the presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significant ways compared with the case where political pressure is assumed to be exogenous.
-As in \Textcite{mrc2007}, I show that one of the uses of tariff caps is to incentivize the lobby to engage in the political process after the trade agreement is in place.
+Here, I find that the presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significant ways compared with the case where political pressure is assumed to be exogenous, and that it is harder to incentivize cooperation through repeated game incentives when political pressure is endogenous.
+As in \Textcite{mrc2007}, I show in this simpler framework that allows for easy comparison to the exogenous shocks case that one of the uses of tariff caps is to incentivize the lobby to engage in the political process after the trade agreement is in place.
 This, perhaps, provides an explanation for the ubiquity of tariff caps.
@@
 In order to make this point clear, I compare political welfare under the standard \Textcite{baldwin}-style government objective function to political welfare under the government objective function I introduce that takes account of endogenous politics.
-In my formulation, the government is not necessarily better off as political pressure climbs higher and higher, while the standard government objective function is everywhere increasing in political pressure.
+In my formulation, the government is not necessarily always better off as political pressure climbs higher and higher, while the standard government objective function is everywhere increasing in political pressure.
 \Textcite{gh94} and its extension to trade agreements in \Textcite{gh95} provide microfoundations for this government objective function that is everywhere increasing in political pressure, where a fixed weight of $1+a$ is attached to the surplus of groups who lobby, and $a$ to the surplus of those who don't lobby.\footnote{Note that $a$ is the weight the government places on social welfare.} This cannot microfound a flexible model as in \Textcite{longvousden} or in which weights are the result of shocks to any parameter other than the weight the government places on social welfare.
 
-More importantly, this functional form leads inexorably to the conclusion that governments do not want to discourage special interest groups from exerting pressure except in special cases such as those demonstrated by \Textcite{mitra} and \Textcite{mrc2007}.
+More importantly, this standard functional form leads inexorably to the conclusion that governments do not want to discourage political pressure except in special cases such as those demonstrated by \Textcite{mitra} and \Textcite{mrc2007}.
 The modification to the government objective function I propose leads to the result that governments may indeed want to use trade agreements to reduce lobbying in many environments.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
@@
 The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role for which it was intended.
-We miss important dynamics like this one when we assume political pressure is exogenous.
 Moreover, if the purpose of contingent protection measures is actually to provide a political safety valve, this could help explain the infrequency with which safeguards are invoked under the WTO, as I argue that the WTO's escape clause has been designed to provide relief only from economic shocks.
+We miss important dynamics like this one when we assume political pressure is exogenous as is done for tractability in almost all the literature that studies trading rules and institutions (see \Textcite{mrc2007} and \Textcite{lt} for notable exceptions).
+The main contribution of this paper is to provide a tractable framework for integrating lobbying incentives into these analyses.
+This framework can be used to study a wide range of questions, including dispute settlement design, optimal retaliation schemes and depth of integration among others.
+It's well-positioned to examine the interactions between exogenous shocks and lobbying incentives and should be useful for shedding light on questions surrounding how governments choose among protective measures over time.
 
-Through the lens of these examples, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective function have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.
+Through the lens of the examples considered in this paper, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective function have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.
 It is important to consider how these features interact with the question being asked.
 For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the shape of the objective function is not crucial, there may be no loss in employing a simple form with exogenous political pressure.
-However, for other questions in which lobbying incentives might be impacted or the government's desire for a commitment mechanism is central, a more general treatment may be prudent.
+However, for other questions in which lobbying incentives might be impacted or the government's desire for a commitment mechanism is central, a more general treatment seems prudent.
 

### 2017-01-17 to-do.tex
finish revising inro

@@
 
-\section{Motivation}
-		\begin{itemize}
-			\item Intro doesn't reflect body of paper
-			\item What kinds of empirical questions might we ask with this model?
-			\item Articulate the question / paradox well
-			\item What would be different because we use this model?
-				\begin{itemize}
-					\item Make sure the story is clear: I establish baseline tariff cap case, repeated-game implications. We can already see a story of $\ldots$ emerge... Then add escape clause.
-				\end{itemize}
-			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper.
-However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government.
-This should be integrated with the previous object from the beginning.
-			\item I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
-			\item I also find the introduction too long.
-You could/should split the literature review that begins on page 4 to a separate section.
-And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it.
-Again, your main point seems to get lost into the discussion instead of standing out.
-			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics.
-		\end{itemize}
-
 \section{Lit review}
 \begin{itemize}
-	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
-		\begin{itemize}
-			\item Bagwell and Staiger 1999 AER
-		\end{itemize}
 	\item Have to make connection to endogenous papers clearer
@@
 
-Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
-
+\section{Lit review}
+\begin{itemize}
+	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
+	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
+		\begin{itemize}
+			\item Bagwell and Staiger 1999 AER
+		\end{itemize}
+\end{itemize}
 \end{document}

### 2017-01-17 draft.tex
make absolutely essential changes to lit review

@@
 It still does not seem feasible to provide this flexibility at precisely the level that the shock impacts the government's preferences $\ga(s)$ if the WTO does not know the mapping from $s$ to $\ga$.
-What's more, this may provide an answer to why Safeguard measures have fallen out of use.
-If the real purpose is to provide an outlet for political pressure, I have demonstrated above that the current WTO-style Safeguards rules seem designed explicitly to discourage such rent-seeking behavior.
+What's more, this may provide an answer to why Safeguard measures have fallen out of use.\footnote{This is by no means the only proposed explanation.
+See \Textcite{bown2002s} and \Textcite{sykes2003} for example.} If the real purpose is to provide an outlet for political pressure, I have demonstrated above that the current WTO-style Safeguards rules seem designed explicitly to discourage such rent-seeking behavior.
 Thus if other measures such as Anti-Dumping Duties are more susceptible to manipulation, it seems natural that lobbies' efforts would migrate there.

### 2017-01-17 intro.tex
make absolutely essential changes to lit review

@@
 In order to examine this question and others, I use a simple model that is standard in the literature (e.g.
-\Textcite{bown2002}, \Textcite{bown2004}, \Textcite{bs2005}, \Textcite{martinvergote}, \Textcite{bagwell2009}, \Textcite{beshkar2010a}, \Textcite{ms2012a}, and \Textcite{ms2013}) and add an endogenously-determined element to the usual exogenously-determined political economy weights.
+\Textcite{bs2001}, \Textcite{bown2002}, \Textcite{bown2004}, \Textcite{bs2005}, \Textcite{martinvergote}, \Textcite{bagwell2009}, \Textcite{beshkar2010a}, \Textcite{ms2012a}, and \Textcite{ms2013}) and add an endogenously-determined element to the usual exogenously-determined political economy weights.
 Because of its tractability, this model can address a wide range of institutional design questions while also allowing endogenous politics to be taken into account in a repeated-game setting.\footnote{Repeated non-cooperative-game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{coatesludema}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.

### 2017-01-18 draft.tex
reduce abstract to 100 words, include titlepage

@@
-\documentclass[12pt]{article}
+\documentclass[12pt,titlepage]{article}
 
@@
 \begin{quote}
-{\small Political pressure is undoubtedly an important influence in the formulation of trade agreements and the institutions that govern them.
-Much of the literature that speaks to the design of trade policy institutions takes the political pressure that governments face as resulting from an exogenous, stochastic process.
-This paper shows that when political pressure arises endogenously, important results may be overturned and new insights into the motivation for features of the trade agreements we observe and rules of organizations such as the WTO come to light.
-Developing a model that integrates both exogenous and endogenous political pressure and taking a general approach to the government objective function, I show that governments may want to use tariff caps both to force special interest groups to continue lobbying after a trade agreement is signed and to reduce the magnitude of that lobbying effort.
-The presence of endogenous politics can also destroy the ability of a traditionally-defined escape clause to provide flexibility in times of large negative political shocks when lobbies use the flexibility to seek rents.
-This can explain why the use of WTO Safeguards are conditioned on measurable economic indicators.}
+{\small Much of the literature on the design of trade policy institutions takes the political pressure that governments face as exogenous.
+This paper integrates exogenous and endogenous political pressure into a single model, showing that the presence of endogenous political pressure overturns important results and provides insight into observed features of trade agreements as well as rules of organizations like the WTO.
+Endogenous politics destroys a traditionally-defined escape clause's ability to provide flexibility in the face of political shocks when lobbies use the flexibility to seek rents.
+This can explain why WTO Safeguard use is conditioned on measurable economic indicators.}
 \end{quote}

### 2017-01-18 draft.tex
proofread intro and model sections

@@
 
-$P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff.
-This gives rise to a terms-of-trade externality.
-Profits in a sector are increasing in the price of its good and also in the domestic tariff.
-This fact, combined with the assumptions on specific factor ownership, motivates political activity by import-competing lobbies.
-
 Note that $P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective importing country's tariff.
@@
 I next describe the politically-relevant actors, which are a government and import-competing lobby in each country.
-In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization.\footnote{Adding a pro-trade lobby for the exporting industry would serve to strengthen some results.} Each country's government can set trade policy either unilaterally or through a negotiated trade agreement and then choose an applied tariff that is potentially different from that agreed upon in a trade negotiation.
+In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization.
+Each country's government can set trade policy either unilaterally or through a negotiated trade agreement and then choose an applied tariff that is potentially different from that agreed upon in a trade negotiation.
 
-The stage-game timing is shown in Figure~\ref{fig:ext}.
+The stage-game timing is shown in Figure~\ref{fig:ext} for the home country.
 First, the governments cooperatively form a trade agreement.\footnote{See Section~\ref{sec:rigid} for a discussion of why lobbies are assumed not to be involved at the trade agreement formation stage.} After the trade agreement is concluded, any exogenous shock is realized.
@@
 	\end{center}
-	\caption{Extensive Form of the Stage Game\label{fig:ext}}
+	\caption{Extensive Form of the Stage Game (Home Country)\label{fig:ext}}
 \end{figure}
@@
 See Section~\ref{sec:objfcn} for a detailed discussion.} I will use ``$\ga$'' to represent the political economy parameter in a general sense, ``$\ga(s)$'' when treating the case of pressure arising from exogenous shocks only, and ``$\ga(e)$'' when referring to the case where political pressure  results from the rent-seeking behavior of lobbies only.
-I subtract the cost of lobbying effort from the government's objective function because if stands to reason that if the government cares about lobbying gross profits, it should also care about the costs.\footnote{Aside from making the government's optimum interior in $\ga$, taking account of the cost of lobbying changes the results that follow quantitatively but not qualitatively because there is no marginal impact on the government's decision-making regarding the optimal applied tariff.} This is often abstracted from in the closely-related literature, but it seems important to take into account in this environment of non-transferable utility and decreasing returns to lobbying.
-Note that there is no cost of lobbying effort in models that assume political pressure is exogenous only.
+I subtract the cost of lobbying effort from the government's objective function because if stands to reason that if the government cares about the lobby's gross profits, it should also care about the costs.\footnote{Aside from making the government's optimum interior in $\ga$ in most specifications, taking account of the cost of lobbying changes the results that follow quantitatively but not qualitatively because there is no marginal impact on the government's decision-making regarding the optimal applied tariff.} This is often abstracted from in the closely-related literature, but it seems important to take into account in this environment of non-transferable utility and decreasing returns to lobbying.
+Note that there can be no cost of lobbying effort in models that assume political pressure is exogenous only.
 
@@
 We can thus interpret the fundamental structure of preferences in a couple of ways.
-We can think of lobbying effort as a kind of persuasion than affects decision making of a single actor or we can think of lobbying effort as influencing the identify of the ex-post decision-maker.
+We can think of lobbying effort as a kind of persuasion than affects decision making of a single policy maker or we can think of lobbying effort as influencing the identify of the ex-post decision-maker.
 In either case, $\ga(s,e)$ is the weight the decision-maker places on the lobbying industry's profits ex-post, and the realization of this weight changes with the realization of the shock and lobbying effort even while the mapping---the preferences---do not.\footnote{The model also admits an interpretation in which the decision-making body is non-unitary, such as a large legislature.
@@
 \end{equation}
-I refer to this as the Nash tariff because it is the result of Nash equilibrium in the non-cooperative game between the governments.
-$\tau^N$ is increasing in $\ga$ and I assume the realization of $\ga$ is below the threshold need to ensure the second order condition is satisfied.
+I refer to this as the Nash tariff because it is the result of Nash equilibrium play in the non-cooperative game between the governments.
+$\tau^N$ is increasing in $\ga$ and I assume the realization of $\ga$ is below the threshold needed to ensure the second order condition is satisfied.
 This assumption does not in general rule out that the solution to Expression~\ref{eq:nash} could be the prohibitive tariff.
@@
 The implication is that the lobby can directly influence only the home government, and so the influence of one country's lobby on the other country's government occurs only through the tariffs selected.\footnote{cfr.
-\Textcite{gh95}, page 685.} $\ga(s,e)$ is assumed to be private information of the government, with specifics about its distributions given below.
+\Textcite{gh95}, page 685.} $\ga(s,e)$ is assumed to be private information of the government.
 This makes the environment one of asymmetric information, with asymmetries potentially coming from two sources: the exogenous shocks and the endogenous behavior of the special interest group.

### 2017-01-18 intro.tex
proofread intro and model sections

@@
-Much of the work on the political economy of trade agreements focuses on questions of the optimal design of trade agreements, trade agreement negotiations, and trade dispute settlement that arise in the presence of asymmetric information about shocks to an exogenous political economy parameter.
-The assumption that political economy forces are entirely driven by exogenous forces is a rather drastic simplifying assumption.
+Much of the work on the political economy of trade agreements focuses on questions regarding the optimal design of trade agreements, trade agreement negotiations, and trade dispute settlement that arise in the presence of asymmetric information about shocks to an exogenous political economy parameter.
+That political economy forces are entirely driven by exogenous forces is a rather drastic simplifying assumption.
 Do the predictions of our models hold up if this simplifying assumption is relaxed?
@@
 I first establish the baseline case with tariff caps and no escape clause.
-Here, I find that the presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significant ways compared with the case where political pressure is assumed to be exogenous, and that it is harder to incentivize cooperation through repeated game incentives when political pressure is endogenous.
+Here, I find that the presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significant ways compared with the case where political pressure is assumed to be exogenous, and that endogenous political pressure makes it harder to incentivize cooperation through repeated game incentives.
 As in \Textcite{mrc2007}, I show in this simpler framework that allows for easy comparison to the exogenous shocks case that one of the uses of tariff caps is to incentivize the lobby to engage in the political process after the trade agreement is in place.
-This, perhaps, provides an explanation for the ubiquity of tariff caps.
 Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.
@@
 In order to make this point clear, I compare political welfare under the standard \Textcite{baldwin}-style government objective function to political welfare under the government objective function I introduce that takes account of endogenous politics.
-In my formulation, the government is not necessarily always better off as political pressure climbs higher and higher, while the standard government objective function is everywhere increasing in political pressure.
-\Textcite{gh94} and its extension to trade agreements in \Textcite{gh95} provide microfoundations for this government objective function that is everywhere increasing in political pressure, where a fixed weight of $1+a$ is attached to the surplus of groups who lobby, and $a$ to the surplus of those who don't lobby.\footnote{Note that $a$ is the weight the government places on social welfare.} This cannot microfound a flexible model as in \Textcite{longvousden} or in which weights are the result of shocks to any parameter other than the weight the government places on social welfare.
+In my formulation, the government is not necessarily always better off as political pressure climbs higher and higher, while the \Textcite{baldwin}-style  government objective function is everywhere increasing in political pressure.
+\Textcite{gh94} and its extension to trade agreements in \Textcite{gh95} provide microfoundations for this \Textcite{baldwin}-style government objective function, where a fixed weight of $1+a$ is attached to the surplus of groups who lobby, and $a$ to the surplus of those who don't lobby.\footnote{Note that $a$ is the weight the government places on social welfare.} This cannot microfound a flexible model as in \Textcite{longvousden} or in which shocks to the political-economy weights come from shocks to any parameter other than the weight the government places on social welfare.
 
-More importantly, this standard functional form leads inexorably to the conclusion that governments do not want to discourage political pressure except in special cases such as those demonstrated by \Textcite{mitra} and \Textcite{mrc2007}.
-The modification to the government objective function I propose leads to the result that governments may indeed want to use trade agreements to reduce lobbying in many environments.\footnote{In the environment with only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
+More importantly, this standard functional form leads inexorably to the conclusion that governments do not want to discourage political pressure except in the presence of long-run inefficiencies such as in \Textcite{mitra} and \Textcite{mrc2007}.
+The modification to the government objective function I propose leads to the result that governments may indeed want to use trade agreements to reduce lobbying in many environments.\footnote{With only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
 
@@
 The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role for which it was intended.
-Moreover, if the purpose of contingent protection measures is actually to provide a political safety valve, this could help explain the infrequency with which safeguards are invoked under the WTO, as I argue that the WTO's escape clause has been designed to provide relief only from economic shocks.
+On the other hand, if the purpose of contingent protection measures is actually to provide a political safety valve, this analysis could help explain the infrequency with which safeguards are invoked under the WTO, as I argue that the WTO's escape clause has been designed to provide relief only from economic shocks.
+
 We miss important dynamics like this one when we assume political pressure is exogenous as is done for tractability in almost all the literature that studies trading rules and institutions (see \Textcite{mrc2007} and \Textcite{lt} for notable exceptions).
@@
 This framework can be used to study a wide range of questions, including dispute settlement design, optimal retaliation schemes and depth of integration among others.
-It's well-positioned to examine the interactions between exogenous shocks and lobbying incentives and should be useful for shedding light on questions surrounding how governments choose among protective measures over time.
+It's well-positioned to examine the interactions between exogenous shocks and lobbying incentives and will be useful for shedding light on questions surrounding how governments choose among protective measures over time.
 
-Through the lens of the examples considered in this paper, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective function have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.
+Through the lens of the examples considered in this paper, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective functions have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.
 It is important to consider how these features interact with the question being asked.
-For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the shape of the objective function is not crucial, there may be no loss in employing a simple form with exogenous political pressure.
+For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the government's objective function is not crucial, there may be no loss in employing a simple form with exogenous political pressure.
 However, for other questions in which lobbying incentives might be impacted or the government's desire for a commitment mechanism is central, a more general treatment seems prudent.
@@
 \Textcite{bc} find empirical support for this theory.
-\Textcite{beshkar2010b} compares the GATT escape clause to the WTO Safeguards agreement and shows that the DSB as a non-binding arbitrator can assist governments in self-enforcing their trade agreements.
+\Textcite{beshkar2010b} compares the GATT escape clause to the WTO Safeguards agreement and shows that the Dispute Settlement Body as a non-binding arbitrator can assist governments in self-enforcing their trade agreements.
 

### 2017-01-18 draft.tex
add references to my previous work per AEJ:micro guidance

@@
 \begin{document}
-\title{Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Renee Bowen, Chad Bown, Paola Conconi, David DeRemer, James Lake, Nuno Limao, Giovanni Maggi, Devashish Mitra, Peter Rosendorff, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at McGill, Rochester, Ryerson, Wesleyan, ETSG, InsTED, the Midwest International Trade Meetings, PEIO and SITE for very helpful comments and discussions.}}
+\title{Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Renee Bowen, Chad Bown, Paola Conconi, David DeRemer, James Lake, Nuno Lim\~ao, Giovanni Maggi, Devashish Mitra, Peter Rosendorff, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at McGill, Rochester, Ryerson, Wesleyan, ETSG, InsTED, the Midwest International Trade Meetings, PEIO and SITE for very helpful comments and discussions.}}
 \author{Kristy Buzard\thanks{Syracuse University, Economics Department. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard.}}
@@
 \label{sec:economic}
-I begin by describing the economic setting within which trade occurs.\footnote{Note this is the economic environment of \Textcite{gh95} specialized to two tradable goods with symmetry assumed for simplicity of exposition.} It is a three-good model with two countries: home (no asterisk) and foreign (asterisk).
+I begin by describing the economic setting within which trade occurs.\footnote{Note this is the economic environment of \Textcite{gh95} specialized to two tradable goods with symmetry assumed for simplicity of exposition.
+The basic economic and political structure is similar in many ways to that in \Textcite{buzard2013b}.} It is a three-good model with two countries: home (no asterisk) and foreign (asterisk).
 In the home country, the representative consumer maximizes

### 2017-01-18 intro.tex
add references to my previous work per AEJ:micro guidance

@@
 
-We miss important dynamics like this one when we assume political pressure is exogenous as is done for tractability in almost all the literature that studies trading rules and institutions (see \Textcite{mrc2007} and \Textcite{lt} for notable exceptions).
+We miss important dynamics like this one when we assume political pressure is exogenous as is done for tractability in almost all the literature that studies trading rules and institutions (see \Textcite{mrc2007}, \Textcite{lt} and \Textcite{buzard2013a} for exceptions).
 The main contribution of this paper is to provide a tractable framework for integrating lobbying incentives into these analyses.

### 2017-01-18 draft.tex
finish proofreading corrections

@@
 \label{sec:objfcn}
-There are several advantages to using the government objective function in Expression~\ref{eq:wel}, first of all tractability for analyzing institutional design questions.
+There are several advantages to using the government objective function in Expression~\ref{eq:wel}, first among them tractability for analyzing institutional design questions.
 I do not propose this alternative lightly, and it is closely related to objective functions that are well-studied in the literature.
@@
 \Textcite{gh94}'s well-known government objective function both restricts the returns to lobbying to be constant and assumes that utility is transferable between the government and lobby, preventing analysis of distributional questions.
-\Textcite{dgh97} point this out and proposes a more general form where the government objective is a function of a policy vector payments from each interest group (here $\left(\tau,\tau^*\right)$) and $e$, respectively).
+\Textcite{dgh97} point this out and propose a more general form where the government objective is a function of a policy vector and payments from each interest group (here $\left(\tau,\tau^*\right)$) and $e$, respectively).
 When there are no exogenous shocks, Expression~\ref{eq:wel} is therefore a special case of the \Textcite{dgh97} objective function.
@@
 See the \hyperlink{envelope}{Appendix} for details.
-
 An interior solution does not exist to this problem.
@@
 In my setup, the lobby is not active during the trade agreement negotiation; in order to to fit into the framework of \Textcite{mrc2007}, we can assume that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
-Of course, I consider only the case of perfectly immobile capital in the short run, whereas \Textcite{mrc2007} examines all possible levels of capital mobility.
+Of course, I consider only the case of perfectly immobile capital in the short run, whereas \Textcite{mrc2007} examine all possible levels of capital mobility.
 
-Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights in an environment with no ex-ante lobbying.
+Although \Textcite{mrc2007} assume the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights in an environment with no ex-ante lobbying.
 
@@
 Note that this is still consistent with the \Textcite{dgh97} form, although I have the lobby maximizing net profits instead of making a take-it-or-leave-it offer.
-} When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, the monotonicity in $\ga$ inherent in \ref{eq:baldwin} typically breaks down.
+} When the cost of lobbying effort is included in government welfare as in Equation~\ref{eq:wel}, the monotonicity in $\ga$ inherent in Expression~\ref{eq:baldwin} typically breaks down.
 For most specifications of the $\ga(e)$ function, there is an interior optimum.
-If a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin} with $e$ endogenized, the government will likely be interested in using the trade agreement to manipulate the lobby's behavior and therefore the political pressure that it experiences.
+If a government's objective function is closer to Expression~\ref{eq:wel} than Expression~\ref{eq:baldwin} with endogenous political pressure, the government will likely be interested in using the trade agreement to manipulate the lobby's behavior.
 By setting a tariff cap below the lobby's optimal tariff, the government can use the trade agreement to improve welfare by changing lobbying incentives and therefore the political pressure it encounters.\footnote{There are many candidates for the government objective function that are not monotonic in political pressure.
@@
 
-I do not argue that the government welfare function in Expression~\ref{eq:wel} perfectly represents the preferences of those who negotiate trade agreements.
-I only want to make the point that assumptions on the features of the government objective function can be important drivers of results on the role that trade agreements can play.
+I do not argue that the government welfare function in Expression~\ref{eq:wel} perfectly represents the preferences of policy makers.
+I simply want to make the point that assumptions on the features of the government objective function can be important drivers of results on the role that trade agreements can play.
 And this can potentially provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
@@
 Since a central goal here is to determine the impact of relaxing the assumption that political pressure is exogenously determined, I assume the latter to make the comparison most direct to this literature in which by definition there can be no endogenous political pressure at the trade agreement formation stage.
-The former, which is undoubtedly an important possibility in many institutional settings, is left as an important extension.\footnote{Note that the endogenous trade policy literature that involves trade agreements (e.g.
-\Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure.
-Results for the case of ex-ante lobbying are available from the author upon request.}
+Results for the former, which is undoubtedly an important possibility in many institutional settings, are available upon request from the author.\footnote{Note that the endogenous trade policy literature that involves trade agreements (e.g.
+\Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure.}
 
@@
 The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga}$ is small enough that the government's optimal tariff will be non-prohibitive.
-They also assume $\ga$ and $\ga^*$ are private information.
-They also assume linear supply and demand.} In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.
+They also assume $\ga$ and $\ga^*$ are private information and that supply and demand are linear.} In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.
 
@@
 In fact, any strong binding accomplishes this very feat, but is also \textit{unable} to generate lobbying activity because of its effect on lobbying incentives.
-Thus the introduction of endogenous political pressure provides an explanation of the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.
+Thus the introduction of endogenous political pressure provides an explanation for the prevalence of tariff caps: they are a a sort of carrot to encourage political contributions, or perhaps they should be seen as a stick with which to threaten the removal of protection if political support does not continue.
 
@@
 \subsubsection{Repeated Game}
-Because the governments are faced with asymmetric information, perfect public equilibrium (PPE) is the appropriate solution concept. Here attention is restricted to symmetric, stationary PPE.
+Because the governments are faced with asymmetric information in that lobbying effort and political shocks can't be observed across international borders, perfect public equilibrium (PPE) is the appropriate solution concept.
+Here attention is restricted to symmetric, stationary PPE.
 
-In order to establish an equilibrium, we must ensure that both the on-schedule (static) incentive constraint and the off-schedule (repeated-game) incentive constraint is satisfied.
+In order to establish an equilibrium, we must ensure that both the on-schedule (static) incentive constraint and the off-schedule (repeated-game) incentive constraint are satisfied.
 The former is trivially satisfied for the case of rigid bindings, so it is only the repeated-game constraint that must be considered here.
@@
 \label{sec:selfweak}
-When political pressure is exogenous, \Textcite{bs2005} perform a standard repeated-game prisoner's dilemma analysis and show that if governments are patient enough ($\de$ is sufficiently high), the optimal weak binding from the perfect external enforcement setting can always be sustained with repeated-game incentives.
+When political pressure is exogenous, \Textcite{bs2005} perform a standard repeated-game prisoner's dilemma analysis and show that if governments are patient enough, the optimal weak binding from the perfect external enforcement setting can always be sustained with repeated-game incentives.
 
@@
 In order for the government to have the incentive to abide by the agreement, this must be at least as large as the benefit to cheating.
-This is the current period increase in unilateral welfare from the tariff applied when breaking the agreement and is given on the righthand side of Inequality~\ref{exp:govincent}.
+The benefit is the current period increase in unilateral welfare from the tariff applied when breaking the agreement and is given on the righthand side of Inequality~\ref{exp:govincent}.
 
@@
 Here we have the possibility of a starkly different result from that under exogenous political pressure.
-Instead, the addition of a second incentive constraint can make it impossible to sustain the politically efficient tariffs.
+The addition of a second incentive constraint can make it impossible to sustain the politically efficient tariffs.
 
@@
 When taking into account the influence of $\de$ from the government's constraint on $e^B$, for a small part of the parameter space where $e^B$ must be very large, it's possible for the lobby's constraint to eventually loosen as $\de$ increases since $\pi_X(\tau^B) - e^B$ decreases to the right of $\tau^N$.
-
 In general, however, we cannot be guaranteed that even the most patient government can sustain cooperation at the politically optimal trade agreement level.
@@
 \label{sec:escape}
-I begin by following the literature in examining a trade agreement with an escape clause by adding a second negotiated binding to the agreement so that there is one (weak) binding for when political pressure is low and a higher, weak binding for exceptional circumstances of high levels of political pressure.
+I begin by following the literature in examining a trade agreement with an escape clause modeled as a second negotiated binding so that there is one (weak) binding for when political pressure is low and a higher, weak binding for exceptional circumstances of high levels of political pressure.
 Escape clauses are common features of trade agreements in practice, and when one models political pressure as exogenous, an escape clause seems attractive: in the presence of a particularly large negative political shock, being bound to a tariff designed for normal times would cause significant welfare losses to the politician.
@@
 But in the stark case under examination here with purely endogenous politics, an exceptionally high $\ga$ can only derive from higher effort exerted by the lobby.
-If one chooses an optimal tariff cap, a second, higher cap can only encourage excess lobbying that is sub-optimal from the point of view of the government.
+If one chooses an optimal tariff cap for normal times, a second, higher cap can only encourage excess lobbying that is sub-optimal from the point of view of the government.
 
@@
 Suppose that political pressure derives only from endogenous sources but an escape clause is to be implemented anyway.
-As \Textcite{bs2005} point out, in order for an escape clause to be useful to the government, it should be costly so that its use can be made incentive compatible.
-They suggest several avenues for introducing a cost for the use of the escape clause.
+In order for an escape clause to be useful to the government, it should be costly so that its use can be made incentive compatible.
+\Textcite{bs2005}  suggest several avenues for introducing a cost for the use of the escape clause.
 
@@
 Thus, in this case the strong binding does not create an incentive-compatibility-inducing cost, but again the central problem is that of lobbying incentives.
-This could explain, as Bagwell and Staiger point out, why the WTO does not incorporate strong bindings for the escape clause.
+This could explain, as \Textcite{bs2005} point out, why the WTO does not incorporate strong bindings for the escape clause.
 
@@
 In efficiency terms, this will only redistribute surplus from the trading partner who invokes the escape clause to the partner whose goods are targeted by the higher tariff.
-This is not entirely realistic as cash transfers are rarely observed in the WTO and other trade agreements; instead compensation is in the form of retaliatory tariffs---which are not efficient.\footnote{Note also that the WTO Safeguards agreement has a dynamic use constraint but removes the requirement for compensatory action in the first three years.}
+This is not entirely realistic as cash transfers are rarely observed in the WTO and other trade agreements; instead compensation is in the form of retaliatory tariffs, which are not efficient.\footnote{Note also that the WTO Safeguards agreement has a dynamic use constraint but removes the requirement for compensatory action in the first three years.}
 
-\Textcite{bs2005} show that, again, the transfer function must satisfy two incentive compatibility conditions: the static condition that makes it worthwhile to truthfully reveal one's true $\ga$, and the repeated-game incentive constraint required for self-enforcement.
-Given these requirements, their Proposition 5 establishes that an appropriate transfer scheme can make a trade agreement with an escape clause incentive compatible.
+Recall that the transfer function must satisfy two incentive compatibility conditions: the static condition that makes it worthwhile to truthfully reveal one's true $\ga$, and the repeated-game incentive constraint required for self-enforcement.
+Given these requirements, \Textcite{bs2005}'s Proposition 5 establishes that an appropriate transfer scheme can make a trade agreement with an escape clause incentive compatible.
 
@@
 \label{sec:escape2}
-When $\ga$ is exogenous, the purpose of the escape clause is clear: it improves welfare by allowing the government to apply a higher tariff when the realization of $\ga$ is particularly high.
+When $\ga$ is exogenous, the purpose of the escape clause is clear: it improves welfare in expectation by allowing the government to apply a higher tariff when the realization of $\ga$ is particularly high.
 In this case, the government would suffer significant welfare losses from abiding by the lower, normal tariff binding, and may find its interests are better served by abrogating the trade agreement if no escape were permitted.
@@
 In reality, it is likely that both exogenous shocks and endogenous forces contribute to the pressure to which policy makers are subject.
-Thus, the need for flexibility that derives from the purely exogenous case remains and we desire to see if it is possible to implement an escape clause in the face of the additional endogenous source of political pressure.
+Thus, the need for flexibility that derives from the purely exogenous case remains and we would like to know if it is possible to implement an escape clause in the face of the additional endogenous source of political pressure.
 
@@
 Imagine first that $s$ and its mapping $\ga(s)$ are completely unverifiable so as to make the analogy to the exogenous shocks examined above complete.
-There is a large parameter space over which an escape clause could not function in the way it is intended.
 When the realization of the shock is high, the government will report the high shock and apply the escape clause tariff.
@@
 
-From the point of view of maximizing the political welfare of the governments, this is only problematic if, as demonstrated in Section~\ref{sec:objfcn}, the governments' ex-ante welfare is reduced by excess lobbying.
+From the point of view of maximizing the political welfare of the governments, this is only problematic if the governments' ex-ante welfare is reduced by excess lobbying.
 
@@
 Given the inability of achieving the first best in this environment, what can be done?
-Unless political pressure that comes from a shock can be distinguished from political pressure that is purely rent-seeking, there appears to be little hope that the two can be disentangled through the investigation of a body such as the WTO's DSB.
-In fact, in its rulings the DSB does not appear to take into account political conditions at all.
+Unless political pressure that comes from a shock can be distinguished from political pressure that is purely rent-seeking, there appears to be little hope that the two can be disentangled through the investigation of a body such as the WTO's Dispute Settlement Body.
+In fact, in its rulings the Dispute Settlement Body does not appear to take into account political conditions at all.
 
@@
 This would appear to be the only way to prevent lobbies from exploiting the escape clause for uses for which it was not intended.
-However, once a shock has occurred, given that the DSB cannot distinguish between $\ga(s)$ and $\ga(e)$, determining the tariff that should be applied could only be done if (a) the DSB has the mapping $s \rightarrow \ga$ \textit{and} this strong form of separability holds.
-These conditions seem unlikely to be met in reality and we observe, again, that the DSB makes no attempt to determine the political parameter.
+However, once a shock has occurred, given that the Dispute Settlement Body cannot distinguish between $\ga(s)$ and $\ga(e)$, determining the tariff that should be applied could only be done if (a) the Dispute Settlement Body has the mapping $s \rightarrow \ga$ \textit{and} this strong form of separability holds.
+These conditions seem unlikely to be met in reality and we observe, again, that the Dispute Settlement Body makes no attempt to determine the political mapping.
 Without this information, it thus can only certify the legitimacy of a Safeguard measure and perhaps the level of tariff that is necessary to counteract $s$.
@@
 \begin{enumerate}
-	\item The government sets the trade agreement tariff $\tau_e^F$ cooperatively. I assume this is a weak binding. The subscript $F$ denotes the environment with flexibility to deal with shocks.
+	\item The government sets the trade agreement tariff $\tau_e^F$ cooperatively. I assume this is a weak binding. The superscript $F$ denotes the environment with flexibility to deal with shocks.
 	\item The exogenous shock $s$ is realized. Its level is assumed costlessly verifiable for simplicity.
@@
 Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau_e^R\right)$ so that in some cases the lobby's incentive constraint is violated.
-This would lead to the government's constraint being violated as well, creating a dispute.
+This would lead to the government's constraint being violated as well, creating a dispute when the government applies the tariff $\tau^B(\ga(s,e^F))$.
 
@@
 
-The addition of the lobby can make it harder to sustain cooperation and complicates the incentive compatibility problem associated with the use of escape clauses.
+The addition of the lobby can make it harder to sustain cooperation and complicates the incentive compatibility problem associated with both self-enforcement and the use of escape clauses.
 The temptation for lobbies to exploit the opportunity presented by an escape clause provides a justification for the WTO requirements that verifiable economic conditions are met in order to legally invoke these protections.
 
-I introduce a fully-weighted government utility function that is a slight modification of the standard Baldwin-style government objective function to demonstrate that it may be in even a politically-motivated government's interest to use a trade agreement to restrain political activity.
+I introduce a government utility function that incorporates endogenous political pressure into the standard Baldwin-style government objective function to demonstrate that it may be in even a politically-motivated government's interest to use a trade agreement to restrain political activity.
 The fact that the WTO Safeguards Agreement replaced compensation in the first three years of the invocation of a safeguard with a dynamic use constraint is consistent with the idea that the organization and the governments that constitute it are working to reduce rent seeking by lobbies.
@@
 In particular, here there is no scope for the government to encourage lobbying above the level that is optimal from the special interest group's point of view; adding additional lobbies may create insights to this effect.
-There is also strong potential to gain insights concerning very realistic dynamics between political and economics shocks and lobbying, especially insofar as they impact questions of enforcement and administered protection.
+There is also strong potential to gain insights concerning very realistic dynamics between political and economics shocks and lobbying, especially insofar as they impact questions of the choice between various forms of protection over time.
 

### 2017-01-19 draft.tex
add costly state verification reference

@@
 The new element compared to self-enforcement without the escape clause is that the government can cheat on both its tariff and its transfer payment.
-With no lobby, the Bagwell and Staiger result shows that this temptation can be overcome; the question is whether the lobby makes this more difficult.
+With no lobby, the Bagwell and Staiger result shows that this temptation can be overcome; the question is whether the presence of a lobby makes this more difficult.
 
 The answer is that it does not.
-Assuming the lobby represents a negligible share of the population, it does not bear the cost of the transfer.
-In fact, because it receives all the benefit of the tariff increase under the escape clause with none of the burden of paying for it, the lobby is much better off under the trade agreement with escape than the trade agreement without escape.
+Assuming the lobby represents a negligible share of the population, the lobby does not bear the cost of the transfer.
+In fact, because it receives all the benefit of the tariff increase under the escape clause with none of the burden of paying for it, the lobby is much better off under the trade agreement with escape and transfers than the trade agreement without escape.
 
@@
 Here, we have a mechanism that is incentive compatible despite the presence of lobbying.
-It makes the lobby exert effort in order to receive the higher level of protection under the escape clause just as in the other mechanism.
-And although the burden of the transfer falls on others in the economy, at least the trading partner is compensated.
+And although the burden of paying for the transfer falls on others in the economy, at least the trading partner is compensated.
 Here the government is truthfully revealing the level of political pressure it faces, but that pressure is created by endogenous forces that are intensified by the existence of the escape clause.
-The escape clause is incentive compatible, but as will be clarified in the next section, the spirit of the escape clause is not upheld.
+The escape clause is incentive compatible, but again the spirit of the escape clause is not upheld because the high binding is---always---used to accommodate pure rent-seeking instead of invoked only when necessary to accommodate a political shock.
 
@@
 Unless political pressure that comes from a shock can be distinguished from political pressure that is purely rent-seeking, there appears to be little hope that the two can be disentangled through the investigation of a body such as the WTO's Dispute Settlement Body.
-In fact, in its rulings the Dispute Settlement Body does not appear to take into account political conditions at all.
+`Costly state verification' as in \Textcite{bb} would help only if it could separately verify the part of $\ga$ that comes from shocks.
+This seems prohibitively difficult in practice and, in fact, the Dispute Settlement Body does not appear to take into account political conditions at all in its rulings.
 

### 2017-01-19 to-do.tex
make absolutely essential changes to lit review
cut down to-do for final proof-read
add costly state verification reference

@@
 
+\section{Existence proofs?}
+
+\newpage
+Current structure of paper (August 25, 2016)
+\begin{enumerate}
+	\item Introduction
+	\item Model
+	\item Rigid Tariffs with Endogenous Political Pressure
+		\begin{itemize}
+			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
+			\item[3.2] Self-Enforcing Trade Agreements
+				\begin{itemize}
+					\item[3.2.1] Repeated Game
+					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
+				\end{itemize}
+		\end{itemize}
+	\item Endogenous Political Pressure and the Escape Clause
+		\begin{itemize}
+			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
+			\item[4.2] Side payments: spirit not upheld, but IC
+			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
+			\item[4.4] EC for endogenous politics
+		\end{itemize}
+	\item Conclusion
+\end{enumerate}
+
+\newpage
+\begin{center}
+\textbf{\large{For Later}}
+\end{center}
+
+\section{Dynamic use constraint}
+When would lobby exert effort to top up?
+\begin{itemize}
+	\item Do I want correlation between endogenous/exogenous parts? 
+\end{itemize}
+
+\section{Add export lobby}
+		\begin{itemize}
+			\item Extension or appendix b/c want to make comparison to existing literature.
+		\end{itemize}
+
 \section{Lit review}
@@
 	\item Review Bown, Bagwell $\&$ Staiger
-	\item Do I include papers like BS2001 (non-repeated) in lit review?
-	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
-	 contributions `felonious'
+	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with contributions `felonious'
 	\item David DeRemer: Safeguards have dropped off in use around the world
@@
 		\end{itemize}
+	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
+	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
+		\begin{itemize}
+			\item Bagwell and Staiger 1999 AER
+		\end{itemize}
 \end{itemize}
 
-\section{Costly state verification}
+\section{Misc}
 \begin{itemize}
-				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?
-This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.
-How does your approach relate to costly state verification?
-				\item I say that I'm going to provide a result along the lines of Beshkar 2010b. Same or different from costly state verification? Which is more helpful?
-				\item Need to check this for sure in new section. I don't have cost to verification.
-		\end{itemize}
-
-\section{Misc.}
-\begin{itemize}
-	\item Specialize to BS2005 when necessary (tariff cap example)
-	\item They don't like gov't choosing its own preferences. May just need to change wording.
-	\item Define $\ga$ before using it
-	\item If add back in assumption 1, add Ethier footnote
-	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
-	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
 	\item Ben: I'm wondering how you view the time-frame over which your model is set.
@@
 But on the other hand you have an infinite time horizon which suggests very long run.
-	\item Clarify information structure
-	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together
-	\item Add ref to JIE R$\&$R to fn:krw ?
 	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
-	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
 \end{itemize}
-
-\section{Existence proofs?}
 
@@
 
-\newpage
-Current structure of paper (August 25, 2016)
-\begin{enumerate}
-	\item Introduction
-	\item Model
-	\item Rigid Tariffs with Endogenous Political Pressure
-		\begin{itemize}
-			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
-			\item[3.2] Self-Enforcing Trade Agreements
-				\begin{itemize}
-					\item[3.2.1] Repeated Game
-					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
-				\end{itemize}
-		\end{itemize}
-	\item Endogenous Political Pressure and the Escape Clause
-		\begin{itemize}
-			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
-			\item[4.2] Side payments: spirit not upheld, but IC
-			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
-			\item[4.4] EC for endogenous politics
-		\end{itemize}
-	\item Conclusion
-\end{enumerate}
-
-\newpage
-\begin{center}
-\textbf{\large{For Later}}
-\end{center}
-
-\section{Dynamic use constraint}
-When would lobby exert effort to top up?
-\begin{itemize}
-	\item Do I want correlation between endogenous/exogenous parts? 
-\end{itemize}
-
-\section{Add export lobby}
-		\begin{itemize}
-			\item Extension or appendix b/c want to make comparison to existing literature.
-		\end{itemize}
-
-\section{Lit review}
-\begin{itemize}
-	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
-	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
-		\begin{itemize}
-			\item Bagwell and Staiger 1999 AER
-		\end{itemize}
-\end{itemize}
 \end{document}

### 2017-01-19 draft.tex
clarify existence conditions

@@
 I refer to this as the Nash tariff because it is the result of Nash equilibrium play in the non-cooperative game between the governments.
-$\tau^N$ is increasing in $\ga$ and I assume the realization of $\ga$ is below the threshold needed to ensure the second order condition is satisfied.
-This assumption does not in general rule out that the solution to Expression~\ref{eq:nash} could be the prohibitive tariff.
 
@@
 where $\pi(\cdot)$ is the current-period profit and $\tau$ is the home country's tariff on the import good.
-%I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.
+
+$\tau^N$ is increasing in $\ga$ and, given Assumption~\ref{as:ga_c3}, the lobby's incentives imply that the realization of $\ga$ is such that the government's second order condition is satisfied.\footnote{See Page 28 of \Textcite{buzard2013b} for a proof of this point.
+In the case of exogenous shocks only, an assumption that the upper limit of the support of the shocks is bounded by the value of $\ga(s)$ that would lead to the prohibitive tariff is sufficent.
+Note that for an interior solution to the lobby's problem, we need a limit on the convexity of the profit function relative to $\tau(\ga(s,e))$.} This does not in general rule out that the solution to Expression~\ref{eq:nash} could be the prohibitive tariff.
 
@@
 
-Note that in this symmetric environment, this problem can be solved by applying the Nash bargaining solution where the disagreement point is the welfare resulting from the Nash equilibrium in the non-cooperative game.
+Note that in this symmetric environment, this problem is equivalent to solving the Nash bargaining solution where the disagreement point is the welfare resulting from the Nash equilibrium in the non-cooperative game.
 

### 2017-03-03 intro.tex
add costly state verification reference
add reference to Cole, Lake and Zissimos

--- 2017-01-18 intro.tex
@@
 
-We miss important dynamics like this one when we assume political pressure is exogenous as is done for tractability in almost all the literature that studies trading rules and institutions (see \Textcite{mrc2007} and \Textcite{lt} for notable exceptions).
-The main contribution of this paper is to provide a tractable framework for integrating lobbying incentives into these analyses.
+We miss important dynamics like this one when we assume political pressure is exogenous as is done for tractability in almost all the literature that studies trading rules and institutions (see \Textcite{mrc2007}, \Textcite{lt} and \Textcite{buzard2013a} for exceptions).\footnote{\Textcite{clz} also endogenize lobbying using a new approach, that of a contest model.} The main contribution of this paper is to provide a tractable framework for integrating lobbying incentives into these analyses.
 This framework can be used to study a wide range of questions, including dispute settlement design, optimal retaliation schemes and depth of integration among others.
@@
 \Textcite{beshkar2010b} compares the GATT escape clause to the WTO Safeguards agreement and shows that the Dispute Settlement Body as a non-binding arbitrator can assist governments in self-enforcing their trade agreements.
+\Textcite{ms2012a} study the optimal form of the trade agreement in the presence of renegotiation where compensation between governments is inefficient.
 

### 2017-07-10 contests.tex
add contests.tex with lit suggestions from Dan Kovenock

--- 2017-01-19 to-do.tex
@@
 \begin{center}
-PE Shocks To-Do list
+Contests literature
 \end{center}
 
-\section{Existence proofs?}
-
-\newpage
-Current structure of paper (August 25, 2016)
-\begin{enumerate}
-	\item Introduction
-	\item Model
-	\item Rigid Tariffs with Endogenous Political Pressure
+Suggestions from Dan Kovenock, July 2017
 \begin{itemize}
-			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
-			\item[3.2] Self-Enforcing Trade Agreements
+	\item Roberson and Kovenock survey on Contests with Multiple Battlefields, Oxford Handbook of the Economics of Peace and Conflict, Michele Garfinkel and Stergios Skaperdas eds.
+	\item Snyder, with looks at both the expected sum of battlefields won objective and the majoritarian objective.
+	\item Klumpp and Polborn: cases (with 7+  targets) when the equilibrium in the majoritarian case is only in nondegenerate mixed strategies.
+	\item Several by Fletcher and Slutsky 
 		\begin{itemize}
-					\item[3.2.1] Repeated Game
-					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
+			\item they may have more since these, you might look and see
 		\end{itemize}
-		\end{itemize}
-	\item Endogenous Political Pressure and the Escape Clause
-		\begin{itemize}
-			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
-			\item[4.2] Side payments: spirit not upheld, but IC
-			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
-			\item[4.4] EC for endogenous politics
-		\end{itemize}
-	\item Conclusion
-\end{enumerate}
-
-\newpage
-\begin{center}
-\textbf{\large{For Later}}
-\end{center}
-
-\section{Dynamic use constraint}
-When would lobby exert effort to top up?
-\begin{itemize}
-	\item Do I want correlation between endogenous/exogenous parts? 
-\end{itemize}
-
-\section{Add export lobby}
-		\begin{itemize}
-			\item Extension or appendix b/c want to make comparison to existing literature.
-		\end{itemize}
-
-\section{Lit review}
-\begin{itemize}
-	\item Have to make connection to endogenous papers clearer
-		\begin{itemize}
-			\item More on Coates $\&$ Ludema (2001)
-			\item Check out Ben and James new paper
-		\end{itemize}
-	\item Paola: How do I want to sell the paper?
-		\begin{itemize}
-			\item In relation to what literature?
-				\begin{itemize}
-					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
-					\item What literature \textit{doesn't}?
-				\end{itemize}
-			\end{itemize}
-	\item Review Bown, Bagwell $\&$ Staiger
-	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with contributions `felonious'
-	\item David DeRemer: Safeguards have dropped off in use around the world
-		\begin{itemize}
-			\item Look at U.S. vs. ROW (he sent papers)
-			\item Wouters and Zissimos (2016 WP): Appellate Body failed to give clear direction in cases in early 2000s as to when Safeguards can be applied. See Sykes (2003).
-		\end{itemize}
-	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
-	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
-		\begin{itemize}
-			\item Bagwell and Staiger 1999 AER
-		\end{itemize}
-\end{itemize}
-
-\section{Misc}
-\begin{itemize}
-	\item Ben: I'm wondering how you view the time-frame over which your model is set.
-On one hand a specific-factors set-up is normally associated with the short-run.
-But on the other hand you have an infinite time horizon which suggests very long run.
-	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
-\end{itemize}
-
-\section{More on escape clause / WEC}
-		\begin{itemize}
-			\item The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.
-In fact I was itching to get to that part right from when you discussed it in the introduction.
-				\begin{itemize}
-					\item The paper makes the extremely important point almost in passing.
-I would be interested to see a more detailed examination of how this works.
-It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.
-I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
-				\end{itemize}
-			\item `Withdrawal of equivalent concessions' (WEC).
-				\begin{itemize}
-					\item Zissimos (2007): a government \textbf{`chooses the severity of its own punishment'} by the extent of its initial deviation.
-I'm wondering whether your framework offers an answer for why WEC made sense as an approach to punishment.
-					\item Under WEC, it is worth deviating from the agreement in proportion to $s$ because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
-					\item But WEC might eliminate the incentive to respond to lobbying pressure $e$ because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
-						\begin{itemize}
-							\item \textbf{Think about how WEC may mitigate LOBBY'S incentives through reaction of gov't}
-							\item Ben doesn't have any need for escape; have to put WEC into my framework
-							\item Government feels $\ga$ the same whether it's elevated because of $s$ or $e$
-							\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
-							\item WEC should work against TOT, not PE shock?
-						\end{itemize}
-					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated.
-				\end{itemize}
-			\item Puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round, even though no retaliation for 3 yrs
-				\begin{itemize}
-					\item ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), %\url{C:\Users\Kristy\Dropbox\Research\PEshocks\Literature\2016}
-					\item Need to understand better what changed, both with safeguard and with other policy instruments
-					\item Note that safeguards agreement states that safeguard can only be at level necessary to remedy injury, and must liberalize as possible
-					\item \textbf{perhaps better able to respond to $\ga(e)$ before dispute settlement?}
-						\begin{itemize}
-							\item Lack of effective enforcement before WTO?
-						\end{itemize}
-				\end{itemize}
-			\item What does WTO \textit{really} want? To discourage rent-seeking lobbying but allow governments to escape when there's a real shock.
-				\begin{itemize}
-					\item There may be legitimate lobbying to communicate about the shock, so can't look at the presence of lobbying as a sufficient statistic
-				\end{itemize}
-			\item Does WEC help or hurt vs. grim trigger, $T$ period Nash reversion?
+	\item Alexander Robson (student of Stergios): multibattle contests with Tullock CSF. 
 \end{itemize}

### 2018-02-25 presentation.tex
finalize presentation for 2017 ASSAs
replace presentation with SITE version from 6/23/16

--- 2017-01-01 presentation.tex
@@
 %\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
-\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
-\date{January 6, 2017}
+\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
+\date{June 30, 2016}
 \begin{document}
@@
 \begin{frame}
-\includegraphics[height=2.75in, width=4.25in]{weight.jpg}
+\includegraphics[height=2.75in, width=4.25in]{weight.jpeg}
 \end{frame}

### 2018-02-25 to-do.tex
update to-do list
add ideas from IER referee reports to to-do.tex

@@
 
-\section{Existence proofs?}
-
-\newpage
-Current structure of paper (August 25, 2016)
-\begin{enumerate}
-	\item Introduction
-	\item Model
-	\item Rigid Tariffs with Endogenous Political Pressure
+\section{From IER Referee Reports}
 \begin{itemize}
-			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
-			\item[3.2] Self-Enforcing Trade Agreements
+	\item Is there a proposition to support each of my main points?
+	\item Make results more formal
 		\begin{itemize}
-					\item[3.2.1] Repeated Game
-					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
+			\item explicitly provide mappings from exogenous parameter regions to predictions about endogenous variables.
+For example, Proposition 2 simply indicates that a certain prediction ``may'' emerge without providing a detailed formal characterization.
+		\end{itemize}
+	\item Result 2 is interesting but not new
+	\item Clarifying points
+		\begin{itemize}
+			\item I would recommend saying on p. 7 what it means to “form a trade agreement.” I believe at this point in the paper the meaning is a tariff binding of some kind.
+			\item I would recommend formally defining the Wx function on p.
+9.
+This is used often in the sequel, and I was unclear as to whether this function includes the lobbying cost “-e” that is found in expression (1).
+			\item In expression (3) on p.
+10, it would be helpful to discuss the tariff function that is embedded in the profit function.
+Is this the “trade agreement”?
+By using this function, are we assuming that political pressure is observed here?
+I was a little uncertain as to exactly what was being represented here.
+			\item In expression (5) on p.
+11, where is the expectation over s and s*?
+Isn’t the trade agreement formed in anticipation of possible economic shocks?
+Since political pressure is privately observed, should the reader understand this as a complete-information benchmark of some kind?
+If so, is this the same “trade agreement” tariff that was mentioned previously?
+			\item On p.
+14, rigid tariffs are defined as corresponding to a situation in which governments have no provision for flexibility, and the section is entitled “Rigid tariffs with Endogenous Political Pressure.” But in the first subsection (on p.
+15) we are told that attention is restricted to the case of weak bindings throughout.
+Weak bindings seem to provide flexibility to apply lower tariffs, so they would not seem to be rigid tariffs, which makes the title to the section seem odd.
+Some additional clarity about terminology would be helpful here.
+			\item On p.
+17, I am not sure what τN in the third paragraph is.
+Do you have some specific pressure in mind that is suppressed?
+Or is this the tariff that is associated with eL?
+Or perhaps it was defined somewhere and I missed it.
+Similar remarks apply to expression (8) on p.
+20.
+What is the definition of eN there?
+Is it different from eL?
+			\item On p. 21, line 5, I don’t think you meant to say that the left hand side is decreasing in $\delta$, or maybe I am misunderstanding something.
+		\end{itemize}
+	\item Does costly state verification create threshold for endogenous $e$ even with verifiable shock conditionality?
+	\item Differentiate from BS2005: they have exogenous only, so can't get my insight on 'free pass' on page 2; any paper that gets it with exogenous shocks only are fundamentally different
+		\begin{itemize}
+			\item I'm showing something new: that endogenizing politics in these models \textit{matters}
+		\end{itemize}
+	\item Is my write-up very different from the presentation? What are my main points?
+	\item Clarify that it's not the $\gamma(e)$ that replaces Mitra / MRC inefficiencies; these are very complicated ways to get the government welfare function to be not everywhere increasing in lobbying effort.
+		\begin{itemize}
+			\item What I've done is provide a tractable way to get it
+			\item May need to add more: Mitra 1999, Mitra 2002, Krishna and Mitra 2005
+			\item Yes, relative to GH, mechanism by which greater lobbying activity leads government to give more weight to lobbying industry's profits is more of a black box.
+But GH functional form isn't very reasonable, and this is much better than what is usually done in the literature with exogenous shocks only.
+How would you put shocks into GH framework?
+				\begin{itemize}
+					\item Need to emphasize this being able to do shocks at same time as endogenous pressure
 				\end{itemize}
 		\end{itemize}
-	\item Endogenous Political Pressure and the Escape Clause
-		\begin{itemize}
-			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
-			\item[4.2] Side payments: spirit not upheld, but IC
-			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
-			\item[4.4] EC for endogenous politics
+	\item Clarify that by long-run inefficiencies I also mean the lobbying inefficiencies of Mitra
+	\item Do I really restrict attention to infinitely patient governments in Proposition 2?
 \end{itemize}
-	\item Conclusion
-\end{enumerate}
-
-\newpage
-\begin{center}
-\textbf{\large{For Later}}
-\end{center}
+\section{Existence proofs?}
 
@@
 
+\newpage
+Current structure of paper (August 25, 2016)
+\begin{enumerate}
+	\item Introduction
+	\item Model
+	\item Rigid Tariffs with Endogenous Political Pressure
+		\begin{itemize}
+			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
+			\item[3.2] Self-Enforcing Trade Agreements
+				\begin{itemize}
+					\item[3.2.1] Repeated Game
+					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
+				\end{itemize}
+		\end{itemize}
+	\item Endogenous Political Pressure and the Escape Clause
+		\begin{itemize}
+			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
+			\item[4.2] Side payments: spirit not upheld, but IC
+			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
+			\item[4.4] EC for endogenous politics
+		\end{itemize}
+	\item Conclusion
+\end{enumerate}
+
 \end{document}

### 2018-03-01 presentation.tex
significant revisions to presentation for Nebraska / new pitch

--- 2017-01-01 presentation.tex
@@
 
-\usepackage{appendixnumberbeamer}
+%\usepackage{appendixnumberbeamer}
 
@@
 \author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
-\date{January 6, 2017}
+\date{March 2, 2018}
 \begin{document}
@@
 \begin{enumerate}[<+->]
+\item Can trade agreements be used to manipulate domestic lobbying incentives?
+	\begin{itemize}
+		\item Government objective function
+	\end{itemize}
+\vskip.1in
 	\item When is endogenizing political pressure important for answering optimal design questions?
@@
 		\item Exogenous vs. endogenous politics
-	\end{itemize}
-	\vskip.1in
-\item Can trade agreements be used to manipulate domestic lobbying incentives?
-	\begin{itemize}
-		\item Government objective function
 	\end{itemize}
@@
 	\item endogenize politics into a standard model for studying TA design questions
+		\begin{itemize}
+			\item use this to examine gov't objective
+		\end{itemize}
 	\item carefully distinguish between dynamics induced by exogenous and endogenous politics for
@@
 \begin{itemize}[<+->]
-	\item Show that TAs may be used to manipulate domestic political actors (no long-run distortions)
-	\item For both tariff caps and escape clauses, outcomes are very different with endogenous politics
-	\item Demonstrate that (standard, theoretical) escape clause can't work in the presence of endogenous political pressure
+	\item TAs may be used to manipulate domestic political actors (no long-run distortions)
+	\item Simple modeling framework can capture results from models of both exogenous and endogenous politics
+		\begin{itemize}
+			\item For both tariff caps and escape clauses, endogenous politics changes outcomes dramatically
+		\end{itemize}
+	\item Standard, theoretical escape clause can't work in the presence of endogenous political pressure
 		\begin{itemize}
@@
 \begin{frame}
-\includegraphics[height=2.75in, width=4.25in]{weight.jpg}
+\includegraphics[height=2.75in, width=4.25in]{weight.jpeg}
+\end{frame}
+
+\begin{frame}{Comparison to the Literature}
+Others derive non-monotonicity in lobbying effort/tariff 
+\begin{itemize}
+	\item Maggi $\&$ Rodriguez-Clare (1998/2007): dynamic firm investment disortions
+	\item Mitra (2002): lobbies pay investment cost to form
+\end{itemize}
+
+\vskip.15in
+Here, I acknowledge the gov't objective function may be fundamentally non-monotonic
+\begin{itemize}
+	\item Achieve same results with simpler model
+	\item Endogenous politics in a wider range of questions
+	\item Can have both endogenous / exogenous at the same time $\Rightarrow$ unify the exogenous and endogenous politics literatures
+\end{itemize}
 \end{frame}
@@
 \begin{itemize}[<+->]
-	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): Negotiated weak bindings (a) are higher than those gov'ts would choose if they instead negotiated strong bindings and (b) imply that governments with low realizations of $\ga$ set their applied tariffs strictly below the bound level.
+	\item $\ga(s)$, i.e.
+exogenous: Negotiated weak bindings (a) are higher than those gov'ts would choose if they instead negotiated strong bindings and (b) imply that governments with low realizations of $\ga$ set their applied tariffs strictly below the bound level.
+		\begin{itemize}
+			\item Bagwell $\&$ Staiger 2005 result
+		\end{itemize}
 
-	\item $\ga$ endogenous: Governments will not set applied tariffs strictly below the bound level.
+	\item $\ga(e)$, i.e.
+endogenous: Governments will not set applied tariffs strictly below the bound level.
 They may use the weak tariff binding either to encourage and/or restrain endogenous political pressure.
+		\begin{itemize}
+			\item Maggi $\&$ Rodriguez-Clare (1998/2007) result
+		\end{itemize}
 
@@
 \begin{itemize}[<+->]
-	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): if governments patient enough ($\de$ high enough), optimal externally-enforced weak binding can be self-enforced
+	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): if governments patient enough (discount factor $\de$ high enough), optimal externally-enforced weak binding can be self-enforced
   \item $\ga$ endogenous: optimal externally-enforced weak binding may not be self-enforcing
@@
 \pause
-\hyperlink{repeated<1>}{\beamergotobutton{Repeated Game Intuition}}
+%\hyperlink{repeated<1>}{\beamergotobutton{Repeated Game Intuition}}
 \end{frame}
@@
 \end{comment}
+
+\begin{frame}[label=repeated]
+\frametitle{Repeated Game Intuition}
+\pause
+Legislature: break agreement if punishment not strong enough
+\pause
+\begin{itemize}
+	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
+\end{itemize}
+
+\vskip.1in
+\pause
+Lobby: solve for lowest effort ($\ov{e}_b$) that breaks this constraint
+\pause
+\begin{itemize}
+	\item pay $\ov{e}_b$ if it's less than gain from $T$ periods of trade-war profits
+\end{itemize}
+
+\pause
+\vskip.1in
+Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
+\pause
+\begin{itemize}
+	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
+	\pause
+	\item High tariffs, no lobbying, no trade disruptions
+\end{itemize}
+%\hyperlink{self<1>}{\beamergotobutton{Go Back}}
+\end{frame}
+
+\begin{frame}{When the world is more complicated...}
+Now suppose political pressure is a result of both endogenous and exogenous forces (i.e. $\ga(s,e)$):
+\pause
+\vskip.1in
+\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Water in the Bindings with $\ga(s,e)$}
+  Assume $\ga(s,e) = \ga(s) + \ga(e)$.
+In order for governments to set applied tariffs strictly below the weak binding, the low shock must be realized \textit{and} the lobby must not have the incentive to `top up' political pressure to the level associated with the binding.
+\end{beamerboxesrounded}
+
+\pause
+\begin{itemize}[<+->]
+	\item Can happen if gov't mis-judges lobby's incentives
+	\item In general, gov't prefers cap because lobby will `fill in' for low shock up to gov'ts optimal level of $\ga$
+\end{itemize}
+\end{frame}
 
@@
 	\item Assuming lower binding is set to maximize political welfare, escape clause encourages inefficiently high lobbying effort / protection
-	\item (Incentive compatibility still an issue, but often not the central one)
+	\item Incentive compatibility still an issue, but often not the central one
 		\begin{itemize}[<+->]
-			\item (If lobby's preferred tariff $\geq$ escape clause binding, gov't experiences high $\ga$, no need to lie)
+			\item If lobby's preferred tariff $\geq$ escape clause binding, gov't experiences high $\ga$, no need to lie
 		\end{itemize}
@@
 \begin{frame}{Conclusion}
-Taking into account endogenous political forces alongside exogenous ones...
+Taking into account endogenous political forces alongside exogenous ones in this simplified modeling framework
 \pause
 \begin{itemize}[<+->]
+		\item demonstrates that TAs can be used to discourage lobbing activity in general
+		\item can nest established results and provide new insights
+		\item can answer questions about optimal design of trading institutions more fully
+			\begin{itemize}
+				\item provides additional general explanation for tariff caps
 				\item helps explain the structure and enforcement of the WTO Safeguards measure
-		\item can help us think about optimal design of trading insitutions
-		\item demonstrates that TAs can be used to discourage lobbing activity in general
-		\item provides additional general explanation for tariff caps
+			\end{itemize}
 \end{itemize}
@@
 
-\appendix
-\begin{frame}[label=repeated]
-\frametitle{Repeated Game Intuition}
-\pause
-Legislature: break agreement if punishment not strong enough
-\pause
-\begin{itemize}
-	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
-\end{itemize}
-
-\vskip.1in
-\pause
-Lobby: solve for lowest effort ($\ov{e}_b$) that breaks this constraint
-\pause
-\begin{itemize}
-	\item pay $\ov{e}_b$ if it's less than gain from $T$ periods of trade-war profits
-\end{itemize}
-
-\pause
-\vskip.1in
-Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
-\pause
-\begin{itemize}
-	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
-	\pause
-	\item High tariffs, no lobbying, no trade disruptions
-\end{itemize}
-\hyperlink{self<1>}{\beamergotobutton{Go Back}}
-\end{frame}
-
 \end{document}

### 2018-03-02 presentation.tex
final presentation (handout) for UNL

@@
 \begin{enumerate}[<+->]
-\item Can trade agreements be used to manipulate domestic lobbying incentives?
+\item Can trade agreements (TAs) be used to manipulate domestic lobbying incentives?
 	\begin{itemize}
@@
 \vskip.1in
-	\item When is endogenizing political pressure important for answering optimal design questions?
+	\item What is the optimal design of various trade agreement properties?
 	\begin{itemize}
@@
 \begin{itemize}[<+->]
-	\item TAs may be used to manipulate domestic political actors (no long-run distortions)
+	\item TAs may be used to manipulate domestic political actors (even with no long-run distortions)
 	\item Simple modeling framework can capture results from models of both exogenous and endogenous politics
@@
 \pause
+\vskip.1in
 \begin{itemize}[<+->]
-	\item Standard \textit{except} weight on import-competing industry profits $\ga(s,e)$: 
+	\item Standard \textit{except} weight on import-competing profits: 
 		\begin{itemize}
@@
 \begin{frame}
-\frametitle{Domestic Political Pressure}
-Two potential sources
+\frametitle{Domestic Political Pressure: Two potential sources}
+
 \pause
@@
 			\item Shock directly to $\ga$ as in Bagwell $\&$ Staiger (2005): $\ga$, $\ga^*$ with CDF $H(\ga)$ on support $\left[\un{\ga},\ov{\ga}\right]$; or
-			\item Can take $\ga$ as a function of $s$: $\ga(s)$
+			\item Can take $\ga$ as a function of shock $s$: $\ga(s)$
 		\end{itemize}
@@
 
-	\item If weights must sum to 1, welfare no longer monotonic in $\ga$
+	\item If lobbying effort subtracted as cost from $W$, welfare no longer monotonic in $\ga$
+	\pause
+		\begin{itemize}
+			\item If weights must sum to 1, welfare also not monotonic in $\ga$
+		\end{itemize}
 \end{itemize}
@@
 \pause
-Must set tariff at or below specified level (aka tariff cap)
+Must set tariff at or below specified level (aka weak binding)
 \pause
@@
 \begin{frame}{When the world is more complicated...}
-Now suppose political pressure is a result of both endogenous and exogenous forces (i.e. $\ga(s,e)$):
+Now suppose $\ga(s,e)$, i.e. political pressure is a result of both endogenous and exogenous forces:
 \pause
@@
 	\item Can happen if gov't mis-judges lobby's incentives
-	\item In general, gov't prefers cap because lobby will `fill in' for low shock up to gov'ts optimal level of $\ga$
+	\item In general, gov't prefers cap because lobby will `fill in' for low shock up to gov's optimal level of $\ga$
 \end{itemize}
@@
 
-\begin{frame}{An Escape Clause for Endogenous Politics}
+\begin{frame}{An Escape Clause for a Complicated World}
 \pause
@@
 \section{Conclusion}
-
+\subsection{}
 \begin{comment}

### 2018-03-02 presentation.tex
UNL presentation with pauses

@@
-%\documentclass{beamer} 
-\documentclass[handout]{beamer} 
+\documentclass{beamer} 
+%\documentclass[handout]{beamer} 
 \usetheme{Ilmenau}
@@
 	\item If lobbying effort subtracted as cost from $W$, welfare no longer monotonic in $\ga$
-	\pause
 		\begin{itemize}
@@
 \begin{frame}{Comparison to the Literature}
+\pause
 Others derive non-monotonicity in lobbying effort/tariff 
+\pause
 \begin{itemize}
 	\item Maggi $\&$ Rodriguez-Clare (1998/2007): dynamic firm investment disortions
+	\pause
 	\item Mitra (2002): lobbies pay investment cost to form
@@
 \vskip.15in
+\pause
 Here, I acknowledge the gov't objective function may be fundamentally non-monotonic
-\begin{itemize}
+\pause
+\begin{itemize}[<+->]
 	\item Achieve same results with simpler model
@@
 \end{itemize}
-\pause
+%\pause
 %\hyperlink{repeated<1>}{\beamergotobutton{Repeated Game Intuition}}
@@
 
+\vskip.2in
 \pause
-\vskip.2in
-
 If $\ga$ is only endogenous, escape clause causes problems, provides no benefits

### 2018-03-03 draft.tex
recompile without weird title page for web posting
re-write abstract; replace intro with relevant slides from UNL talk

@@
-\documentclass[12pt,titlepage]{article}
+%\documentclass[12pt,titlepage]{article} %for submission to AEJ-micro
+\documentclass[12pt]{article}
 
@@
 \begin{document}
-\title{Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Renee Bowen, Chad Bown, Paola Conconi, David DeRemer, James Lake, Nuno Lim\~ao, Giovanni Maggi, Devashish Mitra, Peter Rosendorff, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at McGill, Rochester, Ryerson, Wesleyan, ETSG, InsTED, the Midwest International Trade Meetings, PEIO and SITE for very helpful comments and discussions.}}
+\title{Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Renee Bowen, Chad Bown, Paola Conconi, David DeRemer, James Lake, Nuno Lim\~ao, Giovanni Maggi, Devashish Mitra, Peter Rosendorff, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at McGill, Rochester, Ryerson, Wesleyan, ETSG, InsTED, the Midwest International Trade Meetings, PEIO, SITE, and the University of Nebraska for very helpful comments and discussions.}}
 \author{Kristy Buzard\thanks{Syracuse University, Economics Department. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard.}}
 \date{\vskip-.1in \today}
+
 \maketitle
@@
 \begin{center} {\bf Abstract} \end{center}
-
 \begin{quote}
-{\small Much of the literature on the design of trade policy institutions takes the political pressure that governments face as exogenous.
-This paper integrates exogenous and endogenous political pressure into a single model, showing that the presence of endogenous political pressure overturns important results and provides insight into observed features of trade agreements as well as rules of organizations like the WTO.
-Endogenous politics destroys a traditionally-defined escape clause's ability to provide flexibility in the face of political shocks when lobbies use the flexibility to seek rents.
+{\small While most of the literature on the design of trade agreements and trading institutions takes the political pressure governments face to be exogenous, this paper endogenizes politics in a standard model for studying trade policy design questions.
+This permits a careful examination of the impact of assumptions about the policy-maker's objective function on conclusions about the role that trade agreements play.
+With this simple modeling framework, one can carefully distinguish between the dynamics induced by exogenous political shocks and endogenous incentives of political actors, unifying these two distinct strands of the literature.
+The modeling framework can also provide fuller answers to trade policy design questions by elucidating the interactions between exogenous and endogenous political forces.
+With applications to tariff caps and escape clauses, the modeling framework shows that important insights are missed if attention is restricted to exogenous political shocks.
+Most notably, endogenous politics destroys a traditionally-defined escape clause's ability to provide flexibility in the face of political shocks when lobbies use the flexibility to seek rents.
 This can explain why WTO Safeguard use is conditioned on measurable economic indicators.}
@@
 \section{Introduction}
-\input{intro}
+
+\begin{enumerate}
+\item Can trade agreements (TAs) be used to manipulate domestic lobbying incentives?
+	\begin{itemize}
+		\item Government objective function
+	\end{itemize}
+\vskip.1in
+	\item What is the optimal design of various trade agreement properties?
+	\begin{itemize}
+		\item Exogenous vs. endogenous politics
+	\end{itemize}
+\end{enumerate}
+
+Political Economy of Trade Institutions\\
+
+With a few exceptions, TA design literature has taken political economy forces to be exogenous. I:
+\begin{itemize}
+	\item endogenize politics into a standard model for studying TA design questions
+		\begin{itemize}
+			\item use this to examine gov't objective
+		\end{itemize}
+	\item carefully distinguish between dynamics induced by exogenous and endogenous politics for
+		\begin{itemize}
+			\item base case with tariff caps
+			\item tariff caps with escape clause
+		\end{itemize}
+	\item examine escape clause design when both exogenous and endogenous forces are present
+\end{itemize}
+
+Results
+\begin{itemize}
+	\item TAs may be used to manipulate domestic political actors (even with no long-run distortions)
+	\item Simple modeling framework can capture results from models of both exogenous and endogenous politics
+		\begin{itemize}
+			\item For both tariff caps and escape clauses, endogenous politics changes outcomes dramatically
+		\end{itemize}
+	\item Standard, theoretical escape clause can't work in the presence of endogenous political pressure
+		\begin{itemize}
+			\item Points to real-world design of WTO Agreement on Safeguards
+			\item May explain why escape clause has fallen out of use
+		\end{itemize}
+\end{itemize}
+
+Objective Function
+
+Restraining Political Pressure through TAs
+\begin{itemize}
+	\item Will TA be used to discourage lobbying? Depends on how gov't welfare varies in $\ga$
+	\item With standard Baldwin-style objective function, welfare always increases with $\ga$
+\[
+  W = \mathit{CS}_X(\tau) + \ga \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
+\]
+
+	\begin{itemize}
+		\item Isomorphic to `Protection for Sale' objective function
+	\end{itemize}
+
+	\item If lobbying effort subtracted as cost from $W$, welfare no longer monotonic in $\ga$
+		\begin{itemize}
+			\item If weights must sum to 1, welfare also not monotonic in $\ga$
+		\end{itemize}
+\end{itemize}
+
+Comparison to the Literature\\
+
+Others derive non-monotonicity in lobbying effort/tariff
+
+\begin{itemize}
+	\item Maggi $\&$ Rodriguez-Clare (1998/2007): dynamic firm investment disortions
+	\item Mitra (2002): lobbies pay investment cost to form
+\end{itemize}
+
+\vskip.15in
+
+Here, I acknowledge the gov't objective function may be fundamentally non-monotonic
+
+\begin{itemize}
+	\item Achieve same results with simpler model
+	\item Endogenous politics in a wider range of questions
+	\item Can have both endogenous / exogenous at the same time $\Rightarrow$ unify the exogenous and endogenous politics literatures
+\end{itemize}
+
+%\input{intro}
 

### 2018-03-03 draft.tex
rewrite intro for submission to Midwest trade

--- 2017-01-19 draft.tex
@@
 \begin{center} {\bf Abstract} \end{center}
-
 \begin{quote}
-{\small Much of the literature on the design of trade policy institutions takes the political pressure that governments face as exogenous.
-This paper integrates exogenous and endogenous political pressure into a single model, showing that the presence of endogenous political pressure overturns important results and provides insight into observed features of trade agreements as well as rules of organizations like the WTO.
-Endogenous politics destroys a traditionally-defined escape clause's ability to provide flexibility in the face of political shocks when lobbies use the flexibility to seek rents.
+{\small While most of the literature on the design of trade agreements and trading institutions takes the political pressure governments face to be exogenous, this paper endogenizes politics in a standard model for studying trade policy design questions.
+One can use this simple modeling framework to distinguish between the dynamics induced by exogenous political shocks and endogenous incentives of political actors, unifying these two strands of literature.
+The modeling framework can also provide fuller answers to trade policy design questions by elucidating the interactions between exogenous and endogenous political forces.
+Applications to tariff caps and the escape clause show that important insights are missed if attention is restricted to exogenous political shocks.
+Most notably, endogenous politics destroys a traditionally-defined escape clause's ability to provide flexibility in the face of political shocks when lobbies use the flexibility to seek rents.
 This can explain why WTO Safeguard use is conditioned on measurable economic indicators.}
@@
 \section{Introduction}
-\input{intro}
+The intensity and ubiquity of lobbying is a salient feature of trade policy making.
+The United States--Korea Free Trade Agreement was famously delayed by over four years, mainly by concerns from the U.S.
+auto industry.
+Over the last two decades in the United States alone, anywhere between 700 and 1200 entities reported lobbying on trade issues each year.
+\Textcite{bt} find that total annual lobbying expenditure on trade issues in 1999-2001 when Lobbying Disclosure Act data first became available was about $\$$200 million.
+Campaign contributions are harder to attribute cleanly to a specific issue but are likely an order of magnitude larger.
+
+Despite the obvious importance of interest groups in the trade policy-making process, relatively few papers have examined lobbying incentives in trade agreement and trade institution design.
+This is understandable, as models of endogenous politics following \Textcite{gh94} are complex.
+It can be quite challenging to endogenize politics along these lines while also taking account of complex institutional features.
+Thus we have very important (e.g.
+\Textcite{mitra}, \Textcite{mrc2007}) but also very few models in this literature in which political pressure is modeled as endogenous.
+
+However, it seems likely that many questions of trade agreement and trade institution design are impacted in important ways by decisions taken by political pressure groups, who themselves respond to features of the designed environment.
+In order to take account of endogenous politics when asking questions of trade policy design, I use a simple model that is standard in the literature (e.g.
+\Textcite{bs2001}, \Textcite{bown2002}, \Textcite{bown2004}, \Textcite{bs2005}, \Textcite{martinvergote}, \Textcite{bagwell2009}, \Textcite{beshkar2010a}, \Textcite{ms2012a}, and \Textcite{ms2013}) and add an endogenously-determined element to the usual exogenously-determined political economy weights.
+This framework is more tractable than those in the \Textcite{gh94} tradition and allows the incorporation of endogenous politics when asking a rich set of institutional design questions, including in repeated-game settings.\footnote{Repeated non-cooperative-game models of trade agreements have been considered by \Textcite{mcm86,mcm89}, \Textcite{dixit1987}, \Textcite{bs1990, bs1997a, bs1997b, bs2002}, \Textcite{kovthurs}, \Textcite{maggi99}, \Textcite{coatesludema}, \Textcite{ederington}, \Textcite{rosendorff}, \Textcite{bagwell2009}, and \Textcite{park}.
+\Textcite{buzard2013a} features a repeated-game model similar in spirit to the model under consideration here but focusing on questions of optimal punishments when the government is non-unitary.}
+
+After introducing the modeling framework, I use it to carefully examine the assumptions implicit in the government objective functions that are most commonly used in the literature.
+The \Textcite{baldwin}-style  government objective function is everywhere increasing in political pressure.
+\Textcite{gh94} and its extension to trade agreements in \Textcite{gh95} provide microfoundations for this \Textcite{baldwin}-style government objective function, where a fixed weight of $1+a$ is attached to the surplus of groups who lobby, and $a$ to the surplus of those who don't lobby.\footnote{Note that $a$ is the weight the government places on social welfare.} This cannot microfound a flexible model as in \Textcite{longvousden} or in which shocks to the political-economy weights come from shocks to any parameter other than the weight the government places on social welfare.
+
+More importantly, this standard functional form leads inexorably to the conclusion that governments do not want to discourage political pressure except in the presence of some kind of dynamic inefficiency such as in \Textcite{mitra} or \Textcite{mrc2007}.
+The modification to the government objective function I propose leads to the result that governments may indeed want to use trade agreements to reduce lobbying in a much wider set of circumstances than those identified in the previous literature.\footnote{With only one lobby in each country, the governments will be unable to encourage lobbying in excess of that which is optimal from the lobby's point of view.} Examining this alternative welfare function in combination with endogenous lobbying can provide a bridge between the theoretical literature and the claims of trade policy practitioners that an important role of trade agreements is to rein in protectionist pressure.
+
+Having looked carefully at the government's objective function, I then use the model to carefully distinguish between the dynamics induced by exogenous and endogenous politics for two example policies: tariff caps and escape clauses.
+I show that the model captures results from both the exogenous politics (i.e.
+\Textcite{bs2005} and the endogenous politics (i.e.
+\Textcite{mrc2007}) literatures on tariff caps.
+The presence of endogenous political pressure changes the optimal choice of tariff bindings within a trade agreement in quite significant ways compared with the case where political pressure is assumed to be exogenous.
+Additionally, endogenous political pressure makes it harder to incentivize cooperation through repeated game incentives.
+As in \Textcite{mrc2007}, I show in this simpler framework that allows for easy comparison to the exogenous shocks case that one of the uses of tariff caps is to incentivize the lobby to engage in the political process after the trade agreement is in place.
+Combined with the idea that governments can use tariff caps to restrain endogenous political pressure, a story emerges in which governments employ trade agreements to carefully manipulate lobbying incentives in order to maximize their political objectives.
+
+The changes induced by endogenous politics are even more stark for the escape clause.
+Here one finds that the standard, theoretical escape clause can't work in the presence of endogenous political pressure.
+The idea of the escape clause is that it is often advantageous to grant governments a period of relief from trade commitments when there is asymmetric information about the strength of ex-post political economy shocks.
+That is, one would rather allow a short period of ``escape'' from the agreement rather than have the agreement violated because domestic political opposition is temporarily too strong to be resisted.
+Unfortunately, the logic breaks down in the presence of endogenous political pressure.
+An escape clause allows a government to apply a higher tariff barrier when it experiences intense political pressure.
+When political pressure is chosen endogenously by special interest groups \textit{and} the government gets a free pass at the WTO whenever it feels sufficient political pressure from domestic interest groups, those interest groups have a strong incentive to exert the required level of pressure regardless of the underlying state of the world, eviscerating the escape clause.
+I argue on this basis that if design choices might affect lobbying incentives, one should carefully consider when it is appropriate to assume that political pressure is given exogenously.
+
+When we turn to another attractive feature of the model---that one can study the simultaneous effects and interactions of exogenous and endogenous political forces---we derive a clear explanation for why the conditions for invoking the WTO Safeguards measure rely on purely observable economic variables, and why the level of protection governments can choose when invoking a Safeguard is related to the injury imposed by the shock and not its political impact.
+The model formalizes the intuition that if the WTO's Dispute Settlement Body actually followed the procedures suggested by the literature---that is, it enforced commitments based on signals of the political pressure experienced by governments---pressure groups would simply exert more pressure and the Safeguard clause could not play the escape-providing role for which it was intended.
+On the other hand, if the purpose of contingent protection measures is actually to provide a political safety valve, this analysis could help explain the infrequency with which safeguards are invoked under the WTO, as I argue below that the WTO's escape clause appears to have been designed to provide relief only from economic shocks.
+
+We miss important dynamics like this one when we assume political pressure is exogenous as is done for tractability in almost all the literature that studies trading rules and institutions (see \Textcite{mrc2007}, \Textcite{lt} and \Textcite{buzard2013a} for exceptions).\footnote{\Textcite{clz} also endogenize lobbying using a new approach, that of a contest model.} The main contribution of this paper is to provide a tractable framework for integrating lobbying incentives into these analyses.
+While others (cfr.
+\Textcite{mrc2007}, \Textcite{mitra}) have formulated models in which some dynamic inefficiency induces a government objective that is non-monotonic in lobbying effort, I can achieve results with the same flavor in a more tractable framework.
+The framework can therefore be used to study a wider range of questions, including dispute settlement design, optimal retaliation schemes and the depth of integration among others.
+It can also examine exogenous shocks to political weights and lobbying incentives simultaneously, unifying the two distinct literatures.
+
+Through the lens of the examples considered in this paper, one can see that modeling choices concerning both the source of political pressure and the form of governments' political objective functions have important impacts on questions of the design of and motives behind trade agreements as well as the institutions that help to enforce them.
+It is important to consider how these features interact with the question being asked.
+For some inquiries, such as those that are the focus of \Textcite{ms2011, ms2012a} in which the impact of the political economy parameter on the government's objective function is not crucial, there may be no loss in employing a simple form with exogenous political pressure.
+However, for other questions in which lobbying incentives might be impacted or the government's desire for a commitment mechanism is central, a more general treatment seems prudent.
+
+\textbf{A few additional comments on the literature}.
+Tariff caps have received particular attention in the literature that takes political pressure to be exogenous.
+The basic result on the political efficiency of tariff caps established in \Textcite{bs2005} has been extended in numerous directions.
+\Textcite{hms} show that the basic logic concerning tariff caps is unaltered in a model with contracting costs and multiple policy instruments, while \Textcite{ls} show that the result holds when fines are allowed as punishments, although the politically efficient tariffs may not be achievable when enforcement of the punishment is taken into consideration.
+
+\Textcite{ab2012} consider that the weight that tariff revenue receives in the government welfare function may be private information and show that tariff caps remain optimal under certain conditions, while \Textcite{ab2013} show that the result goes through when the \Textcite{bs2005} model is generalized to include monopolistic competition as well as more general payoff and distribution function.
+\Textcite{bb} extend the theory to an environment with asymmetric country sizes with costly state verification and show that tariff caps and escape clauses can both emerge endogenously in optimal trade agreements.
+\Textcite{bbr} and \Textcite{nos} provide evidence on the relationship between tariff caps, binding overhang and market power that confirms the predictions of the terms of trade theory.
+
+On escape clauses, \Textcite{bs1990} is similar in spirit to \Textcite{bs2005}.
+Here, there is no private information and trade volume shocks in the presence of self-enforcement constraints are behind the need for an escape clause.
+\Textcite{bc} find empirical support for this theory.
+\Textcite{beshkar2010b} compares the GATT escape clause to the WTO Safeguards agreement and shows that the Dispute Settlement Body as a non-binding arbitrator can assist governments in self-enforcing their trade agreements.
+\Textcite{ms2012a} study the optimal form of the trade agreement in the presence of renegotiation where compensation between governments is inefficient.
+
+In the next section, I present the model and an in-depth discussion of the political welfare function of the government.
+Section~\ref{sec:rigid} contains an analysis of tariff caps without the flexibility of escape and introduces the repeated-game environment.
+Section~\ref{sec:escape} explores questions regarding the design of the escape clause.
+Section~\ref{sec:concl} concludes.
+
+%\input{intro}
 

### 2024-08-08 draft.tex
comment out addbibresource lines that threw warnings
Add .bib file from my hard drive and recompile

@@
 %dell laptop
-\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
+%\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
+
+%Surface Book 2
+\addbibresource{tradeagreements.bib}
 
 %campus computer
-\addbibresource{C:/Apps-SU/Dropbox/Research/xBibs/tradeagreements.bib}
+%\addbibresource{C:/Apps-SU/Dropbox/Research/xBibs/tradeagreements.bib}
 \renewcommand{\newunitpunct}{,}
@@
 One can use this simple modeling framework to distinguish between the dynamics induced by exogenous political shocks and endogenous incentives of political actors, unifying these two strands of literature.
-The modeling framework can also provide fuller answers to trade policy design questions by elucidating the interactions between exogenous and endogenous political forces.
+The modeling framework can also provide fuller answers to trade policy design questions by clarifying the interactions between exogenous and endogenous political forces.
 Applications to tariff caps and the escape clause show that important insights are missed if attention is restricted to exogenous political shocks.
